### 2014-12-24 paper.tex
adicionados arquivos iniciais

% ****** Start of file aipsamp.tex ******
%
%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
%   Version 4.1 of REVTeX, October 2009
%
%   Copyright (c) 2009 American Institute of Physics.

% Use this file as a source of example code for your aip document.
% Use the file aiptemplate.tex as a template for your document.
\documentclass[%
 aip,
 jmp,%
 amsmath,amssymb,
%preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{grffile}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{multirow}
\usepackage{color} % for the notes
\usepackage{etex}
\reserveinserts{58}
\usepackage{morefloats}

\maxdeadcycles=1000

\begin{document}

\preprint{XXXXX (preprint)}

%\title[Evolution of interaction networks]{On the evolution of interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
%\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominance of measures and activity statistics}% Force line breaks with \\

\author{Renato Fabbri}%
 \homepage{http://ifsc.usp.br/~fabbri/}
 \email{fabbri@usp.br}
  \affiliation{ 
S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP)%\\This line break forced with \textbackslash\textbackslash
}%
%
%\author{Camilo Akimushkin Valencia}
%  \homepage{http://ifsc.usp.br/}
%  \email{akimushkin@ifsc.usp.br}
%  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
%

\author{Vilson Vieira da Silva Junior}
  \homepage{http://automata.cc/}
  \email{vilson@void.cc}
  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\

\author{Ricardo Fabbri}
  \homepage{http://www.lems.brown.edu/~rfabbri/}
  \email{rfabbri@iprj.uerj.br}
 \altaffiliation{
Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
}%Lines break automatically or can be forced with \\

\author{Deborah Christina Antunes}
  \homepage{http://lattes.cnpq.br/1065956470701739}
  \email{deborahantunes@gmail.com}
  \altaffiliation{
Curso de Psicologia, Universidade Federal do Cer\'a (UFC)
}%Lines break automatically or can be forced with \\

\author{Marilia Mello Pisani}
  \homepage{http://lattes.cnpq.br/6738980149860322}
  \email{marilia.m.pisani@gmail.com}
 \altaffiliation{
Centro de Ciências Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
}%Lines break automatically or can be forced with \\

\author{Luciano da Fontoura Costa}
  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
  \email{ldfcosta@gmail.com}
  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\

\author{Osvaldo N. Oliveira Jr.}
  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
  \email{chu@ifsc.usp.br}
 \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
    This article reports a minimal and general characterization of interaction networks evolution. Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants, 2) a sound and stable classification of vertex: peripheral, intermediary and hub sectors, 3) composition of basic measures into components with greater dispersion. While time patterns of activity are not obvious, participant activity follow concentrations expected by scale-free networks. Comparison with ideal Erd\"os-R\'enyi network with the same number of edges and vertexes revealed as a sound criterion  for distinguishing sectors on the networks. Principal components in basic measures spaces revealed interesting and regular patterns of independence and dispersion. This includes a ranking of measures that most contribute to dispersion: 1) degree and strength measures, 2) symmetry related quantization, and 3) clusterization. Results suggested typologies for these networks and participants. Further work include considerations of text production, psychoanalysis inspired typologies, participatory democracy exploitation of observed properties, and better visualization support for network evolution.
\end{abstract}

\pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
\keywords{complex networks, social network analysis, pattern recognition, statistics}%Use showkeys class option if keyword
\maketitle

\begin{quotation}
% `The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social. The fact that human society has been up to now divided into classes affects more than the external relations of men. The marks of social repression are left within the individual soul.' 
% 
`The conception of personality structure is the
best safeguard against the
inclination to attribute persistent trends in the
individual to something
"innate" or "basic" or "racial" within him. The
Nazi allegation that natural, biological traits decide the total being of a person
would not have been such
a successful political device
had it not been possible to point to numerous
instances of relative fixity in human behavior and to
challenge those who
thought to explain them on any basis other than a biological one.'

\emph{- Adorno et al, 1969, p. 747}
\end{quotation}

\section{\label{sec:into}Introduction}
The present work is aimed at finding common characteristics among (email) interaction networks. This includes observations along time, which imply network evolution, a field that has received dedicated attention from the research community
for more than a decade~\cite{barabasiEvo,newmanEvolving}.

While significant measures will depend on the model and system characteristics~\cite{newmanStru, newmanWeight},
this work considers only directed, weighted and human interaction networks. Undirected and unweighted representation of such networks is also seen in the literature and can be obtained by simplification~\cite{GMANE2}.

Text production and typologies of online participants benefit from results here presented~\cite{rcText,rcTipo}. 
Although all networks considered originated from email lists,
 coherence with literature suggests that results hold for a more general class of interaction networks,
 such as observed in online platforms (LinkedIn, Facebook, Twitter).

\subsection{Related work}
 Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}. Moreover, selected exceptions are repported in this section.
The evolution considered in the present article is characterized by a constant number of messages, which is also present in literature, but was less explored to date.

 The evolution of interaction networks was addressed with focus on community evolution, a work that ignores the direction of edges~\cite{barabasiEvo}. Two topologically different networks are repported to emerge, depending on the frequency of interactions, which present a generalized power law or an exponencial connectivity distribution~\cite{barabasiTopologicalEv}. 
Recent work on email lists consider an isolated snapshot in order to verify or draw hypothesis. In such, free-scale properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.

 Such results are in accordance with phenomena observed in this work and linguistic characterization is being described in a deriving article~\cite{rcText}.

\section{Data description}
    \subsection{Email lists and messages}
Email list messages were obtained from
the GMANE email archive~\cite{GMANE}, which consists of more than 20,000 email lists and more than 130,000,000 messages~\cite{GMANEwikipedia}. These lists cover a variety of topics, mostly technology-related. It can be described as a corpus with metadata of its messages, like send time, place, sender name, sender email adderss etc.
GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}. Appendix~\ref{scripts} is dedicated scripts for gathering and processing GMANE email messages.

    \subsection{Chosen dataset}
The four lists below were selected for their diversity, easing initial observance of natural and general properties.
\begin{itemize}
    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE archive.}. Participantes are from different countries, and English is the language used the most. More technical and less active version of LAU. Abbreviated LAD from now on.
    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE archive.}. Dominated by specialized computer programmers. Participants are from different countries, and English is the language used the most. Abbreviated as CPP from now on.
    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE archive.}. Dominated by Brazilian activists and digital culture interests. Participants are mostly Brazilians, and Portuguese is the most used language, although Spanish and English usage is commonplace. Abbreviated MET from now on.
    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE archive.}. Dominated by participants with hybrid artistic and technological interests. Participantes are from different countries, and English is the language used the most. Abbreviated as LAU from now on.
\end{itemize}

 The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages exposed in Table~\ref{geralListas}.

\begin{table}
  \centering
  \begin{tabular}{|l|c|c|c|c|c|}\hline
list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
LAU  & Jun/29/2003 & Jul/23/2005 & 1183 & 3373 & 5 \\
LAD  & Jun/30/2003 & Oct/07/2009 & 1268 & 3113 & 4 \\
MET  & Ago/01/2005 & Mar/07/2008 & 492  & 4607 & 23 \\
CPP  & Mar/13/2002 & Aug/25/2009 & 1052 & 4506 & 7 \\ \hline
  \end{tabular}
  \caption{Columns $date_1$ and $date_M$ have first and last messages dates from the 20,000 messages considered.
$N$ is the number of participants (number of different email addresses).
$\Gamma$ is the number of threads (count of messages without antecedent).
$\overline{M}$ is messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
MET notably has the fewer participants and the larger number of threads.
This relation holds for each pair of the lists considered: as the number of participants increases,
 the number of threads decreases.}
  \label{geralListas}
\end{table}

\section{Characterization methods}
After immersion in the data and on appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological ellaborations of networks and participants.
Each of these methods are described bellow, as other handy alternatives.

    \subsection{Temporal activity statistics}
Number of messages along time with respect to seconds, minutes, hours, days of the week or the month, and months of the year. These were displayed as tables in Appendix~\ref{tabTime}, results are outlined in Section~\ref{constDisc}.

    \subsection{Interaction network}\label{intNet}
Regarding literature~\cite{bird,newmanCommunityDirected,newmanCommunity2013}, interaction networks might be both weighted or unweighted, both directed or undirected. Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, and undirected weighted and undirected unweighted).
More precisely, the networks repported are erected as follows: a direct response from participant B to participant A forms an edge from A to B, as information went from A to B. The reasoning is: if B wrote a response to A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$. Inverting edge direction yields the status network, as B read the message and considered what A wrote worth of responding, giving status to A, thus $B\rightarrow A$. This article uses the information network described above and depicted in Figure~\ref{formationNetwork}. Edges in both directions are allowed. Each time an interaction occurs, one is added to edge weight. Self-loops were regarded as non-informative and discarded. This networks are described as exhibiting free-scale and small world properties, as expected for a social network~\cite{bird}.

\begin{figure}[hb]
    \centering
    \includegraphics[width=0.5\textwidth]{figs2/criaRede_}
    \caption{Formation of interaction network from email messages. Each vertex represents a participant. If participant B replyes participant A, that is regarded as evidence that B received information from A. Multiple messages add ``weight'' to directed edge. Further details are in Section~\ref{intNet}.}
    \label{formationNetwork}
\end{figure}

Previous messages on the thread create directed edges from their author to the observed message's author. Edges can be created from all antecedent messages on the message-response thread. Is this work, only immediate predecessors are linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$. Potential interpretations for this weaker connection are usually common sense, such as: double length, half weight or with one more ``obstacle''. This suggests the adoption of other centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.

        \subsubsection{Sectioning network in periphery, intermediary and hubs classes}\label{sectioning}

\begin{figure}[hb]
    \centering
    \includegraphics[width=0.5\textwidth]{figs2/fser}
    \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The later has more
        intermediary, as the former has more peripheral and hubs. Sections are
        given by the two intersections of the distributions $k_1$ and $k_2$. Characteristic degrees
    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the three sections considered (periphery, intermediary and hubs).}
    \label{fig:setores}
\end{figure}

Because of social networks' tendency to have scale-free distribution of properties,
one can compare it to an Erd\"os-R\'enyi random graph and consider peripheral, intermediary and hub
sectors~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
The degree distribution $\widetilde{P}(k)$ of an ideal
scale-free network $\mathcal{N}_f$ with $N$ vertexes and $z$ edges, has less
average degree nodes when compared with the distribution $P(k)$ of an Erd\"os-R\'enyi
random graph with the same number of vertexes and edges:

\begin{equation}\label{criterio}
    \widetilde{P}(k)<P(k) \Rightarrow \text{k is intermediary degree}
\end{equation}

If $\mathcal{N}_f$ is directed and has self-loops, the probability
of the presence of an unknown edge is $p=\frac{z}{N(N-1)}$, where $N(N-1)$ is the maximum number of edges for a network with $N$ vertexes, directed edges and without selfloops.
A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertexes and edges, and thus the same probability $p$ for the presence of an edge, will have degree $k$ with probability:

\begin{equation}
    P(k)=\binom{2(N-1)}{k}p^k(1-p)^{2(N-1)-k}
\end{equation}

The lower degree fat tail constitute the border
vertexes or peripheral sector. 
The higher degree fat tail is the hub sector.

The arguments behind this classification are: 1) vertexes so connected that they are virtually inexistent in networks connected at pure chance, specially without preferential attachment, are correctly associated to hubs sector. Vertexes with very few connections, which are way more abundant than expected by pure chance, are correctly associated to periphery. Degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in free-scale phenomena, are correctly associated to intermediary vertexes.

To assure statistical validity, bins can be chosen to span the average of $\eta$ gaps between measure values. Thus, each bin, starting at degree $k_i$, spans $\Delta_i=\frac{\sum_{i=1}^{\eta}(k_{i+1}-k_i)}{\eta}$ values, with borders $k_i$ and $k_i+\Delta_{i}$. This changes equation~\ref{criterio} to:

\begin{equation}\label{criterio2}
    \sum_{x=k_i}^{k_i+\Delta_i} \widetilde{P}(x) < \sum_{x=k_i}^{k_i+\Delta_i} P(x) \Rightarrow \text{i is intermediary}
\end{equation}

If instead strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}$ the average weight of an edge and $s_i$ the vertex strength. For in and out degrees and strengths, comparisons should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$. Results of these segmentations are discussed in subsection~\ref{subsec:pih}.

As a further refinement of the network segmentation, compound criteria is used for classification of vertex, considering all measures: total, in and out degree and strength. After a careful inspection of possible combinations, these were abbreviated to six:

\begin{itemize}
    \item Exclusivist criteria: vertex are only classified if the class is the same with respect to all measures. In this case, total vertex classified (usually) does not reach 100\%, which is specified by a black line in Appendix~\ref{figures}.
    \item Inclusivist criteria: vertex has class given by any of the measures. In this case, a vertex can have more than one class and the fraction of class attribution beyond total number of vertexes is also represented by a black line in Appendix~\ref{figures}.
    \item Exclusivist cascade: hubs are only hubs if classified as hub with respect to all measures. Intermediary are the vertexes classified as intermediary or hub with respect to all measures. Vertexes left are regarded as peripheral vertex.
    \item Inclusivist cascade: hubs are vertexes classified as hubs by any of the measures. From the vertexes left, if any is classified as intermediary by any measure, than it is intermediary. The rest of the vertexes are peripheral.
    \item Exclusivist externals: hubs have unanimous classification with respect to all measures. Of vertexes left, peripheral vertexes are the ones classified as hub or peripheral by simple criterion. The rest represent intermediary sector.
    \item Inclusivist externals: hubs are vertexes classified as hubs with respect to any measure. From vertexes left, if a vertex is classified as peripheral with respect to any measure, than it is peripheral. The rest is regarded intermediary sector.
\end{itemize}

These compound criteria, and reduction of possibilities to them, can be formalized in strict mathematical terms. This was considered out of the scope of the present article.

Results from applying this classification method is further reported in Section~\ref{subsec:pih}.

        \subsubsection{Prevalence of topological measures}
This article
restricts this analisys to a small selection of the most basic
measures of each vertex:

\begin{itemize}
    \item Degree $d_i$: number edges linked to node $i$.
    \item In-degree $d_i^{in}$: number of edges ending at node $i$.
    \item Out-degree $d_i^{out}$: number of edges departing from node $i$.
    \item Strength $s$: sum of weights of all edges linked to node $i$.
    \item In-strength $s_i^{in}$: sum of weights of all edges ending at node $i$.
    \item Out-strength $s_i^{out}$: sum of weights of all edges departing from node $i$.
    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked. Standard clustering coefficient for undirected graphs was used.
    \item Betweenness centrality $bt_i$: fraction of geodesics that contain the node $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
\end{itemize}

In order to capture asymmetries in the activity of participants, the following metrics were introduced (see subsection~\ref{prevalence}):

\begin{itemize}
    \item asymmetry of note $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
    \item mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$. Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise. $|J_i|$ is the number of neighbors of vertex $i$.
    \item standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
    \item disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
    \item mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
    \item standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
\end{itemize}

   \subsection{Evolutive observations}
Evolution of network is observed within a fixed number of messages (window size: $ws$) that shifts in the message timeline.
 All $ws=$50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000 were used. Within a same $ws$, the number vertex and edges vary in time, as do other network characteristics. Further work should deepen inspection of measure interdependence, this article holds to measures in Section~\ref{measures}.

        \subsubsection*{Visualization of network evolution}
In refining hypotheses, visualization of the network was crucial. Animations, image galleries and an online gadgets were made~\cite{animacoes,galGMANE,appGMANE}. Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research. Furthermore, stable aspects of measures prominance along time are captured through mean and standard deviation (see Section~\ref{}). Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{}.

    \subsection{Typological deepening}
There are other ways to split a network. To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius\footnote{Radius is the minimum maximum distance to all nodes. Equivalently, the radius is the minimum eccentricity.}. 
In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter\footnote{Diameter is the maximum geodesic on the network.}. Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery. Interestingly, in the email networks analyzed, with these criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceed 93\% of the nodes~\cite{networkx}.

Models of human dynamics can be used to predict and classify activity. In this case, agent activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time. Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with of bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}. Emails are reported as having a heavy tailed distribution with $\alpha=1$, together with web browsing and library loans~\cite{barabasiHumanDyn}.

Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories. 
Concerning empirical research,
Theodor Adorno was a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
personality traces\footnote{Some of them related to Nazism adoption, antisemitism and potential fascists.}, sometimes depicted as an authoritarian syndrome.
%Influenced by Social and Psychoanalytic Theories, Adorno et all applied a questionnaire to individuals, from which they reached a position in the the ``F Scale'', to verify etnocentric, conservatory and antidemocratic trends~\cite{adorno}. From psychoanalitic interviews and the F Scale, they derived a typology, which gathers prejudice-inclined traces in personality. Both, low and high scores are considered with prejudicial traces. This typology has nine authoritarian types, the six types with high score in the F Scale: surface resentment, conventional, authoritarian, rebel and psychopath, crank, manipulative; and three of the five types with low score in the F Scale: rigid, protesting, impulsive, easygoing, genuine liberal. Each side of the dipole has a rank of intensity that increases as the order written above.
 Other typologies include Jung's extroversion-introversion trait with four modes of orientation. This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling)~\cite{jung}. Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}. Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.

It was considered central to benefit from key human typologies, both by adding descriptions to a type and by further characterizing classes in the terms encountered.

%    \subsection{Other analisys of interest}
%
%        \subsubsection{Textual content}
%        \subsubsection{Geographical localization}

\section{Results and discussion}
    \subsection{Constancy and discrepancy of activity along time}\label{constDisc}
        \subsubsection{Seconds and minutes}
The incidence of messages at each second in a minute and at each minute in an hour is compatible with uniform distribution tests.
If compared to simulations using an uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function.}, messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$. Simulations reach these values, but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$ (see subsection~\ref{consistency}).

        \subsubsection{Hours of the day}
Table~\ref{dia} shows how the four lists distribute activity along the day.
Afternoon was the most active period 6h of the day. Second 12h more active than first 12h. Even so, activity peak occurs around midday, with a slight skew towards earlier hours.

        \subsubsection{Days of the week}
 Weekdays also exhibit an interesting pattern, to which is dedicated Table~\ref{semana}.
Days exhibit pattern on weekdays, with a decrease of at least one third and reaching two thirds on weekends.
        \subsubsection{Days along the month}
 Table~\ref{mes} shows activity along the month.

        \subsubsection{Months and larger divisions of the year}
Table~\ref{ano} is dedicated to activity in months and larger divisions of the year.

As for months along the year, there seems to be two periods of more prominent activities: Jun-Aug (MET and LAD), and Dec-Mar (CPP, LAU and LAD). These observations fit academic calendars, vacations and end-of-year holidays.

 Variation of activity in the days along the month are less prominent, one cannot point much more than a (probably not statistically relevant) tendency to first and second weeks to be more active. The most important result in the days along the month is their homogeneity with respect to activity. Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.

    \subsection{Scalable fat-tail structure}\label{subsec:pih}

There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this distribution of activity among participants.

As specified in Section~\ref{sectioning}, in order to classify nodes as hubs, intermediary or peripheral, $\widetilde{P}(k)$ is compared
to $P(k)$. For the networks studied here, i.e. derived from public email lists, if degree distribution is used classification, hubs sector size reaches peaks with $10\%$ of all vertexes. 
 If strength is used for comparrison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
to $P(k)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. 
This results hold for in and out degrees and strengths,

Classification criteria exposed in subsection~\ref{sectioning} was used efficiently with windows with at least 200 messages. Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex. The compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement. This is specially useful in dealing with fewer messages, in which case the structure degenerates with respect to some of the degree and strength measures, but not all.

For the networks analysed, differences of using this smoothing process were not significant. There were between 20 and 1200 participants, so each participant were between $5\%$ and $0.08\%$ of all participants.  The bottom line is: for network sizes considered, if connectivity intensity would most proprably not exist in an Ed\"os Renyi network, than it is not an intermediary intensity. As peripheral vertex are abundant, this statistic discussion has no relevance.

A reasonable window size for observation can be inferred by monitoring the giant component size and the degeneration of the hub, intermediary and peripheral sections. This degeneration is critical in the span of 50-100 messages. With compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the network seems to hold basic structure even with as few as 20-50 messages. This indicates that concentration of activity and low-activity participants occurs even with very few messages.

Appendix~\ref{figures} is dedicated to figures on these networks and their evolution.

    \subsection{Prevalence of centrality over asymmetries and asymetries over clusterization}\label{measures}

The principal component exhibit ponderation of centrality measures: degrees, strengthes and betweenness centrality. Clustering coefficient is presented in almost perfect orthogonality. Dispersion is more prevalent in symmetry related measures than clustering coefficient. This holds for all network snapshots observed, even with as few messages as to degenerate structure. Symmetric and asymmetric edges have been reported as bounded to different roles played by participants and relations~\cite{newmanEvolving}. Principal components formation from original measures can be observed in Tables~\ref{compPCA0},~\ref{compPCA} and~\ref{compPCA2}. Individual vertexes relation with top two principal components can be appreciated in Figures~\ref{PCA} and~\ref{PCA2}.
This peculiar first component that consists of the averaged sum of degree, strength and betweenness measures was verified to be incident in virtually all networks with 500 or more messages and most smaller networks (degeneration of basic structure is critical with $ws \approx 50-100$ messages). This composition of principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.

As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for window larger sizes ($ws>1000$); and high degree is associated with low clustering coefficient, as can be observed in Figure~\ref{clust}.

When symmetry of node connectivity is considered, the first component remains mostly the same, but clustering coefficient is only relevant to third and fourth components. A snapshot of vertexes with respect to these first two principal components are in Figure~\ref{PCA2}. This asymmetry and disequilibrium measures revealed as more proper measures to characterization of hubs and intermediaries, as seen in greater spreading of second PCA plot. Both symmetry of node overal activity and of individual relations play important role in second component, as can be observed in Table~\ref{compPCA2}.

\begin{figure} 
   \centering
        \includegraphics[width=\columnwidth]{figs2/ev0pr3PCA}
    \caption{PCA of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}. Table~\ref{compPCA} has the composition of principle components from the original measures. First principle component is a pondered sum of centrality measures: degrees, strengthes and betweenness centrality. Second component is mostly clustering coefficient in this figure, but asymetries holds second component if also considered. Similarity to plot in Figure~\ref{clust} was verified with all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
    \label{PCA}
\end{figure}

\begin{figure} 
   \centering
        \includegraphics[width=\columnwidth]{figs2/ev0pr1PCA}
    \caption{Degree and strength, clustering coefficient, betweenness centrality and symmetry related measures are used for this scatter plot of principal components. Compositions of first three components are in table~\ref{compPCA2} and measure details in subsection~\ref{measures}. Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.}
    \label{PCA2}
\end{figure}

\begin{figure} 
   \centering
        \includegraphics[width=\columnwidth]{figs2/ev0pr11CC}
    \caption{Clustering versus degree of vertex in interaction network observed, $ws = 1000$ email messages, LAU list. General layout is in accordance with literature. With ver connected vertexes with low clusterization and high clusterization being gradually more incident as number of connections is lowered.}
    \label{clust}
\end{figure}

    \subsection{Primitive typology}

This work aimed at finding common characteristics among (email) interaction networks. The analysis  involved primary measures observance and a formal criteria for coherent ratios of hub, intermediary and hub sectors. Nevertheless, inspection done by visualizations and raw data manipulations suggest peculiarities of interest, specially:

\begin{itemize}
    \item Core hubs usually have intermittent activity. Very stable activity was found on MET hubs, which motivated its integration to this work. Literature repports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered a direct cause of stable hubs activity.
    \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone. Peripheral vertex activity also follows a simple patters: they will interact very rarelly, in very few occasions. Intermediary vertexes seem responsible for network structure.
    \item Network operation modes dictated by intermediary vertexes behavior. These can exhibit preferential communication to peripheral or hub vertex.
    \item Some of the most active participants receive many responses with relative few messages sent, and are never top hubs. These seem as authorities and contrast with participants that respond way more than receive responses.
    \item Most obvious community structure, as observed by hung clustering coefficient, is found only in peripheral and intermediary sectors.
\end{itemize}

This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts involved in other typologies, such Meyer-Briggs, Pavlov or F-Scale. This has no pretension of being a direct result from numeric analysis, it is a refinement the description of found structure and classes considered.

\section{Conclusions and future work}
In fact, some characterization of chosen networks resulted from stable observations. Along temporal activity statistics, this work repports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning relative sizes (evident with the comparrison with the Erd\"os-Renyi model).

The task of delivering a first and general characterization of chosen interaction networks involved starting a larger effort. The different aspects covered requires not only different analytical background, but also considerations about textual production and social psychology. These are receiving attention within dedicated works and are summarized in this section.

    \subsection{Further work}
        \subsubsection{Constancy of general characteristics, easing tipologization of outiliers and mundane occurences}

Regarding topological aspects of interaction networks, further work should inspect other measures (e.g. closeness centrality, accessibility), and statistics in each of the three conective sectors: hubs, intermediary and peripheral.

 Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space. Another possibility, specially for a brute-force characterization of such sections, is to remove vertexes with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.

        \subsubsection{Interaction network characterization}

Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns. Even so, verification of results in other virtual environment, such as Facebook, Twitter and LinkedIn, should verify the generality of this repport.

        \subsubsection{Textual productions}
Further work should observe textual production of network sectors. Resulting knowledge purposes to network and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.

        \subsubsection{Tipology enhancements}
From the stable characteristics, outliers can be pointed and further developed in terms of networks and participants typologization.

        \subsubsection{Results exploitation}
Usage of such characteristics are takingn place in linked data and electronic goverment technologies~\cite{ops,opa}. Further steps involve ellaboration and tests of social dynamics that takes advantages of these results

\begin{acknowledgments}
Renato Fabbri is grateful to CNPq (process: 140860/2013-4,
project 870336/1997-5), United Nations Development Program (PNUD/ONU, contract: 2013/000566; project BRA/12/018)  and 
the Postgraduate Committee of the IFSC/USP. This author is also grateful for
the American Jewish Committee for maintaining an online copy of the Adorno book
used on the epigraph~\cite{adorno}. Authors thanks Gmane creators and keepers, specifically: Gmane is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R. Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B. Krause, Nicolas Bareil, Raymond Scholz, and Adam Sjøgren. Authors thank refered email lists communities and welcome feedback as core contribution to this, and similar, research.
\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\section{Tables}\label{sectables}
\subsection{PCA tables}

\begin{table}
  \centering
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
  \end{tabular}
  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. First component is a pondered sum of degree and betweenness centrality measures. Second component is mostly clustering coefficient. First and second components sum more than 95\% of total dispersion.}
  \label{compPCA0}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
  \end{tabular}
  \caption{Principal components' composition in percentages. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. First component is a pondered sum of degree and strength measures and betweenness centrality. Second component is mostly clustering coefficient. First and second components sum more than 90\% of dispersion.}
  \label{compPCA}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
  \end{tabular}
  \caption{Distribution of components, added measures of symmetry described in subsection~\ref{measures}. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. In this case, clusterization is pushed to third component, with disequilibrium measures. Second component is primarily symmetry measures, but also some out degree and strength contribution. Betweenness again has a role similar to degree, but weaker. Clusterization component combines with disequilibrium, while asymmetry is related to out degree and strength. Three components has in average 80.36\% of dispersion.}
  \label{compPCA2}
\end{table}

\subsection{Tables for activity along time}\label{tabTime}
\begin{table*}
%\tiny
    \caption{Hours of the day and percentage of activity ($\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours. Maximum activity rates are in bold. In hour columns, minimum activity is also bold. The less active period of the day is around 4-6h. Maximum activity is between 10-13h. Afternoon is most active in 6h division of the day. The noon has $\approx \frac{2}{3}$ of 24h activity. }
\begin{center}
    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
        \hline
    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
    & 1h   & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h \\ \hline\hline
0h  & 3.66 & \multirow{6}{*}{10.67} & \multirow{12}{*}{33.76} & 2.87  & \multirow{6}{*}{7.15} & \multirow{12}{*}{29.33} & 3.58 & \multirow{6}{*}{10.14} &  \multirow{12}{*}{36.88} & 4.00 & \multirow{6}{*}{10.77} & \multirow{12}{*}{33.13} \\
1h  & 2.76 &                        &                         & 1.77  & & &  2.22 & & & 2.52 & & \\
2h  & 1.79 &                        &                         & 1.04  & & & 1.63 & & & 1.79 & & \\
3h  & 1.10 &                        &                         & 0.64  & & & 1.06 & & & 1.06 & & \\
4h  & {\bf 0.68} &                        &                         & 0.47  & & &  0.84 & & & 0.75 & & \\
5h  & 0.69 &                        &                         & {\bf 0.38}  & & & {\bf 0.82} & & & {\bf 0.66} & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
6h  & 0.83 & \multirow{6}{*}{23.09} &                         & 0.72  & \multirow{6}{*}{22.18} & & 1.17 & \multirow{6}{*}{26.74} & & 0.85 & \multirow{6}{*}{22.36} &  \\
7h  & 1.24 &                        &                         & 1.33  & & & 2.37 & & & 1.56 & & \\
8h  & 2.28 &                        &                         & 2.67  & & & 3.54 & & & 2.96 & & \\
9h  & 4.52 &                        &                         & 4.40  & & & 6.04 & & & 4.68 & & \\
10h & 6.62 &                        &                         & 6.29  & & & {\bf 6.83} & & & 5.93 & & \\
11h & {\bf 7.61} &                        &                         & 6.78  & & & 6.79 & & & 6.40 & & \\\hline
12h & 6.44 & \multirow{6}{*}{\bf 37.63} & \multirow{12}{*}{\bf 66.24} & {\bf 7.33}  & \multirow{6}{*}{\bf 42.22} & \multirow{12}{*}{ \bf 70.66} & 6.11 & \multirow{6}{*}{\bf 35.65} &  \multirow{12}{*}{ \bf 63.12} & {\bf 6.41} & \multirow{6}{*}{\bf 37.25} &  \multirow{12}{*}{\bf 66.87} \\
13h & 6.04 &                        &                         & 7.08  & & & 6.26 & & & 6.12 & & \\
14h & 6.47 &                        &                         & 7.09  & & & 6.38 & & & 6.33 & & \\
15h & 6.10 &                        &                         & 7.14  & & & 5.93 & & & 5.98 & & \\
16h & 6.22 &                        &                         & 6.68  & & & 5.52 & & & 6.40 & & \\
17h & 6.36 &                        &                         & 6.89  & & & 5.46 & & & 6.02 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}

18h & 6.01 & \multirow{6}{*}{28.61} &                         & 5.99  & \multirow{6}{*}{28.44} & & 5.24 & \multirow{6}{*}{27.46} & & 5.99 & \multirow{6}{*}{29.63} & \\
19h & 5.02 &                        &                         & 5.23  & & & 4.52 & & & 5.03 & & \\
20h & 4.85 &                        &                         & 4.98  & & & 4.55 & & & 4.63 & & \\
21h & 4.38 &                        &                         & 4.37  & & & 4.42 & & & 4.59 & & \\
22h & 4.06 &                        &                         & 4.24  & & & 4.51 & & & 4.88 & & \\
23h & 4.30 &                        &                         & 3.64  & & & 4.23 & & & 4.53 & & \\\hline
    \end{tabular}
\end{center}
\label{dia}
\end{table*}

\begin{table}[h]
    \caption{Concentration of activity on days along the week. Weekend days are at least $\frac{1}{3}$ less active and can reach $\frac{1}{3}$ of activity. MET concentrates activity in weekdays the most, leaving only 13.98\% of total activity to Saturday and Sunday. LAU is the one that less concentrates activity in weekdays, reaching 20.94\% of total activity in weekends. These might suggest professional relation of CPP and MET participants to the topics of interest, or a hobby relation of LAU and LAD participants.}
\begin{center}
    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
        \hline
        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
        CPP & 17.06 & 17.43 & 17.61 & 17.13 & 16.30 & 6.81 & 7.67 \\ \hline
        MET & 17.53 & 17.54 & 16.43 & 17.06 & 17.46 & 7.92 & 6.06 \\ \hline
        LAU & 15.71 & 15.80 & 15.88 & 16.43 & 15.13 & 10.13 & 10.91 \\ \hline
        LAD & 14.91 & 17.73 & 17.01 & 15.40 & 14.25 & 10.39 & 10.30 \\\hline
    \end{tabular}
\end{center}
\label{semana}
\end{table}

\begin{table*}
%\tiny
    \caption{Activity along the days of the month. As can be noted, the pattern is to have no clear prevalent period. One might point a slight tendency for the first two weeks to be more active, although this table does not present statistical foundation for such an assumption. For the scope of this study, differences of activity along the month is assumed to be non existent.}
\begin{center}
    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
        \hline
    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
day    & 1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days \\ \hline\hline
1  & 3.19 & \multirow{7}{*}{23.05} & \multirow{14}{*}{45.63} & 3.01 & \multirow{7}{*}{25.16} & \multirow{14}{*}{48.08} & 3.34 & \multirow{7}{*}{23.06} &  \multirow{14}{*}{47.31} & 3.22 & \multirow{7}{*}{21.96} & \multirow{14}{*}{46.70} \\
2  & 3.07 &                        &                         & 3.38  & & &  3.38 & & & 3.42 & & \\
3  & 3.20 &                        &                         & 3.55  & & & 3.20 & & & 2.87 & & \\
4  & 3.63 &                        &                         & 4.34  & & & 3.52 & & & 2.91 & & \\
5  & 2.85 &                        &                         & 3.93  & & &  2.68 & & & 3.30 & & \\
6  & 3.67 &                        &                         & 3.76  & & & 3.18 & & & 3.52 & & \\
7  & 3.45 &                        &                         & 3.18  & & & 3.77 & & & 2.27 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
8  & 3.12 & \multirow{7}{*}{22.57} &                         & 3.36  & \multirow{7}{*}{22.92} & & 3.62 & \multirow{7}{*}{24.25} & & 3.72 & \multirow{7}{*}{24.73} &  \\
9  & 2.57 &                        &                         & 3.44  & & & 3.82 & & & 3.97 & & \\
10  & 2.92 &                        &                         & 3.17  & & & 3.06 & & & 3.77 & & \\
11  & 3.54 &                        &                         & 3.88  & & & 3.11 & & & 3.27 & & \\
12 & 3.23 &                        &                         & 2.94  & & & 3.40 & & & 2.75 & & \\
13 & 3.39 &                        &                         & 3.29  & & & 3.55 & & & 3.34 & & \\
14 & 3.81 &                        &                         & 2.83  & & & 3.69 & & & 3.93 & & \\\hline
15 & 3.35 & \multirow{7}{*}{23.02} & \multirow{14}{*}{46.31} & 2.72  & \multirow{7}{*}{21.87} & \multirow{14}{*}{ 43.56} & 3.23 & \multirow{7}{*}{22.84} &  \multirow{14}{*}{ 44.01 } & 3.37 & \multirow{7}{*}{22.82} &  \multirow{14}{*}{46.00} \\
16 & 3.77 &                        &                         & 2.96  & & & 2.94 & & & 3.37 & & \\
17 & 3.45 &                        &                         & 3.01  & & & 3.02 & & & 2.95 & & \\
18 & 3.47 &                        &                         & 3.39  & & & 3.63 & & & 3.22 & & \\
19 & 2.90 &                        &                         & 3.42  & & & 3.16 & & & 3.59 & & \\
20 & 2.80 &                        &                         & 3.09  & & & 3.25 & & & 3.21 & & \\
21 & 3.29 &                        &                         & 3.27  & & & 3.61 & & & 3.13 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}

22 & 2.88 & \multirow{7}{*}{23.29} &                         & 2.92  & \multirow{7}{*}{21.69} & & 3.80 & \multirow{7}{*}{21.17} & & 3.07 & \multirow{7}{*}{23.18} & \\
23 & 4.01 &                        &                         & 3.27  & & & 3.03 & & & 3.06 & & \\
24 & 3.13 &                        &                         & 2.92  & & & 2.31 & & & 2.72 & & \\
25 & 3.57 &                        &                         & 2.83  & & & 2.38 & & & 3.16 & & \\
26 & 3.27 &                        &                         & 2.97  & & & 3.49 & & & 3.57 & & \\
27 & 3.27 &                        &                         & 3.41  & & & 2.92 & & & 3.92 & & \\
28 & 3.17 &                        &                         & 3.36  & & & 3.26 & & & 3.69 & & \\\hline
29 & 3.68 &  \multirow{3}{*}{8.06} &  \multirow{3}{*}{8.06}  & 2.93  & \multirow{3}{*}{8.36} & \multirow{3}{*}{8.36} & 3.34 & \multirow{3}{*}{8.68} & \multirow{3}{*}{8.68} & 3.15 & \multirow{3}{*}{7.30} & \multirow{3}{*}{7.30} \\
30 & 2.76 &                        &                         & 3.14  & & & 3.75 & & & 2.71 & & \\
31 & 1.63 &                        &                         & 2.29  & & & 1.60 & & & 1.45 & & \\\hline
    \end{tabular}
\end{center}
\label{mes}
\end{table*}

\begin{table*}[t]
\scriptsize
    \caption{Activity along the year, in months, trimesters, quadrimesters and semesters. Engagement in list participation seem to concentrate in two periods: middle of the year (Jun-Aug, lists MET and LAD), and transition from years (Dec-Mar, lists CPP, LAU and LAD). Messages were considered as to complete 12 months slots, so every month has the same time of occurrences.}
\begin{center}
    \begin{tabular}{ |l|| c|c|c|c|c|  c|c|c|c|c| c|c|c|c|c| c|c|c|c|c|}
        \hline
    & \multicolumn{5}{c|}{CPP} & \multicolumn{5}{c|}{MET} & \multicolumn{5}{c|}{LAU} & \multicolumn{5}{c|}{LAD}  \\ \hline
    & m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. \\ \hline\hline
Jan  & 8.70 & \multirow{2}{*}{17.00} & \multirow{3}{*}{\bf 27.23} &  \multirow{4}{*}{\bf 36.48} & \multirow{6}{*}{\bf 54.26} & 4.88 &  \multirow{2}{*}{11.01} & \multirow{3}{*}{16.90} & \multirow{4}{*}{23.32} & \multirow{6}{*}{47.74} & 10.22 & \multirow{2}{*}{\bf 19.56} & \multirow{3}{*}{\bf 28.23} & \multirow{4}{*}{\bf 35.09} & \multirow{6}{*}{49.17} & 11.23 & \multirow{2}{*}{18.49} & \multirow{3}{*}{26.43} & \multirow{4}{*}{36.04} & \multirow{6}{*}{\bf 57.95} \\
Fev  & 8.29 &                        &                         & & & 6.13 & & & & & 9.34  & & & & & 7.26 & & & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
Mar  & {\bf 10.23} & \multirow{2}{*}{\bf 19.49} &                         & & & 5.89 & \multirow{2}{*}{12.31} & & & & 8.67 & \multirow{2}{*}{15.52} & & & & 7.94 & \multirow{2}{*}{17.55} & & & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
Apr  & 9.26 &                        &  \multirow{3}{*}{27.03} & & & 6.42 & & \multirow{3}{*}{30.84} & & & 6.85 & & \multirow{3}{*}{20.94} & & & 9.61 & & \multirow{3}{*}{\bf 31.51} & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
Mai  & 9.41 & \multirow{2}{*}{17.78} &                         & \multirow{4}{*}{33.46} & & 10.46 & \multirow{2}{*}{\bf 24.42} & & \multirow{4}{*}{\bf 47.83} & & 7.27 & \multirow{2}{*}{14.09} & & \multirow{4}{*}{30.37} &  & 8.94 & \multirow{2}{*}{\bf 21.91} & & \multirow{4}{*}{\bf 37.56} & \\
Jun & 8.37 &                        &                         & & & {\bf 13.96} & & & & & 6.81 & & & & & {\bf 12.97} & & & & \\\cline{3-3}\cline{4-4}\cline{6-6}\cline{8-9}\cline{11-11}\cline{13-14}\cline{16-16}\cline{18-19}\cline{21-21}
Jul & 8.70 &  \multirow{2}{*}{15.68} & \multirow{3}{*}{22.94} & & \multirow{6}{*}{45.73} & 13.23 & \multirow{2}{*}{23.41} & \multirow{3}{*}{\bf 31.16} & & \multirow{6}{*}{\bf 52.26} & 8.96 & \multirow{2}{*}{16.28} & \multirow{3}{*}{24.47} & & \multirow{6}{*}{\bf 50.82} & 9.02 & \multirow{2}{*}{15.65} & \multirow{3}{*}{22.29} & & \multirow{6}{*}{42.05} \\
Ago  & 6.98 &                       &                         & & & 10.28 & & & & & 7.31 & & & & & 6.63 & & & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
Set  & 7.26 & \multirow{2}{*}{15.36} &                         & \multirow{4}{*}{30.06} & & 7.75 & \multirow{2}{*}{16.80} & & \multirow{4}{*}{28.86} & & 8.18 & \multirow{2}{*}{16.24} & & \multirow{4}{*}{34.54} & & 6.63 & \multirow{2}{*}{12.38} & & \multirow{4}{*}{26.40} & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
Oct  & 8.10 &                        &  \multirow{3}{*}{22.80} & & & 9.05 & & \multirow{3}{*}{21.10} & & & 8.06 & & \multirow{3}{*}{26.36} & & & 5.74 & & \multirow{3}{*}{19.77} & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
Nov  & 7.86 & \multirow{2}{*}{14.69} &                         & & & 7.46 & \multirow{2}{*}{12.06} & & & & 7.63 & \multirow{2}{*}{18.30} & & & & 7.63 & \multirow{2}{*}{14.02} & & & \\
Dec & 6.81 &                        &                         &  & & 4.59 & & & & & {\bf 10.66} & & & & & 6.39 & & & & \\\hline
    \end{tabular}
\end{center}
\label{ano}
\end{table*}

\begin{table}
    \caption{Distribution of activity among agents. First column is dedicated to percentage of messages sent by the most active participant. Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages. Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages. The last decile $10D$ column has maximum percentage of participants responsible for 10\% of activity (messages).}
\begin{center}
    \begin{tabular}{ | l ||  c | c | c | c | }
        \hline
        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
    \end{tabular}
\end{center}
\label{autores}
\end{table}

\section{Data and scripts further description}\label{scripts}
Messages are downloaded from GMANE database by RSS in the mbox email text format. 
They are requested one by one to avoid reaching maximum size of the requests accepted by
GMANE API.

Every message has about 30 fields, from which the following are crucial
for the present work:
\begin{itemize}
    \item ``From'' field, as it specifies the sender of the message, in the usual format of ``First\_name Last\_Name $<email>$''.
    \item ``Date'' field, which is given with the resolution of a second.
    \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
    \item ``References'', has the ID of the message it is an answer for, if any, and earlier messages in the thread.
\end{itemize}

Field ``In-Reply-To'' has only the ID of the message it replies and can be sometimes
a shortcut or an alternative to ``References''. Also, the textual content of the messages,
accessed through ``payload'' method of the mbox message object, is of central interest and
the authors dedicated an article to include the textual content of the messages to the analysis.

Basic constructs for obtaining all results in this article are described in~\ref{ap:os}. Scripts, written in Python programming language, are publicly available at~\cite{scriptsFim} and very briefly specified below.

\subsection{Third party libraries and software}
Programming resources used were mainly Python and part of the
common scientific bundle for the language. More specifically,
scripts where written for 2.7.3 version of Python,
with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.

\subsection{Python scripts}\label{ap:os}
All results were obtained with scripts writen in the Python programming language. These are kept in a public repository for backup and sharing with research community~\cite{scriptsFim}. Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.

\section{Figures of vertex classification fractions as the network evolves}\label{figures}

Two lists are exhibited in this section, CPP and LAD. These structures are very similar in all
four lists and laying extensively all figures is redundant. Window sizes of $ws =$ 10000, 5000,
1000, 500, 250, 100 and 50 messages were used.

\begin{figure*}[hb]
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/10000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp10000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/10000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp10000_}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/5000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp5000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/5000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp5000_}
\end{figure*}

%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/1000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp1000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/1000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp1000_}
\end{figure*}

%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/500}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp500}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/500_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp500_}
\end{figure*}

%%%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/250}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp250}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/250_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp250_}
\end{figure*}

%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/100}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp100}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/100_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp100_}
\end{figure*}

%%%%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/50}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:cpp50}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/CPP/50_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:cpp50_}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/10000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad10000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/10000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad10000_}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/5000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad5000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/5000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad5000_}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/1000}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad1000}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/1000_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad1000_}
\end{figure*}

%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/500}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad500}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/500_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad500_}
\end{figure*}

%%%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/250}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad250}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/250_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad250_}
\end{figure*}

%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/100}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad100}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/100_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad100_}
\end{figure*}

%%%%%%%%%%
\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/50}
    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths. CPP Std library official mailing list. In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction. On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component. On the same graph, green counts the disconnected vertexes.}
    \label{fig:lad50}
\end{figure*}

\begin{figure*}[hbtp] 
   \centering
        \includegraphics[width=\textwidth]{figs2/LAD/50_2}
    \caption{Distribution of vertex with respect to each measure of activity: compound criteria. Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions. The first two plots picture classifications that are not functions. Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black. On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$. Compound criteria described in subsection~\ref{benchmarks}.}
    \label{fig:lad50_}
\end{figure*}

%
%\begin{figure*}
%    \centering
%        \includegraphics[width=\textwidth]{pcm}
%        \caption{Pulse Code Modulation (PCM) audio: an analogical signal is represented by 25 samples with 4 bits each.}
%        \label{fig:PCM}
%\end{figure*}
%

\section{Further consideration of related work}
Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}. Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.

The seminal Nature Letter by Palla, Bara{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.

Controllability of these networks is also an uncovered issue. These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.

Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}, which can be further investigated to hold in email lists and in evolving terms, as a community oriented, non-private interactions are drawn from public emails groups with hundred of participants.

Considered years altogether, hundreds to thousands of participants post on a list, more rarely dozens or tenths of thousands. The most active lists usually reaches a few thousands of participants. Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to a deeper insights in community-related network evolution.

\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file aipsamp.tex ******

### 2014-12-24 convertScript.sh

@@ -1,0 +1,28 @@
+#!/bin/bash
+#n=0;
+#
+#echo " ";
+#echo "Converting:"
+#echo " ";
+#
+#for pngfile in "$@"; do
+#
+#  jpgfile=${epsfile%.eps}.jpg
+#
+#  echo "$((++n)): $epsfile to $jpgfile."
+#
+#  convert -density 300x300 $epsfile $jpgfile
+#
+#done
+#
+#echo " ";
+#echo "$((n)) files were converted from EPS to JPG format."
+cd figs2
+#for i in $(find . | grep eps); do
+for AFILE in $(find . -iname "*.eps"); do
+#    echo ${AFILE} "${AFILE%.*}".png;
+    convert ${AFILE} "${AFILE%.*}".png;
+#    echo "${AFILE}";
+#    echo "${AFILE%.*}";
+#    convert ${i} "${AFILE%%.*}".png;
+done

### 2014-12-24 paper.tex
adicionado pdf compilado
feito retoque geral do artigo

@@
 
-\author{Luciano da Fontoura Costa}
-  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
-  \email{ldfcosta@gmail.com}
-  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-
-\author{Osvaldo N. Oliveira Jr.}
-  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
-  \email{chu@ifsc.usp.br}
- \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%\author{Luciano da Fontoura Costa}
+%  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
+%  \email{ldfcosta@gmail.com}
+%  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%
+%\author{Osvaldo N. Oliveira Jr.}
+%  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
+%  \email{chu@ifsc.usp.br}
+% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
@@
 \begin{abstract}
-    This article reports a minimal and general characterization of interaction networks evolution.
-Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants, 2) a sound and stable classification of vertex: peripheral, intermediary and hub sectors, 3) composition of basic measures into components with greater dispersion.
-While time patterns of activity are not obvious, participant activity follow concentrations expected by scale-free networks.
-Comparison with ideal Erd\"os-R\'enyi network with the same number of edges and vertexes revealed as a sound criterion  for distinguishing sectors on the networks.
+    This article reports a characterization of interaction networks and its stability.
+Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants, 2) a sound classification of vertex: peripheral, intermediary and hub sectors, 3) combination of basic measures into components with greater dispersion.
+While time patterns of activity are not obvious, participant activity follows a scale-free distribution.
+Comparison with ideal Erd\"os-R\'enyi network with the same number of edges and vertexes was a sound criterion  for distinguishing sectors on the networks.
 Principal components in basic measures spaces revealed interesting and regular patterns of independence and dispersion.
 This includes a ranking of measures that most contribute to dispersion: 1) degree and strength measures, 2) symmetry related quantization, and 3) clusterization.
-Results suggested typologies for these networks and participants.
+Results suggest typologies for these networks and participants.
 Further work include considerations of text production, psychoanalysis inspired typologies, participatory democracy exploitation of observed properties, and better visualization support for network evolution.
@@
 
-Text production and typologies of online participants benefit from results here presented~\cite{rcText,rcTipo}. 
+Text mining and typologies of online participants benefit from results here presented~\cite{rcText,rcTipo}. 
 Although all networks considered originated from email lists,
@@
  Such results are in accordance with phenomena observed in this work and linguistic characterization is being described in a deriving article~\cite{rcText}.
+See Appendix~\ref{sec:fure} for further considerations of related work.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figs2/criaRede_}
+    \includegraphics[width=0.5\textwidth]{figs/criaRede_}
     \caption{Formation of interaction network from email messages.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figs2/fser}
+    \includegraphics[width=0.5\textwidth]{figs/fser}
     \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The later has more
@@
 
-        \subsubsection{Prevalence of topological measures}
+        \subsubsection{Topological measures}\label{measures}
 This article
@@
 Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research.
-Furthermore, stable aspects of measures prominance along time are captured through mean and standard deviation (see Section~\ref{}).
-Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{}.
+Furthermore, stable aspects of measures prominance along time are captured through mean and standard deviation (see Section~\ref{measures}).
+Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{figures}.
 
@@
 If compared to simulations using an uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function.}, messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$.
-Simulations reach these values, but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$ (see subsection~\ref{consistency}).
+Simulations reach these values, but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
 
@@
 
-    \subsection{Prevalence of centrality over asymmetries and asymetries over clusterization}\label{measures}
+    \subsection{Prevalence of centrality over asymmetries and asymetries over clusterization}\label{prevalence}
 
@@
    \centering
-        \includegraphics[width=\columnwidth]{figs2/ev0pr3PCA}
+        \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
     \caption{PCA of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
@@
    \centering
-        \includegraphics[width=\columnwidth]{figs2/ev0pr1PCA}
+        \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
     \caption{Degree and strength, clustering coefficient, betweenness centrality and symmetry related measures are used for this scatter plot of principal components.
@@
    \centering
-        \includegraphics[width=\columnwidth]{figs2/ev0pr11CC}
+        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
     \caption{Clustering versus degree of vertex in interaction network observed, $ws = 1000$ email messages, LAU list.
@@
 used on the epigraph~\cite{adorno}.
-Authors thanks Gmane creators and keepers, specifically: Gmane is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
+Authors thanks Gmane creators and maintainers, specifically: Gmane is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
 Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B.
@@
 \appendix
+\section{Data and scripts further description}\label{scripts}
+Messages are downloaded from GMANE database by RSS in the mbox email text format. 
+They are requested one by one to avoid reaching maximum size of the requests accepted by
+GMANE API.
+
+Every message has about 30 fields, from which the following are crucial
+for the present work:
+\begin{itemize}
+    \item ``From'' field, as it specifies the sender of the message, in the usual format of ``First\_name Last\_Name $<email>$''.
+    \item ``Date'' field, which is given with the resolution of a second.
+    \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
+    \item ``References'', has the ID of the message it is an answer for, if any, and earlier messages in the thread.
+\end{itemize}
+
+Field ``In-Reply-To'' has only the ID of the message it replies and can be sometimes
+a shortcut or an alternative to ``References''. Also, the textual content of the messages,
+accessed through ``payload'' method of the mbox message object, is of central interest and
+the authors dedicated an article to include the textual content of the messages to the analysis.
+
+Basic constructs for obtaining all results in this article are described in~\ref{ap:os}.
+Scripts, written in Python programming language, are publicly available at~\cite{scriptsFim} and very briefly specified below.
+
+\subsection{Third party libraries and software}
+Programming resources used were mainly Python and part of the
+common scientific bundle for the language. More specifically,
+scripts where written for 2.7.3 version of Python,
+with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
+Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
+
+\subsection{Python scripts}\label{ap:os}
+All results were obtained with scripts writen in the Python programming language.
+These are kept in a public repository for backup and sharing with research community~\cite{scriptsFim}.
+Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
+\section{Further consideration of related work}\label{sec:fure}
+Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}.
+Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
+
+The seminal Nature Letter by Palla, Bara{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.
+
+Controllability of these networks is also an uncovered issue.
+These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.
+
+Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}, which can be further investigated to hold in email lists and in evolving terms, as a community oriented, non-private interactions are drawn from public emails groups with hundred of participants.
+
+Considered years altogether, hundreds to thousands of participants post on a list, more rarely dozens or tenths of thousands.
+The most active lists usually reaches a few thousands of participants.
+Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to a deeper insights in community-related network evolution.
 
 \section{Tables}\label{sectables}
+\clearpage
 \subsection{PCA tables}
 
-\begin{table}
+\begin{table}[H]
   \centering
@@
 \end{table}
+
+\clearpage
 
@@
 
-\section{Data and scripts further description}\label{scripts}
-Messages are downloaded from GMANE database by RSS in the mbox email text format. 
-They are requested one by one to avoid reaching maximum size of the requests accepted by
-GMANE API.
-
-Every message has about 30 fields, from which the following are crucial
-for the present work:
-\begin{itemize}
-    \item ``From'' field, as it specifies the sender of the message, in the usual format of ``First\_name Last\_Name $<email>$''.
-    \item ``Date'' field, which is given with the resolution of a second.
-    \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
-    \item ``References'', has the ID of the message it is an answer for, if any, and earlier messages in the thread.
-\end{itemize}
-
-Field ``In-Reply-To'' has only the ID of the message it replies and can be sometimes
-a shortcut or an alternative to ``References''. Also, the textual content of the messages,
-accessed through ``payload'' method of the mbox message object, is of central interest and
-the authors dedicated an article to include the textual content of the messages to the analysis.
-
-Basic constructs for obtaining all results in this article are described in~\ref{ap:os}.
-Scripts, written in Python programming language, are publicly available at~\cite{scriptsFim} and very briefly specified below.
-
-\subsection{Third party libraries and software}
-Programming resources used were mainly Python and part of the
-common scientific bundle for the language. More specifically,
-scripts where written for 2.7.3 version of Python,
-with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
-Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
-
-\subsection{Python scripts}\label{ap:os}
-All results were obtained with scripts writen in the Python programming language.
-These are kept in a public repository for backup and sharing with research community~\cite{scriptsFim}.
-Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
-
+\clearpage
 \section{Figures of vertex classification fractions as the network evolves}\label{figures}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/10000}
+        \includegraphics[width=\textwidth]{figs/CPP/10000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/10000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/10000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp10000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/5000}
+        \includegraphics[width=\textwidth]{figs/CPP/5000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/5000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/5000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp5000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/1000}
+        \includegraphics[width=\textwidth]{figs/CPP/1000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/1000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/1000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp1000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/500}
+        \includegraphics[width=\textwidth]{figs/CPP/500}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/500_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/500_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp500_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/250}
+        \includegraphics[width=\textwidth]{figs/CPP/250}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/250_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/250_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp250_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/100}
+        \includegraphics[width=\textwidth]{figs/CPP/100}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/100_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/100_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp100_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/50}
+        \includegraphics[width=\textwidth]{figs/CPP/50}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/CPP/50_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/CPP/50_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:cpp50_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/10000}
+        \includegraphics[width=\textwidth]{figs/LAD/10000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/10000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/10000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad10000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/5000}
+        \includegraphics[width=\textwidth]{figs/LAD/5000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/5000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/5000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad5000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/1000}
+        \includegraphics[width=\textwidth]{figs/LAD/1000}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/1000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/1000_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad1000_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/500}
+        \includegraphics[width=\textwidth]{figs/LAD/500}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/500_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/500_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad500_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/250}
+        \includegraphics[width=\textwidth]{figs/LAD/250}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/250_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/250_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad250_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/100}
+        \includegraphics[width=\textwidth]{figs/LAD/100}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/100_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/100_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad100_}
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/50}
+        \includegraphics[width=\textwidth]{figs/LAD/50}
     \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
@@
    \centering
-        \includegraphics[width=\textwidth]{figs2/LAD/50_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{benchmarks}.}
+        \includegraphics[width=\textwidth]{figs/LAD/50_2}
+    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
+Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots picture classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Compound criteria described in subsection~\ref{sectioning}.}
     \label{fig:lad50_}
@@
 
-\section{Further consideration of related work}
-Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}.
-Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
-
-The seminal Nature Letter by Palla, Bara{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.
-
-Controllability of these networks is also an uncovered issue.
-These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.
-
-Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}, which can be further investigated to hold in email lists and in evolving terms, as a community oriented, non-private interactions are drawn from public emails groups with hundred of participants.
-
-Considered years altogether, hundreds to thousands of participants post on a list, more rarely dozens or tenths of thousands.
-The most active lists usually reaches a few thousands of participants.
-Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to a deeper insights in community-related network evolution.
-
 \nocite{*}

### 2014-12-24 paper.tex
arrumadas referencias minimamente

@@
 \usepackage{morefloats}
+\usepackage{hyperref}
 
@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominance of measures and activity statistics}% Force line breaks with \\
+\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
 
@@
  Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
-Moreover, selected exceptions are repported in this section.
+Moreover, selected exceptions are reported in this section.
 The evolution considered in the present article is characterized by a constant number of messages, which is also present in literature, but was less explored to date.
@@
  The evolution of interaction networks was addressed with focus on community evolution, a work that ignores the direction of edges~\cite{barabasiEvo}.
-Two topologically different networks are repported to emerge, depending on the frequency of interactions, which present a generalized power law or an exponencial connectivity distribution~\cite{barabasiTopologicalEv}.
+Two topologically different networks are reported to emerge, depending on the frequency of interactions, which present a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
 Recent work on email lists consider an isolated snapshot in order to verify or draw hypothesis.
@@
 These lists cover a variety of topics, mostly technology-related.
-It can be described as a corpus with metadata of its messages, like send time, place, sender name, sender email adderss etc.
+It can be described as a corpus with metadata of its messages, like send time, place, sender name, sender email address etc.
 GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
@@
     \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE archive.}.
-Participantes are from different countries, and English is the language used the most.
+Participants are from different countries, and English is the language used the most.
 More technical and less active version of LAU.
@@
 Dominated by participants with hybrid artistic and technological interests.
-Participantes are from different countries, and English is the language used the most.
+Participants are from different countries, and English is the language used the most.
 Abbreviated as LAU from now on.
@@
 \section{Characterization methods}
-After immersion in the data and on appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological ellaborations of networks and participants.
+After immersion in the data and on appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological elaborations of networks and participants.
 Each of these methods are described bellow, as other handy alternatives.
@@
 Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, and undirected weighted and undirected unweighted).
-More precisely, the networks repported are erected as follows: a direct response from participant B to participant A forms an edge from A to B, as information went from A to B.
+More precisely, the networks reported are erected as follows: a direct response from participant B to participant A forms an edge from A to B, as information went from A to B.
 The reasoning is: if B wrote a response to A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
@@
 Each vertex represents a participant.
-If participant B replyes participant A, that is regarded as evidence that B received information from A.
+If participant B replies participant A, that is regarded as evidence that B received information from A.
 Multiple messages add ``weight'' to directed edge.
@@
 This article
-restricts this analisys to a small selection of the most basic
+restricts this analysis to a small selection of the most basic
 measures of each vertex:
@@
 Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research.
-Furthermore, stable aspects of measures prominance along time are captured through mean and standard deviation (see Section~\ref{measures}).
+Furthermore, stable aspects of measures prominence along time are captured through mean and standard deviation (see Section~\ref{measures}).
 Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{figures}.
@@
 to $P(k)$. For the networks studied here, i.e. derived from public email lists, if degree distribution is used classification, hubs sector size reaches peaks with $10\%$ of all vertexes. 
- If strength is used for comparrison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
+ If strength is used for comparison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
 to $P(k)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. 
@@
 There were between 20 and 1200 participants, so each participant were between $5\%$ and $0.08\%$ of all participants.
-The bottom line is: for network sizes considered, if connectivity intensity would most proprably not exist in an Ed\"os Renyi network, than it is not an intermediary intensity.
+The bottom line is: for network sizes considered, if connectivity intensity would most probably not exist in an Ed\"os Renyi network, than it is not an intermediary intensity.
 As peripheral vertex are abundant, this statistic discussion has no relevance.
@@
 
-    \subsection{Prevalence of centrality over asymmetries and asymetries over clusterization}\label{prevalence}
-
-The principal component exhibit ponderation of centrality measures: degrees, strengthes and betweenness centrality.
+    \subsection{Prevalence of centrality over asymmetries and asymmetries over clusterization}\label{prevalence}
+
+The principal component exhibit ponderation of centrality measures: degrees, strengths and betweenness centrality.
 Clustering coefficient is presented in almost perfect orthogonality.
@@
 This asymmetry and disequilibrium measures revealed as more proper measures to characterization of hubs and intermediaries, as seen in greater spreading of second PCA plot.
-Both symmetry of node overal activity and of individual relations play important role in second component, as can be observed in Table~\ref{compPCA2}.
+Both symmetry of node overall activity and of individual relations play important role in second component, as can be observed in Table~\ref{compPCA2}.
 
@@
 Table~\ref{compPCA} has the composition of principle components from the original measures.
-First principle component is a pondered sum of centrality measures: degrees, strengthes and betweenness centrality.
-Second component is mostly clustering coefficient in this figure, but asymetries holds second component if also considered.
+First principle component is a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
+Second component is mostly clustering coefficient in this figure, but asymmetries holds second component if also considered.
 Similarity to plot in Figure~\ref{clust} was verified with all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
@@
     \caption{Clustering versus degree of vertex in interaction network observed, $ws = 1000$ email messages, LAU list.
-General layout is in accordance with literature.
-With ver connected vertexes with low clusterization and high clusterization being gradually more incident as number of connections is lowered.}
+General layout is in accordance with literature: connected vertexes have low clusterization while higher clusterization is gradually more incident as number of connections is lowered.}
     \label{clust}
@@
 Very stable activity was found on MET hubs, which motivated its integration to this work.
-Literature repports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered a direct cause of stable hubs activity.
+Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered a direct cause of stable hubs activity.
     \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone.
-Peripheral vertex activity also follows a simple patters: they will interact very rarelly, in very few occasions.
+Peripheral vertex activity also follows a simple patters: they will interact very rarely, in very few occasions.
 Intermediary vertexes seem responsible for network structure.
@@
 In fact, some characterization of chosen networks resulted from stable observations.
-Along temporal activity statistics, this work repports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning relative sizes (evident with the comparrison with the Erd\"os-Renyi model).
+Along temporal activity statistics, this work reports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning relative sizes (evident with the comparison with the Erd\"os-Renyi model).
 
@@
     \subsection{Further work}
-        \subsubsection{Constancy of general characteristics, easing tipologization of outiliers and mundane occurences}
+        \subsubsection{Constancy of general characteristics, easing tipologization of outliers and mundane occurrences}
 
 Regarding topological aspects of interaction networks, further work should inspect other measures (e.g.
-closeness centrality, accessibility), and statistics in each of the three conective sectors: hubs, intermediary and peripheral.
+closeness centrality, accessibility), and statistics in each of the three connective sectors: hubs, intermediary and peripheral.
 
@@
 Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns.
-Even so, verification of results in other virtual environment, such as Facebook, Twitter and LinkedIn, should verify the generality of this repport.
+Even so, verification of results in other virtual environment, such as Facebook, Twitter and LinkedIn, should verify the generality of this report.
 
@@
 
-        \subsubsection{Tipology enhancements}
-From the stable characteristics, outliers can be pointed and further developed in terms of networks and participants typologization.
+        \subsubsection{Typology enhancements}
+From the stable characteristics, outliers can be pointed and further developed in terms of networks and participants typology.
 
         \subsubsection{Results exploitation}
-Usage of such characteristics are takingn place in linked data and electronic goverment technologies~\cite{ops,opa}.
-Further steps involve ellaboration and tests of social dynamics that takes advantages of these results
+Usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa}.
+Further steps involve elaboration and tests of social dynamics that takes advantages of these results
 
@@
 used on the epigraph~\cite{adorno}.
-Authors thanks Gmane creators and maintainers, specifically: Gmane is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
+Authors thanks GMANE creators and maintainers, specifically: GMANE is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
 Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B.
 Krause, Nicolas Bareil, Raymond Scholz, and Adam Sjøgren.
-Authors thank refered email lists communities and welcome feedback as core contribution to this, and similar, research.
+Authors thank referred email lists communities and welcome feedback as core contribution to this, and similar, research.
 \end{acknowledgments}
@@
 \subsection{Python scripts}\label{ap:os}
-All results were obtained with scripts writen in the Python programming language.
+All results were obtained with scripts written in the Python programming language.
 These are kept in a public repository for backup and sharing with research community~\cite{scriptsFim}.

### 2015-01-20 paper.tex
revisado todo artigo, menos tabelas, figuras e referencias

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\hypersetup{
+        colorlinks,
+        linkcolor={red!50!black},
+        citecolor={blue!50!black},
+        urlcolor={blue!80!black}
+}
 
@@
     This article reports a characterization of interaction networks and its stability.
-Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants, 2) a sound classification of vertex: peripheral, intermediary and hub sectors, 3) combination of basic measures into components with greater dispersion.
+Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants; 2) a sound classification of vertex: peripheral, intermediary and hub sectors; 3) combination of basic measures into components with greater dispersion (PCA).
 While time patterns of activity are not obvious, participant activity follows a scale-free distribution.
-Comparison with ideal Erd\"os-R\'enyi network with the same number of edges and vertexes was a sound criterion  for distinguishing sectors on the networks.
-Principal components in basic measures spaces revealed interesting and regular patterns of independence and dispersion.
+Comparison of incident networks with ideal Erd\"os-R\'enyi networks bearing the same number of edges and vertexes yields a sound criterion for distinguishing sectors on the networks.
+Principal components in basic measures space revealed interesting and regular patterns of independence and dispersion.
 This includes a ranking of measures that most contribute to dispersion: 1) degree and strength measures, 2) symmetry related quantization, and 3) clusterization.
@@
 this work considers only directed, weighted and human interaction networks.
-Undirected and unweighted representation of such networks is also seen in the literature and can be obtained by simplification~\cite{GMANE2}.
-
-Text mining and typologies of online participants benefit from results here presented~\cite{rcText,rcTipo}. 
+Undirected and unweighted representation of such networks is also found in the literature and can be obtained by simplification~\cite{GMANE2}.
+
+Text mining and typologies of online participants benefit from the results here presented~\cite{rcText,rcTipo}. 
 Although all networks considered originated from email lists,
  coherence with literature suggests that results hold for a more general class of interaction networks,
- such as observed in online platforms (LinkedIn, Facebook, Twitter).
+ such as observed in online platforms (e.g. LinkedIn, Facebook, Twitter).
 
@@
 Moreover, selected exceptions are reported in this section.
-The evolution considered in the present article is characterized by a constant number of messages, which is also present in literature, but was less explored to date.
-
- The evolution of interaction networks was addressed with focus on community evolution, a work that ignores the direction of edges~\cite{barabasiEvo}.
+The evolution considered in the present article is characterized by a constant number of messages, which is also present in literature, but less explored to date~\cite{netWindowEvo}.
+
+ The evolution of interaction networks was addressed with a community focus, in a work that ignores the direction of edges~\cite{barabasiEvo}.
 Two topologically different networks are reported to emerge, depending on the frequency of interactions, which present a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
@@
 
- Such results are in accordance with phenomena observed in this work and linguistic characterization is being described in a deriving article~\cite{rcText}.
+ Such results are in accordance with phenomena observed in this work.
+Linguistic characterization is being described in a deriving article~\cite{rcText}.
 See Appendix~\ref{sec:fure} for further considerations of related work.
@@
 These lists cover a variety of topics, mostly technology-related.
-It can be described as a corpus with metadata of its messages, like send time, place, sender name, sender email address etc.
+It can be described as a corpus with metadata of its messages, such as time, place, sender name, sender email address.
 GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
@@
     \subsection{Chosen dataset}
-The four lists below were selected for their diversity, easing initial observance of natural and general properties.
+The four lists below were selected for the resulting diversity, easing initial observance of natural and general properties.
 \begin{itemize}
-    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE archive.}.
+    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}.
+Dominated by participants with hybrid artistic and technological interests.
 Participants are from different countries, and English is the language used the most.
-More technical and less active version of LAU.
+Abbreviated as LAU from now on.
+    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}.
+Participants are from different countries, and English is the language used the most.
+A more technical and less active version of LAU.
 Abbreviated LAD from now on.
-    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE archive.}.
+    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}.
 Dominated by specialized computer programmers.
@@
 Abbreviated as CPP from now on.
-    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE archive.}.
+    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}.
 Dominated by Brazilian activists and digital culture interests.
-Participants are mostly Brazilians, and Portuguese is the most used language, although Spanish and English usage is commonplace.
+Participants are mostly Brazilians, and Portuguese is the most used language, although Spanish and English are also incident.
 Abbreviated MET from now on.
-    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE archive.}.
-Dominated by participants with hybrid artistic and technological interests.
-Participants are from different countries, and English is the language used the most.
-Abbreviated as LAU from now on.
 \end{itemize}
@@
 \section{Characterization methods}
-After immersion in the data and on appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological elaborations of networks and participants.
-Each of these methods are described bellow, as other handy alternatives.
+After immersion with the data and appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological elaborations of networks and participants.
+Each of these methods are described bellow with supporting structures.
 
     \subsection{Temporal activity statistics}
-Number of messages along time with respect to seconds, minutes, hours, days of the week or the month, and months of the year.
-These were displayed as tables in Appendix~\ref{tabTime}, results are outlined in Section~\ref{constDisc}.
+Number of messages along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
+These are exhibited as tables in Appendix~\ref{tabTime}.
+Results are outlined in Section~\ref{constDisc}.
 
     \subsection{Interaction network}\label{intNet}
-Regarding literature~\cite{bird,newmanCommunityDirected,newmanCommunity2013}, interaction networks might be both weighted or unweighted, both directed or undirected.
-Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, and undirected weighted and undirected unweighted).
-More precisely, the networks reported are erected as follows: a direct response from participant B to participant A forms an edge from A to B, as information went from A to B.
-The reasoning is: if B wrote a response to A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
-Inverting edge direction yields the status network, as B read the message and considered what A wrote worth of responding, giving status to A, thus $B\rightarrow A$.
+Regarding literature~\cite{bird,newmanCommunityDirected,newmanCommunity2013}, interaction networks can be modeled both weighted or unweighted, both directed or undirected.
+Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, undirected weighted, and undirected unweighted).
+More precisely, the networks reported are erected as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
+The reasoning is: if B wrote a response to a message from A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
+Inverting edge direction yields the status network, as B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
 This article uses the information network described above and depicted in Figure~\ref{formationNetwork}.
@@
 Edges can be created from all antecedent messages on the message-response thread.
-Is this work, only immediate predecessors are linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$.
+In this work, only the immediate predecessor are linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$.
 Potential interpretations for this weaker connection are usually common sense, such as: double length, half weight or with one more ``obstacle''.
@@
     \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The later has more
-        intermediary, as the former has more peripheral and hubs. Sections are
+        intermediary, as the former has more peripheral and hub vertexes. Sections are
         given by the two intersections of the distributions $k_1$ and $k_2$. Characteristic degrees
@@
 
-Because of social networks' tendency to have scale-free distribution of properties,
+Because of social networks tendency to have a scale-free distribution of connectivity,
 one can compare it to an Erd\"os-R\'enyi random graph and consider peripheral, intermediary and hub
@@
 
-If $\mathcal{N}_f$ is directed and has self-loops, the probability
-of the presence of an unknown edge is $p=\frac{z}{N(N-1)}$, where $N(N-1)$ is the maximum number of edges for a network with $N$ vertexes, directed edges and without selfloops.
-A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertexes and edges, and thus the same probability $p$ for the presence of an edge, will have degree $k$ with probability:
+If $\mathcal{N}_f$ is directed and has no self-loops, the probability
+of the presence of an edge between two arbitrary vertexes is $p_e=\frac{z}{N(N-1)}$ (see Appendix~\ref{ap:ded}).
+A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertexes and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:
 
@@
 
-The arguments behind this classification are: 1) vertexes so connected that they are virtually inexistent in networks connected at pure chance, specially without preferential attachment, are correctly associated to hubs sector.
-Vertexes with very few connections, which are way more abundant than expected by pure chance, are correctly associated to periphery.
-Degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in free-scale phenomena, are correctly associated to intermediary vertexes.
-
-To assure statistical validity, bins can be chosen to span the average of $\eta$ gaps between measure values.
-Thus, each bin, starting at degree $k_i$, spans $\Delta_i=\frac{\sum_{i=1}^{\eta}(k_{i+1}-k_i)}{\eta}$ values, with borders $k_i$ and $k_i+\Delta_{i}$.
+The arguments behind this classification are: 1) vertexes so connected that they are virtually inexistent in networks whose edges are incident regardless of vertex properties, are correctly associated to the hubs sector.
+Vertexes with very few connections, which are way more abundant than expected in networks whose edges are incident regardless of vertex properties, are correctly associated to periphery.
+Degree values near average, predicted as the most abundant if connections are created without criteria, and less frequent in free-scale phenomena, are correctly associated to intermediary vertexes.
+
+To assure statistical validity, bins can be chosen to contain at least $\eta$ vertexes.
+Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer in which degrees $k_i$-$k_{j}$ contain at least $\eta$ values.
 This changes equation~\ref{criterio} to:
@@
 \begin{equation}\label{criterio2}
-    \sum_{x=k_i}^{k_i+\Delta_i} \widetilde{P}(x) < \sum_{x=k_i}^{k_i+\Delta_i} P(x) \Rightarrow \text{i is intermediary}
+    \sum_{x=k_i}^{k_j} \widetilde{P}(x) < \sum_{x=k_i}^{k_j} P(x) \Rightarrow \text{i is intermediary}
 \end{equation}
@@
 For in and out degrees and strengths, comparisons should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$.
-Results of these segmentations are discussed in subsection~\ref{subsec:pih}.
+Results of these criteria for network segmentation are discussed in subsection~\ref{subsec:pih}.
 
@@
     \item Exclusivist criteria: vertex are only classified if the class is the same with respect to all measures.
-In this case, total vertex classified (usually) does not reach 100\%, which is specified by a black line in Appendix~\ref{figures}.
+In this case, total vertex classified (usually) does not reach 100\%, which is depicted by a black line in Appendix~\ref{figures}.
     \item Inclusivist criteria: vertex has class given by any of the measures.
-In this case, a vertex can have more than one class and the fraction of class attribution beyond total number of vertexes is also represented by a black line in Appendix~\ref{figures}.
+In this case, a vertex can have more than one class and the fraction of class attribution beyond total number of vertexes is also depicted by a black line in Appendix~\ref{figures}.
     \item Exclusivist cascade: hubs are only hubs if classified as hub with respect to all measures.
@@
 This article
-restricts this analysis to a small selection of the most basic
+restricts topological analysis to a small selection of the most basic
 measures of each vertex:
@@
     \item Out-strength $s_i^{out}$: sum of weights of all edges departing from node $i$.
-    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked. Standard clustering coefficient for undirected graphs was used.
+    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.
+This is the standard clustering coefficient for undirected graphs.
+Usage of other clustering coefficients (e.g.
+for directed graphs) was considering out of scope, as such clustering coefficients are usually employed in special contexts.
     \item Betweenness centrality $bt_i$: fraction of geodesics that contain the node $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
@@
 \begin{itemize}
-    \item asymmetry of note $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
-    \item mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
+    \item asymmetry of vertex $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
+    \item mean of edge asymmetry of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
 $|J_i|$ is the number of neighbors of vertex $i$.
-    \item standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
-    \item disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
+    \item standard deviation of edge asymmetry of vertex $i$: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
+    \item disequilibrium of vertex $i$: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
+    \item mean of edge disequilibrium of vertex $i$: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
+    \item standard deviation of edge disequilibrium of vertex $i$: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
 \end{itemize}
@@
 Evolution of network is observed within a fixed number of messages (window size: $ws$) that shifts in the message timeline.
- All $ws=$50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000 were used.
-Within a same $ws$, the number vertex and edges vary in time, as do other network characteristics.
+ All $ws=$ 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000 were used.
+Within a same $ws$, the number vertexes and edges vary in time, as do other network characteristics.
 Further work should deepen inspection of measure interdependence, this article holds to measures in Section~\ref{measures}.
@@
 In refining hypotheses, visualization of the network was crucial.
-Animations, image galleries and an online gadgets were made~\cite{animacoes,galGMANE,appGMANE}.
+Animations, image galleries and online gadgets were made~\cite{animacoes,galGMANE,appGMANE}.
 Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research.
-Furthermore, stable aspects of measures prominence along time are captured through mean and standard deviation (see Section~\ref{measures}).
+Furthermore, stable aspects of measures prominence along time are captured through mean and standard deviation (see Section~\ref{measures} and Appendix~\ref{sec:pcat}).
 Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{figures}.
@@
 Table~\ref{dia} shows how the four lists distribute activity along the day.
-Afternoon was the most active period 6h of the day. Second 12h more active than first 12h. Even so, activity peak occurs around midday, with a slight skew towards earlier hours.
+Afternoon was the most active 6h period of the day. Second 12h more active than first 12h. Even so, activity peak occurs around midday, with a skew towards earlier hours.
 
         \subsubsection{Days of the week}
- Weekdays also exhibit an interesting pattern, to which is dedicated Table~\ref{semana}.
-Days exhibit pattern on weekdays, with a decrease of at least one third and reaching two thirds on weekends.
+ Weekdays also exhibit an interesting pattern, to which is dedicated Table~\ref{semana}: a decrease of activity on weekends of at least one third, reaching two thirds in extreme cases.
         \subsubsection{Days along the month}
- Table~\ref{mes} shows activity along the month.
+ Table~\ref{mes} shows activity along the month, in which no prevalent pattern was observed.
+ Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
+The most important trait might be their homogeneity with respect to activity.
+Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
 
@@
 Table~\ref{ano} is dedicated to activity in months and larger divisions of the year.
-
-As for months along the year, there seems to be two periods of more prominent activities: Jun-Aug (MET and LAD), and Dec-Mar (CPP, LAU and LAD).
-These observations fit academic calendars, vacations and end-of-year holidays.
-
- Variation of activity in the days along the month are less prominent, one cannot point much more than a (probably not statistically relevant) tendency to first and second weeks to be more active.
-The most important result in the days along the month is their homogeneity with respect to activity.
-Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
+Observation points two periods of more prominent activities: Jun-Aug (MET and LAD), and Dec-Mar (CPP, LAU and LAD), which fit academic calendars, vacations and end-of-year holidays.
 
@@
 
-As specified in Section~\ref{sectioning}, in order to classify nodes as hubs, intermediary or peripheral, $\widetilde{P}(k)$ is compared
-to $P(k)$. For the networks studied here, i.e. derived from public email lists, if degree distribution is used classification, hubs sector size reaches peaks with $10\%$ of all vertexes. 
+As specified in Section~\ref{sectioning}, in order to classify nodes as hubs, intermediary or peripheral, the incident connectivity distribution $\widetilde{P}(k)$ is compared
+to a connectivity distribution of an ideal Erd\"os R\`nyi $P(k)$ with the same number of vertexes and edges.
+In the networks inspected for this article (see Section~\ref{intNet}), if degree distribution is used for classification, hubs sector size reaches peaks with $10\%$ of all vertexes.
  If strength is used for comparison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
-to $P(k)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. 
+to $P(k_i)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. 
 This results hold for in and out degrees and strengths,
 
-Classification criteria exposed in subsection~\ref{sectioning} was used efficiently with windows with at least 200 messages.
-Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex.
-The compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement.
+Classification criteria exposed in Section~\ref{sectioning} was used efficiently with windows of at least 200 messages.
+Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex, which match literature~\cite{priSect}.
+A compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement.
 This is specially useful in dealing with fewer messages, in which case the structure degenerates with respect to some of the degree and strength measures, but not all.
-
-For the networks analysed, differences of using this smoothing process were not significant.
-There were between 20 and 1200 participants, so each participant were between $5\%$ and $0.08\%$ of all participants.
-The bottom line is: for network sizes considered, if connectivity intensity would most probably not exist in an Ed\"os Renyi network, than it is not an intermediary intensity.
-As peripheral vertex are abundant, this statistic discussion has no relevance.
-
-A reasonable window size for observation can be inferred by monitoring the giant component size and the degeneration of the hub, intermediary and peripheral sections.
-This degeneration is critical in the span of 50-100 messages.
-With compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the network seems to hold basic structure even with as few as 20-50 messages.
+A minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of hub, intermediary and peripheral sections.
+This degeneration is critical in the span of 50-100 messages for window sizes.
+With compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold basic structure even with as few as 20-50 messages.
 This indicates that concentration of activity and low-activity participants occurs even with very few messages.
 
+For the histograms used in the classification process, the usage of at least $\eta$ vertexes for each bin did not yield significant differences.
+That was understood as a consequence of the observation scale:
+\emph{There are between 20 and 1200 participants in the message window sizes used to derive most results ($ws \in [200,1500]$ messages).
+As peripheral vertexes are abundant and span few degrees, there are more than $\eta$ vertexes with each low degree value.
+Higher degrees cases requires a bit more reasoning.
+Specified $ws$ implicates that each participant is $p \in [0.1\%,0.5\%]$ of all participants.
+Therefore, if incident connectivity is very improbable in an Ed\"os R\`enyi network (less than $p$, the probability that a single participant represent when histogram is normalized to density function), than it is not an intermediary connectivity, but a hub.} This concludes the explanation of why using at least $\eta$ vertexes for each bin did not impact results.
+
 Appendix~\ref{figures} is dedicated to figures on these networks and their evolution.
@@
 
-The principal component exhibit ponderation of centrality measures: degrees, strengths and betweenness centrality.
+The principal component (PCA~\cite{pca}) exhibit ponderation of centrality measures: degrees, strengths and betweenness centrality.
 Clustering coefficient is presented in almost perfect orthogonality.
@@
 
-As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for window larger sizes ($ws>1000$); and high degree is associated with low clustering coefficient, as can be observed in Figure~\ref{clust}.
+As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for larger window sizes ($ws>1000$).
+Also trivial: high degree is associated with low clustering coefficient, as can be observed in Figure~\ref{clust}.
 
@@
 This work aimed at finding common characteristics among (email) interaction networks.
-The analysis  involved primary measures observance and a formal criteria for coherent ratios of hub, intermediary and hub sectors.
-Nevertheless, inspection done by visualizations and raw data manipulations suggest peculiarities of interest, specially:
+Analysis involved primary measures observance and formal criteria for coherent ratios of hub, intermediary and hub sectors.
+Nevertheless, inspection done by visualizations and raw data manipulations suggests agents peculiarities of typological character.
+These are initial observations, which should be further developed as posed in Section~\ref{sec:ftyp}:
 
@@
 Very stable activity was found on MET hubs, which motivated its integration to this work.
-Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered a direct cause of stable hubs activity.
+Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered coherent with the stable activity of hubs.
     \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone.
-Peripheral vertex activity also follows a simple patters: they will interact very rarely, in very few occasions.
+Peripheral vertex activity also follows a simple pattern: they will interact very rarely, in very few occasions.
 Intermediary vertexes seem responsible for network structure.
-    \item Network operation modes dictated by intermediary vertexes behavior. These can exhibit preferential communication to peripheral or hub vertex.
-    \item Some of the most active participants receive many responses with relative few messages sent, and are never top hubs.
+    \item Network operation exhibit modes that seem dictated by intermediary vertexes behavior.
+For example, these can exhibit preferential communication to peripheral, intermediary, or hub vertexes; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
+    \item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
 These seem as authorities and contrast with participants that respond way more than receive responses.
-    \item Most obvious community structure, as observed by hung clustering coefficient, is found only in peripheral and intermediary sectors.
+    \item Most obvious community structure, as observed by high clustering coefficient, is found only in peripheral and intermediary sectors.
 \end{itemize}
@@
 This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts involved in other typologies, such Meyer-Briggs, Pavlov or F-Scale.
-This has no pretension of being a direct result from numeric analysis, it is a refinement the description of found structure and classes considered.
+This has no pretension of being a direct result from numeric analysis, it is a description refinement of the found structure, in typological terms.
+Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.
 
 \section{Conclusions and future work}
-In fact, some characterization of chosen networks resulted from stable observations.
-Along temporal activity statistics, this work reports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning relative sizes (evident with the comparison with the Erd\"os-Renyi model).
-
+Characterization of interaction networks resulted from stability observations.
+Along temporal activity statistics, this work reports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning (periphery, intermediary, hubs) relative sizes, evident in the comparison with the Erd\"os-R\'enyi model.
+
+    \subsection{Further work}
 The task of delivering a first and general characterization of chosen interaction networks involved starting a larger effort.
@@
 
-    \subsection{Further work}
-        \subsubsection{Constancy of general characteristics, easing tipologization of outliers and mundane occurrences}
+        \subsubsection{Constancy of general characteristics eases tipologization}
 
@@
 Another possibility, specially for a brute-force characterization of such sections, is to remove vertexes with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
-
-        \subsubsection{Interaction network characterization}
+The subtraction $\widetilde{P}(k)-P(k)$ should result in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertexes.
+This might support classification of the three sectors by clustering, a more traditional approach to classification.
 
 Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns.
-Even so, verification of results in other virtual environment, such as Facebook, Twitter and LinkedIn, should verify the generality of this report.
-
-        \subsubsection{Textual productions}
-Further work should observe textual production of network sectors.
-Resulting knowledge purposes to network and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
-
-        \subsubsection{Typology enhancements}
-From the stable characteristics, outliers can be pointed and further developed in terms of networks and participants typology.
+Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, should verify the generality of this report.
+
+A related work observed textual production of network sectors~\cite{textProd}.
+Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences in a dedicated study.
+Stability reported in this article eases tipologization of outliers and more usual participation patterns.
 
         \subsubsection{Results exploitation}
-Usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa}.
-Further steps involve elaboration and tests of social dynamics that takes advantages of these results
+Usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa,ensaio}.
+Further steps involve elaboration and tests of social dynamics that takes advantages of these results.
 
@@
 \appendix
-\section{Data and scripts further description}\label{scripts}
+\section{Deduction of edge existence probability in a directed network without self-loops}\label{ap:ded}
+Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertexes.
+The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertexes})}$, where $max( \text{number of edges} |\ \text{N vertexes})=2[(N-1)+(N-2)...1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertexes.
+Therefore:
+\begin{align}
+    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertexes})}=\nonumber\\
+       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}}=\nonumber\\
+   p_e &=\frac{z}{N(N-1)}
+\end{align}
+
+\section{Further consideration of related work}\label{sec:fure}
+Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}.
+Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
+
+The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.
+
+Controllability of these networks is also an uncovered issue.
+These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.
+
+Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}.
+This can be further investigated to hold in email lists, and in evolving terms.
+Community oriented, non-private interactions takes place daily in public email groups with dozens, hundreds or thousands of participants.
+
+Considered years altogether, hundreds to tenths of thousands of participants can post on a list.
+The most active lists usually reaches a few thousands of participants.
+Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to deeper insights in community-related network evolution.
+
+\section{Data and scripts}\label{scripts}
 Messages are downloaded from GMANE database by RSS in the mbox email text format. 
@@
     \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
-    \item ``References'', has the ID of the message it is an answer for, if any, and earlier messages in the thread.
+    \item ``References'', has the ID of the message it is an answer to, if any, and earlier messages in the thread.
 \end{itemize}
@@
 accessed through ``payload'' method of the mbox message object, is of central interest and
-the authors dedicated an article to include the textual content of the messages to the analysis.
-
-Basic constructs for obtaining all results in this article are described in~\ref{ap:os}.
-Scripts, written in Python programming language, are publicly available at~\cite{scriptsFim} and very briefly specified below.
+the authors dedicated an article to include the textual content of the messages to the analysis~\cite{textProd}.
 
 \subsection{Third party libraries and software}
-Programming resources used were mainly Python and part of the
-common scientific bundle for the language. More specifically,
+The programming framework used 
+is mainly Python-based, with emphasis on usual
+scientific tools. More specifically,
 scripts where written for 2.7.3 version of Python,
@@
 All results were obtained with scripts written in the Python programming language.
-These are kept in a public repository for backup and sharing with research community~\cite{scriptsFim}.
+These are kept in a public git repository for backup and sharing with research community~\cite{scriptsFim}.
 Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
-\section{Further consideration of related work}\label{sec:fure}
-Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}.
-Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
-
-The seminal Nature Letter by Palla, Bara{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.
-
-Controllability of these networks is also an uncovered issue.
-These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.
-
-Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}, which can be further investigated to hold in email lists and in evolving terms, as a community oriented, non-private interactions are drawn from public emails groups with hundred of participants.
-
-Considered years altogether, hundreds to thousands of participants post on a list, more rarely dozens or tenths of thousands.
-The most active lists usually reaches a few thousands of participants.
-Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to a deeper insights in community-related network evolution.
 
@@
 \clearpage
-\subsection{PCA tables}
+\subsection{PCA tables}\label{sec:pcat}
 

### 2015-01-20 paper.tex
terminada revisao completa do artigo

@@
   \centering
+  \caption{Columns $date_1$ and $date_M$ have first and last message dates from the 20,000 messages considered in each email list.
+$N$ is the number of participants (number of different email addresses).
+$\Gamma$ is the number of threads (count of messages without antecedent).
+$\overline{M}$ is messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
+MET notably has the fewer participants and the larger number of threads.
+This relation holds for each pair of the lists considered: as the number of participants increases,
+ the number of threads decreases.}
+  \label{geralListas}
   \begin{tabular}{|l|c|c|c|c|c|}\hline
@@
   \end{tabular}
-  \caption{Columns $date_1$ and $date_M$ have first and last messages dates from the 20,000 messages considered.
-$N$ is the number of participants (number of different email addresses).
-$\Gamma$ is the number of threads (count of messages without antecedent).
-$\overline{M}$ is messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
-MET notably has the fewer participants and the larger number of threads.
-This relation holds for each pair of the lists considered: as the number of participants increases,
- the number of threads decreases.}
-  \label{geralListas}
 \end{table}
@@
 Each vertex represents a participant.
-If participant B replies participant A, that is regarded as evidence that B received information from A.
+A reply message from participant B to a message from participant A is considered as evidence that B received information from A.
 Multiple messages add ``weight'' to directed edge.
@@
         intermediary, as the former has more peripheral and hub vertexes. Sections are
-        given by the two intersections of the distributions $k_1$ and $k_2$. Characteristic degrees
+        given by the two intersections $k_1$ and $k_2$ of the connectivity distributions. Characteristic degrees
     are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the three sections considered (periphery, intermediary and hubs).}
@@
 Classification criteria exposed in Section~\ref{sectioning} was used efficiently with windows of at least 200 messages.
-Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex, which match literature~\cite{priSect}.
+Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex, which match literature expectations~\cite{secFree}.
 A compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement.
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-    \caption{Clustering versus degree of vertex in interaction network observed, $ws = 1000$ email messages, LAU list.
+    \caption{Clustering versus degree of vertex with a window size of $ws = 1000$ email messages (LAU list).
 General layout is in accordance with literature: connected vertexes have low clusterization while higher clusterization is gradually more incident as number of connections is lowered.}
@@
 Nevertheless, inspection done by visualizations and raw data manipulations suggests agents peculiarities of typological character.
-These are initial observations, which should be further developed as posed in Section~\ref{sec:ftyp}:
+These are initial observations, which should be further developed as posed in Section~\ref{constDisc}:
 
@@
 
-A related work observed textual production of network sectors~\cite{textProd}.
+A related work observed textual production of network sectors~\cite{rcText}.
 Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences in a dedicated study.
@@
 accessed through ``payload'' method of the mbox message object, is of central interest and
-the authors dedicated an article to include the textual content of the messages to the analysis~\cite{textProd}.
+the authors dedicated an article to include the textual content of the messages to the analysis~\cite{rcText}.
 
@@
   \centering
+  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
+LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
+First component is a pondered sum of degree and betweenness centrality measures.
+Second component is mostly clustering coefficient.
+First and second components sum more than 95\% of total dispersion.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
   \end{tabular}
-  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a pondered sum of degree and betweenness centrality measures.
-Second component is mostly clustering coefficient.
-First and second components sum more than 95\% of total dispersion.}
   \label{compPCA0}
@@
   \centering
+  \caption{Principal components composition in percentages.
+LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
+First component is a pondered sum of degree and strength and betweenness centrality measures.
+Second component is mostly clustering coefficient.
+First and second components sum more than 90\% of dispersion.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
   \end{tabular}
-  \caption{Principal components' composition in percentages.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a pondered sum of degree and strength measures and betweenness centrality.
-Second component is mostly clustering coefficient.
-First and second components sum more than 90\% of dispersion.}
   \label{compPCA}
@@
   \centering
+  \caption{Principal components formation with measures of symmetry described (see Section~\ref{measures}).
+LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
+In this case, clusterization is pushed to third components.
+Second component is primarily symmetry measures, but also out degree and strength, and disequilibrium standard deviation.
+Betweenness centrality again has a role similar to degree, but weaker.
+Clusterization component combines with disequilibrium, while asymmetry is combined to out degree and strength.
+Three components has in average 80.36\% of dispersion.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
   \end{tabular}
-  \caption{Distribution of components, added measures of symmetry described in subsection~\ref{measures}.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-In this case, clusterization is pushed to third component, with disequilibrium measures.
-Second component is primarily symmetry measures, but also some out degree and strength contribution.
-Betweenness again has a role similar to degree, but weaker.
-Clusterization component combines with disequilibrium, while asymmetry is related to out degree and strength.
-Three components has in average 80.36\% of dispersion.}
   \label{compPCA2}
@@
 %\tiny
-    \caption{Hours of the day and percentage of activity ($\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
+    \caption{Percentage of activity ($\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
 Maximum activity rates are in bold.
-In hour columns, minimum activity is also bold.
+In 1h columns, minimum activity is also bold.
 The less active period of the day is around 4-6h.
@@
     \caption{Activity along the days of the month.
-As can be noted, the pattern is to have no clear prevalent period.
+The pattern is to have no clear prevalent period.
 One might point a slight tendency for the first two weeks to be more active, although this table does not present statistical foundation for such an assumption.
-For the scope of this study, differences of activity along the month is assumed to be non existent.}
+For the scope of this study, differences of activity along the month is assumed to be inexistent.}
 \begin{center}
@@
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex in equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
+On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
 On the same graph, green counts the disconnected vertexes.}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp10000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp5000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp1000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp500_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp250_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp100_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:cpp50_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad10000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad5000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad1000_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad500_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad250_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad100_}
@@
 On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria described in subsection~\ref{sectioning}.}
+Compound criteria is described in Section~\ref{sectioning}.}
     \label{fig:lad50_}

### 2015-01-21 paper.tex
feita parte de related work

@@
 \begin{abstract}
-    This article reports a characterization of interaction networks and its stability.
+    This article reports a characterization of interaction networks and its temporal stability.
 Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants; 2) a sound classification of vertex: peripheral, intermediary and hub sectors; 3) combination of basic measures into components with greater dispersion (PCA).
-While time patterns of activity are not obvious, participant activity follows a scale-free distribution.
-Comparison of incident networks with ideal Erd\"os-R\'enyi networks bearing the same number of edges and vertexes yields a sound criterion for distinguishing sectors on the networks.
-Principal components in basic measures space revealed interesting and regular patterns of independence and dispersion.
+While time patterns of activity are not obvious, participant activity follows concentrations expected in scale-free networks.
+Comparison of incident networks with ideal Erd\"os-R\'enyi networks bearing the same number of edges and vertexes reveals a sound criterion for distinguishing sectors on the networks.
+Principal components in the basic measures space revealed interesting and regular patterns of independence and dispersion.
 This includes a ranking of measures that most contribute to dispersion: 1) degree and strength measures, 2) symmetry related quantization, and 3) clusterization.
@@
 \subsection{Related work}
+This work approaches network stability through temporal evolution observations.
+This evolution is characterized by a constant number of contiguous messages, of which snapshots are considered to yield a timeline of the network, which is not explored to date, as far as authors know~\cite{netWindowEvo}.
  Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
 Moreover, selected exceptions are reported in this section.
-The evolution considered in the present article is characterized by a constant number of messages, which is also present in literature, but less explored to date~\cite{netWindowEvo}.
 
  The evolution of interaction networks was addressed with a community focus, in a work that ignores the direction of edges~\cite{barabasiEvo}.
-Two topologically different networks are reported to emerge, depending on the frequency of interactions, which present a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-Recent work on email lists consider an isolated snapshot in order to verify or draw hypothesis.
-In such, free-scale properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
+Two topologically different networks are reported to emerge, depending on the frequency of interactions, presenting a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email networks, free-scale properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
 
@@
 Linguistic characterization is being described in a deriving article~\cite{rcText}.
-See Appendix~\ref{sec:fure} for further considerations of related work.
+See Appendix~\ref{sec:fure} for further considerations of related work, with emphasis on the results from research herein reported.
 
@@
 As specified in Section~\ref{sectioning}, in order to classify nodes as hubs, intermediary or peripheral, the incident connectivity distribution $\widetilde{P}(k)$ is compared
-to a connectivity distribution of an ideal Erd\"os R\`nyi $P(k)$ with the same number of vertexes and edges.
+to a connectivity distribution of an ideal Erd\"os R\`enyi network $P(k)$ with the same number of vertexes and edges.
 In the networks inspected for this article (see Section~\ref{intNet}), if degree distribution is used for classification, hubs sector size reaches peaks with $10\%$ of all vertexes.
@@
 Classification criteria exposed in Section~\ref{sectioning} was used efficiently with windows of at least 200 messages.
-Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertex, which match literature expectations~\cite{secFree}.
+Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertexes, which match literature expectations~\cite{secFree}.
 A compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement.
@@
 
-Appendix~\ref{figures} is dedicated to figures on these networks and their evolution.
+Appendix~\ref{figures} is dedicated to figures of these networks and their evolution.
 
@@
 This holds for all network snapshots observed, even with as few messages as to degenerate structure.
-Symmetric and asymmetric edges have been reported as bounded to different roles played by participants and relations~\cite{newmanEvolving}.
+Symmetric and asymmetric edges have been reported as bonded to different roles played by participants and relations~\cite{newmanEvolving}.
 Principal components formation from original measures can be observed in Tables~\ref{compPCA0},~\ref{compPCA} and~\ref{compPCA2}.
@@
     \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone.
-Peripheral vertex activity also follows a simple pattern: they will interact very rarely, in very few occasions.
+Peripheral vertex activity also follows a simple pattern: they interact very rarely, in very few occasions.
 Intermediary vertexes seem responsible for network structure.
@@
 
-Gender related behavior has been notified in mobile phone datasets~\cite{barabasiSex}.
-This can be further investigated to hold in email lists, and in evolving terms.
-Community oriented, non-private interactions takes place daily in public email groups with dozens, hundreds or thousands of participants.
-
-Considered years altogether, hundreds to tenths of thousands of participants can post on a list.
+Gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex}.
+This can be further investigated to hold in email lists and in evolving terms as community oriented, non-private interactions takes are drawn from public email groups with hundreds or thousands of participants.
+
+Considered years altogether, tenths of thousands of participants can post on a list.
 The most active lists usually reaches a few thousands of participants.
-Authors have not checked each list (more than 20 thousand public email groups~\cite{GMANE}), and this might lead to deeper insights in community-related network evolution.
+Analysis of resulting data might lead to deeper insights in community-related network evolution~\cite{GMANE}.
 

### 2015-01-21 paper.tex
terminado de incorporar os apontamentos do chu

@@
 These lists cover a variety of topics, mostly technology-related.
-It can be described as a corpus with metadata of its messages, such as time, place, sender name, sender email address.
+The archive can be described as a corpus with metadata of its messages, including sent time, place, sender name, sender email address.
 GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
-Appendix~\ref{scripts} is dedicated scripts for gathering and processing GMANE email messages.
+The scripts for gathering and processing GMANE email messages are given in Appendix~\ref{scripts}.
 
     \subsection{Chosen dataset}
-The four lists below were selected for the resulting diversity, easing initial observance of natural and general properties.
+Four email lists were selected for their diversity, easing initial observance of natural and general properties.
 \begin{itemize}
@@
   \centering
-  \caption{Columns $date_1$ and $date_M$ have first and last message dates from the 20,000 messages considered in each email list.
+  \caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
 $N$ is the number of participants (number of different email addresses).
 $\Gamma$ is the number of threads (count of messages without antecedent).
-$\overline{M}$ is messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
+$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
 MET notably has the fewer participants and the larger number of threads.
-This relation holds for each pair of the lists considered: as the number of participants increases,
+This relation holds for each of the lists considered: as the number of participants increases,
  the number of threads decreases.}
@@
 Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, undirected weighted, and undirected unweighted).
-More precisely, the networks reported are erected as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
+The networks were considered as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
 The reasoning is: if B wrote a response to a message from A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
 Inverting edge direction yields the status network, as B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
-This article uses the information network described above and depicted in Figure~\ref{formationNetwork}.
+This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}.
 Edges in both directions are allowed.
@@
 Each vertex represents a participant.
-A reply message from participant B to a message from participant A is considered as evidence that B received information from A.
+A reply message from participant B to a message from participant A is regarded as evidence that B received information from A.
 Multiple messages add ``weight'' to directed edge.
@@
 Edges can be created from all antecedent messages on the message-response thread.
-In this work, only the immediate predecessor are linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$.
-Potential interpretations for this weaker connection are usually common sense, such as: double length, half weight or with one more ``obstacle''.
-This suggests the adoption of other centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
+In this work, only the immediate predecessor is linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$.
+Potential interpretations for this weaker connection are usually common sense, such as: double length, half weight or with one more ``obstacles''.
+This suggests the adoption of centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
 
@@
 \begin{equation}
-    P(k)=\binom{2(N-1)}{k}p^k(1-p)^{2(N-1)-k}
+    P(k)=\binom{2(N-1)}{k}p_e^k(1-p_e)^{2(N-1)-k}
 \end{equation}
 
-The lower degree fat tail constitute the border
-vertexes or peripheral sector. 
+The lower degree fat tail is constituted by the border
+vertexes and is called the peripheral sector. 
 The higher degree fat tail is the hub sector.
@@
 The arguments behind this classification are: 1) vertexes so connected that they are virtually inexistent in networks whose edges are incident regardless of vertex properties, are correctly associated to the hubs sector.
-Vertexes with very few connections, which are way more abundant than expected in networks whose edges are incident regardless of vertex properties, are correctly associated to periphery.
-Degree values near average, predicted as the most abundant if connections are created without criteria, and less frequent in free-scale phenomena, are correctly associated to intermediary vertexes.
-
-To assure statistical validity, bins can be chosen to contain at least $\eta$ vertexes.
-Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer in which degrees $k_i$-$k_{j}$ contain at least $\eta$ values.
+Vertexes with very few connections, which are way more abundant than expected in networks whose edges are incident regardless of vertex properties, are correctly assigned to periphery.
+Degree values near average, predicted as the most abundant if connections are created without criteria, and less frequent in free-scale phenomena, are classified to intermediary vertexes.
+
+To ensure statistical validity, bins can be chosen to contain at least $\eta$ vertexes.
+Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer in which degrees $k_i$-$k_{j}$ contain at least $\eta$ vertexes.
 This changes equation~\ref{criterio} to:
@@
 
-If instead strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}$ the average weight of an edge and $s_i$ the vertex strength.
+If instead strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
 For in and out degrees and strengths, comparisons should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$.
-Results of these criteria for network segmentation are discussed in subsection~\ref{subsec:pih}.
-
-As a further refinement of the network segmentation, compound criteria is used for classification of vertex, considering all measures: total, in and out degree and strength.
-After a careful inspection of possible combinations, these were abbreviated to six:
+Results of these criteria for network segmentation are discussed in Section~\ref{subsec:pih}.
+
+Compound criteria is also used for classification of vertex, considering all measures: total, in and out degree and strength.
+After a careful inspection of possible combinations, these were reduced to six:
 
 \begin{itemize}
-    \item Exclusivist criteria: vertex are only classified if the class is the same with respect to all measures.
-In this case, total vertex classified (usually) does not reach 100\%, which is depicted by a black line in Appendix~\ref{figures}.
-    \item Inclusivist criteria: vertex has class given by any of the measures.
-In this case, a vertex can have more than one class and the fraction of class attribution beyond total number of vertexes is also depicted by a black line in Appendix~\ref{figures}.
-    \item Exclusivist cascade: hubs are only hubs if classified as hub with respect to all measures.
-Intermediary are the vertexes classified as intermediary or hub with respect to all measures.
-Vertexes left are regarded as peripheral vertex.
-    \item Inclusivist cascade: hubs are vertexes classified as hubs by any of the measures.
-From the vertexes left, if any is classified as intermediary by any measure, than it is intermediary.
-The rest of the vertexes are peripheral.
-    \item Exclusivist externals: hubs have unanimous classification with respect to all measures.
-Of vertexes left, peripheral vertexes are the ones classified as hub or peripheral by simple criterion.
-The rest represent intermediary sector.
-    \item Inclusivist externals: hubs are vertexes classified as hubs with respect to any measure.
-From vertexes left, if a vertex is classified as peripheral with respect to any measure, than it is peripheral.
-The rest is regarded intermediary sector.
+    \item Exclusivist criteria: vertexes are only classified if the class is the same according to all metrics.
+In this case, vertexes classified does not usually reach 100\%, which is depicted by a black line in Appendix~\ref{figures}.
+    \item Inclusivist criteria: vertex has class given by any of the metrics.
+Therefore, a vertex can belong to more than one class, and class attribution may add more than total number of vertexes, which is indicated by a black line in Appendix~\ref{figures}.
+    \item Exclusivist cascade: a hub is only classified as hub if it is a hub with respect to all metrics.
+Intermediary are the vertexes classified as intermediary or hub with respect to all metrics.
+Remaining vertexes are regarded as peripheral.
+    \item Inclusivist cascade: hubs are vertexes classified as hubs by any of the metrics.
+The remaining vertexes are classified as intermediary if they belong to this category by any of the metrics.
+Vertexes only classified as peripheral are the only considered as such.
+    \item Exclusivist externals: vertexes are only hubs if they are classified as such according to all the metrics.
+The remaining vertexes are classified as peripheral if they fall into hub or peripheral categories by any metric.
+The rest of the vertexes are classified as intermediary.
+    \item Inclusivist externals: hubs are vertexes classified as hubs according to any metric.
+The remaining vertexes will be peripheral if classified as such according to any metric.
+The rest of the vertexes are regarded regarded as intermediary.
 \end{itemize}
 
-These compound criteria, and reduction of possibilities to them, can be formalized in strict mathematical terms. This was considered out of the scope of the present article.
-
-Results from applying this classification method is further reported in Section~\ref{subsec:pih}.
+These compound criteria, and reduction of possibilities to them, can be formalized in strict mathematical terms.
+This was considered out of the scope of the present article.
+Important here is to notice that the compound criteria can be used to examine network sections in the case of a low number of messages, as will be shown later.
+
+Results from applying this classification method is reported in Section~\ref{subsec:pih}.
 
         \subsubsection{Topological measures}\label{measures}
-This article
-restricts topological analysis to a small selection of the most basic
-measures of each vertex:
+The topology of the networks was characterized by a small selection of the most standard
+measurements of each vertex:
 
 \begin{itemize}
-    \item Degree $d_i$: number edges linked to node $i$.
-    \item In-degree $d_i^{in}$: number of edges ending at node $i$.
-    \item Out-degree $d_i^{out}$: number of edges departing from node $i$.
-    \item Strength $s$: sum of weights of all edges linked to node $i$.
-    \item In-strength $s_i^{in}$: sum of weights of all edges ending at node $i$.
-    \item Out-strength $s_i^{out}$: sum of weights of all edges departing from node $i$.
-    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.
+    \item Degree $d_i$: number edges linked to                                 vertex $i$.
+    \item In-degree $d_i^{in}$: number of edges ending at                      vertex $i$.
+    \item Out-degree $d_i^{out}$: number of edges departing from               vertex $i$.
+    \item Strength $s$: sum of weights of all edges linked to                  vertex $i$.
+    \item In-strength $s_i^{in}$: sum of weights of all edges ending at        vertex $i$.
+    \item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
+    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of vertex $i$ that are linked.
 This is the standard clustering coefficient for undirected graphs.
@@
 for directed graphs) was considering out of scope, as such clustering coefficients are usually employed in special contexts.
-    \item Betweenness centrality $bt_i$: fraction of geodesics that contain the node $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
+    \item Betweenness centrality $bt_i$: fraction of geodesics that contain the vertex $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
 \end{itemize}
 
-In order to capture asymmetries in the activity of participants, the following metrics were introduced (see subsection~\ref{prevalence}):
+In order to capture asymmetries in the activity of participants, the following metrics were introduced (see Section~\ref{prevalence}):
 
 \begin{itemize}
-    \item asymmetry of vertex $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
-    \item mean of edge asymmetry of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
+    \item Asymmetry of vertex $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
+    \item Mean of edge asymmetry of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
 $|J_i|$ is the number of neighbors of vertex $i$.
-    \item standard deviation of edge asymmetry of vertex $i$: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
-    \item disequilibrium of vertex $i$: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item mean of edge disequilibrium of vertex $i$: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item standard deviation of edge disequilibrium of vertex $i$: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
+    \item Standard deviation of edge asymmetry of vertex $i$: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
+    \item Disequilibrium of vertex $i$: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
+    \item Mean of edge disequilibrium of vertex $i$: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
+    \item Standard deviation of edge disequilibrium of vertex $i$: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
 \end{itemize}
@@
    \subsection{Evolutive observations}
-Evolution of network is observed within a fixed number of messages (window size: $ws$) that shifts in the message timeline.
+Evolution of network was observed within a fixed number of messages (or window size $ws$) that shifts in the message timeline.
  All $ws=$ 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000 were used.
 Within a same $ws$, the number vertexes and edges vary in time, as do other network characteristics.
-Further work should deepen inspection of measure interdependence, this article holds to measures in Section~\ref{measures}.
+Further work should deepen inspection of measure interdependence.
 
@@
 In refining hypotheses, visualization of the network was crucial.
-Animations, image galleries and online gadgets were made~\cite{animacoes,galGMANE,appGMANE}.
+
+The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
 Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research.
@@
         \subsubsection{Seconds and minutes}
-The incidence of messages at each second in a minute and at each minute in an hour is compatible with uniform distribution tests.
-If compared to simulations using an uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function.}, messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$.
+The incidence of messages at each second of a minute and at each minute of an hour is compatible with uniform distribution simulations\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations.}.
+Messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$.
 Simulations reach these values, but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
@@
 Table~\ref{dia} shows how the four lists distribute activity along the day.
-Afternoon was the most active 6h period of the day. Second 12h more active than first 12h. Even so, activity peak occurs around midday, with a skew towards earlier hours.
+Afternoon was the most active 6h period of the day. Second 12h is $\frac{1}{3}$ more active than first 12h. Even so, activity peak occurs around midday, with a skew toward one hour before noon.
 
@@
  Weekdays also exhibit an interesting pattern, to which is dedicated Table~\ref{semana}: a decrease of activity on weekends of at least one third, reaching two thirds in extreme cases.
+
         \subsubsection{Days along the month}
- Table~\ref{mes} shows activity along the month, in which no prevalent pattern was observed.
+ Table~\ref{mes} shows activity along the month.
 Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait might be their homogeneity with respect to activity.
+The most important trait seems to be homogeneity.
 Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
@@
 to a connectivity distribution of an ideal Erd\"os R\`enyi network $P(k)$ with the same number of vertexes and edges.
-In the networks inspected for this article (see Section~\ref{intNet}), if degree distribution is used for classification, hubs sector size reaches peaks with $10\%$ of all vertexes.
- If strength is used for comparison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
-to $P(k_i)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. 
-This results hold for in and out degrees and strengths,
-
-Classification criteria exposed in Section~\ref{sectioning} was used efficiently with windows of at least 200 messages.
-Specially with 1000 or more messages, criteria yields stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertexes, which match literature expectations~\cite{secFree}.
-A compound criteria, also described in Section~\ref{sectioning}, can be used as a classification refinement.
-This is specially useful in dealing with fewer messages, in which case the structure degenerates with respect to some of the degree and strength measures, but not all.
+In the networks inspected for this article (see Section~\ref{intNet}), if degree distribution is used for classification, hubs can reach $10\%$ of all vertexes.
+ If instead strength is used for comparison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
+to $P(k_i)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. This results hold for in and out degrees and strengths.
+
+Classification criteria exposed in Section~\ref{sectioning} reveled remarkable stability over time, provided that a sample of at least 200 messages is taken as the window of analysis.
+Specially with 1000 or more messages, stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertexes were obtained, which match literature expectations~\cite{secFree}.
+Compound criterion resulted in a greater stability and a window of only 200 messages was already sufficient for the analysis.
 A minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of hub, intermediary and peripheral sections.
-This degeneration is critical in the span of 50-100 messages for window sizes.
-With compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold basic structure even with as few as 20-50 messages.
+This degeneration is critical in the span of 50-100 messages.
+Upon usage of the compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the network seem to hold basic structure even with as few as 20-50 messages.
 This indicates that concentration of activity and low-activity participants occurs even with very few messages.
@@
 Dispersion is more prevalent in symmetry related measures than clustering coefficient.
-This holds for all network snapshots observed, even with as few messages as to degenerate structure.
 Symmetric and asymmetric edges have been reported as bonded to different roles played by participants and relations~\cite{newmanEvolving}.
@@
 This peculiar first component that consists of the averaged sum of degree, strength and betweenness measures was verified to be incident in virtually all networks with 500 or more messages and most smaller networks (degeneration of basic structure is critical with $ws \approx 50-100$ messages).
-This composition of principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-
-As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for larger window sizes ($ws>1000$).
-Also trivial: high degree is associated with low clustering coefficient, as can be observed in Figure~\ref{clust}.
-
-When symmetry of node connectivity is considered, the first component remains mostly the same, but clustering coefficient is only relevant to third and fourth components.
-A snapshot of vertexes with respect to these first two principal components are in Figure~\ref{PCA2}.
-This asymmetry and disequilibrium measures revealed as more proper measures to characterization of hubs and intermediaries, as seen in greater spreading of second PCA plot.
-Both symmetry of node overall activity and of individual relations play important role in second component, as can be observed in Table~\ref{compPCA2}.
+This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
+When symmetry measures are used in PCA, though, the third component has special contributions of out degree and strength.
+
+As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
+Also trivial in the light of complex network literature: high degree is associated with low clustering coefficient, as shown in Figure~\ref{clust}.
+
+When symmetry metrics are included, the first component remains mostly the same, but clustering coefficient is only relevant to third and fourth components.
+A snapshot of vertexes with respect to these first two principal components is in Figure~\ref{PCA2}.
+This asymmetry and disequilibrium measures revealed as more proper measures to characterization of hubs and intermediaries, as seen in greater spreading of second plot.
+These results are summarized in Table~\ref{compPCA2}.
 
@@
     \caption{PCA of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
-Table~\ref{compPCA} has the composition of principle components from the original measures.
-First principle component is a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
-Second component is mostly clustering coefficient in this figure, but asymmetries holds second component if also considered.
-Similarity to plot in Figure~\ref{clust} was verified with all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
+Table~\ref{compPCA} exposes the composition of principle components.
+Principle component is a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
+Second component is mostly clustering coefficient.
+Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
     \label{PCA}
@@
     \caption{Degree and strength, clustering coefficient, betweenness centrality and symmetry related measures are used for this scatter plot of principal components.
-Compositions of first three components are in table~\ref{compPCA2} and measure details in subsection~\ref{measures}.
+Compositions of first three components are in Table~\ref{compPCA2} and measure details in Section~\ref{measures}.
 Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.}
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-    \caption{Clustering versus degree of vertex with a window size of $ws = 1000$ email messages (LAU list).
+    \caption{Clustering coefficient versus degree of vertex with a window size of $ws = 1000$ email messages, LAU list.
 General layout is in accordance with literature: connected vertexes have low clusterization while higher clusterization is gradually more incident as number of connections is lowered.}
@@
 Very stable activity was found on MET hubs, which motivated its integration to this work.
-Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, reason why smaller number of participants in MET was considered coherent with the stable activity of hubs.
+Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
     \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone.
@@
 These seem as authorities and contrast with participants that respond way more than receive responses.
-    \item Most obvious community structure, as observed by high clustering coefficient, is found only in peripheral and intermediary sectors.
+    \item The most obvious community structure, as observed by high clustering coefficient, is found only in peripheral and intermediary sectors.
 \end{itemize}
@@
 
-Regarding topological aspects of interaction networks, further work should inspect other measures (e.g.
-closeness centrality, accessibility), and statistics in each of the three connective sectors: hubs, intermediary and peripheral.
+Regarding topological aspects of interaction networks, further work should inspect other measures in each of the three connective sectors: hubs, intermediary and peripheral.
 
  Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space.
-Another possibility, specially for a brute-force characterization of such sections, is to remove vertexes with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
+Another possibility, specially for a brute-force characterization of such sectors, is to remove vertexes with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
 The subtraction $\widetilde{P}(k)-P(k)$ should result in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertexes.
@@
 Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns.
-Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, should verify the generality of this report.
+Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are convenient uses.
 
 A related work observed textual production of network sectors~\cite{rcText}.
-Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences in a dedicated study.
+Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
 Stability reported in this article eases tipologization of outliers and more usual participation patterns.
@@
 
+\subsection{Python scripts}\label{ap:os}
+Basic constructs for obtaining all results are the product of scripts written in the Python programming language.
+These are kept in a public git repository for backup and sharing with research community~\cite{scriptsFim}.
+Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
+
 \subsection{Third party libraries and software}
@@
 Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
-
-\subsection{Python scripts}\label{ap:os}
-All results were obtained with scripts written in the Python programming language.
-These are kept in a public git repository for backup and sharing with research community~\cite{scriptsFim}.
-Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
-
 \section{Tables}\label{sectables}
@@
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a pondered sum of degree and betweenness centrality measures.
-Second component is mostly clustering coefficient.
+The first component is a pondered sum of degree and betweenness centrality.
+The second component is mostly clustering coefficient.
 First and second components sum more than 95\% of total dispersion.}
@@
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a pondered sum of degree and strength and betweenness centrality measures.
-Second component is mostly clustering coefficient.
-First and second components sum more than 90\% of dispersion.}
+First component is a pondered sum of degree and strength and betweenness centrality.
+The second component is mostly related to clustering coefficient.
+The first and second components sum more than 90\% of dispersion.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
   \centering
-  \caption{Principal components formation with measures of symmetry described (see Section~\ref{measures}).
+  \caption{Principal components formation with measures of symmetry (see Section~\ref{measures}).
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-In this case, clusterization is pushed to third components.
+In this case, clusterization is pushed to third component.
 Second component is primarily symmetry measures, but also out degree and strength, and disequilibrium standard deviation.
@@
 Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $10D$ column has maximum percentage of participants responsible for 10\% of activity (messages).}
+The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
 \begin{center}
@@
         \includegraphics[width=\textwidth]{figs/CPP/10000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/10000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}

### 2015-01-21 paper.tex
arrumado caption das figuras finais

@@
         \includegraphics[width=\textwidth]{figs/CPP/5000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/5000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/1000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/1000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/500}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/500_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/250}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/250_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/100}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/100_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/50}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
         \includegraphics[width=\textwidth]{figs/CPP/50_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/10000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/10000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/5000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/5000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/1000}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/1000_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/500}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/500_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/250}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/250_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/100}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/100_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/LAD/50}
-    \caption{Distribution of vertex with respect to each measure of activity: in and out degrees and strengths.
-CPP Std library official mailing list.
+    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+Linux Audio Users (LAD) official mailing list.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
@@
         \includegraphics[width=\textwidth]{figs/LAD/50_2}
-    \caption{Distribution of vertex with respect to each measure of activity: compound criteria.
-Red, green and blue are for hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots picture classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black is used to depict how much a vertex was classified in more than section: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+    \caption{Distribution of vertex with respect to compound criteria.
+Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
+The first two plots exhibit classifications that are not functions.
+Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}

### 2015-01-21 paper.tex
repassadas das perguntas do chu e colocado ele como ultimo autor

@@
 %
-%\author{Osvaldo N. Oliveira Jr.}
-%  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
-%  \email{chu@ifsc.usp.br}
-% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+\author{Osvaldo N. Oliveira Jr.}
+  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
+  \email{chu@ifsc.usp.br}
+ \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
@@
 That was understood as a consequence of the observation scale:
-\emph{There are between 20 and 1200 participants in the message window sizes used to derive most results ($ws \in [200,1500]$ messages).
+\emph{There are between 20 and 200 participants in the message window sizes used to derive most results ($ws \in [200,1500]$ messages).
 As peripheral vertexes are abundant and span few degrees, there are more than $\eta$ vertexes with each low degree value.
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
-    \caption{PCA of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
-Table~\ref{compPCA} exposes the composition of principle components.
-Principle component is a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
+    \caption{Scatter plot of vertexes using two principal components from a PCA in the measures space of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
+Table~\ref{compPCA} exposes the composition of principal components.
+Principal component is a weighted average of centrality measures: degrees, strengths and betweenness centrality.
 Second component is mostly clustering coefficient.
-Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
+Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.
+Therefore, the outline of this plot is very similar to Figure~\ref{clust}.}
     \label{PCA}
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
-    \caption{Degree and strength, clustering coefficient, betweenness centrality and symmetry related measures are used for this scatter plot of principal components.
+    \caption{Scatter plot of vertexes using two principal components from a PCA in the measures space of degree, strength, clustering coefficient, betweenness centrality and symmetry related measures.
 Compositions of first three components are in Table~\ref{compPCA2} and measure details in Section~\ref{measures}.
-Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.}
+Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
+Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertexes.}
     \label{PCA2}
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-    \caption{Clustering coefficient versus degree of vertex with a window size of $ws = 1000$ email messages, LAU list.
+    \caption{Clustering coefficient versus degree of vertexes with a window size of $ws = 1000$ email messages, LAU list.
 General layout is in accordance with literature: connected vertexes have low clusterization while higher clusterization is gradually more incident as number of connections is lowered.}
@@
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-The first component is a pondered sum of degree and betweenness centrality.
+The first component is a weighted average of degree and betweenness centrality.
 The second component is mostly clustering coefficient.
@@
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a pondered sum of degree and strength and betweenness centrality.
+First component is a weighted average sum of degree and strength and betweenness centrality.
 The second component is mostly related to clustering coefficient.

### 2015-02-15 paper.tex
padronizados nomes dos autores
adicionadas versoes anotadas das revisoes do chu
feita revisao ateh boa parte dos metodos

@@
 S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP)%\\This line break forced with \textbackslash\textbackslash
-}%
-%
-%\author{Camilo Akimushkin Valencia}
-%  \homepage{http://ifsc.usp.br/}
-%  \email{akimushkin@ifsc.usp.br}
-%  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-%
-
-\author{Vilson Vieira da Silva Junior}
+}
+
+\author{Vilson V. da Silva Jr.}
   \homepage{http://automata.cc/}
@@
 
-\author{Deborah Christina Antunes}
+\author{Deborah C. Antunes}
   \homepage{http://lattes.cnpq.br/1065956470701739}
@@
 
-\author{Marilia Mello Pisani}
+\author{Marilia M. Pisani}
   \homepage{http://lattes.cnpq.br/6738980149860322}
@@
 %  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-%
+
 \author{Osvaldo N. Oliveira Jr.}
@@
 \begin{abstract}
-    This article reports a characterization of interaction networks and its temporal stability.
-Such a task involves a selection of aspects to investigate, which lead to: 1) activity distribution in time and among participants; 2) a sound classification of vertex: peripheral, intermediary and hub sectors; 3) combination of basic measures into components with greater dispersion (PCA).
-While time patterns of activity are not obvious, participant activity follows concentrations expected in scale-free networks.
-Comparison of incident networks with ideal Erd\"os-R\'enyi networks bearing the same number of edges and vertexes reveals a sound criterion for distinguishing sectors on the networks.
-Principal components in the basic measures space revealed interesting and regular patterns of independence and dispersion.
-This includes a ranking of measures that most contribute to dispersion: 1) degree and strength measures, 2) symmetry related quantization, and 3) clusterization.
-Results suggest typologies for these networks and participants.
-Further work include considerations of text production, psychoanalysis inspired typologies, participatory democracy exploitation of observed properties, and better visualization support for network evolution.
+ This article reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance.
+We analyzed the temporal activity and topology evolution of networks in four email lists by considering  window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline.
+Activity in terms of seconds, minutes, days and months was remarkably similar for all the networks.
+Participant activity follows concentrations expected in scale-free networks.
+We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
+The fractions of members in these sectors were essentially the same for all networks and, most importantly, stable over time.
+If strength is used as the criterion for classification, for instance, ca.
+5\% of the vertices are hubs, 15-20\% are intermediary and the remainder compose periphery.
+The metrics  most representative for  data dispersion were found to be centrality-related  (degree, strength and betweenness), followed by symmetry-related,  and then clustering coefficient.
+Degree, strength and clustering cope with different symmetry measures to bear secondary components.
+Again, the importance of these metrics to network topology dispersion was practically the same for all email lists examined, and stable over time.
+The research also resulted on a sketch of a physics-based typology of agents, with proper social and psychological speculations.
+Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed  holds for general human interaction networks, and the approach can be applied to other types of networks.
+Current unfoldings include governance and accountability proposals and implementations, anthropological physics experimentations and the report of quantitative differences of textual production as connectivity of agents changes.
 \end{abstract}
@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics}%Use showkeys class option if keyword
+\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}%Use showkeys class option if keyword
 \maketitle
 
+%Obs. Em geral não se incluem citações como esta abaixo nos artigos, e podemos %ter problema de aceitação se um referee não entender o porquê de colocá-la.
+
 \begin{quotation}
-% `The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.
+`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.
 The fact that human society has been up to now divided into classes affects more than the external relations of men.
 The marks of social repression are left within the individual soul.'
-% 
-`The conception of personality structure is the
-best safeguard against the
-inclination to attribute persistent trends in the
-individual to something
-"innate" or "basic" or "racial" within him. The
-Nazi allegation that natural, biological traits decide the total being of a person
-would not have been such
-a successful political device
-had it not been possible to point to numerous
-instances of relative fixity in human behavior and to
-challenge those who
-thought to explain them on any basis other than a biological one.'
-
+
+% `The conception of personality structure is the
+%best safeguard against the
+% inclination to attribute persistent trends in the
+% individual to something
+% "innate" or "basic" or "racial" within him. The
+% Nazi allegation that natural, biological traits decide the total being of a % person
+% would not have been such
+% a successful political device
+% had it not been possible to point to numerous
+% instances of relative fixity in human behavior and to
+% challenge those who
+% thought to explain them on any basis other than a biological one.'
 \emph{- Adorno et al, 1969, p. 747}
@@
 \subsection{Related work}
-This work approaches network stability through temporal evolution observations.
-This evolution is characterized by a constant number of contiguous messages, of which snapshots are considered to yield a timeline of the network, which is not explored to date, as far as authors know~\cite{netWindowEvo}.
+
+This work approaches network stability through temporal evolution observation.
+This evolution is represented by a timeline of activity snapshots with a constant number of contiguous messages.
+These snapshots yield a succession of networks.
+This approach, although coherent and intuitive, seems not to be explored to date~\cite{netWindowEvo}.
+
 Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
-Moreover, selected exceptions are reported in this section.
-
- The evolution of interaction networks was addressed with a community focus, in a work that ignores the direction of edges~\cite{barabasiEvo}.
-Two topologically different networks are reported to emerge, depending on the frequency of interactions, presenting a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email networks, free-scale properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
-
- Such results are in accordance with phenomena observed in this work.
-Linguistic characterization is being described in a deriving article~\cite{rcText}.
-See Appendix~\ref{sec:fure} for further considerations of related work, with emphasis on the results from research herein reported.
+Exceptions are reported in this section, specially those more closely related to the present article.
+
+The evolution of human interaction networks was addressed with a community focus, where the direction of edges was not taken into account~\cite{barabasiEvo}.
+In another article, two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which could either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email list networks, scale-free properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
+These results are corroborated by phenomena reported in this paper.
+See Appendix~\ref{sec:fure} for further considerations of related work, with emphasis on the results described below and deriving articles dedicated to texts produced by each of the primitive connective sectors~\cite{rcText}, a visualization method for network evolution~\cite{versinus} and an ongoing elaboration of the primitive typology.
 
@@
 These lists cover a variety of topics, mostly technology-related.
-The archive can be described as a corpus with metadata of its messages, including sent time, place, sender name, sender email address.
-GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
-The scripts for gathering and processing GMANE email messages are given in Appendix~\ref{scripts}.
-
-    \subsection{Chosen dataset}
-Four email lists were selected for their diversity, easing initial observance of natural and general properties.
+The archive can be described as a corpus with metadata of its messages, including sent time, place, sender name, and sender email address.
+The GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
+The computer scripts for gathering and processing GMANE email messages are given in Appendix~\ref{scripts}.
+
+Four email lists were selected for their diversity, making it easier to infer general properties.
 \begin{itemize}
@@
 
- The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages exposed in Table~\ref{geralListas}.
+ The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages, as indicated in Table~\ref{geralListas}.
 
@@
 MET notably has the fewer participants and the larger number of threads.
-This relation holds for each of the lists considered: as the number of participants increases,
- the number of threads decreases.}
+This relation holds for each of the lists considered here and a derived article (see Appendix~\ref{sec:fure}): as the number of participants increases, the number of threads decreases.}
   \label{geralListas}
@@
 \section{Characterization methods}
-After immersion with the data and appropriated literature, the following methods for network characterization were chosen: 1) statistics of activity along time; 2) division of network in hubs, intermediary and peripheral vertex; 3) prominence of topological measures; 4) evolutive visualizations and quantitative observations; 5) typological elaborations of networks and participants.
-Each of these methods are described bellow with supporting structures.
-
-    \subsection{Temporal activity statistics}
-Number of messages along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
-These are exhibited as tables in Appendix~\ref{tabTime}.
+The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, with a detailed analysis for time durations from seconds to years; 2)  sectioning of the networks in hubs, intermediary and peripheral vertices; 3) prominence of topological measures, including their time dependence; 4) iterative visualization and data mining of networks; 5) typological elaborations of networks and participants.
+Each of these procedures are described below with supporting structures.
+
+Distribution of activity among participants are in Table~\ref{autores} and indirectly through almost all results of this article, as degree and strength are highly correlated to activity.
+
+  \subsection{Time activity statistics}                          
+Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
+These are exhibited as Tables~\ref{dia}-\ref{ano} in Appendix~\ref{tabTime}.
 Results are outlined in Section~\ref{constDisc}.
@@
 Regarding literature~\cite{bird,newmanCommunityDirected,newmanCommunity2013}, interaction networks can be modeled both weighted or unweighted, both directed or undirected.
-Networks in this article are directed and weighted, considered as more informative among possibilities (directed unweighted, undirected weighted, and undirected unweighted).
-The networks were considered as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
-The reasoning is: if B wrote a response to a message from A, he read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
-Inverting edge direction yields the status network, as B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
+Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e.
+we did not observe directed unweighted, undirected weighted, and undirected unweighted representations of the interaction network.
+The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
+The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
+Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
 This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}.
 Edges in both directions are allowed.
-Each time an interaction occurs, one is added to edge weight.
+Each time an interaction occurs, one is added to the edge weight.
 Self-loops were regarded as non-informative and discarded.
-This networks are described as exhibiting free-scale and small world properties, as expected for a social network~\cite{bird}.
+These networks are reported in literature as exhibiting scale-free and small world properties, as expected for  (some) social networks~\cite{bird,newmanBook}.
 
@@
 A reply message from participant B to a message from participant A is regarded as evidence that B received information from A.
-Multiple messages add ``weight'' to directed edge.
-Further details are in Section~\ref{intNet}.}
+Multiple messages add ``weight'' to a directed edge.
+Further details are given in Section~\ref{intNet}.}
     \label{formationNetwork}
@@
 
-Previous messages on the thread create directed edges from their author to the observed message's author.
 Edges can be created from all antecedent messages on the message-response thread.
-In this work, only the immediate predecessor is linked to new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a connection between $x\rightarrow z$.
-Potential interpretations for this weaker connection are usually common sense, such as: double length, half weight or with one more ``obstacles''.
-This suggests the adoption of centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
-
-        \subsubsection{Sectioning network in periphery, intermediary and hubs classes}\label{sectioning}
-
-\begin{figure}[hb]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figs/fser}
-    \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The later has more
-        intermediary, as the former has more peripheral and hub vertexes. Sections are
-        given by the two intersections $k_1$ and $k_2$ of the connectivity distributions. Characteristic degrees
-    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the three sections considered (periphery, intermediary and hubs).}
-    \label{fig:setores}
-\end{figure}
-
-Because of social networks tendency to have a scale-free distribution of connectivity,
-one can compare it to an Erd\"os-R\'enyi random graph and consider peripheral, intermediary and hub
-sectors~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
+We only linked the immediate antecedent to the new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
+Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''.
+This suggests pertinence of centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
+
+        \subsubsection{Sectioning networks in periphery, intermediary and hubs sectors}\label{sectioning}
+
+Social networks tend to have a scale-free distribution of connectivity, and the primitive sectors (periphery, intermediary and hubs) can be derived from comparison with an Erd\"os-R\'enyi  network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
 The degree distribution $\widetilde{P}(k)$ of an ideal
-scale-free network $\mathcal{N}_f$ with $N$ vertexes and $z$ edges, has less
-average degree nodes when compared with the distribution $P(k)$ of an Erd\"os-R\'enyi
-random graph with the same number of vertexes and edges:
+scale-free network $\mathcal{N}_f$ with $N$ vertices and $z$ edges has less
+average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
+network with the same number of vertices and edges.
+Indeed, we define in this work the intermediary sector of a network to be the set of all the nodes whose degree is less abundant in the real network than on the Erd\"os-R\'enyi model:
 
@@
 If $\mathcal{N}_f$ is directed and has no self-loops, the probability
-of the presence of an edge between two arbitrary vertexes is $p_e=\frac{z}{N(N-1)}$ (see Appendix~\ref{ap:ded}).
-A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertexes and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:
+of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$ (see Appendix~\ref{ap:ded}).
+A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertices and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:
 
@@
 
-The lower degree fat tail is constituted by the border
-vertexes and is called the peripheral sector. 
-The higher degree fat tail is the hub sector.
-
-The arguments behind this classification are: 1) vertexes so connected that they are virtually inexistent in networks whose edges are incident regardless of vertex properties, are correctly associated to the hubs sector.
-Vertexes with very few connections, which are way more abundant than expected in networks whose edges are incident regardless of vertex properties, are correctly assigned to periphery.
-Degree values near average, predicted as the most abundant if connections are created without criteria, and less frequent in free-scale phenomena, are classified to intermediary vertexes.
-
-To ensure statistical validity, bins can be chosen to contain at least $\eta$ vertexes.
-Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer in which degrees $k_i$-$k_{j}$ contain at least $\eta$ vertexes.
+The lower degree fat tail represents the border vertices, i.e.
+the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower and any intermediary sector value of $k$.
+The higher degree fat tail is the hub sector, i.e.
+$\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
+The reasoning for this classification is: 1) vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g.
+without preferential attachment) are correctly associated to the hubs sector.
+Vertices with very few connections, which are way more abundant than expected by pure chance, are assigned to the periphery.
+Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in scale-free phenomena, are classified as intermediary.
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figs/fser}
+    \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
+        intermediary vertices, while the former has more peripheral and hub vertices. Sector borders are
+        given by the two intersections $k_1$ and $k_2$ of the connectivity distributions. Characteristic degrees
+    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the three sectors considered (periphery, intermediary and hubs).}
+    \label{fig:setores}
+\end{figure}
+
+To ensure statistical validity, bins can be chosen to contain at least $\eta$ vertices.
+Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than or equal $k_{j}$.
 This changes equation~\ref{criterio} to:
@@
 
-If instead strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees and strengths, comparisons should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$.
-Results of these criteria for network segmentation are discussed in Section~\ref{subsec:pih}.
-
-Compound criteria is also used for classification of vertex, considering all measures: total, in and out degree and strength.
+If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
+For in and out degrees and strengths, comparison should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$.
+Results of these criteria for network segmentation are discussed in Section~\ref{subsec:pih} and exhibited in Figures~\ref{fig:cpp10000} to~\ref{fig:lad50_} of Appendix~\ref{figures}.
+
+Since different metrics can be used in the segmentation to identify the three types of vertices, various criteria can be defined, e.g.
+with a very stringent criterion according to which a vertex will only be classified as hub if it is so for all the metrics.
 After a careful inspection of possible combinations, these were reduced to six:
@@
 \begin{itemize}
-    \item Exclusivist criteria: vertexes are only classified if the class is the same according to all metrics.
-In this case, vertexes classified does not usually reach 100\%, which is depicted by a black line in Appendix~\ref{figures}.
-    \item Inclusivist criteria: vertex has class given by any of the metrics.
-Therefore, a vertex can belong to more than one class, and class attribution may add more than total number of vertexes, which is indicated by a black line in Appendix~\ref{figures}.
-    \item Exclusivist cascade: a hub is only classified as hub if it is a hub with respect to all metrics.
-Intermediary are the vertexes classified as intermediary or hub with respect to all metrics.
-Remaining vertexes are regarded as peripheral.
-    \item Inclusivist cascade: hubs are vertexes classified as hubs by any of the metrics.
-The remaining vertexes are classified as intermediary if they belong to this category by any of the metrics.
-Vertexes only classified as peripheral are the only considered as such.
-    \item Exclusivist externals: vertexes are only hubs if they are classified as such according to all the metrics.
-The remaining vertexes are classified as peripheral if they fall into hub or peripheral categories by any metric.
-The rest of the vertexes are classified as intermediary.
-    \item Inclusivist externals: hubs are vertexes classified as hubs according to any metric.
-The remaining vertexes will be peripheral if classified as such according to any metric.
-The rest of the vertexes are regarded regarded as intermediary.
+\item Exclusivist criterion:  vertices are only classified if the class is the same according to all metrics.
+In this case, vertices classified (usually) does not reach 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
+        \item Inclusivist criterion: a vertex has the class given by any of the metrics.
+Therefore, a vertex can belong to more than one class, and the percentages of members may add to more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
+    \item Exclusivist cascade: vertices are only classified as hubs if they are hubs according to all metrics.
+Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics.
+The remaining vertices are regarded as peripheral.
+    \item Inclusivist cascade: vertices are hubs if they are so classified according to any of the metrics.
+The remaining vertices are classified as intermediary, if they belong to this category for any of the metrics.
+Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics.
+    \item Exclusivist externals: vertices are only hubs if they are classified as such according to all the metrics.
+The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric.
+The rest of the nodes are classified as intermediary.
+    \item Inclusivist externals: hubs are vertices classified as hubs according to any metric.
+The remaining vertices will be peripheral if they are classified as such according to any metric.
+The rest of the vertices will be intermediary vertices.
 \end{itemize}
 
-These compound criteria, and reduction of possibilities to them, can be formalized in strict mathematical terms.
-This was considered out of the scope of the present article.
-Important here is to notice that the compound criteria can be used to examine network sections in the case of a low number of messages, as will be shown later.
-
-Results from applying this classification method is reported in Section~\ref{subsec:pih}.
-
-        \subsubsection{Topological measures}\label{measures}
-The topology of the networks was characterized by a small selection of the most standard
-measurements of each vertex:
+These compound criteria, and the simplification of possibilities listed above (exclusivist/inclusivist, criterion/cascade/externals) , can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+Important here is to notice that the compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
+
+Results from applying this classification method, i.e.
+the achievement of well defined sectors by comparison of the real network with the Erd\"os R\'enyi model, are reported in Section~\ref{subsec:pih} and Appendix~\ref{figures}.
+
+\subsubsection{Topological measurements}\label{measures}
+
+The topology of the networks was characterized with a small selection of the most standard measurements for each vertex, as follows:
 
 \begin{itemize}
-    \item Degree $d_i$: number edges linked to                                 vertex $i$.
-    \item In-degree $d_i^{in}$: number of edges ending at                      vertex $i$.
-    \item Out-degree $d_i^{out}$: number of edges departing from               vertex $i$.
+    \item Degree $k_i$: number edges linked to vertex $i$.
+    \item In-degree $k_i^{in}$: number of edges ending at vertex $i$.
+    \item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
     \item Strength $s$: sum of weights of all edges linked to vertex $i$.
@@
     \item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
-    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of vertex $i$ that are linked.
-This is the standard clustering coefficient for undirected graphs.
-Usage of other clustering coefficients (e.g.
-for directed graphs) was considering out of scope, as such clustering coefficients are usually employed in special contexts.
+    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
   \item Betweenness centrality $bt_i$: fraction of geodesics that contain the vertex $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
@@
 
-In order to capture asymmetries in the activity of participants, the following metrics were introduced (see Section~\ref{prevalence}):
+In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$ (see Section~\ref{prevalence}):
 
 \begin{itemize}
-    \item Asymmetry of vertex $i$: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
-    \item Mean of edge asymmetry of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
+    \item Asymmetry: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
+    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
 $|J_i|$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of edge asymmetry of vertex $i$: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
-    \item Disequilibrium of vertex $i$: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item Mean of edge disequilibrium of vertex $i$: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of edge disequilibrium of vertex $i$: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
+    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
+    \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
+    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
 \end{itemize}
 
-   \subsection{Evolutive observations}
-Evolution of network was observed within a fixed number of messages (or window size $ws$) that shifts in the message timeline.
- All $ws=$ 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000 were used.
-Within a same $ws$, the number vertexes and edges vary in time, as do other network characteristics.
-Further work should deepen inspection of measure interdependence.
-
-        \subsubsection*{Visualization of network evolution}
-In refining hypotheses, visualization of the network was crucial.
-
+   \subsection{Evolution of the networks}
+The evolution of the networks was observed within a fixed number of messages (or window size $ws$) that shifts in the message timeline.
+The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
+
+        \subsection{Visualization of network evolution}
 The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
-Mapping of various topological measures to glyphs and layouts are being further explored as a parallel research.
-Furthermore, stable aspects of measures prominence along time are captured through mean and standard deviation (see Section~\ref{measures} and Appendix~\ref{sec:pcat}).
-Constant sector sizes along time are observed in a timeline fashion in Appendix~\ref{figures}.
+Such visualization was crucial to guide research into the most important features of network evolution, and prompted us to capture the prominence of topological metrics along time using mean and standard deviations (see Section~\ref{measures} and Appendix~\ref{sec:pcat}), in addition to the size of the three sectors in a timeline fashion (Appendix~\ref{figures}).
+Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures as results were driven.
 
     \subsection{Typological deepening}
-There are other ways to split a network.
-To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius\footnote{Radius is the minimum maximum distance to all nodes.
-Equivalently, the radius is the minimum eccentricity.}.
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter\footnote{Diameter is the maximum geodesic on the network.}.
+There are other ways to split and characterize networks.
+To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
+the radius is the minimum eccentricity).
+In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter(diameter being the maximum geodesic on the network).
 Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with these criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceed 93\% of the nodes~\cite{networkx}.
+Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceed 90\% of the nodes~\cite{networkx}.
 
@@
 In this case, agent activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
-Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with of bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
+Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
 Emails are reported as having a heavy tailed distribution with $\alpha=1$, together with web browsing and library loans~\cite{barabasiHumanDyn}.
@@
 Concerning empirical research,
-Theodor Adorno was a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
-personality traces\footnote{Some of them related to Nazism adoption, antisemitism and potential fascists.}, sometimes depicted as an authoritarian syndrome.
-%Influenced by Social and Psychoanalytic Theories, Adorno et all applied a questionnaire to individuals, from which they reached a position in the the ``F Scale'', to verify etnocentric, conservatory and antidemocratic trends~\cite{adorno}.
-From psychoanalitic interviews and the F Scale, they derived a typology, which gathers prejudice-inclined traces in personality.
-Both, low and high scores are considered with prejudicial traces.
-This typology has nine authoritarian types, the six types with high score in the F Scale: surface resentment, conventional, authoritarian, rebel and psychopath, crank, manipulative; and three of the five types with low score in the F Scale: rigid, protesting, impulsive, easygoing, genuine liberal.
-Each side of the dipole has a rank of intensity that increases as the order written above.
- Other typologies include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling)~\cite{jung}.
+Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
+personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
+
+%Influenced by Social and Psychoanalytic Theories, Adorno et all applied a %questionnaire to individuals, from which they reached a position in the the %``F Scale'', to verify etnocentric, conservatory and antidemocratic %trends~\cite{adorno}.
+From psychoanalitic interviews and the F Scale, they %derived a typology, which gathers prejudice-inclined traces in personality.
+%Both, low and high scores are considered with prejudicial traces.
+This %typology has nine authoritarian types, the six types with high score in the F %Scale: surface resentment, conventional, authoritarian, rebel and psychopath, %crank, manipulative; and three of the five types with low score in the F %Scale: rigid, protesting, impulsive, easygoing, genuine liberal.
+Each side of %the dipole has a rank of intensity that increases as the order written above.
+
+Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
+This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four as dominant, and each mode expressed primarily as in an introverted or extraverted form~\cite{jung}.
 Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
@@
 
-It was considered central to benefit from key human typologies, both by adding descriptions to a type and by further characterizing classes in the terms encountered.
+It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
+A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\`enyi model.
+Also, ethic and moral issues are developed by such legacy.
+For example, Adorno et al.
+conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
 
@@
 \section{Results and discussion}
+
 \subsection{Constancy and discrepancy of activity along time}\label{constDisc}
+One remarkable feature from the analysis is that the activity along time is practically the same for all lists.
+
 \subsubsection{Seconds and minutes}
@@
 Messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$.
-Simulations reach these values, but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
+Simulations reach these values  but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
 
 \subsubsection{Hours of the day}
-Table~\ref{dia} shows how the four lists distribute activity along the day.
-Afternoon was the most active 6h period of the day. Second 12h is $\frac{1}{3}$ more active than first 12h. Even so, activity peak occurs around midday, with a skew toward one hour before noon.
-
-        \subsubsection{Days of the week}
- Weekdays also exhibit an interesting pattern, to which is dedicated Table~\ref{semana}: a decrease of activity on weekends of at least one third, reaching two thirds in extreme cases.
+Higher activity was observed between noon and 6 pm, followed by the time period between 6 pm and midnight.
+Around 2/3 of the whole activity takes place from noon to midnight, as can be seen in Table ?.
+Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
+
+\subsubsection{Days of the week and along the month}
+ Higher activity was observed during weekdays, as expected, especially for the CPP and MET lists (see Table~\ref{semana}).
 
@@
 \subsubsection{Months and larger divisions of the year}
-Table~\ref{ano} is dedicated to activity in months and larger divisions of the year.
-Observation points two periods of more prominent activities: Jun-Aug (MET and LAD), and Dec-Mar (CPP, LAU and LAD), which fit academic calendars, vacations and end-of-year holidays.
+ Activity is concentrated in Jun-Aug for MET and LAD, and from Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
 
 \subsection{Scalable fat-tail structure}\label{subsec:pih}
-
-There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this distribution of activity among participants.
-
-As specified in Section~\ref{sectioning}, in order to classify nodes as hubs, intermediary or peripheral, the incident connectivity distribution $\widetilde{P}(k)$ is compared
-to a connectivity distribution of an ideal Erd\"os R\`enyi network $P(k)$ with the same number of vertexes and edges.
-In the networks inspected for this article (see Section~\ref{intNet}), if degree distribution is used for classification, hubs can reach $10\%$ of all vertexes.
- If instead strength is used for comparison ($\widetilde{P}(k_i=\frac{s_i}{\overline{w}})$) is compared
-to $P(k_i)$), hubs account for approximately $5\%$ of all vertex, i.e. strength classification yields half the number of hubs as plain degree. This results hold for in and out degrees and strengths.
-
-Classification criteria exposed in Section~\ref{sectioning} reveled remarkable stability over time, provided that a sample of at least 200 messages is taken as the window of analysis.
-Specially with 1000 or more messages, stable fractions of $\approx 5\%$ of hubs, $\approx [15-20]\%$ of intermediary and $\approx [75-80]\%$ peripheral vertexes were obtained, which match literature expectations~\cite{secFree}.
-Compound criterion resulted in a greater stability and a window of only 200 messages was already sufficient for the analysis.
-A minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of hub, intermediary and peripheral sections.
+There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this well known and expected distribution of activity among participants.
+
+The distribution of vertices in the three sectors defined in Section~\ref{sectioning} (hubs, intermediary, peripheral) is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
+Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}.
+If, for instance, strength is taken as the criterion to define the sectors, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-20]\%$ are intermediary and $\approx [75-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
+If the degree is used for classification, hubs can reach $10\%$ of all vertices, i.e.
+classification with strength yields half the number of hubs as plain degree.
+These results hold for in and out degrees and strengths.
+Stable distributions can also be obtained for as few as 200 messages if classification of the three sectors is performed with one of the compound criteria established in Section ?.
+In fact, a minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
 This degeneration is critical in the span of 50-100 messages.
-Upon usage of the compound criteria, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the network seem to hold basic structure even with as few as 20-50 messages.
-This indicates that concentration of activity and low-activity participants occurs even with very few messages.
-
-For the histograms used in the classification process, the usage of at least $\eta$ vertexes for each bin did not yield significant differences.
+Even so, using a compound criterion such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold their basic structure even with as few as 20-50 messages.
+This indicates that concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
+A minimum window size for the observation of more general properties can be inferred by monitoring the giant component and the degeneration of hub, intermediary and peripheral sections.
+
+For the histograms used in the classification process (see Section~\ref{sectioning}), the use of at least $\eta$ vertices for each bin did not yield significant differences.
 That was understood as a consequence of the observation scale:
-\emph{There are between 20 and 200 participants in the message window sizes used to derive most results ($ws \in [200,1500]$ messages).
-As peripheral vertexes are abundant and span few degrees, there are more than $\eta$ vertexes with each low degree value.
-Higher degrees cases requires a bit more reasoning.
-Specified $ws$ implicates that each participant is $p \in [0.1\%,0.5\%]$ of all participants.
-Therefore, if incident connectivity is very improbable in an Ed\"os R\`enyi network (less than $p$, the probability that a single participant represent when histogram is normalized to density function), than it is not an intermediary connectivity, but a hub.} This concludes the explanation of why using at least $\eta$ vertexes for each bin did not impact results.
-
-Appendix~\ref{figures} is dedicated to figures of these networks and their evolution.
-
-    \subsection{Prevalence of centrality over asymmetries and asymmetries over clusterization}\label{prevalence}
-
-The principal component (PCA~\cite{pca}) exhibit ponderation of centrality measures: degrees, strengths and betweenness centrality.
+\emph{There are between 20 and 200 participants in the message window sizes used to derive most of the results ($ws \in [200,1500]$ messages).
+As peripheral vertices are abundant and span few degrees, there are more than $\eta$ vertices with each low degree value.
+For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
+Therefore, if incident connectivity is very improbable in an Ed\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
+Therefore, using at least $\eta$ vertices for each bin did not impact the results.}
+
+    \subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
+
+The principal component (PCA~\cite{pca}) exhibited ponderation of centrality measures: degrees, strengths and betweenness centrality.
 Clustering coefficient is presented in almost perfect orthogonality.
-Dispersion is more prevalent in symmetry related measures than clustering coefficient.
-Symmetric and asymmetric edges have been reported as bonded to different roles played by participants and relations~\cite{newmanEvolving}.
-Principal components formation from original measures can be observed in Tables~\ref{compPCA0},~\ref{compPCA} and~\ref{compPCA2}.
-Individual vertexes relation with top two principal components can be appreciated in Figures~\ref{PCA} and~\ref{PCA2}.
-This peculiar first component that consists of the averaged sum of degree, strength and betweenness measures was verified to be incident in virtually all networks with 500 or more messages and most smaller networks (degeneration of basic structure is critical with $ws \approx 50-100$ messages).
+Symmetry of edges have been reported as bonded to different roles played by participants and relations~\cite{newmanEvolving}, and dispersion was more prevalent in symmetry related measures than clustering coefficient.
 This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-When symmetry measures are used in PCA, though, the third component has special contributions of out degree and strength.
-
-As expected, degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-Also trivial in the light of complex network literature: high degree is associated with low clustering coefficient, as shown in Figure~\ref{clust}.
-
-When symmetry metrics are included, the first component remains mostly the same, but clustering coefficient is only relevant to third and fourth components.
-A snapshot of vertexes with respect to these first two principal components is in Figure~\ref{PCA2}.
-This asymmetry and disequilibrium measures revealed as more proper measures to characterization of hubs and intermediaries, as seen in greater spreading of second plot.
-These results are summarized in Table~\ref{compPCA2}.
+Most important is that the contribution from the distinct metrics to network topology variance is very similar for all the networks considered, and did not vary with time.
+This stability in network behavior is remarkable, as will be shown by the very small standard deviations of the contributions from the metrics along time.
+
+The variance is mostly presented by degree, strength and betweenness centrality, as indicated in Tables~\ref{compPCA0} and~\ref{compPCA} for the LAU list (similar results are obtained for the other lists and omitted here for simplicity).
+The standard deviations are small, which means that the first and second components varies little with time.
+ Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
+The corresponding PCA plot for the two first components is shown in Figure~\ref{PCA}, where the vertices have been colored according to the sector they belong.
+As expected peripheral vertices have very low values in the first component and greater dispersion in second component.
+
+Since the clustering coefficient is almost entirely orthogonal to the degree (and other centrality measurements), the plot of clustering coefficient versus degree in Figure~\ref{clust} is similar to the PCA in Figure~\ref{PCA}.
+
+The PCA plot in Figure~\ref{PCA2}, where all metrics were considered,  reflects symmetry-related relevance for the variance.
+This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength).
+It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.
 
@@
         \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
-    \caption{Scatter plot of vertexes using two principal components from a PCA in the measures space of in and out degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
-Table~\ref{compPCA} exposes the composition of principal components.
+    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of in- and out- degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
 Principal component is a weighted average of centrality measures: degrees, strengths and betweenness centrality.
 Second component is mostly clustering coefficient.
-Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.
-Therefore, the outline of this plot is very similar to Figure~\ref{clust}.}
+Table~\ref{compPCA} shows the composition of principal components.
+Similar plots were obtained for all window sizes  $ws\;\in\;[500,10000]$), and for the networks of the other email lists.
+Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
     \label{PCA}
@@
    \centering
+        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
+    \caption{Clustering coefficient versus degree of vertices with a window size of $ws = 1000$ email messages, LAU list.
+The general layout is consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.}
+    \label{clust}
+\end{figure}
+
+\begin{figure} 
+   \centering
         \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
-    \caption{Scatter plot of vertexes using two principal components from a PCA in the measures space of degree, strength, clustering coefficient, betweenness centrality and symmetry related measures.
-Compositions of first three components are in Table~\ref{compPCA2} and measure details in Section~\ref{measures}.
+    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
+The composition of the first three components are shown in Table~\ref{compPCA2} and measure details in Section~\ref{measures}.
 Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
@@
 
-\begin{figure} 
-   \centering
-        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-    \caption{Clustering coefficient versus degree of vertexes with a window size of $ws = 1000$ email messages, LAU list.
-General layout is in accordance with literature: connected vertexes have low clusterization while higher clusterization is gradually more incident as number of connections is lowered.}
-    \label{clust}
-\end{figure}
-
-    \subsection{Primitive typology}
-
-This work aimed at finding common characteristics among (email) interaction networks.
+    \subsection{Primitive typology: the activity of participants from different sectors}\label{sec:pty}
+
+This work is aimed at finding common characteristics among (email) interaction networks.
 Analysis involved primary measures observance and formal criteria for coherent ratios of hub, intermediary and hub sectors.
-Nevertheless, inspection done by visualizations and raw data manipulations suggests agents peculiarities of typological character.
+In this process, inspection done by visualizations and raw data manipulations suggests agents typological peculiarities.
 These are initial observations, which should be further developed as posed in Section~\ref{constDisc}:
@@
 Very stable activity was found on MET hubs, which motivated its integration to this work.
-Literature reports greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
-    \item Typically, hub activity is trivial: they interact as much as possible, in every occasion with everyone.
-Peripheral vertex activity also follows a simple pattern: they interact very rarely, in very few occasions.
-Intermediary vertexes seem responsible for network structure.
-    \item Network operation exhibit modes that seem dictated by intermediary vertexes behavior.
-For example, these can exhibit preferential communication to peripheral, intermediary, or hub vertexes; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
+There are reports in the literature of greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
+    \item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
+The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
+Thereafter, intermediary vertices seem responsible for the network structure.
+For example, intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
     \item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
-These seem as authorities and contrast with participants that respond way more than receive responses.
+These seem as authorities and contrast with participants that respond much more than receive responses.
     \item The most obvious community structure, as observed by high clustering coefficient, is found only in peripheral and intermediary sectors.
@@
 
-Regarding topological aspects of interaction networks, further work should inspect other measures in each of the three connective sectors: hubs, intermediary and peripheral.
+Regarding topological aspects of interaction networks, further work should inspect other measures in each of the three connective sectors: hubs, intermediary and periphery.
 
 Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space.
-Another possibility, specially for a brute-force characterization of such sectors, is to remove vertexes with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
-The subtraction $\widetilde{P}(k)-P(k)$ should result in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertexes.
+Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
+The subtraction $\widetilde{P}(k)-P(k)$ should result in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
 This might support classification of the three sectors by clustering, a more traditional approach to classification.
@@
 
-A related work observed textual production of network sectors~\cite{rcText}.
+A paired article  is dedicated to textual production of the network sectors~\cite{rcText}.
 Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
@@
 used on the epigraph~\cite{adorno}.
-Authors thanks GMANE creators and maintainers, specifically: GMANE is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
+Authors thank GMANE creators and maintainers, specifically: GMANE is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
 Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B.
@@
 \appendix
-\section{Deduction of edge existence probability in a directed network without self-loops}\label{ap:ded}
-Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertexes.
-The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertexes})}$, where $max( \text{number of edges} |\ \text{N vertexes})=2[(N-1)+(N-2)...1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertexes.
+\section{Edge existence probability in a directed network without self-loops}\label{ap:ded}
+Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertices.
+The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertices})}$, where $max( \text{number of edges} |\ \text{N vertices})=2[(N-1)+(N-2)...1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertices.
 Therefore:
 \begin{align}
-    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertexes})}=\nonumber\\
-       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}}=\nonumber\\
+    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertices})} \\       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}} \\
    p_e &=\frac{z}{N(N-1)}
@@
 
+\subsection{Paired article to analyse textual production}
+Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
+Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
+
+\subsection{The Versinus visualization method for evolving complex networks}
+To visualize network evolution, the Versinus method was developed.
+The method is simple, and consists of placing each vertex along a sinusoid and a line.
+Even so, fellow researchers repeatedly asked for an article describing Versinus~\cite{versinus}.
+
+\subsection{Typology of online agents}
+A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
+This is a work under construction, but should be online in near future.
+
 \section{Data and scripts}\label{scripts}
-Messages are downloaded from GMANE database by RSS in the mbox email text format. 
+Messages are downloaded from the GMANE database by RSS in the mbox email text format. 
 They are requested one by one to avoid reaching maximum size of the requests accepted by
@@
 The second component is mostly clustering coefficient.
-First and second components sum more than 95\% of total dispersion.}
+The first and second components represent more than 95\% of total variance.
+The $\lambda$ line holds the percentage of total variance attributed to each component.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
 First component is a weighted average sum of degree and strength and betweenness centrality.
-The second component is mostly related to clustering coefficient.
-The first and second components sum more than 90\% of dispersion.}
+The second component is mostly related to the  clustering coefficient.
+The first and second components represent more than 90\% of the variance.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
   \centering
-  \caption{Principal components formation with measures of symmetry (see Section~\ref{measures}).
+  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-In this case, clusterization is pushed to third component.
-Second component is primarily symmetry measures, but also out degree and strength, and disequilibrium standard deviation.
+In this case, clusterization is pushed to the third principal component.
+The second component is primarily derived from symmetry measurements, but also out degree and strength, and disequilibrium standard deviation.
 Betweenness centrality again has a role similar to degree, but weaker.
-Clusterization component combines with disequilibrium, while asymmetry is combined to out degree and strength.
-Three components has in average 80.36\% of dispersion.}
+The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength.
+The three components have in average 80.36\% of the variance.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
 
-\subsection{Tables for activity along time}\label{tabTime}
+\subsection{Tables for activity along time and among participants}\label{tabTime}
+
+\begin{table}
+    \caption{Distribution of activity among agents.
+First column is dedicated to percentage of messages sent by the most active participant.
+Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
+Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
+The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
+\begin{center}
+    \begin{tabular}{ | l ||  c | c | c | c | }
+        \hline
+        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
+    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
+    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
+    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
+    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
+    \end{tabular}
+\end{center}
+\label{autores}
+\end{table}
+
 \begin{table*}
@@
 
-\begin{table}
-    \caption{Distribution of activity among agents.
-First column is dedicated to percentage of messages sent by the most active participant.
-Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
-\begin{center}
-    \begin{tabular}{ | l ||  c | c | c | c | }
-        \hline
-        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
-    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
-    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
-    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
-    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
-    \end{tabular}
-\end{center}
-\label{autores}
-\end{table}
-
 \clearpage
@@
         \includegraphics[width=\textwidth]{figs/CPP/10000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp10000}
@@
 The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Thus, in the first plot, the fraction of vertices with unique classification is plotted in black.
+On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/5000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp5000}
@@
 The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertexes with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertexes that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
+On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 Compound criteria is described in Section~\ref{sectioning}.}
@@
         \includegraphics[width=\textwidth]{figs/CPP/1000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp1000}
@@
         \includegraphics[width=\textwidth]{figs/CPP/500}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp500}
@@
         \includegraphics[width=\textwidth]{figs/CPP/250}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp250}
@@
         \includegraphics[width=\textwidth]{figs/CPP/100}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp100}
@@
         \includegraphics[width=\textwidth]{figs/CPP/50}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:cpp50}
@@
         \includegraphics[width=\textwidth]{figs/LAD/10000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad10000}
@@
         \includegraphics[width=\textwidth]{figs/LAD/5000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad5000}
@@
         \includegraphics[width=\textwidth]{figs/LAD/1000}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad1000}
@@
         \includegraphics[width=\textwidth]{figs/LAD/500}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad500}
@@
         \includegraphics[width=\textwidth]{figs/LAD/250}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad250}
@@
         \includegraphics[width=\textwidth]{figs/LAD/100}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad100}
@@
         \includegraphics[width=\textwidth]{figs/LAD/50}
-    \caption{Distribution of vertexes with respect to each centrality measure: in and out degrees and strengths.
+    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertexes.}
+On the same graph, green counts the disconnected vertices.}
     \label{fig:lad50}

### 2015-02-16 paper.tex
feita revisao ateh os resultados

@@
    \subsection{Evolution of the networks}
-The evolution of the networks was observed within a fixed number of messages (or window size $ws$) that shifts in the message timeline.
+The evolution of the networks was observed within a fixed number of messages, the window size $ws$, that shifts in the message timeline.
 The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
@@
 The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
-Such visualization was crucial to guide research into the most important features of network evolution, and prompted us to capture the prominence of topological metrics along time using mean and standard deviations (see Section~\ref{measures} and Appendix~\ref{sec:pcat}), in addition to the size of the three sectors in a timeline fashion (Appendix~\ref{figures}).
-Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures as results were driven.
+Such visualizations were crucial to guide research into the most important features of network evolution, and prompted us to capture the prominence of topological metrics along time using mean and standard deviations (see Section~\ref{measures} and Appendix~\ref{sec:pcat}), in addition to the size of the three sectors in a timeline fashion (Appendix~\ref{figures}).
+Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures and results were driven.
 
@@
 the radius is the minimum eccentricity).
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter(diameter being the maximum geodesic on the network).
+In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
 Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceed 90\% of the nodes~\cite{networkx}.
+Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
 
@@
 Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four as dominant, and each mode expressed primarily as in an introverted or extraverted form~\cite{jung}.
+This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
 Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
@@
 It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
-A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\`enyi model.
+A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
 Also, ethic and moral issues are developed by such legacy.
@@
 conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
+Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
 
@@
 \subsubsection{Hours of the day}
-Higher activity was observed between noon and 6 pm, followed by the time period between 6 pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight, as can be seen in Table ?.
+Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
+Around 2/3 of the whole activity takes place from noon to midnight, as can be seen in Table~\ref{dia}.
 Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
 
-\subsubsection{Days of the week and along the month}
- Higher activity was observed during weekdays, as expected, especially for the CPP and MET lists (see Table~\ref{semana}).
+\subsubsection{Days of the week}
+Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
+The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
 
@@
 \subsubsection{Months and larger divisions of the year}
- Activity is concentrated in Jun-Aug for MET and LAD, and from Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
+ Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
 
@@
 
-The distribution of vertices in the three sectors defined in Section~\ref{sectioning} (hubs, intermediary, peripheral) is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
+The distribution of vertices in the three sectors defined in Section~\ref{sectioning} (hubs, intermediary, periphery) is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
 Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}.
@@
 These results hold for in and out degrees and strengths.
-Stable distributions can also be obtained for as few as 200 messages if classification of the three sectors is performed with one of the compound criteria established in Section ?.
+Stable distributions can also be obtained for 200 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}.
 In fact, a minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
 This degeneration is critical in the span of 50-100 messages.
-Even so, using a compound criterion such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold their basic structure even with as few as 20-50 messages.
+Even so, using a compound criterion, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold their basic structure with as few as 20-50 messages.
 This indicates that concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
-A minimum window size for the observation of more general properties can be inferred by monitoring the giant component and the degeneration of hub, intermediary and peripheral sections.
 
@@
 For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
-Therefore, if incident connectivity is very improbable in an Ed\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
+Therefore, if incident connectivity is very improbable in an Erd\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
 Therefore, using at least $\eta$ vertices for each bin did not impact the results.}
@@
     \subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-
-The principal component (PCA~\cite{pca}) exhibited ponderation of centrality measures: degrees, strengths and betweenness centrality.
+The principal component (PCA~\cite{pca}) exhibited a ponderation of centrality measures: degrees, strengths and betweenness centrality.
 Clustering coefficient is presented in almost perfect orthogonality.
@@
 This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-Most important is that the contribution from the distinct metrics to network topology variance is very similar for all the networks considered, and did not vary with time.
-This stability in network behavior is remarkable, as will be shown by the very small standard deviations of the contributions from the metrics along time.
-
-The variance is mostly presented by degree, strength and betweenness centrality, as indicated in Tables~\ref{compPCA0} and~\ref{compPCA} for the LAU list (similar results are obtained for the other lists and omitted here for simplicity).
+The contribution from the distinct metrics to network topology variance is very similar for all the networks considered, and did not vary with time.
+This stability in network behavior is remarkable, as shown by the very small standard deviations of the contributions from the metrics along time in Appendix~\ref{sec:pcat}.
+
+The dispersion is mostly presented by degree, strength and betweenness centrality, as indicated in Tables~\ref{compPCA0} and~\ref{compPCA} for the LAU list (similar results are obtained for the other lists and omitted here for simplicity).
 The standard deviations are small, which means that the first and second components varies little with time.
  Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The corresponding PCA plot for the two first components is shown in Figure~\ref{PCA}, where the vertices have been colored according to the sector they belong.
-As expected peripheral vertices have very low values in the first component and greater dispersion in second component.
-
-Since the clustering coefficient is almost entirely orthogonal to the degree (and other centrality measurements), the plot of clustering coefficient versus degree in Figure~\ref{clust} is similar to the PCA in Figure~\ref{PCA}.
-
-The PCA plot in Figure~\ref{PCA2}, where all metrics were considered,  reflects symmetry-related relevance for the variance.
+The corresponding PCA plot for the two first components is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
+As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component.
+The plot of clustering coefficient versus degree in Figure~\ref{clust} is similar to the PCA in Figure~\ref{PCA}.
+
+The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry-related relevance for the variance.
 This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength).
-It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.
+It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.
 
@@
 Table~\ref{compPCA} shows the composition of principal components.
-Similar plots were obtained for all window sizes  $ws\;\in\;[500,10000]$), and for the networks of the other email lists.
-Similarity to plot in Figure~\ref{clust} was verified within all window sizes considered ($ws\;\in\;[100,10000]$), which exposes a common relation is held by degree, strength and betweenness measures to clustering coefficient.}
+Similar plots were obtained for all window sizes  $ws\;\in\;[500,10000]$, and for the networks of the other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.}
     \label{PCA}
@@
 Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
-Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertexes.}
+Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.}
     \label{PCA2}
@@
 This work is aimed at finding common characteristics among (email) interaction networks.
-Analysis involved primary measures observance and formal criteria for coherent ratios of hub, intermediary and hub sectors.
+Analysis involved primary measures observance and a formal criteria for coherent ratios of hub, intermediary and hub sectors.
 In this process, inspection done by visualizations and raw data manipulations suggests agents typological peculiarities.
-These are initial observations, which should be further developed as posed in Section~\ref{constDisc}:
+These are initial observations, which inspired this article and other ongoing research:
 
@@
 This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts involved in other typologies, such Meyer-Briggs, Pavlov or F-Scale.
-This has no pretension of being a direct result from numeric analysis, it is a description refinement of the found structure, in typological terms.
+This has no pretension of being a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
 Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.
@@
 The noon has $\approx \frac{2}{3}$ of 24h activity.
-}
+}\label{dia}
 \begin{center}
@@
 \end{center}
-\label{dia}
 \end{table*}

### 2015-02-16 paper.tex
feita revisao final, falta somente melhorar as figuras e as tabelas de PCA

@@
 Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures and results were driven.
-
-    \subsection{Typological deepening}
+See Appendix~\ref{sec:fure} for further directions about visualization and text mining concerning the results herein presented, with dedicated articles.
+
+    \subsection{Typological deepening}\label{subsec:typ}
 There are other ways to split and characterize networks.
@@
 
-\section{Conclusions and future work}
-Characterization of interaction networks resulted from stability observations.
+Another typology suggested by results is about the networks themselves.
+In accordance with previous research results~\cite{barabasiEvo}, this and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
+
+\section{Conclusions}
+The characterization of interaction networks resulted from stability observations.
 Along temporal activity statistics, this work reports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning (periphery, intermediary, hubs) relative sizes, evident in the comparison with the Erd\"os-R\'enyi model.
+These results suggested typologies for both agents and the networks.
 
     \subsection{Further work}
-The task of delivering a first and general characterization of chosen interaction networks involved starting a larger effort.
+The task of delivering a first and general characterization of chosen interaction networks involved a larger effort.
 The different aspects covered requires not only different analytical background, but also considerations about textual production and social psychology.
@@
 Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
-The subtraction $\widetilde{P}(k)-P(k)$ should result in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
+The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
 This might support classification of the three sectors by clustering, a more traditional approach to classification.
@@
 A paired article  is dedicated to textual production of the network sectors~\cite{rcText}.
-Resulting knowledge purposes networks and participants tipologization, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
-Stability reported in this article eases tipologization of outliers and more usual participation patterns.
+Resulting knowledge purposes networks and participants typologies, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
+Stability reported in this article eases tipologization of outliers and more usual participation patterns. See Appendix~\ref{sec:fure} for further consideration of works related to presented results.
 
         \subsubsection{Results exploitation}
-Usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa,ensaio}.
+On the technological trend, usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa,ensaio}.
 Further steps involve elaboration and tests of social dynamics that takes advantages of these results.
+These results are also being used for anthropological physics experiments~\cite{anPhy} and knowledge acquisition~\cite{rcText}.
 
@@
 
-\section{Further consideration of related work}\label{sec:fure}
-Unreciprocated edges often exceed 50\%, which matches empirical evidence reported in~\cite{newmanEvolving}.
+\section{Work related to the results}\label{sec:fure}
+Unreciprocated edges often exceed 50\%, which matches empirical evidence~\cite{newmanEvolving}.
 Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
@@
 
-Gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex}.
-This can be further investigated to hold in email lists and in evolving terms as community oriented, non-private interactions takes are drawn from public email groups with hundreds or thousands of participants.
-
-Considered years altogether, tenths of thousands of participants can post on a list.
-The most active lists usually reaches a few thousands of participants.
-Analysis of resulting data might lead to deeper insights in community-related network evolution~\cite{GMANE}.
+Gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and can be further investigated in non-private email lists with hundreds or thousands of participants~\cite{GMANE}.
 
@@
 To visualize network evolution, the Versinus method was developed.
-The method is simple, and consists of placing each vertex along a sinusoid and a line.
-Even so, fellow researchers repeatedly asked for an article describing Versinus~\cite{versinus}.
+The method is simple and consists in the placement of each vertex along a sinusoid and a line.
+In the resulting fixed layout, the time evolution takes place.
+Even so, fellow researchers repeatedly asked for an article describing Versinus, and it was written in 2013~\cite{versinus}.
 
@@
 A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
-This is a work under construction, but should be online in near future.
+This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be published in near future.
 
@@
 The first and second components represent more than 95\% of total variance.
-The $\lambda$ line holds the percentage of total variance attributed to each component.}
+The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
   \begin{tabular}{|l|c|c| c|c| c|c|}\hline
@@
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a weighted average sum of degree and strength and betweenness centrality.
+First component is a weighted average of degree and strength and betweenness centrality.
 The second component is mostly related to the  clustering coefficient.
@@
 In this case, clusterization is pushed to the third principal component.
-The second component is primarily derived from symmetry measurements, but also out degree and strength, and disequilibrium standard deviation.
+The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation.
 Betweenness centrality again has a role similar to degree, but weaker.
@@
 %\tiny
-    \caption{Percentage of activity ($\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
+    \caption{Percentage of activity ($100\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
 Maximum activity rates are in bold.
@@
 \begin{table}[h]
-    \caption{Concentration of activity on days along the week.
+    \caption{Percentage of activity on days along the week.
 Weekend days are at least $\frac{1}{3}$ less active and can reach $\frac{1}{3}$ of activity.

### 2015-02-16 paper.tex
retocada equacoes do apendice A

@@
 Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertices.
-The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertices})}$, where $max( \text{number of edges} |\ \text{N vertices})=2[(N-1)+(N-2)...1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertices.
+The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertices})}$, where $max( \text{number of edges} |\ \text{N vertices})=2[(N-1)+(N-2)+...+1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertices.
 Therefore:
 \begin{align}
-    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertices})} \\       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}} \\
+    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertices})} \nonumber \\
+       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}} \nonumber \\
    p_e &=\frac{z}{N(N-1)}

### 2015-02-16 scripts/setoresRede_teorico.py
arrumada figura dos setores

@@ -1,0 +1,55 @@
+#-*- coding: utf8 -*-
+from __future__ import division
+import numpy as n, pylab as p, networkx as x, random as r, collections as c, string
+from scipy import special # special.binom(x,y) ~ binom(x,y)
+
+__doc__="""Script to plot both scale-free network and random network distributions."""
+
+N=1200
+
+# for the free scale network
+# P(k) = \beta *  k^(-\alpha), with \alpha typically in (2, 3)
+kk=n.arange(N)
+Pk=.10*kk**(-10.5)
+
+# for the random network
+# P(k) = (N-1 k)p^k(1-p)^(N-k)
+p_=.1
+Pk_=special.binom(N-1, kk)*(p_**kk)*((1-p_)**(N-kk))
+
+#p.plot(n.log(kk),n.log(Pk));p.plot(n.log(kk),n.log(Pk),"ro",label="free-scale model")
+#p.plot(n.log(kk),n.log(Pk_));p.plot(n.log(kk),n.log(Pk_),"bo",label="Edos-Renyi model")
+
+#F=p.gcf()
+#F.set_size_inches((10.,3.))
+#F.set_figwidth(10.)
+p.figure(figsize=(10.,3.))
+p.subplots_adjust(left=0.06,bottom=0.32,right=0.99,top=0.86)
+p.plot(n.log(kk),n.log(Pk),label="scale-free model", linewidth=3)
+p.plot(n.log(kk),n.log(Pk_),label=u"Erdös-Rényi model", linewidth=3)
+p.text(0.9,-70,"periphery"  ,size=25)
+p.text(3.6,-100,"intermediary",size=25)
+p.text(5.9,-110,"hubs"      ,size=25)
+p.title(u"Three sections of a scale-free network",size=25)
+p.legend()
+p.xlabel(r"$\log(k)\;\rightarrow$",size=25)
+p.ylabel(r"$\log(P(k))\;\rightarrow$",size=25)
+p.yticks(())
+p.xticks(())
+p.xlim((-0.1,n.log(N)+1))
+p.ylim((-150,1))
+
+# pontos de intersecção:
+x1=3.7; y=-41.16-5
+p.plot((x1,x1),(-1000,y),"r--")
+x2=5.4951; y=-60.-5
+p.plot((x2,x2),(-1000,y),"r--")
[…]

### 2015-03-01 paper.tex
arrumada figura dos setores
arrumado quase tudo para mandar para o chu

@@
 
-%\author{Luciano da Fontoura Costa}
-%  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
-%  \email{ldfcosta@gmail.com}
+%
+%%\author{Luciano da Fontoura Costa}
+%%  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
+%%  \email{ldfcosta@gmail.com}
 %  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
-\author{Osvaldo N. Oliveira Jr.}
-  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
-  \email{chu@ifsc.usp.br}
- \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%\author{Osvaldo N. Oliveira Jr.}
+%  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
+%  \email{chu@ifsc.usp.br}
+% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
@@
 Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed  holds for general human interaction networks, and the approach can be applied to other types of networks.
-Current unfoldings include governance and accountability proposals and implementations, anthropological physics experimentations and the report of quantitative differences of textual production as connectivity of agents changes.
+Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments and the report of quantitative differences of textual production as connectivity of agents changes.
 \end{abstract}
@@
 
-\subsection{Related work}
+\subsection{Related work}\label{sec:related}
 
 This work approaches network stability through temporal evolution observation.
-This evolution is represented by a timeline of activity snapshots with a constant number of contiguous messages.
+The evolution is represented by a timeline of activity snapshots with a constant number of contiguous messages.
 These snapshots yield a succession of networks.
@@
 In email list networks, scale-free properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
-These results are corroborated by phenomena reported in this paper.
-See Appendix~\ref{sec:fure} for further considerations of related work, with emphasis on the results described below and deriving articles dedicated to texts produced by each of the primitive connective sectors~\cite{rcText}, a visualization method for network evolution~\cite{versinus} and an ongoing elaboration of the primitive typology.
+
+Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence~\cite{newmanEvolving}.
+No correlation of topological characteristics and geographical position was found~\cite{barabasiGeo}, therefore geographical incidences was discarded in the present article.
+On the other hand, gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and was not considered in this article as the email messages and addresses have no gender related metadata~\cite{GMANE}.
+Controllability of these networks is also an uncovered issue.
+These has unintuitive properties and might bring into forefront crucial properties~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}, which might be the subject of subsequent research.
+
+The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that the smaller size of MET community is responsible for the stronger hubs observed.
+
+These results are corroborated by phenomena reported in this paper, which is an indicative that the stability reported herein is common to general human interaction networks.
+
+\subsection{Paired article to analyse textual production}\label{subsec:text}
+Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
+Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
+
+\subsection{The Versinus visualization method for evolving complex networks}
+To visualize network evolution, the Versinus method was developed.
+The method is simple and consists in the placement of each vertex along a sinusoid and a line.
+In the resulting fixed layout, the time evolution takes place.
+Even so, fellow researchers repeatedly asked for an article describing Versinus, and therefore it was written in 2013~\cite{versinus}.
+
+\subsection{Typology of online agents}
+A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
+This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be extensively published in near future.
 
@@
 MET notably has the fewer participants and the larger number of threads.
-This relation holds for each of the lists considered here and a derived article (see Appendix~\ref{sec:fure}): as the number of participants increases, the number of threads decreases.}
+This relation holds for each of the lists considered here and a derived article (see Section~\ref{subsec:text}): as the number of participants increases, the number of threads decreases.}
   \label{geralListas}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figs/fser}
+    \includegraphics[width=0.5\textwidth]{figs/fser_}
     \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
@@
 Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures and results were driven.
-See Appendix~\ref{sec:fure} for further directions about visualization and text mining concerning the results herein presented, with dedicated articles.
+See Appendix~\ref{sec:related} for further directions about visualization and text mining concerning the results herein presented, with dedicated articles.
 
@@
 
+Remarkable features from the analysis of the four email lists are:
+\begin{itemize}
+    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}).
+    \item The fraction of participants in each connective sector is stable and can be observed even with very few messages (Section~\ref{subsec:pih}).
+    \item The measures combine in principal components in the same way and with symmetry related measures presenting more dispersion than clustering coefficient (Section~\ref{prevalence}).
+    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
+\end{itemize}
+
+Furthermore, the Appendix holds tables and figures from which such results can be observed.
+
 \subsection{Constancy and discrepancy of activity along time}\label{constDisc}
@@
 
-\subsection{Scalable fat-tail structure}\label{subsec:pih}
+\subsection{Scalable fat-tail structure: constancy of participants fraction in the connective sectors}\label{subsec:pih}
 There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this well known and expected distribution of activity among participants.
@@
     \subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
+The contribution from the distinct metrics to network topology 
+is very similar for all the networks considered, and did not vary with time.
+ This stability in network behavior is remarkable, as can be noticed by the very 
+small standard deviations of the contributions from the metrics along time (Tables~\ref{compPCA0}-\ref{compPCA2}).
+
 The principal component (PCA~\cite{pca}) exhibited a ponderation of centrality measures: degrees, strengths and betweenness centrality.
@@
 The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
-Thereafter, intermediary vertices seem responsible for the network structure.
+Therefore, intermediary vertices seem responsible for the network structure.
 For example, intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
@@
 Resulting knowledge purposes networks and participants typologies, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
-Stability reported in this article eases tipologization of outliers and more usual participation patterns. See Appendix~\ref{sec:fure} for further consideration of works related to presented results.
+Stability reported in this article eases tipologization of outliers and more usual participation patterns. See Section~\ref{sec:related} for the consideration of works related to presented results.
 
@@
 \end{align}
-
-\section{Work related to the results}\label{sec:fure}
-Unreciprocated edges often exceed 50\%, which matches empirical evidence~\cite{newmanEvolving}.
-Although no correlation of topological characteristics and geographical position was found in a pertinent study~\cite{barabasiGeo}, geographical incidences should be present in further refinement of the analysis.
-
-The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that smaller size of MET community is responsible for the stronger hubs observed.
-
-Controllability of these networks is also an uncovered issue.
-These has unintuitive properties and might bring into forefront crucial differences between email interaction networks and interaction networks in Facebook or Twitter~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}.
-
-Gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and can be further investigated in non-private email lists with hundreds or thousands of participants~\cite{GMANE}.
-
-\subsection{Paired article to analyse textual production}
-Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
-Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
-
-\subsection{The Versinus visualization method for evolving complex networks}
-To visualize network evolution, the Versinus method was developed.
-The method is simple and consists in the placement of each vertex along a sinusoid and a line.
-In the resulting fixed layout, the time evolution takes place.
-Even so, fellow researchers repeatedly asked for an article describing Versinus, and it was written in 2013~\cite{versinus}.
-
-\subsection{Typology of online agents}
-A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
-This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be published in near future.
 

### 2015-03-01 paper.tex
adicionados apontamentos do chu e anotacoes minhas enviadas de volta com a revisao

@@
 
-\section{Conclusions}
+\section{Conclusions and further work}
 The characterization of interaction networks resulted from stability observations.
@@
 
-    \subsection{Further work}
+%    \subsection{Further work}
 The task of delivering a first and general characterization of chosen interaction networks involved a larger effort.
 The different aspects covered requires not only different analytical background, but also considerations about textual production and social psychology.
-These are receiving attention within dedicated works and are summarized in this section.
-
-        \subsubsection{Constancy of general characteristics eases tipologization}
+These are receiving attention within dedicated works and are summarized in Section~\ref{sec:related}.
+
+%        \subsubsection{Constancy of general characteristics eases tipologization}
 
 Regarding topological aspects of interaction networks, further work should inspect other measures in each of the three connective sectors: hubs, intermediary and periphery.
-
 Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space.
@@
 Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns.
-Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are convenient uses.
-
+Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are proper uses.
+
+        \subsection{Results exploitation}
 A paired article is dedicated to textual production of the network sectors~\cite{rcText}.
 Resulting knowledge purposes networks and participants typologies, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
-Stability reported in this article eases tipologization of outliers and more usual participation patterns. See Section~\ref{sec:related} for the consideration of works related to presented results.
-
-        \subsubsection{Results exploitation}
+Stability reported in this article eases tipologization of outliers and more usual participation patterns (see Sections~\ref{sec:related} and~\ref{sec:pty}).
+
 On the technological trend, usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa,ensaio}.
-Further steps involve elaboration and tests of social dynamics that takes advantages of these results.
-These results are also being used for anthropological physics experiments~\cite{anPhy} and knowledge acquisition~\cite{rcText}.
+Further steps involve elaboration and tests of social dynamics.
+These results are also being used for anthropological physics experiments~\cite{anPhy}.
 

### 2015-03-01 paper.tex
alteração no titulo de subseção

@@
 
-\subsection{Scalable fat-tail structure: constancy of participants fraction in the connective sectors}\label{subsec:pih}
+\subsection{Scalable fat-tail structure: constancy of membership fractions in the connective sectors}\label{subsec:pih}
 There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this well known and expected distribution of activity among participants.

### 2015-03-31 paper.tex
terminada revisao da introducao

@@
  We analyzed the temporal activity and topology evolution of networks in four email lists by considering window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline.
-Activity in terms of seconds, minutes, days and months was remarkably similar for all the networks.
-Participant activity follows concentrations expected in scale-free networks.
+Activity in terms of seconds and minutes exhibit an uniform pattern, while hours, days and months exhibit stable concentrations.
+ Participant activity follows the expected distribution of scale-free networks. 
  We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
-The fractions of members in these sectors were essentially the same for all networks and, most importantly, stable over time.
-If strength is used as the criterion for classification, for instance, ca.
-5\% of the vertices are hubs, 15-20\% are intermediary and the remainder compose periphery.
-The metrics  most representative for  data dispersion were found to be centrality-related  (degree, strength and betweenness), followed by symmetry-related,  and then clustering coefficient.
-Degree, strength and clustering cope with different symmetry measures to bear secondary components.
-Again, the importance of these metrics to network topology dispersion was practically the same for all email lists examined, and stable over time.
-The research also resulted on a sketch of a physics-based typology of agents, with proper social and psychological speculations.
-Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed  holds for general human interaction networks, and the approach can be applied to other types of networks.
-Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments and the report of quantitative differences of textual production as connectivity of agents changes.
+ Approximately 5\% of the vertices are hubs, 15-35\% are intermediary and the remainder belongs periphery. 
+ The metrics that most contribute to data dispersion were found to be centrality-related (degree, strength and betweenness), followed by symmetry-related, and then clustering coefficient. 
+ The results also embrace a sketch of a physics-based typology of agents, with proper social and psychological speculations.
+ Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed are also present in general human interaction networks.
+ Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments, audiovisual reconnaissance, and the report of quantitative differences of textual production as connectivity of agents changes.
 \end{abstract}
@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}%Use showkeys class option if keyword
+\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}
 \maketitle
-
-%Obs. Em geral não se incluem citações como esta abaixo nos artigos, e podemos %ter problema de aceitação se um referee não entender o porquê de colocá-la.
 
@@
 The marks of social repression are left within the individual soul.'
-
 % `The conception of personality structure is the
@@
 
-\section{\label{sec:into}Introduction}
-The present work is aimed at finding common characteristics among (email) interaction networks.
-This includes observations along time, which imply network evolution, a field that has received dedicated attention from the research community
+\section{Introduction}\label{sec:into}
+Human interaction networks have been studied long before modern computers.
+Antecedents can be found in nineteenth century works, while the foundation of
+the social network analysis field is generally attributed to the psychiatrist Jacob Moreno~\cite{newmanBook}.
+More recently, the continuous increase of stored data related to human interaction has stimulated research, and overall characterization is scattered in literature.
+
+The present work is aimed at finding common and stable characteristics among (email) interaction networks.
+This includes observations along time, which imply network evolution, a field that has received attention from the research community
 for more than a decade~\cite{barabasiEvo,newmanEvolving}.
-
-While significant measures will depend on the model and system characteristics~\cite{newmanStru, newmanWeight},
-this work considers only directed, weighted and human interaction networks.
-Undirected and unweighted representation of such networks is also found in the literature and can be obtained by simplification~\cite{GMANE2}.
-
-Text mining and typologies of online participants benefit from the results here presented~\cite{rcText,rcTipo}. 
+We consider only directed, weighted and human interaction networks.
+Undirected and unweighted representations of such networks are also found in the literature and can be obtained by simplification~\cite{GMANE2}.
+Meaningful measures depend heavily on the model and system characteristics~\cite{newmanStru, newmanWeight}, therefore the characterization is restricted to most basic metrics of centrality, clustering and symmetry.
+
 Although all networks considered originated from email lists,
@@
  such as observed in online platforms (e.g. LinkedIn, Facebook, Twitter).
+Indeed, text mining routines and typologies of online participants benefit from the results here presented~\cite{rcText,rcTipo}.
+
+ Next subsection exposes related work.
+Section~\ref{sec:data} details data while Section~\ref{sec:carac} is dedicated to the characterization methods.
+Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work.
+Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity.
+Detailed tables and plots are bundled in Support Information, and are cited upon need.
 
@@
 
-\section{Data description}
+\section{Data description}\label{sec:data}
     \subsection{Email lists and messages}
@@
 
-\section{Characterization methods}
+\section{Characterization methods}\label{sec:carac}
 The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, with a detailed analysis for time durations from seconds to years; 2)  sectioning of the networks in hubs, intermediary and peripheral vertices; 3) prominence of topological measures, including their time dependence; 4) iterative visualization and data mining of networks; 5) typological elaborations of networks and participants.
@@
 \begin{itemize}
-    \item Degree $k_i$: number edges linked to vertex $i$.
-    \item In-degree $k_i^{in}$: number of edges ending at vertex $i$.
-    \item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
+    \item Degree $d_i$: number edges linked to vertex $i$.
+    \item In-degree $d_i^{in}$: number of edges ending at vertex $i$.
+    \item Out-degree $d_i^{out}$: number of edges departing from vertex $i$.
     \item Strength $s$: sum of weights of all edges linked to vertex $i$.
@@
     \item Asymmetry: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
-    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$.
+    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=d_i}$.
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
-$|J_i|$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$
+$|J_i|=d_i$ is the number of neighbors of vertex $i$.
+    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{d_i}  }$
     \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$
+    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{d_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{d_i}}$
 \end{itemize}
@@
 
-\section{Results and discussion}
+\section{Results and discussion}\label{sec:results}
 
@@
 
-\section{Conclusions and further work}
+\section{Conclusions and further work}\label{sec:conc}
 The characterization of interaction networks resulted from stability observations.
@@
 Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
+\section{Circular statistics}\label{ap:circular}
 \section{Tables}\label{sectables}

### 2015-03-31 paper.tex
arrumado mais um pouco a intro

@@
 the social network analysis field is generally attributed to the psychiatrist Jacob Moreno~\cite{newmanBook}.
-More recently, the continuous increase of stored data related to human interaction has stimulated research, and overall characterization is scattered in literature.
+More recently, the continuous increase of stored data related to human interaction has stimulated research,
+and overall characterization is scattered in literature.
+
+Reports on complex networks are usually focused on a particular measure or functionality, such
+as accessibility measure or detecting communities~\cite{access,newmanModularity}.
+Although there are research on networks arisen from specific phenomena, such as social or protein networks, there are few attempts for a general characterization and even less endeavors that take in account the temporal dynamic.
 
 The present work is aimed at finding common and stable characteristics among (email) interaction networks.
-This includes observations along time, which imply network evolution, a field that has received attention from the research community
+This is achieved through observations along time, which imply network evolution, a field that has received attention from the research community
 for more than a decade~\cite{barabasiEvo,newmanEvolving}.
+The evolution is represented herein by a timeline of activity snapshots with a constant number of contiguous messages.
+These snapshots yield a succession of networks.
+This approach, albeit coherent and intuitive, seems not to be explored to date~\cite{netWindowEvo}.
+
 We consider only directed, weighted and human interaction networks.
 Undirected and unweighted representations of such networks are also found in the literature and can be obtained by simplification~\cite{GMANE2}.
-Meaningful measures depend heavily on the model and system characteristics~\cite{newmanStru, newmanWeight}, therefore the characterization is restricted to most basic metrics of centrality, clustering and symmetry.
+Meaningful measures depend heavily on the model and system characteristics~\cite{newmanStru, newmanWeight}, therefore the most basic (and ubiquitous, and fundamental) metrics of centrality, clustering and symmetry are analyzed altogether in principal components formation.
 
@@
 \subsection{Related work}\label{sec:related}
-
-This work approaches network stability through temporal evolution observation.
-The evolution is represented by a timeline of activity snapshots with a constant number of contiguous messages.
-These snapshots yield a succession of networks.
-This approach, although coherent and intuitive, seems not to be explored to date~\cite{netWindowEvo}.
 

### 2015-04-01 paper.tex
revisao feita ateh tipologia na metodologia

@@
 as accessibility measure or detecting communities~\cite{access,newmanModularity}.
-Although there are research on networks arisen from specific phenomena, such as social or protein networks, there are few attempts for a general characterization and even less endeavors that take in account the temporal dynamic.
+Although there are research on networks arisen from specific phenomenon, such as social or protein networks, there are few attempts for a general characterization and even less endeavors that take in account the temporal dynamic.
 
@@
 Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity.
-Detailed tables and plots are bundled in Support Information, and are cited upon need.
+Detailed tables and figures are bundled in Support Information, and are cited upon need.
 
 \subsection{Related work}\label{sec:related}
-
 Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
-Exceptions are reported in this section, specially those more closely related to the present article.
+Exceptions are reported in this section, with emphasis on those more closely related to the present article.
 
@@
 
-Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence~\cite{newmanEvolving}.
-No correlation of topological characteristics and geographical position was found~\cite{barabasiGeo}, therefore geographical incidences was discarded in the present article.
+Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence from literature~\cite{newmanEvolving}.
+No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were discarded in the present article.
 On the other hand, gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and was not considered in this article as the email messages and addresses have no gender related metadata~\cite{GMANE}.
-Controllability of these networks is also an uncovered issue.
-These has unintuitive properties and might bring into forefront crucial properties~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}, which might be the subject of subsequent research.
-
-The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that the smaller size of MET community is responsible for the stronger hubs observed.
-
-These results are corroborated by phenomena reported in this paper, which is an indicative that the stability reported herein is common to general human interaction networks.
-
-\subsection{Paired article to analyse textual production}\label{subsec:text}
-Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
-Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
-
-\subsection{The Versinus visualization method for evolving complex networks}
-To visualize network evolution, the Versinus method was developed.
-The method is simple and consists in the placement of each vertex along a sinusoid and a line.
-In the resulting fixed layout, the time evolution takes place.
-Even so, fellow researchers repeatedly asked for an article describing Versinus, and therefore it was written in 2013~\cite{versinus}.
-
-\subsection{Typology of online agents}
-A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
-This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be extensively published in near future.
-
-\section{Data description}\label{sec:data}
-    \subsection{Email lists and messages}
+
+There are research reports based on the results presented in this article, which are exposed in Section~\ref{sec:further}.
+
+\section{Data description: email lists and messages}\label{sec:data}
+
 Email list messages were obtained from
@@
 The GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
-The computer scripts for gathering and processing GMANE email messages are given in Appendix~\ref{scripts}.
-
-Four email lists were selected for their diversity, making it easier to infer general properties.
+
+Although many lists were analyzed, four email lists were selected for thorough observation, making it easier to infer general properties.
+These lists, picked as representing both a diverse set and usual lists, are:
 \begin{itemize}
-    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}.
-Dominated by participants with hybrid artistic and technological interests.
-Participants are from different countries, and English is the language used the most.
+    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries.
+English is the language used the most.
 Abbreviated as LAU from now on.
-    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}.
-Participants are from different countries, and English is the language used the most.
+    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants are from different countries, and English is the language used the most.
 A more technical and less active version of LAU.
 Abbreviated LAD from now on.
-    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}.
-Dominated by specialized computer programmers.
-Participants are from different countries, and English is the language used the most.
+    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries.
+English is the language used the most.
 Abbreviated as CPP from now on.
-    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}.
-Dominated by Brazilian activists and digital culture interests.
-Participants are mostly Brazilians, and Portuguese is the most used language, although Spanish and English are also incident.
+    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests.
+Portuguese is the most used language, although Spanish and English are also incident.
 Abbreviated MET from now on.
@@
 
- The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages, as indicated in Table~\ref{geralListas}.
+ The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
 
@@
 $\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
-MET notably has the fewer participants and the larger number of threads.
-This relation holds for each of the lists considered here and a derived article (see Section~\ref{subsec:text}): as the number of participants increases, the number of threads decreases.}
-  \label{geralListas}
+A relation holds for all lists carefully considered: as the number of participants increases, the number of threads decreases.
+This underpins a typology sketch of networks, as exposed in Section~\ref{sec:pty}.}
+  \label{tab:genLists}
   \begin{tabular}{|l|c|c|c|c|c|}\hline
@@
 \section{Characterization methods}\label{sec:carac}
-The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, with a detailed analysis for time durations from seconds to years; 2)  sectioning of the networks in hubs, intermediary and peripheral vertices; 3) prominence of topological measures, including their time dependence; 4) iterative visualization and data mining of networks; 5) typological elaborations of networks and participants.
-Each of these procedures are described below with supporting structures.
-
-Distribution of activity among participants are in Table~\ref{autores} and indirectly through almost all results of this article, as degree and strength are highly correlated to activity.
+The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, from seconds to years; 2) sectioning of the networks in hubs, intermediary and periphery vertices; 3) topological measures prominence regarding dispersion; 4) iterative visualization and data mining; 5) typological speculation about networks and participants.
+Each of these procedures are described below.
+
+Distribution of activity among participants received little dedicated attention, but is considered through almost all results of this article, as degree and strength are highly correlated to activity.
 
@@
   Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
-These are exhibited as Tables~\ref{dia}-\ref{ano} in Appendix~\ref{tabTime}.
-Results are outlined in Section~\ref{constDisc}.
+This resulted in histograms from which obvious patterns can be observed.
+Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution.
+Mean and standard deviations in time were taken using circular statistics, as described in Appendix~\ref{ap:circular}.
 
 \subsection{Interaction network}\label{intNet}
-Regarding literature~\cite{bird,newmanCommunityDirected,newmanCommunity2013}, interaction networks can be modeled both weighted or unweighted, both directed or undirected.
+Interaction networks can be modeled both weighted or unweighted, both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
 Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e.
-we did not observe directed unweighted, undirected weighted, and undirected unweighted representations of the interaction network.
+we did not observe directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks.
 The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
@@
 Self-loops were regarded as non-informative and discarded.
-These networks are reported in literature as exhibiting scale-free and small world properties, as expected for  (some) social networks~\cite{bird,newmanBook}.
-
-\begin{figure}[hb]
+These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}.
+
+\begin{figure}[!h]
     \centering
@@
 Each vertex represents a participant.
-A reply message from participant B to a message from participant A is regarded as evidence that B received information from A.
+A reply message from participant B to a message from participant A is regarded as evidence that B received information from A and yields a directed edge.
 Multiple messages add ``weight'' to a directed edge.
@@
 
-Edges can be created from all antecedent messages on the message-response thread.
-We only linked the immediate antecedent to the new message's author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
+Edges can be created from all antecedent message authors on the message-response thread to each message author.
+We only linked the immediate antecedent to the new message author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
 Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''.
@@
 
-        \subsubsection{Sectioning networks in periphery, intermediary and hubs sectors}\label{sectioning}
-
-Social networks tend to have a scale-free distribution of connectivity, and the primitive sectors (periphery, intermediary and hubs) can be derived from comparison with an Erd\"os-R\'enyi  network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
+%\subsubsection{Sectioning of networks in periphery, intermediary and hubs sectors}\label{sectioning}
+\subsubsection{Erd\"os sectioning}\label{sectioning}
+In scale-free networks, the periphery, intermediary and hubs sectors can be derived from comparison with an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
+As a shorthand, this procedure is from now on called \emph{Erd\"os sectioning}, and the resulting sectors the \emph{Erd\"os sectors} or \emph{primitive sectors}.
+
 The degree distribution $\widetilde{P}(k)$ of an ideal
@@
 The lower degree fat tail represents the border vertices, i.e.
-the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower and any intermediary sector value of $k$.
+the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
 The higher degree fat tail is the hub sector, i.e.
@@
 Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in scale-free phenomena, are classified as intermediary.
+
 \begin{figure}[!h]
@@
     \includegraphics[width=0.5\textwidth]{figs/fser_}
-    \caption{Degree distribution on scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
+    \caption{Degree distribution of scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
         intermediary vertices, while the former has more peripheral and hub vertices. Sector borders are
         given by the two intersections $k_1$ and $k_2$ of the connectivity distributions. Characteristic degrees
-    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the three sectors considered (periphery, intermediary and hubs).}
+    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
     \label{fig:setores}
@@
 
-To ensure statistical validity, bins can be chosen to contain at least $\eta$ vertices.
+To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network.
 Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than or equal $k_{j}$.
@@
 
-Since different metrics can be used in the segmentation to identify the three types of vertices, various criteria can be defined, e.g.
-with a very stringent criterion according to which a vertex will only be classified as hub if it is so for all the metrics.
-After a careful inspection of possible combinations, these were reduced to six:
+Since different metrics can be used to identify the three types of vertices, compound criteria can be defined.
+For example, a very stringent criterion can be used, according to which a vertex will only be regarded as pertaining to a sector if it is so for all the metrics.
+After a careful consideration of possible combinations, these were reduced to six:
 
@@
     \item Inclusivist criterion: a vertex has the class given by any of the metrics.
-Therefore, a vertex can belong to more than one class, and the percentages of members may add to more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
+Therefore, a vertex may belong to more than one class, and total members may add more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
     \item Exclusivist cascade: vertices are only classified as hubs if they are hubs according to all metrics.
@@
 The remaining vertices are regarded as peripheral.
-    \item Inclusivist cascade: vertices are hubs if they are so classified according to any of the metrics.
-The remaining vertices are classified as intermediary, if they belong to this category for any of the metrics.
+    \item Inclusivist cascade: vertices are hubs if they are classified as so according to any of the metrics.
+The remaining vertices are classified as intermediary if they belong to this category for any of the metrics.
 Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics.
@@
 
-These compound criteria, and the simplification of possibilities listed above (exclusivist/inclusivist, criterion/cascade/externals) , can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
-Important here is to notice that the compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
-
-Results from applying this classification method, i.e.
+These compound criteria, and the simplification of possibilities listed above (exclusivist/inclusivist, criterion/cascade/externals), can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+Important here is to notice that compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
+
+Results from applying the Erd\"os sectioning, i.e.
 the achievement of well defined sectors by comparison of the real network with the Erd\"os R\'enyi model, are reported in Section~\ref{subsec:pih} and Appendix~\ref{figures}.
 
-\subsubsection{Topological measurements}\label{measures}
-
-The topology of the networks was characterized with a small selection of the most standard measurements for each vertex, as follows:
+\subsubsection{Topology metrics}\label{measures}
+
+The topology of the networks was characterized with a small selection of the most basic and fundamental measurements for each vertex, as follows:
 
 \begin{itemize}
-    \item Degree $d_i$: number edges linked to vertex $i$.
+    \item Degree $d_i$: number of edges linked to vertex $i$.
     \item In-degree $d_i^{in}$: number of edges ending at vertex $i$.
@@
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
-$|J_i|=d_i$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{d_i}  }$
+$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=d_i$ is the number of neighbors of vertex $i$.
+    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{d_i}  }$.
     \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
     \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{d_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{d_i}}$
+    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{d_i}}$.
 \end{itemize}
@@
    \subsection{Evolution of the networks}
-The evolution of the networks was observed within a fixed number of messages, the window size $ws$, that shifts in the message timeline.
+   The evolution of the networks was observed within a fixed number of messages, the window size $ws$.
+This same number of contiguous messages is considered with different shifts in the message timeline to obtain snapshots.
+Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and take topological measures (see Section~\ref{measures}).
+Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological thinking (see Section~\ref{subsec:typ}).
 The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
 
-        \subsection{Visualization of network evolution}
+\subsection{Visualization of network evolution}\label{sec:viz}
 The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
@@
 In accordance with previous research results~\cite{barabasiEvo}, this and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
+The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that the smaller size of MET community is responsible for the stronger hubs observed.
 
@@
 
-%    \subsection{Further work}
 The task of delivering a first and general characterization of chosen interaction networks involved a larger effort.
@@
 These are receiving attention within dedicated works and are summarized in Section~\ref{sec:related}.
-
-%        \subsubsection{Constancy of general characteristics eases tipologization}
 
@@
 
-        \subsection{Results exploitation}
+\subsection{Further work}\label{sec:further}
 A paired article is dedicated to textual production of the network sectors~\cite{rcText}.
@@
 These results are also being used for anthropological physics experiments~\cite{anPhy}.
+
+Controllability of these networks have unintuitive properties and might bring into forefront crucial properties~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}, and might be the subject of subsequent research.
+
+Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
+Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
+
+To visualize network evolution, the Versinus method was developed.
+The method is simple and consists in the placement of each vertex along a sinusoid and a line.
+In the resulting fixed layout, the time evolution takes place.
+Even so, fellow researchers repeatedly asked for an article describing Versinus, and therefore it was written in 2013~\cite{versinus}.
+
+A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
+This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be extensively published in near future.
 

### 2015-04-02 paper.tex
feito basico do seccionamento de erdos

@@
 
+Typologies are the canon of scientific literature for classification of human agents~\cite{typCanon}.
+Consequently, it was considered central to benefit from human typologies.
+This is a complex matter carefully addressed in methodology, results and appendix.
+
 Although all networks considered originated from email lists,
@@
 Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work.
-Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity.
+Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity, and Appendix~\ref{ap:typ} further introduces the typological approach.
 Detailed tables and figures are bundled in Support Information, and are cited upon need.
@@
 The evolution of human interaction networks was addressed with a community focus, where the direction of edges was not taken into account~\cite{barabasiEvo}.
-In another article, two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which could either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email list networks, scale-free properties were verified~\cite{bird}, and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
+In another article, two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email list networks, scale-free properties were reported with $\alpha=1$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
 
@@
         intermediary vertices, while the former has more peripheral and hub vertices. Sector borders are
-        given by the two intersections $k_1$ and $k_2$ of the connectivity distributions. Characteristic degrees
-    are in compact intervals of degree: $[0,k_1]$, $(k_1,k_2]$, $(k_2,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
+        given by the two intersections $k_\Leftarrow$ and $k_\Rightarrow$ of the connectivity distributions. Characteristic degrees
+    are in compact intervals of degree: $[0,k_\Leftarrow]$, $(k_\Leftarrow,k_\Rightarrow]$, $(k_\Rightarrow,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
     \label{fig:setores}
@@
 If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees and strengths, comparison should be made with $\kappa_i=2k_i^{in}$, $\kappa_i=2k_i^{out}$, $\kappa_i=2\frac{s_i^{in}}{\overline{w}}$ and $\kappa_i=2\frac{s_i^{out}}{\overline{w}}$.
+For in and out degrees and strengths, comparison of the real network should be made with:
+\begin{equation}
+\hat{P}(k^{way})=\binom{N-1}{k^{way}}p_e^k(1-p_e)^{N-1-k^{way}}
+\end{equation}
+
+\noindent where \emph{way} can be \emph{in} or \emph{out}.
+In and out strengths are divided by $\overline{w}$ and compared also using $\hat{P}$.
+Notice that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$.
+
+In other words, let $\sigma$ and $\phi$ be integers in the intervals $1 \leq \sigma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\sigma}\}$ have Erd\"os sectors $E_{\sigma}= \{e_{\sigma, \phi} \}$ defined as:
+
+\begin{alignat}{3}
+e_{\sigma,1}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}\geq&&\overline{k}_{\sigma,i}\} \\
+e_{\sigma,2}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}<\;&&\overline{k}_{\sigma,i}\leq\overline{k}_{\sigma,\Rightarrow}\} \\ 
+e_{\sigma,3}&=\{\;i\;|\;&&\overline{k}_{\sigma,i}<\overline{k}_{\sigma,\Rightarrow}\}
+\end{alignat}
+
+\noindent where $\{\overline{k}_{\sigma,i}\}$ is defined as:
+
+\begin{equation}
+\begin{split}
+\overline{k}_{1,i}&=k_i \\
+\overline{k}_{2,i}&=k_i^{in} \\
+\overline{k}_{3,i}&=k_i^{out} \\
+\overline{k}_{4,i}&=\frac{s_i}{\overline{w}} \\
+\overline{k}_{5,i}&=\frac{s_i^{in}}{\overline{w}} \\
+\overline{k}_{6,i}&=\frac{s_i^{out}}{\overline{w}} \\
+\end{split}
+\end{equation}
+
 Results of these criteria for network segmentation are discussed in Section~\ref{subsec:pih} and exhibited in Figures~\ref{fig:cpp10000} to~\ref{fig:lad50_} of Appendix~\ref{figures}.
@@
 
-These compound criteria, and the simplification of possibilities listed above (exclusivist/inclusivist, criterion/cascade/externals), can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+These compound criteria, and the simplification of all the compound possibilities as listed above (into exclusivist/inclusivist criterion/cascade/externals), can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (inverted order of precedence).
 Important here is to notice that compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
@@
 
-\subsubsection{Topology metrics}\label{measures}
-
-The topology of the networks was characterized with a small selection of the most basic and fundamental measurements for each vertex, as follows:
+\subsubsection{Topology metrics for Principal Component Analysis}\label{measures}
+
+The topology of the networks was observed by Principal Component Analysis (PCA~\cite{pca}) outcomes
+with a small selection of the most basic and fundamental measurements for each vertex:
 
 \begin{itemize}
-    \item Degree $d_i$: number of edges linked to vertex $i$.
-    \item In-degree $d_i^{in}$: number of edges ending at vertex $i$.
-    \item Out-degree $d_i^{out}$: number of edges departing from vertex $i$.
+    \item Degree     $k_i$: number of edges linked to vertex $i$.
+    \item In-degree  $k_i^{in}$: number of edges ending at vertex $i$.
+    \item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
     \item Strength $s$: sum of weights of all edges linked to vertex $i$.
@@
 \begin{itemize}
-    \item Asymmetry: $asy_i=\frac{d_i^{in}-d_i^{out}}{d_i}$.
-    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=d_i}$.
+    \item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
+    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$.
 Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
-$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=d_i$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{d_i}  }$.
+$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
+    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
     \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{d_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{d_i}}$.
+    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
 \end{itemize}
@@
 This same number of contiguous messages is considered with different shifts in the message timeline to obtain snapshots.
-Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and take topological measures (see Section~\ref{measures}).
+Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and to perform PCA with topological measures (see Section~\ref{measures}).
 Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological thinking (see Section~\ref{subsec:typ}).
@@
 
-    \subsection{Typological deepening}\label{subsec:typ}
-There are other ways to split and characterize networks.
-To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
-the radius is the minimum eccentricity).
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
-Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
-
-Models of human dynamics can be used to predict and classify activity.
-In this case, agent activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
-Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
-Emails are reported as having a heavy tailed distribution with $\alpha=1$, together with web browsing and library loans~\cite{barabasiHumanDyn}.
-
-Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
-Concerning empirical research,
-Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
-personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
-
-%Influenced by Social and Psychoanalytic Theories, Adorno et all applied a %questionnaire to individuals, from which they reached a position in the the %``F Scale'', to verify etnocentric, conservatory and antidemocratic %trends~\cite{adorno}.
-From psychoanalitic interviews and the F Scale, they %derived a typology, which gathers prejudice-inclined traces in personality.
-%Both, low and high scores are considered with prejudicial traces.
-This %typology has nine authoritarian types, the six types with high score in the F %Scale: surface resentment, conventional, authoritarian, rebel and psychopath, %crank, manipulative; and three of the five types with low score in the F %Scale: rigid, protesting, impulsive, easygoing, genuine liberal.
-Each side of %the dipole has a rank of intensity that increases as the order written above.
-
-Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
-Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
-Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
-
-It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
-A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
-Also, ethic and moral issues are developed by such legacy.
-For example, Adorno et al.
-conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
-Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
-
-%    \subsection{Other analisys of interest}
-%
-%        \subsubsection{Textual content}
-%        \subsubsection{Geographical localization}
+\subsection{Typology speculation}\label{subsec:typ}
+Qualitative, typological speculations were the result of all the methodology.
+More specifically, the Erd\"os sectors was regarded as yielding a \emph{primitive typology} of human agents in social contexts with at least dozens of participants.
+Visualization and data mining were efficient in fueling speculations about the qualities of each type. 
+The stability found in principal components of topological measures suggested that this primitive typology is general for (practically) all human contexts.
+Also, the first author is a developer with cultural interests and has been directly and indirectly part of these communities for more than a decade, which gave further safety about assumptions.
+Care was taken not to regard this acquaintanceship as primary source or as definitive evidence.
+The interested reader should see Appendix~\ref{ap:typ} for further context, 
+as it might be considered audacious to bridge from a physics-based, 
+quantitative classification to a qualitative and speculative approach.
 
@@
 \begin{itemize}
-    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}).
-    \item The fraction of participants in each connective sector is stable and can be observed even with very few messages (Section~\ref{subsec:pih}).
-    \item The measures combine in principal components in the same way and with symmetry related measures presenting more dispersion than clustering coefficient (Section~\ref{prevalence}).
+    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and exhibited noticeable patterns.
+    \item The fraction of participants in each Erd\"os sector is stable and can be observed even with very few messages (Section~\ref{subsec:pih}).
+    \item The measures combine into principal components in the same way for all lists and all snapshots. 
+	    Furthermore, symmetry related measures, defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
     \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
@@
 
-Furthermore, the Appendix holds tables and figures from which such results can be observed.
+The reminder of this section contains a detailed discussion of these results and
+Support Information holds tables and figures from which such results can be observed.
 
 \subsection{Constancy and discrepancy of activity along time}\label{constDisc}
-One remarkable feature from the analysis is that the activity along time is practically the same for all lists.
+One remarkable feature from the analysis is that the activity 
+along time is practically the same for all lists. Circular mean and standard deviation
+for each time scale is given in Table~\ref{tab:circ}.
+
+\begin{table*}[t]
+\scriptsize
+\caption{Circular mean and standard deviation of sent messages for each time scale. The Appendix~\ref{ap:circular} exposes the theoretical background considered.}
+\begin{center}
+    \begin{tabular}{ |l|| c|c|}
+        \hline
+scale & mean & standard deviation \\ \hline
+seconds & - & - \\ \hline
+minutes & - & - \\ \hline
+    \end{tabular}
+\end{center}
+\label{ano}
+\end{table*}
 
 \subsubsection{Seconds and minutes}
-The incidence of messages at each second of a minute and at each minute of an hour is compatible with uniform distribution simulations\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations.}.
-Messages were slightly more evenly distributed in all lists: for both seconds and minutes  $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$.
-Simulations reach these values  but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
+Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
+Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e. no trend was detected and the pattern is the uniformity.
 
@@
 
-\subsection{Scalable fat-tail structure: constancy of membership fractions in the connective sectors}\label{subsec:pih}
+\subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
 There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this well known and expected distribution of activity among participants.
@@
 Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space.
-Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_1$ or $k_2$ depicted in figure~\ref{fig:setores}.
+Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_\Leftarrow$ or $k_\Rightarrow$ depicted in figure~\ref{fig:setores}.
 The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
@@
 \section{Circular statistics}\label{ap:circular}
+\section{Typologies}\label{ap:typ}
+There are other ways to split and characterize networks.
+To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
+the radius is the minimum eccentricity).
+In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
+Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
+Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
+
+Models of human dynamics can be used to predict and classify activity.
+In this case, human activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
+Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
+
+Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
+Concerning empirical research,
+Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
+personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
+
+Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
+This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
+Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
+Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
+
+It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
+A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
+Also, ethic and moral issues are developed by such legacy.
+For example, Adorno et al.
+conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
+Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
+
 \section{Tables}\label{sectables}

### 2015-04-02 paper.tex
terminada parte principal da formalizacao da setorializacao de Erdos

@@
 ]{revtex4-1}
-
 \usepackage{graphicx}% Include figure files
@@
 \usepackage{xcolor}
+\usepackage{amsmath}
 \hypersetup{
@@
 
-\begin{alignat}{3}
-e_{\sigma,1}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}\geq&&\overline{k}_{\sigma,i}\} \\
+\begin{alignat}{3}\label{eq:part}
+e_{\sigma,1}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}\geq&&\overline{k}_{\sigma,i}\} \nonumber \\
 e_{\sigma,2}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}<\;&&\overline{k}_{\sigma,i}\leq\overline{k}_{\sigma,\Rightarrow}\} \\ 
-e_{\sigma,3}&=\{\;i\;|\;&&\overline{k}_{\sigma,i}<\overline{k}_{\sigma,\Rightarrow}\}
+e_{\sigma,3}&=\{\;i\;|\;&&\overline{k}_{\sigma,i}<\overline{k}_{\sigma,\Rightarrow}\} \nonumber
 \end{alignat}
@@
 
+Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1<\delta<6$ can be described as:
+
+%\begin{alignat}{3}
+\begin{equation}
+\begin{split}
+%\begin{multline}
+C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi}, \;\forall\; \sigma\}\right\} \\
+C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\sigma : i \in e_{\sigma,\phi}\}\right\} \\
+C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\geq \phi\}\right\} \\
+C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\leq \phi\}\right\} \\
+C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
+	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
+C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
+	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
+%\end{multline}
+\end{split}
+\end{equation}
+%\end{alignat}
+
 These compound criteria, and the simplification of all the compound possibilities as listed above (into exclusivist/inclusivist criterion/cascade/externals), can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
 Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (inverted order of precedence).
-Important here is to notice that compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
+Compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
 

### 2015-04-02 scripts/setoresRede_teorico.py

@@ -44,10 +44,10 @@ x1=3.7; y=-41.16-5
 p.plot((x1,x1),(-1000,y),"r--")
 x2=5.4951; y=-60.-5
 p.plot((x2,x2),(-1000,y),"r--")
-p.xticks((x1,x2),(r"$k_1$",r"$k_2$"),size=25)
+p.xticks((x1,x2),(r"$(k_\Leftarrow)$",r"$(k_\Rightarrow)$"),size=25)
 
-
-p.show()
+p.savefig("../figs/fser_.png")
+#p.show()
 
 
 #p.plot(kk,Pk) ;p.plot(kk,Pk,"ro")

### 2015-04-03 paper.tex
finished erdos partitioning revision

@@
 If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees and strengths, comparison of the real network should be made with:
+For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with:
 \begin{equation}
@@
 \noindent where \emph{way} can be \emph{in} or \emph{out}.
-In and out strengths are divided by $\overline{w}$ and compared also using $\hat{P}$.
-Notice that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$.
+In and out strengths ($s^{in}$, $s^{out}$) are divided by $\overline{w}$ and compared also using $\hat{P}$.
+Notice that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with total degree.
 
@@
 
-Results of these criteria for network segmentation are discussed in Section~\ref{subsec:pih} and exhibited in Figures~\ref{fig:cpp10000} to~\ref{fig:lad50_} of Appendix~\ref{figures}.
+\noindent and both $\overline{k}_{\sigma,\Leftarrow}$ and $\overline{k}_{\sigma,\Rightarrow}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.
 
 Since different metrics can be used to identify the three types of vertices, compound criteria can be defined.
-For example, a very stringent criterion can be used, according to which a vertex will only be regarded as pertaining to a sector if it is so for all the metrics.
+For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
 After a careful consideration of possible combinations, these were reduced to six:
@@
 \begin{itemize}
-    \item Exclusivist criterion:  vertices are only classified if the class is the same according to all metrics.
+    \item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics.
 In this case, vertices classified (usually) does not reach 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
-    \item Inclusivist criterion: a vertex has the class given by any of the metrics.
+    \item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics.
 Therefore, a vertex may belong to more than one class, and total members may add more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
-    \item Exclusivist cascade: vertices are only classified as hubs if they are hubs according to all metrics.
+    \item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics.
 Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics.
 The remaining vertices are regarded as peripheral.
-    \item Inclusivist cascade: vertices are hubs if they are classified as so according to any of the metrics.
+    \item Inclusivist cascade $C_4$: vertices are hubs if they are classified as so according to any of the metrics.
 The remaining vertices are classified as intermediary if they belong to this category for any of the metrics.
 Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics.
-    \item Exclusivist externals: vertices are only hubs if they are classified as such according to all the metrics.
+    \item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics.
 The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric.
 The rest of the nodes are classified as intermediary.
-    \item Inclusivist externals: hubs are vertices classified as hubs according to any metric.
+    \item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric.
 The remaining vertices will be peripheral if they are classified as such according to any metric.
@@
 
-These compound criteria, and the simplification of all the compound possibilities as listed above (into exclusivist/inclusivist criterion/cascade/externals), can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
-Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (inverted order of precedence).
+The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (with inverted order of sectors precedence).
 Compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.

### 2015-04-03 paper.tex
pequenas melhoras no texto do seccionamento

@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
+\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics}% Force line breaks with \\
 
@@
 
-\noindent where $\{\overline{k}_{\sigma,i}\}$ is defined as:
+\noindent where $\{\overline{k}_{\sigma,i}\}$ is:
 
@@
 
-The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope of the present article.
+The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests.
 Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (with inverted order of sectors precedence).

### 2015-04-07 paper.tex
revisada resultado do seccionamento

@@
 The stability found in principal components of topological measures suggested that this primitive typology is general for (practically) all human contexts.
-Also, the first author is a developer with cultural interests and has been directly and indirectly part of these communities for more than a decade, which gave further safety about assumptions.
-Care was taken not to regard this acquaintanceship as primary source or as definitive evidence.
+Also, the first author is a developer with cultural interests and has been directly and indirectly part of these communities for more than a decade.
+This gave further safety about assumptions although care was taken not to regard this acquaintanceship as primary source or as definitive evidence.
 The interested reader should see Appendix~\ref{ap:typ} for further context, 
@@
 \begin{itemize}
-    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and exhibited noticeable patterns.
+    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and suggests stable patterns.
     \item The fraction of participants in each Erd\"os sector is stable and can be observed even with very few messages (Section~\ref{subsec:pih}).
@@
 \end{center}
-\label{ano}
+\label{tab:circ}
 \end{table*}
@@
 Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
-Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e. no trend was detected and the pattern is the uniformity.
+Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
+no trend was detected and the pattern is the uniformity.
+Circular dispersion is maximized and mean has little meaning.
 
@@
 Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight, as can be seen in Table~\ref{dia}.
+Around 2/3 of the whole activity takes place from noon to midnight.
 Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
+See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
 
@@
 \subsubsection{Days along the month}
-Table~\ref{mes} shows activity along the month.
 Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
@@
 Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
+See Table~\ref{mes} for further insights.
 
@@
 
+\subsubsection{Activity long years}
+Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
+Also, activity should follow trends such as economic, climate and technological.
+Periodicity should be observed to some extent.
+Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:further}).
+
 \subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
-There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} is dedicated to exposing this well known and expected distribution of activity among participants.
-
-The distribution of vertices in the three sectors defined in Section~\ref{sectioning} (hubs, intermediary, periphery) is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
+There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
+
+The distribution of vertices in the  hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
 Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}. 
-If, for instance, strength is taken as the criterion to define the sectors, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-20]\%$ are intermediary and $\approx [75-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
-If the degree is used for classification, hubs can reach $10\%$ of all vertices, i.e.
-classification with strength yields half the number of hubs as plain degree.
-These results hold for in and out degrees and strengths.
-Stable distributions can also be obtained for 200 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}.
-In fact, a minimum window size for observation of more general properties can be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
-This degeneration is critical in the span of 50-100 messages.
-Even so, using a compound criterion, such as exclusive cascade of Figure~\ref{fig:cpp250_}, the networks seem to hold their basic structure with as few as 20-50 messages.
-This indicates that concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
+Typically, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-35]\%$ are intermediary and $\approx [60-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
+These results hold for total, in and out degrees and strengths.
+Stable distributions can also be obtained for 100 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
+The networks hold their basic structure with as few as 10-50 messages, concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
+A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
 
 For the histograms used in the classification process (see Section~\ref{sectioning}), the use of at least $\eta$ vertices for each bin did not yield significant differences.
-That was understood as a consequence of the observation scale:
-\emph{There are between 20 and 200 participants in the message window sizes used to derive most of the results ($ws \in [200,1500]$ messages).
-As peripheral vertices are abundant and span few degrees, there are more than $\eta$ vertices with each low degree value.
-For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
-Therefore, if incident connectivity is very improbable in an Erd\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
-Therefore, using at least $\eta$ vertices for each bin did not impact the results.}
+That was understood as a consequence of the observation scale (see Appendix~\ref{ap:ded}).
 
@@
 \appendix
-\section{Edge existence probability in a directed network without self-loops}\label{ap:ded}
+\section{More graph-related results}
+\subsection{Edge existence probability in a directed network without self-loops}\label{ap:ded}
 Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertices.
@@
 \end{align}
+
+\subsection{Effect of scale in histogram smoothing}\label{ap:ded}
+There are between 20 and 200 participants in the message window sizes used to derive most of the results ($ws \in [200,1500]$ messages).
+As peripheral vertices are abundant and span few degrees, there are more than $\eta$ vertices with each low degree value.
+For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
+Therefore, if incident connectivity is very improbable in an Erd\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
+Therefore, using at least $\eta$ vertices for each bin did not impact the results.
 

### 2015-04-07 paper.tex
finished prevalence of measures results

@@
 \subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-The contribution from the distinct metrics to network topology 
-is very similar for all the networks considered, and did not vary with time.
- This stability in network behavior is remarkable, as can be noticed by the very 
-small standard deviations of the contributions from the metrics along time (Tables~\ref{compPCA0}-\ref{compPCA2}).
-
-The principal component (PCA~\cite{pca}) exhibited a ponderation of centrality measures: degrees, strengths and betweenness centrality.
+We have analyzed the topology of the email networks using standard, well-established metrics based on centrality and clustering.
+We also introduced metrics related to symmetry because there is evidence that such metrics are relevant in a social context~\cite{newmanEvolving}.
+The contribution from the distinct metrics to network topology is very similar for all the networks considered, and did not vary with time.
+This stability in network behavior is remarkable, as can be noticed by the small standard deviations of the contribution of each metric to principal components (from usual Principal Component Analysis~\cite{pca}) along time of Tables~\ref{compPCA0}-\ref{compPCA2}.
+
+The principal component exhibited a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
+This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
 Clustering coefficient is presented in almost perfect orthogonality.
-Symmetry of edges have been reported as bonded to different roles played by participants and relations~\cite{newmanEvolving}, and dispersion was more prevalent in symmetry related measures than clustering coefficient.
-This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-The contribution from the distinct metrics to network topology variance is very similar for all the networks considered, and did not vary with time.
-This stability in network behavior is remarkable, as shown by the very small standard deviations of the contributions from the metrics along time in Appendix~\ref{sec:pcat}.
-
-The dispersion is mostly presented by degree, strength and betweenness centrality, as indicated in Tables~\ref{compPCA0} and~\ref{compPCA} for the LAU list (similar results are obtained for the other lists and omitted here for simplicity).
-The standard deviations are small, which means that the first and second components varies little with time.
+Dispersion was more prevalent in symmetry related measures than clustering coefficient.
+
 Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The corresponding PCA plot for the two first components is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
+The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA1} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
 As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component.
-The plot of clustering coefficient versus degree in Figure~\ref{clust} is similar to the PCA in Figure~\ref{PCA}.
-
-The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry-related relevance for the variance.
+The plot of clustering coefficient versus degree in Figure~\ref{clust} is very similar to the PCA in Figure~\ref{PCA}.
+The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
 This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 

### 2015-04-08 paper.tex
atualizando para repassar

@@
 
-\subsection{Prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-We have analyzed the topology of the email networks using standard, well-established metrics based on centrality and clustering.
-We also introduced metrics related to symmetry because there is evidence that such metrics are relevant in a social context~\cite{newmanEvolving}.
+\subsection{Variance prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
+We have analyzed the topology of the email networks using standard, well-established metrics of centrality and clustering.
+We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
 The contribution from the distinct metrics to network topology is very similar for all the networks considered, and did not vary with time.
@@
 Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA1} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
+The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
 As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component.
@@
 
-    \subsection{Primitive typology: the activity of participants from different sectors}\label{sec:pty}
-
-This work is aimed at finding common characteristics among (email) interaction networks.
-Analysis involved primary measures observance and a formal criteria for coherent ratios of hub, intermediary and hub sectors.
-In this process, inspection done by visualizations and raw data manipulations suggests agents typological peculiarities.
-These are initial observations, which inspired this article and other ongoing research:
+    \subsection{Primitive typology: the activity of participants from Erd\"os sectors}\label{sec:pty}
+
+Visualizations and raw data manipulations suggests agents typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:
 
@@
 
-This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts involved in other typologies, such Meyer-Briggs, Pavlov or F-Scale.
+This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such Meyer-Briggs, Pavlov or F-Scale.
 This has no pretension of being a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
@@
 Another typology suggested by results is about the networks themselves.
-In accordance with previous research results~\cite{barabasiEvo}, this and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
-The seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo} has strong confluence with this work, suggesting that the smaller size of MET community is responsible for the stronger hubs observed.
+This and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
+This network typology is endorsed by the seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo}, which suggests that the smaller size of MET community is responsible for the stronger hubs observed.
 

### 2015-04-09 paper.tex
terminada revisao do artigo (menos apendices)

@@
 %\subsubsection{Sectioning of networks in periphery, intermediary and hubs sectors}\label{sectioning}
-\subsubsection{Erd\"os sectioning}\label{sectioning}
+\subsection{Erd\"os sectioning}\label{sectioning}
 In scale-free networks, the periphery, intermediary and hubs sectors can be derived from comparison with an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
@@
 
-In other words, let $\sigma$ and $\phi$ be integers in the intervals $1 \leq \sigma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\sigma}\}$ have Erd\"os sectors $E_{\sigma}= \{e_{\sigma, \phi} \}$ defined as:
+In other words, let $\sigma$ and $\phi$ be integers in the intervals $1 \leq \sigma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\sigma}\}$ have three Erd\"os sectors $E_{\sigma}= \{e_{\sigma, \phi} \}$ defined as:
 
@@
 
-\subsubsection{Topology metrics for Principal Component Analysis}\label{measures}
+\subsection{Topology metrics for Principal Component Analysis}\label{measures}
 
@@
 Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and to perform PCA with topological measures (see Section~\ref{measures}).
-Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological thinking (see Section~\ref{subsec:typ}).
+Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological speculations (see Section~\ref{subsec:typ}).
 The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
@@
 \subsection{Variance prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-We have analyzed the topology of the email networks using standard, well-established metrics of centrality and clustering.
+The topology was analyzed using standard, well-established metrics of centrality and clustering.
 We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
-The contribution from the distinct metrics to network topology is very similar for all the networks considered, and did not vary with time.
+The contribution from the distinct metrics to network topology is very similar for all the networks, and did not vary with time.
 This stability in network behavior is remarkable, as can be noticed by the small standard deviations of the contribution of each metric to principal components (from usual Principal Component Analysis~\cite{pca}) along time of Tables~\ref{compPCA0}-\ref{compPCA2}.
@@
     \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
-The composition of the first three components are shown in Table~\ref{compPCA2} and measure details in Section~\ref{measures}.
+The composition of the first three components are shown in Table~\ref{compPCA2} and metrics details are given in Section~\ref{measures}.
 Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
@@
 
-    \subsection{Primitive typology: the activity of participants from Erd\"os sectors}\label{sec:pty}
+    \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
 
@@
 These seem as authorities and contrast with participants that respond much more than receive responses.
-    \item The most obvious community structure, as observed by high clustering coefficient, is found only in peripheral and intermediary sectors.
+    \item The most obvious community structure, as observed by a high clustering coefficient, is found only in peripheral and intermediary sectors.
 \end{itemize}
@@
 This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such Meyer-Briggs, Pavlov or F-Scale.
-This has no pretension of being a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
+This is not a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
 Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.
@@
 The characterization of interaction networks resulted from stability observations.
-Along temporal activity statistics, this work reports the stability of the principal components (in the concentration of dispersion and composition) and of the ternary partitioning (periphery, intermediary, hubs) relative sizes, evident in the comparison with the Erd\"os-R\'enyi model.
+Along temporal activity statistics, this work reports the stability of the formation of principal components and of the relative sizes of periphery, intermediary and hubs Erd\"os sectors.
 These results suggested typologies for both agents and the networks.
 
-The task of delivering a first and general characterization of chosen interaction networks involved a larger effort.
-The different aspects covered requires not only different analytical background, but also considerations about textual production and social psychology.
-These are receiving attention within dedicated works and are summarized in Section~\ref{sec:related}.
-
-Regarding topological aspects of interaction networks, further work should inspect other measures in each of the three connective sectors: hubs, intermediary and periphery.
+Observed systems were coherent with literature in different aspects, such as concentration of activity or such as the clusterization versus connectivity pattern.
+Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are proper uses.
+
+Further work should inspect other topological measures in each sector.
+The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
+This might support classification of the three sectors by clustering, a more traditional approach to classification.
 Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space. 
 Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_\Leftarrow$ or $k_\Rightarrow$ depicted in figure~\ref{fig:setores}.
-The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
-This might support classification of the three sectors by clustering, a more traditional approach to classification.
-
-Observed networks were coherent with literature in different aspects, such as concentration of activity, and clusterization versus connectivity patterns.
-Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are proper uses.
-
-\subsection{Further work}\label{sec:further}
-A paired article is dedicated to textual production of the network sectors~\cite{rcText}.
-Resulting knowledge purposes networks and participants typologies, and both topological and textual analysis should foster characterization of interaction networks and participation incidences.
-Stability reported in this article eases tipologization of outliers and more usual participation patterns (see Sections~\ref{sec:related} and~\ref{sec:pty}).
-
-On the technological trend, usage of such characteristics are taking place in linked data and electronic government technologies~\cite{ops,opa,ensaio}.
-Further steps involve elaboration and tests of social dynamics.
-These results are also being used for anthropological physics experiments~\cite{anPhy}.
-
-Controllability of these networks have unintuitive properties and might bring into forefront crucial properties~\cite{barabasiControlCapacity,barabasiControlCentrality,barabasiControllability}, and might be the subject of subsequent research.
-
-Significant differences in the textual production of each connective sector of the network (periphery, intermediary, hubs) is reported in an article by the same research group~\cite{rcText}.
-Most importantly, linguistic differences are more prominent between the sectors of a same network than between the same sector in different lists.
-
-To visualize network evolution, the Versinus method was developed.
-The method is simple and consists in the placement of each vertex along a sinusoid and a line.
-In the resulting fixed layout, the time evolution takes place.
-Even so, fellow researchers repeatedly asked for an article describing Versinus, and therefore it was written in 2013~\cite{versinus}.
-
-A core purpose of the present article, dedicated to stability in interaction networks, and the article about textual production~\cite{rcText} is to derive a sound typology of online agents, with quantitative criteria driven from physics (to the extent possible).
-This is a work under construction, with preliminary elaborations in Section~\ref{sec:pty}, but should be extensively published in near future.
+
+Stability here reported eases tipologization of both outliers and usual participation patterns (see Sections~\ref{sec:related} and~\ref{sec:pty}).
+A paired article with this one is reports significant differences in the textual production of each Ed\"os sector~\cite{rcText}. Resulting knowledge purposes networks and participant typologies.
+Another paired article describes a network time evolution visualization method~\cite{versinus}.
+The usage of such results are taking place in software packages, linked data, electronic government technologies, and anthropological physics experiments~\cite{gmanePack,ops,opa,ensaio,anPhy}.
 

### 2015-04-13 paper.tex
terminados acrescimos mais urgentes

@@
 
-There are research reports based on the results presented in this article, which are exposed in Section~\ref{sec:further}.
+There are research reports based on the results presented in this article, which are exposed in Section~\ref{sec:conc}.
 
@@
 list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
-LAU  & Jun/29/2003 & Jul/23/2005 & 1183 & 3373 & 5 \\
-LAD  & Jun/30/2003 & Oct/07/2009 & 1268 & 3113 & 4 \\
-MET  & Ago/01/2005 & Mar/07/2008 & 492  & 4607 & 23 \\
-CPP  & Mar/13/2002 & Aug/25/2009 & 1052 & 4506 & 7 \\ \hline
+	  \input{tables/tab1Geral}
   \end{tabular}
@@
 \begin{table*}[t]
-\scriptsize
-\caption{Circular mean and standard deviation of sent messages for each time scale. The Appendix~\ref{ap:circular} exposes the theoretical background considered.}
+	\caption{Circular mean, standard deviation, variance, circular dispersion and the relation of maximum and minimum incidence at each time unit.
+The Appendix~\ref{ap:circular} exposes the theoretical background considered of directional (or circular) statistics.}
 \begin{center}
-    \begin{tabular}{ |l|| c|c|}
+    \begin{tabular}{ |l|| c|c|c|c|c| }
         \hline
-scale & mean & standard deviation \\ \hline
-seconds & - & - \\ \hline
-minutes & - & - \\ \hline
+	scale & mean & standard deviation & variance & circular dispersion & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
+	\input{tables/tab2TimeLAU}
     \end{tabular}
@@
 Periodicity should be observed to some extent.
-Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:further}).
+Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
 
 \subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
+
 There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
@@
 
+\begin{figure*} 
+   \centering
+        \includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
+	\caption{Fraction of agents in each Erd\"os sector.
+Hubs, intermediary and periphery fractions are represented in red, green and blue.
+In this figure, simple criteria employed using degree and strength.
+Exclusivist and inclusivist compound criteria have black lines for fraction of vertices without class and with more than one class, respectively.
+See Appendix~\ref{figures} for a collection of such timeline figures with all simple and compound criteria and selected measures.}
+    \label{tab:sectIL}
+\end{figure*}
+
 For the histograms used in the classification process (see Section~\ref{sectioning}), the use of at least $\eta$ vertices for each bin did not yield significant differences.
@@
 Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The corresponding PCA plot for the two first components of the PCA exhibited in Table~\ref{compPCA} is shown in Figure~\ref{PCA}, where each vertex is colored according to the sector they belong.
-As expected, peripheral vertices have very low values in the first component and greater dispersion in the second component.
-The plot of clustering coefficient versus degree in Figure~\ref{clust} is very similar to the PCA in Figure~\ref{PCA}.
-The PCA plot in Figure~\ref{PCA2}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
-This is shown in Table~\ref{compPCA2}  where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
-It is concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices, which are more dispersed in Figure~\ref{PCA2} than in Figure~\ref{PCA}.
-
-\begin{figure} 
-   \centering
-        \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
-    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of in- and out- degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
+The typical PCA plot for the two first components exhibited in Table~\ref{compPCA} is shown in the second system of Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong.
+As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the second component (clustering related).
+The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
+This can be observed in Table~\ref{compPCA2} where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
+We concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices.
+
+%\begin{figure} 
+%   \centering
+%        \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
+%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of in- and out- degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
 Principal component is a weighted average of centrality measures: degrees, strengths and betweenness centrality.
@@
 Similar plots were obtained for all window sizes $ws\;\in\;[500,10000]$, and for the networks of the other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.}
-    \label{PCA}
-\end{figure}
-
-\begin{figure} 
-   \centering
-        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-    \caption{Clustering coefficient versus degree of vertices with a window size of $ws = 1000$ email messages, LAU list.
+%    \label{PCA}
+%\end{figure}
+
+\begin{figure*} 
+   \centering
+        \includegraphics[width=\textwidth,height=12cm]{figs/im13PCAPLOT}
+	\caption{Prevalence of symmetry measures over clustering for data dispersion.
+This was considered an important result of PCA analysis, together with stability on component formation.
+First plot holds the usual degree versus clustering coefficient layout.
+Second plot is similar, but the first component is an average of centrality measures (see Table~\ref{compPCA}).
+Third plot resulted from applying PCA with symmetry measures (see Table~\ref{compPCA2}), where greater dispersion can be observed in all sectors, although more pronounced in intermediary and hubs sectors.
+    For this figure, a window size of $ws = 1000$ messages was used from LAU list.
+The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
+Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.
+    In the third plot, clustering coefficient is omitted as it is only relevant for third component.
+In this case, second component is representative of symmetry measurements of vertex interactions.
+Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.
+    }
+    \label{fig:sym}
+\end{figure*}
+
+%\begin{figure} 
+%   \centering
+%        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
+%    \caption{Clustering coefficient versus degree of vertices with a window size of $ws = 1000$ email messages, LAU list.
 The general layout is consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.}
-    \label{clust}
-\end{figure}
-
-\begin{figure} 
-   \centering
-        \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
-    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
+%    \label{clust}
+%\end{figure}
+
+%\begin{figure} 
+%   \centering
+%        \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
+%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
 The composition of the first three components are shown in Table~\ref{compPCA2} and metrics details are given in Section~\ref{measures}.
@@
 Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.}
-    \label{PCA2}
-\end{figure}
+%    \label{PCA2}
+%\end{figure}
 
     \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
-
-Visualizations and raw data manipulations suggests agents typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:
+Consider the sector to which a participant belongs as her/his type. General considerations of this typology are:
+\begin{itemize}
+	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
+	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
+	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for inclusive compound criteria given in Section~\ref{sectioning}).
+	\item The participant often transitions from one sector to another within a network.
+	\item Reported stability of network structure arises from activity with continuous change of the participant types.
+	\item Type often changes with the scale of the activity window. For example, two snapshots of 200 and 10000 messages will probably reveal participants with more than one type. 
+\end{itemize}
+
+This ephemerality of the human type is in accordance with canonical literature on psychological types.
+The observance of often transitions between types is even though of as an argument against prejudice~\cite{adorno}.
+Visualizations and raw data manipulations suggests further typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:
 
@@
 
-This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such Meyer-Briggs, Pavlov or F-Scale.
+This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such as Meyer-Briggs, Pavlov or F-Scale.
 This is not a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.

### 2015-04-15 paper.tex
recauchutada revisao com estatistica circular e +++

@@
 Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work.
-Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity, and Appendix~\ref{ap:typ} further introduces the typological approach.
+Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:typ} further introduces the typological approach.
 Detailed tables and figures are bundled in Support Information, and are cited upon need.
@@
 
-  \subsection{Time activity statistics}                          
+\subsection{Time activity statistics}\label{sec:mtime}
   Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
@@
 Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution.
-Mean and standard deviations in time were taken using circular statistics, as described in Appendix~\ref{ap:circular}.
+
+  Measures of mean and dispersion were taken using circular statistics.
+For such, each $measure$ (data point) is represented as a complex number of unit magnitude $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measure\frac{2\pi}{period}$.
+Moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:
+
+\begin{align}\label{eq:cmom}
+	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
+	R_n&=|m_n|\\
+	\theta_\mu&=Arg(m_1) \nonumber \\
+	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
+\end{align}
+
+The $\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:
+
+\begin{align}\label{eq:cmd}
+	Var(z)&=1 - R_1 \nonumber\\
+	S(z)&= \sqrt{-2\ln(R_1)}\\
+	\delta(z)&=\frac{1-R_2}{2 R_1^2} \nonumber
+\end{align}
+
+Although very simple, such measures enables both comparison of the email networks exposed in this article, and the deliverance of a benchmark to which other interaction networks can be compared.
 
@@
 
-In other words, let $\sigma$ and $\phi$ be integers in the intervals $1 \leq \sigma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\sigma}\}$ have three Erd\"os sectors $E_{\sigma}= \{e_{\sigma, \phi} \}$ defined as:
+In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as:
 
 \begin{alignat}{3}\label{eq:part}
-e_{\sigma,1}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}\geq&&\overline{k}_{\sigma,i}\} \nonumber \\
-e_{\sigma,2}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}<\;&&\overline{k}_{\sigma,i}\leq\overline{k}_{\sigma,\Rightarrow}\} \\ 
-e_{\sigma,3}&=\{\;i\;|\;&&\overline{k}_{\sigma,i}<\overline{k}_{\sigma,\Rightarrow}\} \nonumber
+e_{\gamma,1}&=\{\;i\;|\;\overline{k}_{\gamma,L}\geq&&\overline{k}_{\gamma,i}\} \nonumber \\
+e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
+e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber
 \end{alignat}
 
-\noindent where $\{\overline{k}_{\sigma,i}\}$ is:
+\noindent where $\{\overline{k}_{\gamma,i}\}$ is:
 
@@
 
-\noindent and both $\overline{k}_{\sigma,\Leftarrow}$ and $\overline{k}_{\sigma,\Rightarrow}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.
+\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.
 
@@
 %\begin{multline}
-C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi}, \;\forall\; \sigma\}\right\} \\
-C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\sigma : i \in e_{\sigma,\phi}\}\right\} \\
-C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\geq \phi\}\right\} \\
-C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\leq \phi\}\right\} \\
-C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
+C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi}, \;\forall\; \gamma\}\right\} \\
+C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\gamma: i \in e_{\gamma,\phi}\}\right\} \\
+C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\geq \phi\}\right\} \\
+C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\leq \phi\}\right\} \\
+C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
 	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
-C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
+C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
 	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
@@
 along time is practically the same for all lists. Circular mean and standard deviation
-for each time scale is given in Table~\ref{tab:circ}.
+for each time scale are given in Table~\ref{tab:circ}.
 
 \begin{table*}[t]
-	\caption{Circular mean, standard deviation, variance, circular dispersion and the relation of maximum and minimum incidence at each time unit.
-The Appendix~\ref{ap:circular} exposes the theoretical background considered of directional (or circular) statistics.}
+	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
+The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
 \begin{center}
@@
         \hline
-	scale & mean & standard deviation & variance & circular dispersion & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
+	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
 	\input{tables/tab2TimeLAU}
@@
    \centering
-        \includegraphics[width=\textwidth,height=12cm]{figs/im13PCAPLOT}
+        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT}
 	\caption{Prevalence of symmetry measures over clustering for data dispersion.
@@
 	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
-	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for inclusive compound criteria given in Section~\ref{sectioning}).
+	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for the compound inclusive and exclusive criterion given in Section~\ref{sectioning}).
 	\item The participant often transitions from one sector to another within a network.
@@
 
+\section{Typologies}\label{ap:typ}
+There are other ways to split and characterize networks.
+To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
+the radius is the minimum eccentricity).
+In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
+Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
+Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
+
+Models of human dynamics can be used to predict and classify activity.
+In this case, human activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
+Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
+
+Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
+Concerning empirical research,
+Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
+personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
+
+Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
+This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
+Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
+Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
+
+It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
+A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
+Also, ethic and moral issues are developed by such legacy.
+For example, Adorno et al.
+conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
+Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
+
 \section{Data and scripts}\label{scripts}
@@
 Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
-\section{Circular statistics}\label{ap:circular}
-\section{Typologies}\label{ap:typ}
-There are other ways to split and characterize networks.
-To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
-the radius is the minimum eccentricity).
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
-Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
-
-Models of human dynamics can be used to predict and classify activity.
-In this case, human activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
-Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
-
-Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
-Concerning empirical research,
-Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
-personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
-
-Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
-Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
-Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
-
-It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
-A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
-Also, ethic and moral issues are developed by such legacy.
-For example, Adorno et al.
-conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
-Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
-
+\clearpage
 \section{Tables}\label{sectables}
-\clearpage
 \subsection{PCA tables}\label{sec:pcat}
-
-\begin{table}[H]
+\begin{table}[!h]
   \centering
@@
 
-\begin{table}
+\begin{table}[!h]
   \centering
@@
 
-\clearpage
-
 \subsection{Tables for activity along time and among participants}\label{tabTime}
 
-\begin{table}
+\begin{table}[!h]
     \caption{Distribution of activity among agents.

### 2015-04-15 scripts/setoresRede_teorico.py

@@ -44,7 +44,7 @@ x1=3.7; y=-41.16-5
 p.plot((x1,x1),(-1000,y),"r--")
 x2=5.4951; y=-60.-5
 p.plot((x2,x2),(-1000,y),"r--")
-p.xticks((x1,x2),(r"$(k_\Leftarrow)$",r"$(k_\Rightarrow)$"),size=25)
+p.xticks((x1,x2),(r"$(k_L)$",r"$(k_R)$"),size=25)
 
 p.savefig("../figs/fser_.png")
 #p.show()

### 2015-04-17 paper.tex
feita tabela inline, adiantado processo
adicionadas tabelas no corpo do artigo

@@
 
+\begin{table}[!h!]
+    \caption{Distribution of activity among agents.
+First column is dedicated to percentage of messages sent by the most active participant.
+Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
+Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
+The last decile $-10D$ column has maximum percentage of participants responsible for 10\% of messages.}
+\begin{center}
+    \begin{tabular}{ | l ||  c | c | c | c | }
+        \hline
+        list & hub & $ 1Q $ & $ 1-3Q $ & $-10D$ \\ \hline
+	\input{tables/userTab}
+    \end{tabular}
+\end{center}
+\label{autores}
+\end{table}
+
 The reminder of this section contains a detailed discussion of these results and
@@
 	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
-	\input{tables/tab2TimeLAU}
+	\input{tables/tab2TimeCPP}
     \end{tabular}
@@
 
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabHoursCPP_}
+\label{tab:hin}
+\end{table}
+
 \subsubsection{Days of the week}
@@
 The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
+
+\begin{table}[h]
+	\caption{.}
+\begin{center}
+    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
+        \hline
+        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
+	\input{tables/tabWeekdays}
+    \end{tabular}
+\end{center}
+\label{tab:win}
+\end{table}
 
@@
 
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabMonthdaysCPP}
+\label{tab:min}
+\end{table}
+
 \subsubsection{Months and larger divisions of the year}
  Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
+
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabMonthsLAD}
+\label{tab:min}
+\end{table}
 
@@
 The contribution from the distinct metrics to network topology is very similar for all the networks, and did not vary with time.
-This stability in network behavior is remarkable, as can be noticed by the small standard deviations of the contribution of each metric to principal components (from usual Principal Component Analysis~\cite{pca}) along time of Tables~\ref{compPCA0}-\ref{compPCA2}.
+This stability in network behavior is remarkable, with small standard deviations of the contribution of each metric to principal components in all lists analyzed.
+Table~\ref{fig:pcain} exemplifies the principal components formation with all measures considered, while Tables~\ref{compPCA0}-\ref{compPCA2} of Support Information gives further assistance to the interested reader.
+
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabPCA3MET}
+\label{tab:min}
+\end{table}
 
@@
     \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
-Consider the sector to which a participant belongs as her/his type. General considerations of this typology are:
+Such stability of activity along time, Erd\"os sector relative sizes, principal components formation and concentration of variance suggests that understanding types of networks and participants might benefit from a characterization within such framework.
+We considered the most trivial case, in which the sector to which a participant belongs is regarded as her/his type.
+This is useful in deriving more elaborate typologies, and yield a sketch of or a ``primitive'' typology.
+General considerations:
 \begin{itemize}

### 2015-04-17 paper.tex
terminada adequacao de algumas tabelas

@@
     \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and suggests stable patterns.
-    \item The fraction of participants in each Erd\"os sector is stable and can be observed even with very few messages (Section~\ref{subsec:pih}).
-    \item The measures combine into principal components in the same way for all lists and all snapshots. 
+    \item The fraction of participants in each Erd\"os sector is stable along time and can be observed even with very few messages (Section~\ref{subsec:pih}).
+    \item The topological measures (regarded as most basic and still representative for a first analysis) combine into principal components in the same way for all lists and all snapshots. 
 	    Furthermore, symmetry related measures, defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
@@
 
+The reminder of this section contains a detailed discussion of these results and Support Information holds further tables and figures from which such results can be better observed.
+The user activity exhibited in Table~\ref{autores} is
+
+\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
+One remarkable feature from the analysis is that the activity 
+along time is practically the same for all lists. Circular mean and standard deviation
+for each time scale are given in Table~\ref{tab:circ}.
+
+\begin{table*}[t]
+	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
+The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
+\begin{center}
+    \begin{tabular}{ |l|| c|c|c|c|c| }
+        \hline
+	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
+	\input{tables/tab2TimeCPP}
+    \end{tabular}
+\end{center}
+\label{tab:circ}
+\end{table*}
+
+\subsubsection{Seconds and minutes}
+Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
+Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
+no trend was detected and the pattern is the uniformity.
+Circular dispersion is maximized and mean has little meaning.
+
+\subsubsection{Hours of the day}
+Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
+Around 2/3 of the whole activity takes place from noon to midnight.
+Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
+See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
+
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabHoursCPP_}
+\label{tab:hin}
+\end{table}
+
+\subsubsection{Days of the week}
+Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
+The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
+
+\begin{table}[h]
+	\caption{.}
+\begin{center}
+    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
+        \hline
+        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
+	\input{tables/tabWeekdays}
+    \end{tabular}
+\end{center}
+\label{tab:win}
+\end{table}
+
+\subsubsection{Days along the month}
+Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
+The most important trait seems to be homogeneity.
+Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
+See Table~\ref{mes} for further insights.
+
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabMonthdaysCPP}
+\label{tab:min}
+\end{table}
+
+\subsubsection{Months and larger divisions of the year}
+ Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
+
+\begin{table}[!h]
+	\caption{.}
+	\footnotesize
+	\input{tables/tabMonthsLAD}
+\label{tab:min}
+\end{table}
+
+\subsubsection{Activity long years}
+Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
+Also, activity should follow trends such as economic, climate and technological.
+Periodicity should be observed to some extent.
+Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
+
+\subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
+
+There is a concentration of hub activity and of vertex with few connections.
+Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
+As we have seen, one way to observe such distribution of participants is through a comparison of the real network with an Erd\"os R\'enyi model.
+
 \begin{table}[!h!]
     \caption{Distribution of activity among agents.
-First column is dedicated to percentage of messages sent by the most active participant.
+First column presents the percentage of messages sent by the most active participant.
 Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
@@
 \end{table}
-
-The reminder of this section contains a detailed discussion of these results and
-Support Information holds tables and figures from which such results can be observed.
-
-\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
-One remarkable feature from the analysis is that the activity 
-along time is practically the same for all lists. Circular mean and standard deviation
-for each time scale are given in Table~\ref{tab:circ}.
-
-\begin{table*}[t]
-	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
-The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c| }
-        \hline
-	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
-	\input{tables/tab2TimeCPP}
-    \end{tabular}
-\end{center}
-\label{tab:circ}
-\end{table*}
-
-\subsubsection{Seconds and minutes}
-Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
-Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
-no trend was detected and the pattern is the uniformity.
-Circular dispersion is maximized and mean has little meaning.
-
-\subsubsection{Hours of the day}
-Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
-See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabHoursCPP_}
-\label{tab:hin}
-\end{table}
-
-\subsubsection{Days of the week}
-Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
-The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
-
-\begin{table}[h]
-	\caption{.}
-\begin{center}
-    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
-        \hline
-        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
-	\input{tables/tabWeekdays}
-    \end{tabular}
-\end{center}
-\label{tab:win}
-\end{table}
-
-\subsubsection{Days along the month}
-Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait seems to be homogeneity.
-Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
-See Table~\ref{mes} for further insights.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabMonthdaysCPP}
-\label{tab:min}
-\end{table}
-
-\subsubsection{Months and larger divisions of the year}
- Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabMonthsLAD}
-\label{tab:min}
-\end{table}
-
-\subsubsection{Activity long years}
-Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
-Also, activity should follow trends such as economic, climate and technological.
-Periodicity should be observed to some extent.
-Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
-
-\subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
-
-There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
 

### 2015-04-18 paper.tex
andado bem no sentido de integrar as figuras e tabelas no texto

@@
 \reserveinserts{58}
-\usepackage{morefloats}
+%\usepackage{morefloats}
 \usepackage{hyperref}
@@
 The reminder of this section contains a detailed discussion of these results and Support Information holds further tables and figures from which such results can be better observed.
-The user activity exhibited in Table~\ref{autores} is
 
 \subsection{Constancy and discrepancy of activity along time}\label{constDisc}
-One remarkable feature from the analysis is that the activity 
-along time is practically the same for all lists. Circular mean and standard deviation
-for each time scale are given in Table~\ref{tab:circ}.
 
@@
 \caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
-The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
+Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
+Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
 \begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c| }
+    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
         \hline
-	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
-	\input{tables/tab2TimeCPP}
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeLAD}
     \end{tabular}
@@
 
-\subsubsection{Seconds and minutes}
-Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
+One remarkable feature from the analysis is that the activity 
+along time is practically the same for all lists. Circular mean and standard deviation
+for each time scale are given in Table~\ref{tab:circ}.
+
+Each time scale also revealed patterns through histograms in Tables~\ref{tab:hin},~\ref{tab:win},~\ref{tab:min} and~\ref{tab:min2}.
+The exceptions of the seconds, minutes and years scales received dedicated Sections~\ref{sec:secmin} and~\ref{sec:years}.
+
+\subsubsection{Seconds and minutes}\label{sec:secmin}
+Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists.
+Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
 Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
@@
 
-\subsubsection{Hours of the day}
+\begin{table}[!h]
+	\caption{Example of activity percentages along the hours of the day.
 Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
 Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
-See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
-
-\begin{table}[!h]
-	\caption{.}
+Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.}
 	\footnotesize
@@
 
-\subsubsection{Days of the week}
-Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
-The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
-
-\begin{table}[h]
-	\caption{.}
+\begin{table}[!h]
+	\caption{Example of activity percentages along the days of the week.
+Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
 \begin{center}
@@
 
-\subsubsection{Days along the month}
-Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait seems to be homogeneity.
-Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
-See Table~\ref{mes} for further insights.
-
 \begin{table}[!h]
-	\caption{.}
+	\caption{Variation of activity in the days along the month was not observed, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
+The most important trait seems to be homogeneity.}
 	\footnotesize
@@
 
-\subsubsection{Months and larger divisions of the year}
- Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
-
 \begin{table}[!h]
-	\caption{.}
+	\caption{Example of activity percentages of the months along the year.
+Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}).
+These observations fit academic calendars, vacations and end-of-year holidays.}
 	\footnotesize
 	\input{tables/tabMonthsLAD}
-\label{tab:min}
+\label{tab:min2}
 \end{table}
 
-\subsubsection{Activity long years}
+\subsubsection{Activity along years}\label{sec:years}
 Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
-Also, activity should follow trends such as economic, climate and technological.
-Periodicity should be observed to some extent.
+Also, activity should follow trends such as economic, climate and technological and periodicity should be observed to some extent.
 Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
@@
 
-\begin{table}[!h!]
-    \caption{Distribution of activity among agents.
+\begin{table}[!h]
+    \caption{Distribution of activity among participants.
 First column presents the percentage of messages sent by the most active participant.
@@
 
-The distribution of vertices in the  hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
-Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}. 
+Accordingly, the distribution of vertices in the  hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
+Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Support Information. 
 Typically, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-35]\%$ are intermediary and $\approx [60-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
 These results hold for total, in and out degrees and strengths.
-Stable distributions can also be obtained for 100 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
-The networks hold their basic structure with as few as 10-50 messages, concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
-A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
+Stable distributions can also be obtained for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
+The networks hold their basic structure with as few as 10-50 messages, concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Support Information.
+A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the Erd\"os sectors.
 
@@
 We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
-The contribution from the distinct metrics to network topology is very similar for all the networks, and did not vary with time.
-This stability in network behavior is remarkable, with small standard deviations of the contribution of each metric to principal components in all lists analyzed.
-Table~\ref{fig:pcain} exemplifies the principal components formation with all measures considered, while Tables~\ref{compPCA0}-\ref{compPCA2} of Support Information gives further assistance to the interested reader.
+The contribution of each metric to the variance is very similar for all the networks, and did not vary with time.
+In applying PCA to the snapshots, the contribution of each metric to principal components resulted in very small standard deviation.
+Table~\ref{tab:pcain} exemplifies the principal components formation with all measures considered, while Support Information gives further assistance to the interested reader.
 
 \begin{table}[!h]
-	\caption{.}
+	\caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
+MET list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
+In this case, clusterization is pushed to the third principal component.
+The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation.
+Betweenness centrality again has a role similar to degree, but weaker.
+The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength.
+The three components have in average 80.36\% of the variance.}
 	\footnotesize
-	\input{tables/tabPCA3MET}
-\label{tab:min}
+	\input{tables/tabPCA3LAU}
+\label{tab:pcain}
 \end{table}
 
-The principal component exhibited a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
-This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
+The principal component is an average of centrality measures: degrees, strengths and betweenness centrality.
+This composition of the principal component suggests that all degree, strength and betweenness centrality measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
 Clustering coefficient is presented in almost perfect orthogonality.
@@
 Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The typical PCA plot for the two first components exhibited in Table~\ref{compPCA} is shown in the second system of Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong.
-As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the second component (clustering related).
+Plots of vertices in the components exhibited in Table~\ref{tab:pcain} are shown in the Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong.
+As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the third component (clustering related).
 The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects symmetry metrics relevance for the variance.
-This can be observed in Table~\ref{compPCA2} where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
+This can be observed in Table~\ref{tab:pcain} where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength).
 We concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices.
@@
    \centering
-        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT}
+        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
 	\caption{Prevalence of symmetry measures over clustering for data dispersion.
@@
 \subsection{PCA tables}\label{sec:pcat}
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
+%\begin{table}[!h]
+%  \centering
+%  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
@@
 The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
-$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
-$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
-$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
-  \end{tabular}
-  \label{compPCA0}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in percentages.
+%  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
+% & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
+%       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
+%$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
+%$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
+%$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
+%$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
+%  \end{tabular}
+%  \label{compPCA0}
+%\end{table}
+%
+%\begin{table}[!h]
+%  \centering
+%  \caption{Principal components composition in percentages.
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
@@
 The first and second components represent more than 90\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
-$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
-$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
-$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
-$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
-$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
-$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
-$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
-$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
-  \end{tabular}
-  \label{compPCA}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
+%  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
+% & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
+%       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
+%$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
+%$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
+%$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
+%$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
+%$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
+%$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
+%$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
+%$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
+%$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
+%  \end{tabular}
+%  \label{compPCA}
+%\end{table}
+%
+%
+%\begin{table}[!h]
+%  \centering
+%  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
 LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
@@
 The three components have in average 80.36\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
-$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
-$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
-$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
-$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
-$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
-$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
-$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
-$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
-$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
-$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
-$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
-$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
-$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
-$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
-  \end{tabular}
-  \label{compPCA2}
-\end{table}
-
-\subsection{Tables for activity along time and among participants}\label{tabTime}
-
-\begin{table}[!h]
-    \caption{Distribution of activity among agents.
+%  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
+% & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
+%       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
+%$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
+%$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
+%$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
+%$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
+%$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
+%$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
+%$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
+%$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
+%$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
+%$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
+%$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
+%$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
+%$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
+%$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
+%$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
+%  \end{tabular}
+%  \label{compPCA2}
+%\end{table}
+%
+%\subsection{Tables for activity along time and among participants}\label{tabTime}
+%
+%\begin{table}[!h]
+%    \caption{Distribution of activity among agents.
 First column is dedicated to percentage of messages sent by the most active participant.
@@
 The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
-\begin{center}
-    \begin{tabular}{ | l ||  c | c | c | c | }
-        \hline
-        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
-    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
-    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
-    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
-    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
-    \end{tabular}
-\end{center}
-\label{autores}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Percentage of activity ($100\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
+%\begin{center}
+%    \begin{tabular}{ | l ||  c | c | c | c | }
+%        \hline
+%        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
+%    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
+%    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
+%    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
+%    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
+%    \end{tabular}
+%\end{center}
+%\label{autores}
+%\end{table}
+%
+%
+%\begin{table*}
+%%\tiny
+%    \caption{Percentage of activity ($100\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
 Maximum activity rates are in bold.
@@
 }\label{dia}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-    & 1h   & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h \\ \hline\hline
-0h  & 3.66 & \multirow{6}{*}{10.67} & \multirow{12}{*}{33.76} & 2.87  & \multirow{6}{*}{7.15} & \multirow{12}{*}{29.33} & 3.58 & \multirow{6}{*}{10.14} &  \multirow{12}{*}{36.88} & 4.00 & \multirow{6}{*}{10.77} & \multirow{12}{*}{33.13} \\
-1h  & 2.76 &                        &                         & 1.77  & & &  2.22 & & & 2.52 & & \\
-2h  & 1.79 &                        &                         & 1.04  & & & 1.63 & & & 1.79 & & \\
-3h  & 1.10 &                        &                         & 0.64  & & & 1.06 & & & 1.06 & & \\
-4h  & {\bf 0.68} &                        &                         & 0.47  & & &  0.84 & & & 0.75 & & \\
-5h  & 0.69 &                        &                         & {\bf 0.38}  & & & {\bf 0.82} & & & {\bf 0.66} & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-6h  & 0.83 & \multirow{6}{*}{23.09} &                         & 0.72  & \multirow{6}{*}{22.18} & & 1.17 & \multirow{6}{*}{26.74} & & 0.85 & \multirow{6}{*}{22.36} &  \\
-7h  & 1.24 &                        &                         & 1.33  & & & 2.37 & & & 1.56 & & \\
-8h  & 2.28 &                        &                         & 2.67  & & & 3.54 & & & 2.96 & & \\
-9h  & 4.52 &                        &                         & 4.40  & & & 6.04 & & & 4.68 & & \\
-10h & 6.62 &                        &                         & 6.29  & & & {\bf 6.83} & & & 5.93 & & \\
-11h & {\bf 7.61} &                        &                         & 6.78  & & & 6.79 & & & 6.40 & & \\\hline
-12h & 6.44 & \multirow{6}{*}{\bf 37.63} & \multirow{12}{*}{\bf 66.24} & {\bf 7.33}  & \multirow{6}{*}{\bf 42.22} & \multirow{12}{*}{ \bf 70.66} & 6.11 & \multirow{6}{*}{\bf 35.65} &  \multirow{12}{*}{ \bf 63.12} & {\bf 6.41} & \multirow{6}{*}{\bf 37.25} &  \multirow{12}{*}{\bf 66.87} \\
-13h & 6.04 &                        &                         & 7.08  & & & 6.26 & & & 6.12 & & \\
-14h & 6.47 &                        &                         & 7.09  & & & 6.38 & & & 6.33 & & \\
-15h & 6.10 &                        &                         & 7.14  & & & 5.93 & & & 5.98 & & \\
-16h & 6.22 &                        &                         & 6.68  & & & 5.52 & & & 6.40 & & \\
-17h & 6.36 &                        &                         & 6.89  & & & 5.46 & & & 6.02 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-18h & 6.01 & \multirow{6}{*}{28.61} &                         & 5.99  & \multirow{6}{*}{28.44} & & 5.24 & \multirow{6}{*}{27.46} & & 5.99 & \multirow{6}{*}{29.63} & \\
-19h & 5.02 &                        &                         & 5.23  & & & 4.52 & & & 5.03 & & \\
-20h & 4.85 &                        &                         & 4.98  & & & 4.55 & & & 4.63 & & \\
-21h & 4.38 &                        &                         & 4.37  & & & 4.42 & & & 4.59 & & \\
-22h & 4.06 &                        &                         & 4.24  & & & 4.51 & & & 4.88 & & \\
-23h & 4.30 &                        &                         & 3.64  & & & 4.23 & & & 4.53 & & \\\hline
-    \end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table}[h]
-    \caption{Percentage of activity on days along the week.
+%\begin{center}
+%    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
+%        \hline
+%    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
+%    & 1h   & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h \\ \hline\hline
+%0h  & 3.66 & \multirow{6}{*}{10.67} & \multirow{12}{*}{33.76} & 2.87  & \multirow{6}{*}{7.15} & \multirow{12}{*}{29.33} & 3.58 & \multirow{6}{*}{10.14} &  \multirow{12}{*}{36.88} & 4.00 & \multirow{6}{*}{10.77} & \multirow{12}{*}{33.13} \\
+%1h  & 2.76 &                        &                         & 1.77  & & &  2.22 & & & 2.52 & & \\
+%2h  & 1.79 &                        &                         & 1.04  & & & 1.63 & & & 1.79 & & \\
+%3h  & 1.10 &                        &                         & 0.64  & & & 1.06 & & & 1.06 & & \\
+%4h  & {\bf 0.68} &                        &                         & 0.47  & & &  0.84 & & & 0.75 & & \\
+%5h  & 0.69 &                        &                         & {\bf 0.38}  & & & {\bf 0.82} & & & {\bf 0.66} & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
+%6h  & 0.83 & \multirow{6}{*}{23.09} &                         & 0.72  & \multirow{6}{*}{22.18} & & 1.17 & \multirow{6}{*}{26.74} & & 0.85 & \multirow{6}{*}{22.36} &  \\
+%7h  & 1.24 &                        &                         & 1.33  & & & 2.37 & & & 1.56 & & \\
+%8h  & 2.28 &                        &                         & 2.67  & & & 3.54 & & & 2.96 & & \\
+%9h  & 4.52 &                        &                         & 4.40  & & & 6.04 & & & 4.68 & & \\
+%10h & 6.62 &                        &                         & 6.29  & & & {\bf 6.83} & & & 5.93 & & \\
+%11h & {\bf 7.61} &                        &                         & 6.78  & & & 6.79 & & & 6.40 & & \\\hline
+%12h & 6.44 & \multirow{6}{*}{\bf 37.63} & \multirow{12}{*}{\bf 66.24} & {\bf 7.33}  & \multirow{6}{*}{\bf 42.22} & \multirow{12}{*}{ \bf 70.66} & 6.11 & \multirow{6}{*}{\bf 35.65} &  \multirow{12}{*}{ \bf 63.12} & {\bf 6.41} & \multirow{6}{*}{\bf 37.25} &  \multirow{12}{*}{\bf 66.87} \\
+%13h & 6.04 &                        &                         & 7.08  & & & 6.26 & & & 6.12 & & \\
+%14h & 6.47 &                        &                         & 7.09  & & & 6.38 & & & 6.33 & & \\
+%15h & 6.10 &                        &                         & 7.14  & & & 5.93 & & & 5.98 & & \\
+%16h & 6.22 &                        &                         & 6.68  & & & 5.52 & & & 6.40 & & \\
+%17h & 6.36 &                        &                         & 6.89  & & & 5.46 & & & 6.02 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
+%
+%
+%18h & 6.01 & \multirow{6}{*}{28.61} &                         & 5.99  & \multirow{6}{*}{28.44} & & 5.24 & \multirow{6}{*}{27.46} & & 5.99 & \multirow{6}{*}{29.63} & \\
+%19h & 5.02 &                        &                         & 5.23  & & & 4.52 & & & 5.03 & & \\
+%20h & 4.85 &                        &                         & 4.98  & & & 4.55 & & & 4.63 & & \\
+%21h & 4.38 &                        &                         & 4.37  & & & 4.42 & & & 4.59 & & \\
+%22h & 4.06 &                        &                         & 4.24  & & & 4.51 & & & 4.88 & & \\
+%23h & 4.30 &                        &                         & 3.64  & & & 4.23 & & & 4.53 & & \\\hline
+%    \end{tabular}
+%\end{center}
+%\end{table*}
+%
+%
+%
+%\begin{table}[h]
+%    \caption{Percentage of activity on days along the week.
 Weekend days are at least $\frac{1}{3}$ less active and can reach $\frac{1}{3}$ of activity.
@@
 These might suggest professional relation of CPP and MET participants to the topics of interest, or a hobby relation of LAU and LAD participants.}
-\begin{center}
-    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
-        \hline
-        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
-        CPP & 17.06 & 17.43 & 17.61 & 17.13 & 16.30 & 6.81 & 7.67 \\ \hline
-        MET & 17.53 & 17.54 & 16.43 & 17.06 & 17.46 & 7.92 & 6.06 \\ \hline
-        LAU & 15.71 & 15.80 & 15.88 & 16.43 & 15.13 & 10.13 & 10.91 \\ \hline
-        LAD & 14.91 & 17.73 & 17.01 & 15.40 & 14.25 & 10.39 & 10.30 \\\hline
-    \end{tabular}
-\end{center}
-\label{semana}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Activity along the days of the month.
+%\begin{center}
+%    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
+%        \hline
+%        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
+%        CPP & 17.06 & 17.43 & 17.61 & 17.13 & 16.30 & 6.81 & 7.67 \\ \hline
+%        MET & 17.53 & 17.54 & 16.43 & 17.06 & 17.46 & 7.92 & 6.06 \\ \hline
+%        LAU & 15.71 & 15.80 & 15.88 & 16.43 & 15.13 & 10.13 & 10.91 \\ \hline
+%        LAD & 14.91 & 17.73 & 17.01 & 15.40 & 14.25 & 10.39 & 10.30 \\\hline
+%    \end{tabular}
+%\end{center}
+%\label{semana}
+%\end{table}
+%
+%
+%\begin{table*}
+%%\tiny
+%    \caption{Activity along the days of the month.
 The pattern is to have no clear prevalent period.
@@
 For the scope of this study, differences of activity along the month is assumed to be inexistent.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-day    & 1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days \\ \hline\hline
-1  & 3.19 & \multirow{7}{*}{23.05} & \multirow{14}{*}{45.63} & 3.01 & \multirow{7}{*}{25.16} & \multirow{14}{*}{48.08} & 3.34 & \multirow{7}{*}{23.06} &  \multirow{14}{*}{47.31} & 3.22 & \multirow{7}{*}{21.96} & \multirow{14}{*}{46.70} \\
-2  & 3.07 &                        &                         & 3.38  & & &  3.38 & & & 3.42 & & \\
-3  & 3.20 &                        &                         & 3.55  & & & 3.20 & & & 2.87 & & \\
-4  & 3.63 &                        &                         & 4.34  & & & 3.52 & & & 2.91 & & \\
-5  & 2.85 &                        &                         & 3.93  & & &  2.68 & & & 3.30 & & \\
-6  & 3.67 &                        &                         & 3.76  & & & 3.18 & & & 3.52 & & \\
-7  & 3.45 &                        &                         & 3.18  & & & 3.77 & & & 2.27 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-8  & 3.12 & \multirow{7}{*}{22.57} &                         & 3.36  & \multirow{7}{*}{22.92} & & 3.62 & \multirow{7}{*}{24.25} & & 3.72 & \multirow{7}{*}{24.73} &  \\
-9  & 2.57 &                        &                         & 3.44  & & & 3.82 & & & 3.97 & & \\
-10  & 2.92 &                        &                         & 3.17  & & & 3.06 & & & 3.77 & & \\
-11  & 3.54 &                        &                         & 3.88  & & & 3.11 & & & 3.27 & & \\
-12 & 3.23 &                        &                         & 2.94  & & & 3.40 & & & 2.75 & & \\
-13 & 3.39 &                        &                         & 3.29  & & & 3.55 & & & 3.34 & & \\
-14 & 3.81 &                        &                         & 2.83  & & & 3.69 & & & 3.93 & & \\\hline
-15 & 3.35 & \multirow{7}{*}{23.02} & \multirow{14}{*}{46.31} & 2.72  & \multirow{7}{*}{21.87} & \multirow{14}{*}{ 43.56} & 3.23 & \multirow{7}{*}{22.84} &  \multirow{14}{*}{ 44.01 } & 3.37 & \multirow{7}{*}{22.82} &  \multirow{14}{*}{46.00} \\
-16 & 3.77 &                        &                         & 2.96  & & & 2.94 & & & 3.37 & & \\
-17 & 3.45 &                        &                         & 3.01  & & & 3.02 & & & 2.95 & & \\
-18 & 3.47 &                        &                         & 3.39  & & & 3.63 & & & 3.22 & & \\
-19 & 2.90 &                        &                         & 3.42  & & & 3.16 & & & 3.59 & & \\
-20 & 2.80 &                        &                         & 3.09  & & & 3.25 & & & 3.21 & & \\
-21 & 3.29 &                        &                         & 3.27  & & & 3.61 & & & 3.13 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-22 & 2.88 & \multirow{7}{*}{23.29} &                         & 2.92  & \multirow{7}{*}{21.69} & & 3.80 & \multirow{7}{*}{21.17} & & 3.07 & \multirow{7}{*}{23.18} & \\
-23 & 4.01 &                        &                         & 3.27  & & & 3.03 & & & 3.06 & & \\
-24 & 3.13 &                        &                         & 2.92  & & & 2.31 & & & 2.72 & & \\
-25 & 3.57 &                        &                         & 2.83  & & & 2.38 & & & 3.16 & & \\
-26 & 3.27 &                        &                         & 2.97  & & & 3.49 & & & 3.57 & & \\
-27 & 3.27 &                        &                         & 3.41  & & & 2.92 & & & 3.92 & & \\
-28 & 3.17 &                        &                         & 3.36  & & & 3.26 & & & 3.69 & & \\\hline
-29 & 3.68 &  \multirow{3}{*}{8.06} &  \multirow{3}{*}{8.06}  & 2.93  & \multirow{3}{*}{8.36} & \multirow{3}{*}{8.36} & 3.34 & \multirow{3}{*}{8.68} & \multirow{3}{*}{8.68} & 3.15 & \multirow{3}{*}{7.30} & \multirow{3}{*}{7.30} \\
-30 & 2.76 &                        &                         & 3.14  & & & 3.75 & & & 2.71 & & \\
-31 & 1.63 &                        &                         & 2.29  & & & 1.60 & & & 1.45 & & \\\hline
-    \end{tabular}
-\end{center}
-\label{mes}
-\end{table*}
-
-\begin{table*}[t]
-\scriptsize
-    \caption{Activity along the year, in months, trimesters, quadrimesters and semesters.
+%\begin{center}
+%    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
+%        \hline
+%    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
+%day    & 1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days \\ \hline\hline
+%1  & 3.19 & \multirow{7}{*}{23.05} & \multirow{14}{*}{45.63} & 3.01 & \multirow{7}{*}{25.16} & \multirow{14}{*}{48.08} & 3.34 & \multirow{7}{*}{23.06} &  \multirow{14}{*}{47.31} & 3.22 & \multirow{7}{*}{21.96} & \multirow{14}{*}{46.70} \\
+%2  & 3.07 &                        &                         & 3.38  & & &  3.38 & & & 3.42 & & \\
+%3  & 3.20 &                        &                         & 3.55  & & & 3.20 & & & 2.87 & & \\
+%4  & 3.63 &                        &                         & 4.34  & & & 3.52 & & & 2.91 & & \\
+%5  & 2.85 &                        &                         & 3.93  & & &  2.68 & & & 3.30 & & \\
+%6  & 3.67 &                        &                         & 3.76  & & & 3.18 & & & 3.52 & & \\
+%7  & 3.45 &                        &                         & 3.18  & & & 3.77 & & & 2.27 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
+%8  & 3.12 & \multirow{7}{*}{22.57} &                         & 3.36  & \multirow{7}{*}{22.92} & & 3.62 & \multirow{7}{*}{24.25} & & 3.72 & \multirow{7}{*}{24.73} &  \\
+%9  & 2.57 &                        &                         & 3.44  & & & 3.82 & & & 3.97 & & \\
+%10  & 2.92 &                        &                         & 3.17  & & & 3.06 & & & 3.77 & & \\
+%11  & 3.54 &                        &                         & 3.88  & & & 3.11 & & & 3.27 & & \\
+%12 & 3.23 &                        &                         & 2.94  & & & 3.40 & & & 2.75 & & \\
+%13 & 3.39 &                        &                         & 3.29  & & & 3.55 & & & 3.34 & & \\
+%14 & 3.81 &                        &                         & 2.83  & & & 3.69 & & & 3.93 & & \\\hline
+%15 & 3.35 & \multirow{7}{*}{23.02} & \multirow{14}{*}{46.31} & 2.72  & \multirow{7}{*}{21.87} & \multirow{14}{*}{ 43.56} & 3.23 & \multirow{7}{*}{22.84} &  \multirow{14}{*}{ 44.01 } & 3.37 & \multirow{7}{*}{22.82} &  \multirow{14}{*}{46.00} \\
+%16 & 3.77 &                        &                         & 2.96  & & & 2.94 & & & 3.37 & & \\
+%17 & 3.45 &                        &                         & 3.01  & & & 3.02 & & & 2.95 & & \\
+%18 & 3.47 &                        &                         & 3.39  & & & 3.63 & & & 3.22 & & \\
+%19 & 2.90 &                        &                         & 3.42  & & & 3.16 & & & 3.59 & & \\
+%20 & 2.80 &                        &                         & 3.09  & & & 3.25 & & & 3.21 & & \\
+%21 & 3.29 &                        &                         & 3.27  & & & 3.61 & & & 3.13 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
+%
+%22 & 2.88 & \multirow{7}{*}{23.29} &                         & 2.92  & \multirow{7}{*}{21.69} & & 3.80 & \multirow{7}{*}{21.17} & & 3.07 & \multirow{7}{*}{23.18} & \\
+%23 & 4.01 &                        &                         & 3.27  & & & 3.03 & & & 3.06 & & \\
+%24 & 3.13 &                        &                         & 2.92  & & & 2.31 & & & 2.72 & & \\
+%25 & 3.57 &                        &                         & 2.83  & & & 2.38 & & & 3.16 & & \\
+%26 & 3.27 &                        &                         & 2.97  & & & 3.49 & & & 3.57 & & \\
+%27 & 3.27 &                        &                         & 3.41  & & & 2.92 & & & 3.92 & & \\
+%28 & 3.17 &                        &                         & 3.36  & & & 3.26 & & & 3.69 & & \\\hline
+%29 & 3.68 &  \multirow{3}{*}{8.06} &  \multirow{3}{*}{8.06}  & 2.93  & \multirow{3}{*}{8.36} & \multirow{3}{*}{8.36} & 3.34 & \multirow{3}{*}{8.68} & \multirow{3}{*}{8.68} & 3.15 & \multirow{3}{*}{7.30} & \multirow{3}{*}{7.30} \\
+%30 & 2.76 &                        &                         & 3.14  & & & 3.75 & & & 2.71 & & \\
+%31 & 1.63 &                        &                         & 2.29  & & & 1.60 & & & 1.45 & & \\\hline
+%    \end{tabular}
+%\end{center}
+%\label{mes}
+%\end{table*}
+%
+%
+%\begin{table*}[t]
+%\scriptsize
+%    \caption{Activity along the year, in months, trimesters, quadrimesters and semesters.
 Engagement in list participation seem to concentrate in two periods: middle of the year (Jun-Aug, lists MET and LAD), and transition from years (Dec-Mar, lists CPP, LAU and LAD).
 Messages were considered as to complete 12 months slots, so every month has the same time of occurrences.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c|  c|c|c|c|c| c|c|c|c|c| c|c|c|c|c|}
-        \hline
-    & \multicolumn{5}{c|}{CPP} & \multicolumn{5}{c|}{MET} & \multicolumn{5}{c|}{LAU} & \multicolumn{5}{c|}{LAD}  \\ \hline
-    & m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. \\ \hline\hline
-Jan  & 8.70 & \multirow{2}{*}{17.00} & \multirow{3}{*}{\bf 27.23} &  \multirow{4}{*}{\bf 36.48} & \multirow{6}{*}{\bf 54.26} & 4.88 &  \multirow{2}{*}{11.01} & \multirow{3}{*}{16.90} & \multirow{4}{*}{23.32} & \multirow{6}{*}{47.74} & 10.22 & \multirow{2}{*}{\bf 19.56} & \multirow{3}{*}{\bf 28.23} & \multirow{4}{*}{\bf 35.09} & \multirow{6}{*}{49.17} & 11.23 & \multirow{2}{*}{18.49} & \multirow{3}{*}{26.43} & \multirow{4}{*}{36.04} & \multirow{6}{*}{\bf 57.95} \\
-Fev  & 8.29 &                        &                         & & & 6.13 & & & & & 9.34  & & & & & 7.26 & & & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Mar  & {\bf 10.23} & \multirow{2}{*}{\bf 19.49} &                         & & & 5.89 & \multirow{2}{*}{12.31} & & & & 8.67 & \multirow{2}{*}{15.52} & & & & 7.94 & \multirow{2}{*}{17.55} & & & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Apr  & 9.26 &                        &  \multirow{3}{*}{27.03} & & & 6.42 & & \multirow{3}{*}{30.84} & & & 6.85 & & \multirow{3}{*}{20.94} & & & 9.61 & & \multirow{3}{*}{\bf 31.51} & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Mai  & 9.41 & \multirow{2}{*}{17.78} &                         & \multirow{4}{*}{33.46} & & 10.46 & \multirow{2}{*}{\bf 24.42} & & \multirow{4}{*}{\bf 47.83} & & 7.27 & \multirow{2}{*}{14.09} & & \multirow{4}{*}{30.37} &  & 8.94 & \multirow{2}{*}{\bf 21.91} & & \multirow{4}{*}{\bf 37.56} & \\
-Jun & 8.37 &                        &                         & & & {\bf 13.96} & & & & & 6.81 & & & & & {\bf 12.97} & & & & \\\cline{3-3}\cline{4-4}\cline{6-6}\cline{8-9}\cline{11-11}\cline{13-14}\cline{16-16}\cline{18-19}\cline{21-21}
-Jul & 8.70 &  \multirow{2}{*}{15.68} & \multirow{3}{*}{22.94} & & \multirow{6}{*}{45.73} & 13.23 & \multirow{2}{*}{23.41} & \multirow{3}{*}{\bf 31.16} & & \multirow{6}{*}{\bf 52.26} & 8.96 & \multirow{2}{*}{16.28} & \multirow{3}{*}{24.47} & & \multirow{6}{*}{\bf 50.82} & 9.02 & \multirow{2}{*}{15.65} & \multirow{3}{*}{22.29} & & \multirow{6}{*}{42.05} \\
-Ago  & 6.98 &                       &                         & & & 10.28 & & & & & 7.31 & & & & & 6.63 & & & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Set  & 7.26 & \multirow{2}{*}{15.36} &                         & \multirow{4}{*}{30.06} & & 7.75 & \multirow{2}{*}{16.80} & & \multirow{4}{*}{28.86} & & 8.18 & \multirow{2}{*}{16.24} & & \multirow{4}{*}{34.54} & & 6.63 & \multirow{2}{*}{12.38} & & \multirow{4}{*}{26.40} & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Oct  & 8.10 &                        &  \multirow{3}{*}{22.80} & & & 9.05 & & \multirow{3}{*}{21.10} & & & 8.06 & & \multirow{3}{*}{26.36} & & & 5.74 & & \multirow{3}{*}{19.77} & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Nov  & 7.86 & \multirow{2}{*}{14.69} &                         & & & 7.46 & \multirow{2}{*}{12.06} & & & & 7.63 & \multirow{2}{*}{18.30} & & & & 7.63 & \multirow{2}{*}{14.02} & & & \\
-Dec & 6.81 &                        &                         &  & & 4.59 & & & & & {\bf 10.66} & & & & & 6.39 & & & & \\\hline
-    \end{tabular}
-\end{center}
-\label{ano}
-\end{table*}
-
-\clearpage
-\section{Figures of vertex classification fractions as the network evolves}\label{figures}
-
-Two lists are exhibited in this section, CPP and LAD. These structures are very similar in all
-four lists and laying extensively all figures is redundant. Window sizes of $ws =$ 10000, 5000,
-1000, 500, 250, 100 and 50 messages were used.
-
-\begin{figure*}[hb]
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%\begin{center}
+%    \begin{tabular}{ |l|| c|c|c|c|c|  c|c|c|c|c| c|c|c|c|c| c|c|c|c|c|}
+%        \hline
+%    & \multicolumn{5}{c|}{CPP} & \multicolumn{5}{c|}{MET} & \multicolumn{5}{c|}{LAU} & \multicolumn{5}{c|}{LAD}  \\ \hline
+%    & m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. \\ \hline\hline
+%Jan  & 8.70 & \multirow{2}{*}{17.00} & \multirow{3}{*}{\bf 27.23} &  \multirow{4}{*}{\bf 36.48} & \multirow{6}{*}{\bf 54.26} & 4.88 &  \multirow{2}{*}{11.01} & \multirow{3}{*}{16.90} & \multirow{4}{*}{23.32} & \multirow{6}{*}{47.74} & 10.22 & \multirow{2}{*}{\bf 19.56} & \multirow{3}{*}{\bf 28.23} & \multirow{4}{*}{\bf 35.09} & \multirow{6}{*}{49.17} & 11.23 & \multirow{2}{*}{18.49} & \multirow{3}{*}{26.43} & \multirow{4}{*}{36.04} & \multirow{6}{*}{\bf 57.95} \\
+%Fev  & 8.29 &                        &                         & & & 6.13 & & & & & 9.34  & & & & & 7.26 & & & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
+%Mar  & {\bf 10.23} & \multirow{2}{*}{\bf 19.49} &                         & & & 5.89 & \multirow{2}{*}{12.31} & & & & 8.67 & \multirow{2}{*}{15.52} & & & & 7.94 & \multirow{2}{*}{17.55} & & & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
+%Apr  & 9.26 &                        &  \multirow{3}{*}{27.03} & & & 6.42 & & \multirow{3}{*}{30.84} & & & 6.85 & & \multirow{3}{*}{20.94} & & & 9.61 & & \multirow{3}{*}{\bf 31.51} & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
+%Mai  & 9.41 & \multirow{2}{*}{17.78} &                         & \multirow{4}{*}{33.46} & & 10.46 & \multirow{2}{*}{\bf 24.42} & & \multirow{4}{*}{\bf 47.83} & & 7.27 & \multirow{2}{*}{14.09} & & \multirow{4}{*}{30.37} &  & 8.94 & \multirow{2}{*}{\bf 21.91} & & \multirow{4}{*}{\bf 37.56} & \\
+%Jun & 8.37 &                        &                         & & & {\bf 13.96} & & & & & 6.81 & & & & & {\bf 12.97} & & & & \\\cline{3-3}\cline{4-4}\cline{6-6}\cline{8-9}\cline{11-11}\cline{13-14}\cline{16-16}\cline{18-19}\cline{21-21}
+%Jul & 8.70 &  \multirow{2}{*}{15.68} & \multirow{3}{*}{22.94} & & \multirow{6}{*}{45.73} & 13.23 & \multirow{2}{*}{23.41} & \multirow{3}{*}{\bf 31.16} & & \multirow{6}{*}{\bf 52.26} & 8.96 & \multirow{2}{*}{16.28} & \multirow{3}{*}{24.47} & & \multirow{6}{*}{\bf 50.82} & 9.02 & \multirow{2}{*}{15.65} & \multirow{3}{*}{22.29} & & \multirow{6}{*}{42.05} \\
+%Ago  & 6.98 &                       &                         & & & 10.28 & & & & & 7.31 & & & & & 6.63 & & & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
+%Set  & 7.26 & \multirow{2}{*}{15.36} &                         & \multirow{4}{*}{30.06} & & 7.75 & \multirow{2}{*}{16.80} & & \multirow{4}{*}{28.86} & & 8.18 & \multirow{2}{*}{16.24} & & \multirow{4}{*}{34.54} & & 6.63 & \multirow{2}{*}{12.38} & & \multirow{4}{*}{26.40} & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
+%Oct  & 8.10 &                        &  \multirow{3}{*}{22.80} & & & 9.05 & & \multirow{3}{*}{21.10} & & & 8.06 & & \multirow{3}{*}{26.36} & & & 5.74 & & \multirow{3}{*}{19.77} & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
+%Nov  & 7.86 & \multirow{2}{*}{14.69} &                         & & & 7.46 & \multirow{2}{*}{12.06} & & & & 7.63 & \multirow{2}{*}{18.30} & & & & 7.63 & \multirow{2}{*}{14.02} & & & \\
+%Dec & 6.81 &                        &                         &  & & 4.59 & & & & & {\bf 10.66} & & & & & 6.39 & & & & \\\hline
+%    \end{tabular}
+%\end{center}
+%\label{ano}
+%\end{table*}
+%
+%
+%\clearpage
+%\section{Figures of vertex classification fractions as the network evolves}\label{figures}
+%
+%Two lists are exhibited in this section, CPP and LAD. These structures are very similar in all
+%four lists and laying extensively all figures is redundant. Window sizes of $ws =$ 10000, 5000,
+%1000, 500, 250, 100 and 50 messages were used.
+%
+%\begin{figure*}[hb]
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/10000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp10000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/10000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp10000_}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/5000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp5000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/5000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp5000_}
-\end{figure*}
-
-%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp5000_}
+%\end{figure*}
+%
+%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/1000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp1000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/1000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp1000_}
+%\end{figure*}
+%
+%
+%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/500}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp500}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/500_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp500_}
+%\end{figure*}
+%
+%
+%%%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/250}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp250}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/250_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp250_}
+%\end{figure*}
+%
+%
+%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/100}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp100}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/100_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp100_}
+%\end{figure*}
+%
+%%%%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/50}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 CPP Std library official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:cpp50}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/CPP/50_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp50_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:cpp50_}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/10000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad10000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/10000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad10000_}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/5000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad5000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/5000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad5000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad5000_}
+%\end{figure*}
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/1000}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad1000}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/1000_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad1000_}
+%\end{figure*}
+%
+%
+%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/500}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad500}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/500_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad500_}
+%\end{figure*}
+%
+%
+%%%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/250}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad250}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/250_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad250_}
+%\end{figure*}
+%
+%
+%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/100}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad100}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/100_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
+%    \label{fig:lad100_}
+%\end{figure*}
+%
+%%%%%%%%%%%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/50}
+%    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
 Linux Audio Users (LAD) official mailing list.
@@
 On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
+%    \label{fig:lad50}
+%\end{figure*}
+%
+%
+%\begin{figure*}[hbtp] 
+%   \centering
+%        \includegraphics[width=\textwidth]{figs/LAD/50_2}
+%    \caption{Distribution of vertex with respect to compound criteria.
 Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
@@
 Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad50_}
-\end{figure*}
+%    \label{fig:lad50_}
+%\end{figure*}
+%
+%
+%
 

### 2015-04-18 paper.tex
pequenas melhoras

@@
 
-Each time scale also revealed patterns through histograms in Tables~\ref{tab:hin},~\ref{tab:win},~\ref{tab:min} and~\ref{tab:min2}.
-The exceptions of the seconds, minutes and years scales received dedicated Sections~\ref{sec:secmin} and~\ref{sec:years}.
+Time scale also revealed patterns through histograms as exposed in Tables~\ref{tab:hin},~\ref{tab:win},~\ref{tab:min} and~\ref{tab:min2}.
+Seconds, minutes and years scales received the dedicated Sections~\ref{sec:secmin} and~\ref{sec:years} instead of tables.
 
@@
 	\footnotesize
-	\input{tables/tabPCA3LAU}
+	\input{tables/tabPCA3CPP}
 \label{tab:pcain}

### 2015-04-26 supportingInformation.tex
added initial supporting information file

--- 2015-04-17 paper.tex
@@
 }
+\usepackage{xr}
+\externaldocument{paper}
 
@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics}% Force line breaks with \\
+\title[Stability of interaction networks, SUPPORTING INFORMATION]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics, SUPPORTING INFORMATION}% Force line breaks with \\
 
@@
 \begin{abstract}
- This article reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance.
- We analyzed the temporal activity and topology evolution of networks in four email lists by considering window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline.
-Activity in terms of seconds and minutes exhibit an uniform pattern, while hours, days and months exhibit stable concentrations.
- Participant activity follows the expected distribution of scale-free networks. 
- We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
- Approximately 5\% of the vertices are hubs, 15-35\% are intermediary and the remainder belongs periphery. 
- The metrics that most contribute to data dispersion were found to be centrality-related (degree, strength and betweenness), followed by symmetry-related, and then clustering coefficient. 
- The results also embrace a sketch of a physics-based typology of agents, with proper social and psychological speculations.
- Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed are also present in general human interaction networks.
- Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments, audiovisual reconnaissance, and the report of quantitative differences of textual production as connectivity of agents changes.
+ This is the supporting information of the article that reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance.
 \end{abstract}
@@
 
-\begin{quotation}
-`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.
-The fact that human society has been up to now divided into classes affects more than the external relations of men.
-The marks of social repression are left within the individual soul.'
-% `The conception of personality structure is the
-%best safeguard against the
-% inclination to attribute persistent trends in the
-% individual to something
-% "innate" or "basic" or "racial" within him. The
-% Nazi allegation that natural, biological traits decide the total being of a % person
-% would not have been such
-% a successful political device
-% had it not been possible to point to numerous
-% instances of relative fixity in human behavior and to
-% challenge those who
-% thought to explain them on any basis other than a biological one.'
-\emph{- Adorno et al, 1969, p. 747}
-\end{quotation}
+These results were produced with the Gmane public domain data and an open python package designed for attaining
+these, and related, results. The interested reader should follow Appendix~\ref{scripts} to access both data and rotines.
+Inline are results for 4 emails lists: LAD, LAU, MET and CPP, as described in Section~\ref{sec:data}.
+Similar results can be reproduced for any number of (Gmane) email lists.
+To avoid repeating text of each table for each list, the text is given inline.
 
-\section{Introduction}\label{sec:into}
-Human interaction networks have been studied long before modern computers.
-Antecedents can be found in nineteenth century works, while the foundation of
-the social network analysis field is generally attributed to the psychiatrist Jacob Moreno~\cite{newmanBook}.
-More recently, the continuous increase of stored data related to human interaction has stimulated research,
-and overall characterization is scattered in literature.
-
-Reports on complex networks are usually focused on a particular measure or functionality, such
-as accessibility measure or detecting communities~\cite{access,newmanModularity}.
-Although there are research on networks arisen from specific phenomenon, such as social or protein networks, there are few attempts for a general characterization and even less endeavors that take in account the temporal dynamic.
-
-The present work is aimed at finding common and stable characteristics among (email) interaction networks.
-This is achieved through observations along time, which imply network evolution, a field that has received attention from the research community
-for more than a decade~\cite{barabasiEvo,newmanEvolving}.
-The evolution is represented herein by a timeline of activity snapshots with a constant number of contiguous messages.
-These snapshots yield a succession of networks.
-This approach, albeit coherent and intuitive, seems not to be explored to date~\cite{netWindowEvo}.
-
-We consider only directed, weighted and human interaction networks.
-Undirected and unweighted representations of such networks are also found in the literature and can be obtained by simplification~\cite{GMANE2}.
-Meaningful measures depend heavily on the model and system characteristics~\cite{newmanStru, newmanWeight}, therefore the most basic (and ubiquitous, and fundamental) metrics of centrality, clustering and symmetry are analyzed altogether in principal components formation.
-
-Typologies are the canon of scientific literature for classification of human agents~\cite{typCanon}.
-Consequently, it was considered central to benefit from human typologies.
-This is a complex matter carefully addressed in methodology, results and appendix.
-
-Although all networks considered originated from email lists,
- coherence with literature suggests that results hold for a more general class of interaction networks,
- such as observed in online platforms (e.g. LinkedIn, Facebook, Twitter).
-Indeed, text mining routines and typologies of online participants benefit from the results here presented~\cite{rcText,rcTipo}.
-
-Next subsection exposes related work.
-Section~\ref{sec:data} details data while Section~\ref{sec:carac} is dedicated to the characterization methods.
-Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work.
-Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:typ} further introduces the typological approach.
-Detailed tables and figures are bundled in Support Information, and are cited upon need.
-
-\subsection{Related work}\label{sec:related}
-Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
-Exceptions are reported in this section, with emphasis on those more closely related to the present article.
-
-The evolution of human interaction networks was addressed with a community focus, where the direction of edges was not taken into account~\cite{barabasiEvo}.
-In another article, two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email list networks, scale-free properties were reported with $\alpha=1$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
-
-Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence from literature~\cite{newmanEvolving}.
-No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were discarded in the present article.
-On the other hand, gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and was not considered in this article as the email messages and addresses have no gender related metadata~\cite{GMANE}.
-
-There are research reports based on the results presented in this article, which are exposed in Section~\ref{sec:conc}.
-
-\section{Data description: email lists and messages}\label{sec:data}
-
-Email list messages were obtained from
-the GMANE email archive~\cite{GMANE}, which consists of more than 20,000 email lists and more than 130,000,000 messages~\cite{GMANEwikipedia}.
-These lists cover a variety of topics, mostly technology-related.
-The archive can be described as a corpus with metadata of its messages, including sent time, place, sender name, and sender email address.
-The GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
-
-Although many lists were analyzed, four email lists were selected for thorough observation, making it easier to infer general properties.
-These lists, picked as representing both a diverse set and usual lists, are:
-\begin{itemize}
-    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries.
-English is the language used the most.
-Abbreviated as LAU from now on.
-    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants are from different countries, and English is the language used the most.
-A more technical and less active version of LAU.
-Abbreviated LAD from now on.
-    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries.
-English is the language used the most.
-Abbreviated as CPP from now on.
-    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests.
-Portuguese is the most used language, although Spanish and English are also incident.
-Abbreviated MET from now on.
-\end{itemize}
-
- The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
-
-\begin{table}
-  \centering
-  \caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
-$N$ is the number of participants (number of different email addresses).
-$\Gamma$ is the number of threads (count of messages without antecedent).
-$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
-A relation holds for all lists carefully considered: as the number of participants increases, the number of threads decreases.
-This underpins a typology sketch of networks, as exposed in Section~\ref{sec:pty}.}
-  \label{tab:genLists}
-  \begin{tabular}{|l|c|c|c|c|c|}\hline
-list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
-	  \input{tables/tab1Geral}
-  \end{tabular}
-\end{table}
-
-\section{Characterization methods}\label{sec:carac}
-The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, from seconds to years; 2) sectioning of the networks in hubs, intermediary and periphery vertices; 3) topological measures prominence regarding dispersion; 4) iterative visualization and data mining; 5) typological speculation about networks and participants.
-Each of these procedures are described below.
-
-Distribution of activity among participants received little dedicated attention, but is considered through almost all results of this article, as degree and strength are highly correlated to activity.
-
-\subsection{Time activity statistics}\label{sec:mtime}
-  Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
-This resulted in histograms from which obvious patterns can be observed.
-Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution.
-
-  Measures of mean and dispersion were taken using circular statistics.
-For such, each $measure$ (data point) is represented as a complex number of unit magnitude $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measure\frac{2\pi}{period}$.
-Moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:
-
-\begin{align}\label{eq:cmom}
-	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
-	R_n&=|m_n|\\
-	\theta_\mu&=Arg(m_1) \nonumber \\
-	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
-\end{align}
-
-The $\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:
-
-\begin{align}\label{eq:cmd}
-	Var(z)&=1 - R_1 \nonumber\\
-	S(z)&= \sqrt{-2\ln(R_1)}\\
-	\delta(z)&=\frac{1-R_2}{2 R_1^2} \nonumber
-\end{align}
-
-Although very simple, such measures enables both comparison of the email networks exposed in this article, and the deliverance of a benchmark to which other interaction networks can be compared.
-
-\subsection{Interaction network}\label{intNet}
-Interaction networks can be modeled both weighted or unweighted, both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
-Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e.
-we did not observe directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks.
-The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
-The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
-Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
-This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}.
-Edges in both directions are allowed.
-Each time an interaction occurs, one is added to the edge weight.
-Self-loops were regarded as non-informative and discarded.
-These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figs/criaRede_}
-    \caption{Formation of interaction network from email messages.
-Each vertex represents a participant.
-A reply message from participant B to a message from participant A is regarded as evidence that B received information from A and yields a directed edge.
-Multiple messages add ``weight'' to a directed edge.
-Further details are given in Section~\ref{intNet}.}
-    \label{formationNetwork}
-\end{figure}
-
-Edges can be created from all antecedent message authors on the message-response thread to each message author.
-We only linked the immediate antecedent to the new message author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
-Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''.
-This suggests pertinence of centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
-
-%\subsubsection{Sectioning of networks in periphery, intermediary and hubs sectors}\label{sectioning}
-\subsection{Erd\"os sectioning}\label{sectioning}
-In scale-free networks, the periphery, intermediary and hubs sectors can be derived from comparison with an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
-As a shorthand, this procedure is from now on called \emph{Erd\"os sectioning}, and the resulting sectors the \emph{Erd\"os sectors} or \emph{primitive sectors}.
-
-The degree distribution $\widetilde{P}(k)$ of an ideal
-scale-free network $\mathcal{N}_f$ with $N$ vertices and $z$ edges has less
-average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
-network with the same number of vertices and edges.
-Indeed, we define in this work the intermediary sector of a network to be the set of all the nodes whose degree is less abundant in the real network than on the Erd\"os-R\'enyi model:
-
-\begin{equation}\label{criterio}
-    \widetilde{P}(k)<P(k) \Rightarrow \text{k is intermediary degree}
-\end{equation}
-
-If $\mathcal{N}_f$ is directed and has no self-loops, the probability
-of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$ (see Appendix~\ref{ap:ded}).
-A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertices and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:
-
-\begin{equation}
-    P(k)=\binom{2(N-1)}{k}p_e^k(1-p_e)^{2(N-1)-k}
-\end{equation}
-
-The lower degree fat tail represents the border vertices, i.e.
-the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
-The higher degree fat tail is the hub sector, i.e.
-$\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
-The reasoning for this classification is: 1) vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g.
-without preferential attachment) are correctly associated to the hubs sector.
-Vertices with very few connections, which are way more abundant than expected by pure chance, are assigned to the periphery.
-Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in scale-free phenomena, are classified as intermediary.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figs/fser_}
-    \caption{Degree distribution of scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
-        intermediary vertices, while the former has more peripheral and hub vertices. Sector borders are
-        given by the two intersections $k_\Leftarrow$ and $k_\Rightarrow$ of the connectivity distributions. Characteristic degrees
-    are in compact intervals of degree: $[0,k_\Leftarrow]$, $(k_\Leftarrow,k_\Rightarrow]$, $(k_\Rightarrow,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
-    \label{fig:setores}
-\end{figure}
-
-To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network.
-Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than or equal $k_{j}$.
-This changes equation~\ref{criterio} to:
-
-\begin{equation}\label{criterio2}
-    \sum_{x=k_i}^{k_j} \widetilde{P}(x) < \sum_{x=k_i}^{k_j} P(x) \Rightarrow \text{i is intermediary}
-\end{equation}
-
-If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with:
-\begin{equation}
-\hat{P}(k^{way})=\binom{N-1}{k^{way}}p_e^k(1-p_e)^{N-1-k^{way}}
-\end{equation}
-
-\noindent where \emph{way} can be \emph{in} or \emph{out}.
-In and out strengths ($s^{in}$, $s^{out}$) are divided by $\overline{w}$ and compared also using $\hat{P}$.
-Notice that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with total degree.
-
-In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as:
-
-\begin{alignat}{3}\label{eq:part}
-e_{\gamma,1}&=\{\;i\;|\;\overline{k}_{\gamma,L}\geq&&\overline{k}_{\gamma,i}\} \nonumber \\
-e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
-e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber
-\end{alignat}
-
-\noindent where $\{\overline{k}_{\gamma,i}\}$ is:
-
-\begin{equation}
-\begin{split}
-\overline{k}_{1,i}&=k_i \\
-\overline{k}_{2,i}&=k_i^{in} \\
-\overline{k}_{3,i}&=k_i^{out} \\
-\overline{k}_{4,i}&=\frac{s_i}{\overline{w}} \\
-\overline{k}_{5,i}&=\frac{s_i^{in}}{\overline{w}} \\
-\overline{k}_{6,i}&=\frac{s_i^{out}}{\overline{w}} \\
-\end{split}
-\end{equation}
-
-\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.
-
-Since different metrics can be used to identify the three types of vertices, compound criteria can be defined.
-For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
-After a careful consideration of possible combinations, these were reduced to six:
-
-\begin{itemize}
-    \item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics.
-In this case, vertices classified (usually) does not reach 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
-    \item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics.
-Therefore, a vertex may belong to more than one class, and total members may add more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
-    \item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics.
-Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics.
-The remaining vertices are regarded as peripheral.
-    \item Inclusivist cascade $C_4$: vertices are hubs if they are classified as so according to any of the metrics.
-The remaining vertices are classified as intermediary if they belong to this category for any of the metrics.
-Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics.
-    \item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics.
-The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric.
-The rest of the nodes are classified as intermediary.
-    \item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric.
-The remaining vertices will be peripheral if they are classified as such according to any metric.
-The rest of the vertices will be intermediary vertices.
-\end{itemize}
-
-Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1<\delta<6$ can be described as:
-
-%\begin{alignat}{3}
-\begin{equation}
-\begin{split}
-%\begin{multline}
-C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi}, \;\forall\; \gamma\}\right\} \\
-C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\gamma: i \in e_{\gamma,\phi}\}\right\} \\
-C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\geq \phi\}\right\} \\
-C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\leq \phi\}\right\} \\
-C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
-	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
-C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
-	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
-%\end{multline}
-\end{split}
-\end{equation}
-%\end{alignat}
-
-The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests.
-Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (with inverted order of sectors precedence).
-Compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
-
-Results from applying the Erd\"os sectioning, i.e.
-the achievement of well defined sectors by comparison of the real network with the Erd\"os R\'enyi model, are reported in Section~\ref{subsec:pih} and Appendix~\ref{figures}.
-
-\subsection{Topology metrics for Principal Component Analysis}\label{measures}
-
-The topology of the networks was observed by Principal Component Analysis (PCA~\cite{pca}) outcomes
-with a small selection of the most basic and fundamental measurements for each vertex:
-
-\begin{itemize}
-    \item Degree     $k_i$: number of edges linked to vertex $i$.
-    \item In-degree  $k_i^{in}$: number of edges ending at vertex $i$.
-    \item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
-    \item Strength $s$: sum of weights of all edges linked to vertex $i$.
-    \item In-strength $s_i^{in}$: sum of weights of all edges ending at vertex $i$.
-    \item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
-    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
-    \item Betweenness centrality $bt_i$: fraction of geodesics that contain the vertex $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
-\end{itemize}
-
-In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$ (see Section~\ref{prevalence}):
-
-\begin{itemize}
-    \item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$.
-Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
-$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
-    \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
-\end{itemize}
-
-\subsection{Evolution of the networks}
-   The evolution of the networks was observed within a fixed number of messages, the window size $ws$.
-This same number of contiguous messages is considered with different shifts in the message timeline to obtain snapshots.
-Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and to perform PCA with topological measures (see Section~\ref{measures}).
-Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological speculations (see Section~\ref{subsec:typ}).
-The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
-
-\subsection{Visualization of network evolution}\label{sec:viz}
-The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
-Such visualizations were crucial to guide research into the most important features of network evolution, and prompted us to capture the prominence of topological metrics along time using mean and standard deviations (see Section~\ref{measures} and Appendix~\ref{sec:pcat}), in addition to the size of the three sectors in a timeline fashion (Appendix~\ref{figures}).
-Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures and results were driven.
-See Appendix~\ref{sec:related} for further directions about visualization and text mining concerning the results herein presented, with dedicated articles.
-
-\subsection{Typology speculation}\label{subsec:typ}
-Qualitative, typological speculations were the result of all the methodology.
-More specifically, the Erd\"os sectors was regarded as yielding a \emph{primitive typology} of human agents in social contexts with at least dozens of participants.
-Visualization and data mining were efficient in fueling speculations about the qualities of each type. 
-The stability found in principal components of topological measures suggested that this primitive typology is general for (practically) all human contexts.
-Also, the first author is a developer with cultural interests and has been directly and indirectly part of these communities for more than a decade.
-This gave further safety about assumptions although care was taken not to regard this acquaintanceship as primary source or as definitive evidence.
-The interested reader should see Appendix~\ref{ap:typ} for further context, 
-as it might be considered audacious to bridge from a physics-based, 
-quantitative classification to a qualitative and speculative approach.
-
-\section{Results and discussion}\label{sec:results}
-
-Remarkable features from the analysis of the four email lists are:
-\begin{itemize}
-    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and suggests stable patterns.
-    \item The fraction of participants in each Erd\"os sector is stable along time and can be observed even with very few messages (Section~\ref{subsec:pih}).
-    \item The topological measures (regarded as most basic and still representative for a first analysis) combine into principal components in the same way for all lists and all snapshots. 
-	    Furthermore, symmetry related measures, defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
-    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
-\end{itemize}
-
-The reminder of this section contains a detailed discussion of these results and Support Information holds further tables and figures from which such results can be better observed.
-The user activity exhibited in Table~\ref{autores} is
-
-\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
-One remarkable feature from the analysis is that the activity 
-along time is practically the same for all lists. Circular mean and standard deviation
-for each time scale are given in Table~\ref{tab:circ}.
-
+\section{Time tables in different scales}\label{sec:time}
+Theory presented in Section~\ref{sec:mtime} and results exposed in Section~\ref{constDisc} of the paper~\cite{tpaper}.
+\subsection{Circular measures}
+The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
+Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
+Section~\ref{sec:mtime} describes the theoretical background of directional (or circular) statistics.
 \begin{table*}[t]
-	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
-The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
+	\caption{LAD circular measures}
 \begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c| }
+    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
         \hline
-	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
-	\input{tables/tab2TimeCPP}
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeLAD}
     \end{tabular}
@@
 
-\subsubsection{Seconds and minutes}
-Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
-Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
-no trend was detected and the pattern is the uniformity.
-Circular dispersion is maximized and mean has little meaning.
+\subsection{Histograms}
 
-\subsubsection{Hours of the day}
-Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
-See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
+\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:frac}
+\subsection{LAD}
+\subsection{CPP}
 
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabHoursCPP_}
-\label{tab:hin}
-\end{table}
-
-\subsubsection{Days of the week}
-Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
-The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
-
-\begin{table}[h]
-	\caption{.}
-\begin{center}
-    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
-        \hline
-        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
-	\input{tables/tabWeekdays}
-    \end{tabular}
-\end{center}
-\label{tab:win}
-\end{table}
-
-\subsubsection{Days along the month}
-Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait seems to be homogeneity.
-Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
-See Table~\ref{mes} for further insights.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabMonthdaysCPP}
-\label{tab:min}
-\end{table}
-
-\subsubsection{Months and larger divisions of the year}
- Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabMonthsLAD}
-\label{tab:min}
-\end{table}
-
-\subsubsection{Activity long years}
-Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
-Also, activity should follow trends such as economic, climate and technological.
-Periodicity should be observed to some extent.
-Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
-
-\subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
-
-There is a concentration of hub activity and of vertex with few connections.
-Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
-As we have seen, one way to observe such distribution of participants is through a comparison of the real network with an Erd\"os R\'enyi model.
-
-\begin{table}[!h!]
-    \caption{Distribution of activity among agents.
-First column presents the percentage of messages sent by the most active participant.
-Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $-10D$ column has maximum percentage of participants responsible for 10\% of messages.}
-\begin{center}
-    \begin{tabular}{ | l ||  c | c | c | c | }
-        \hline
-        list & hub & $ 1Q $ & $ 1-3Q $ & $-10D$ \\ \hline
-	\input{tables/userTab}
-    \end{tabular}
-\end{center}
-\label{autores}
-\end{table}
-
-The distribution of vertices in the  hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
-Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}. 
-Typically, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-35]\%$ are intermediary and $\approx [60-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
-These results hold for total, in and out degrees and strengths.
-Stable distributions can also be obtained for 100 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
-The networks hold their basic structure with as few as 10-50 messages, concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
-A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
-
-\begin{figure*} 
-   \centering
-        \includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
-	\caption{Fraction of agents in each Erd\"os sector.
-Hubs, intermediary and periphery fractions are represented in red, green and blue.
-In this figure, simple criteria employed using degree and strength.
-Exclusivist and inclusivist compound criteria have black lines for fraction of vertices without class and with more than one class, respectively.
-See Appendix~\ref{figures} for a collection of such timeline figures with all simple and compound criteria and selected measures.}
-    \label{tab:sectIL}
-\end{figure*}
-
-For the histograms used in the classification process (see Section~\ref{sectioning}), the use of at least $\eta$ vertices for each bin did not yield significant differences.
-That was understood as a consequence of the observation scale (see Appendix~\ref{ap:ded}).
-
-\subsection{Variance prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-The topology was analyzed using standard, well-established metrics of centrality and clustering.
-We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
-The contribution from the distinct metrics to network topology is very similar for all the networks, and did not vary with time.
-This stability in network behavior is remarkable, with small standard deviations of the contribution of each metric to principal components in all lists analyzed.
-Table~\ref{fig:pcain} exemplifies the principal components formation with all measures considered, while Tables~\ref{compPCA0}-\ref{compPCA2} of Support Information gives further assistance to the interested reader.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabPCA3MET}
-\label{tab:min}
-\end{table}
-
-The principal component exhibited a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
-This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-Clustering coefficient is presented in almost perfect orthogonality.
-Dispersion was more prevalent in symmetry related measures than clustering coefficient.
-
-Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The typical PCA plot for the two first components exhibited in Table~\ref{compPCA} is shown in the second system of Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong.
-As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the second component (clustering related).
-The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
-This can be observed in Table~\ref{compPCA2} where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
-We concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices.
-
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
-%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of in- and out- degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
-Principal component is a weighted average of centrality measures: degrees, strengths and betweenness centrality.
-Second component is mostly clustering coefficient.
-Table~\ref{compPCA} shows the composition of principal components.
-Similar plots were obtained for all window sizes $ws\;\in\;[500,10000]$, and for the networks of the other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.}
-%    \label{PCA}
-%\end{figure}
-
-\begin{figure*} 
-   \centering
-        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT}
-	\caption{Prevalence of symmetry measures over clustering for data dispersion.
-This was considered an important result of PCA analysis, together with stability on component formation.
-First plot holds the usual degree versus clustering coefficient layout.
-Second plot is similar, but the first component is an average of centrality measures (see Table~\ref{compPCA}).
-Third plot resulted from applying PCA with symmetry measures (see Table~\ref{compPCA2}), where greater dispersion can be observed in all sectors, although more pronounced in intermediary and hubs sectors.
-    For this figure, a window size of $ws = 1000$ messages was used from LAU list.
-The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
-Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.
-    In the third plot, clustering coefficient is omitted as it is only relevant for third component.
-In this case, second component is representative of symmetry measurements of vertex interactions.
-Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.
-    }
-    \label{fig:sym}
-\end{figure*}
-
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-%    \caption{Clustering coefficient versus degree of vertices with a window size of $ws = 1000$ email messages, LAU list.
-The general layout is consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.}
-%    \label{clust}
-%\end{figure}
-
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
-%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
-The composition of the first three components are shown in Table~\ref{compPCA2} and metrics details are given in Section~\ref{measures}.
-Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
-Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.}
-%    \label{PCA2}
-%\end{figure}
-
-    \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
-Such stability of activity along time, Erd\"os sector relative sizes, principal components formation and concentration of variance suggests that understanding types of networks and participants might benefit from a characterization within such framework.
-We considered the most trivial case, in which the sector to which a participant belongs is regarded as her/his type.
-This is useful in deriving more elaborate typologies, and yield a sketch of or a ``primitive'' typology.
-General considerations:
-\begin{itemize}
-	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
-	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
-	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for the compound inclusive and exclusive criterion given in Section~\ref{sectioning}).
-	\item The participant often transitions from one sector to another within a network.
-	\item Reported stability of network structure arises from activity with continuous change of the participant types.
-	\item Type often changes with the scale of the activity window. For example, two snapshots of 200 and 10000 messages will probably reveal participants with more than one type. 
-\end{itemize}
-
-This ephemerality of the human type is in accordance with canonical literature on psychological types.
-The observance of often transitions between types is even though of as an argument against prejudice~\cite{adorno}.
-Visualizations and raw data manipulations suggests further typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:
-
-\begin{itemize}
-    \item Core hubs usually have intermittent activity.
-Very stable activity was found on MET hubs, which motivated its integration to this work.
-There are reports in the literature of greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
-    \item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
-The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
-Therefore, intermediary vertices seem responsible for the network structure.
-For example, intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
-    \item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
-These seem as authorities and contrast with participants that respond much more than receive responses.
-    \item The most obvious community structure, as observed by a high clustering coefficient, is found only in peripheral and intermediary sectors.
-\end{itemize}
-
-This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such as Meyer-Briggs, Pavlov or F-Scale.
-This is not a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
-Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.
-
-Another typology suggested by results is about the networks themselves.
-This and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
-This network typology is endorsed by the seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo}, which suggests that the smaller size of MET community is responsible for the stronger hubs observed.
-
-\section{Conclusions and further work}\label{sec:conc}
-The characterization of interaction networks resulted from stability observations.
-Along temporal activity statistics, this work reports the stability of the formation of principal components and of the relative sizes of periphery, intermediary and hubs Erd\"os sectors.
-These results suggested typologies for both agents and the networks.
-
-Observed systems were coherent with literature in different aspects, such as concentration of activity or such as the clusterization versus connectivity pattern.
-Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are proper uses.
-
-Further work should inspect other topological measures in each sector.
-The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
-This might support classification of the three sectors by clustering, a more traditional approach to classification.
-Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space. 
-Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_\Leftarrow$ or $k_\Rightarrow$ depicted in figure~\ref{fig:setores}.
-
-Stability here reported eases tipologization of both outliers and usual participation patterns (see Sections~\ref{sec:related} and~\ref{sec:pty}).
-A paired article with this one is reports significant differences in the textual production of each Ed\"os sector~\cite{rcText}. Resulting knowledge purposes networks and participant typologies.
-Another paired article describes a network time evolution visualization method~\cite{versinus}.
-The usage of such results are taking place in software packages, linked data, electronic government technologies, and anthropological physics experiments~\cite{gmanePack,ops,opa,ensaio,anPhy}.
-
-\begin{acknowledgments}
-Renato Fabbri is grateful to CNPq (process: 140860/2013-4,
-project 870336/1997-5), United Nations Development Program (PNUD/ONU, contract: 2013/000566; project BRA/12/018)  and 
-the Postgraduate Committee of the IFSC/USP. This author is also grateful for
-the American Jewish Committee for maintaining an online copy of the Adorno book
-used on the epigraph~\cite{adorno}.
-Authors thank GMANE creators and maintainers, specifically: GMANE is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
-Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B.
-Krause, Nicolas Bareil, Raymond Scholz, and Adam Sjøgren.
-Authors thank referred email lists communities and welcome feedback as core contribution to this, and similar, research.
-\end{acknowledgments}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{More graph-related results}
-\subsection{Edge existence probability in a directed network without self-loops}\label{ap:ded}
-Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertices.
-The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertices})}$, where $max( \text{number of edges} |\ \text{N vertices})=2[(N-1)+(N-2)+...+1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertices.
-Therefore:
-\begin{align}
-    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertices})} \nonumber \\
-       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}} \nonumber \\
-   p_e &=\frac{z}{N(N-1)}
-\end{align}
-
-\subsection{Effect of scale in histogram smoothing}\label{ap:ded}
-There are between 20 and 200 participants in the message window sizes used to derive most of the results ($ws \in [200,1500]$ messages).
-As peripheral vertices are abundant and span few degrees, there are more than $\eta$ vertices with each low degree value.
-For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
-Therefore, if incident connectivity is very improbable in an Erd\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
-Therefore, using at least $\eta$ vertices for each bin did not impact the results.
-
-\section{Typologies}\label{ap:typ}
-There are other ways to split and characterize networks.
-To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
-the radius is the minimum eccentricity).
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
-Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
-
-Models of human dynamics can be used to predict and classify activity.
-In this case, human activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
-Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
-
-Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
-Concerning empirical research,
-Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
-personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
-
-Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
-Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
-Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
-
-It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
-A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
-Also, ethic and moral issues are developed by such legacy.
-For example, Adorno et al.
-conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
-Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
-
-\section{Data and scripts}\label{scripts}
-Messages are downloaded from the GMANE database by RSS in the mbox email text format. 
-They are requested one by one to avoid reaching maximum size of the requests accepted by
-GMANE API.
-
-Every message has about 30 fields, from which the following are crucial
-for the present work:
-\begin{itemize}
-    \item ``From'' field, as it specifies the sender of the message, in the usual format of ``First\_name Last\_Name $<email>$''.
-    \item ``Date'' field, which is given with the resolution of a second.
-    \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
-    \item ``References'', has the ID of the message it is an answer to, if any, and earlier messages in the thread.
-\end{itemize}
-
-Field ``In-Reply-To'' has only the ID of the message it replies and can be sometimes
-a shortcut or an alternative to ``References''. Also, the textual content of the messages,
-accessed through ``payload'' method of the mbox message object, is of central interest and
-the authors dedicated an article to include the textual content of the messages to the analysis~\cite{rcText}.
-
-\subsection{Python scripts}\label{ap:os}
-Basic constructs for obtaining all results are the product of scripts written in the Python programming language.
-These are kept in a public git repository for backup and sharing with research community~\cite{scriptsFim}.
-Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
-
-\subsection{Third party libraries and software}
-The programming framework used 
-is mainly Python-based, with emphasis on usual
-scientific tools. More specifically,
-scripts where written for 2.7.3 version of Python,
-with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
-Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
-\clearpage
-\section{Tables}\label{sectables}
-\subsection{PCA tables}\label{sec:pcat}
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-The first component is a weighted average of degree and betweenness centrality.
-The second component is mostly clustering coefficient.
-The first and second components represent more than 95\% of total variance.
-The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
-$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
-$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
-$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
-  \end{tabular}
-  \label{compPCA0}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in percentages.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a weighted average of degree and strength and betweenness centrality.
-The second component is mostly related to the  clustering coefficient.
-The first and second components represent more than 90\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
-$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
-$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
-$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
-$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
-$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
-$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
-$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
-$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
-  \end{tabular}
-  \label{compPCA}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-In this case, clusterization is pushed to the third principal component.
-The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation.
-Betweenness centrality again has a role similar to degree, but weaker.
-The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength.
-The three components have in average 80.36\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
-$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
-$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
-$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
-$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
-$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
-$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
-$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
-$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
-$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
-$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
-$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
-$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
-$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
-$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
-  \end{tabular}
-  \label{compPCA2}
-\end{table}
-
-\subsection{Tables for activity along time and among participants}\label{tabTime}
-
-\begin{table}[!h]
-    \caption{Distribution of activity among agents.
-First column is dedicated to percentage of messages sent by the most active participant.
-Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
-\begin{center}
-    \begin{tabular}{ | l ||  c | c | c | c | }
-        \hline
-        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
-    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
-    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
-    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
-    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
-    \end{tabular}
-\end{center}
-\label{autores}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Percentage of activity ($100\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
-Maximum activity rates are in bold.
-In 1h columns, minimum activity is also bold.
-The less active period of the day is around 4-6h.
-Maximum activity is between 10-13h.
-Afternoon is most active in 6h division of the day.
-The noon has $\approx \frac{2}{3}$ of 24h activity.
-}\label{dia}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-    & 1h   & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h \\ \hline\hline
-0h  & 3.66 & \multirow{6}{*}{10.67} & \multirow{12}{*}{33.76} & 2.87  & \multirow{6}{*}{7.15} & \multirow{12}{*}{29.33} & 3.58 & \multirow{6}{*}{10.14} &  \multirow{12}{*}{36.88} & 4.00 & \multirow{6}{*}{10.77} & \multirow{12}{*}{33.13} \\
-1h  & 2.76 &                        &                         & 1.77  & & &  2.22 & & & 2.52 & & \\
-2h  & 1.79 &                        &                         & 1.04  & & & 1.63 & & & 1.79 & & \\
-3h  & 1.10 &                        &                         & 0.64  & & & 1.06 & & & 1.06 & & \\
-4h  & {\bf 0.68} &                        &                         & 0.47  & & &  0.84 & & & 0.75 & & \\
-5h  & 0.69 &                        &                         & {\bf 0.38}  & & & {\bf 0.82} & & & {\bf 0.66} & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-6h  & 0.83 & \multirow{6}{*}{23.09} &                         & 0.72  & \multirow{6}{*}{22.18} & & 1.17 & \multirow{6}{*}{26.74} & & 0.85 & \multirow{6}{*}{22.36} &  \\
-7h  & 1.24 &                        &                         & 1.33  & & & 2.37 & & & 1.56 & & \\
-8h  & 2.28 &                        &                         & 2.67  & & & 3.54 & & & 2.96 & & \\
-9h  & 4.52 &                        &                         & 4.40  & & & 6.04 & & & 4.68 & & \\
-10h & 6.62 &                        &                         & 6.29  & & & {\bf 6.83} & & & 5.93 & & \\
-11h & {\bf 7.61} &                        &                         & 6.78  & & & 6.79 & & & 6.40 & & \\\hline
-12h & 6.44 & \multirow{6}{*}{\bf 37.63} & \multirow{12}{*}{\bf 66.24} & {\bf 7.33}  & \multirow{6}{*}{\bf 42.22} & \multirow{12}{*}{ \bf 70.66} & 6.11 & \multirow{6}{*}{\bf 35.65} &  \multirow{12}{*}{ \bf 63.12} & {\bf 6.41} & \multirow{6}{*}{\bf 37.25} &  \multirow{12}{*}{\bf 66.87} \\
-13h & 6.04 &                        &                         & 7.08  & & & 6.26 & & & 6.12 & & \\
-14h & 6.47 &                        &                         & 7.09  & & & 6.38 & & & 6.33 & & \\
-15h & 6.10 &                        &                         & 7.14  & & & 5.93 & & & 5.98 & & \\
-16h & 6.22 &                        &                         & 6.68  & & & 5.52 & & & 6.40 & & \\
-17h & 6.36 &                        &                         & 6.89  & & & 5.46 & & & 6.02 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-18h & 6.01 & \multirow{6}{*}{28.61} &                         & 5.99  & \multirow{6}{*}{28.44} & & 5.24 & \multirow{6}{*}{27.46} & & 5.99 & \multirow{6}{*}{29.63} & \\
-19h & 5.02 &                        &                         & 5.23  & & & 4.52 & & & 5.03 & & \\
-20h & 4.85 &                        &                         & 4.98  & & & 4.55 & & & 4.63 & & \\
-21h & 4.38 &                        &                         & 4.37  & & & 4.42 & & & 4.59 & & \\
-22h & 4.06 &                        &                         & 4.24  & & & 4.51 & & & 4.88 & & \\
-23h & 4.30 &                        &                         & 3.64  & & & 4.23 & & & 4.53 & & \\\hline
-    \end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table}[h]
-    \caption{Percentage of activity on days along the week.
-Weekend days are at least $\frac{1}{3}$ less active and can reach $\frac{1}{3}$ of activity.
-MET concentrates activity in weekdays the most, leaving only 13.98\% of total activity to Saturday and Sunday.
-LAU is the one that less concentrates activity in weekdays, reaching 20.94\% of total activity in weekends.
-These might suggest professional relation of CPP and MET participants to the topics of interest, or a hobby relation of LAU and LAD participants.}
-\begin{center}
-    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
-        \hline
-        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
-        CPP & 17.06 & 17.43 & 17.61 & 17.13 & 16.30 & 6.81 & 7.67 \\ \hline
-        MET & 17.53 & 17.54 & 16.43 & 17.06 & 17.46 & 7.92 & 6.06 \\ \hline
-        LAU & 15.71 & 15.80 & 15.88 & 16.43 & 15.13 & 10.13 & 10.91 \\ \hline
-        LAD & 14.91 & 17.73 & 17.01 & 15.40 & 14.25 & 10.39 & 10.30 \\\hline
-    \end{tabular}
-\end{center}
-\label{semana}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Activity along the days of the month.
-The pattern is to have no clear prevalent period.
-One might point a slight tendency for the first two weeks to be more active, although this table does not present statistical foundation for such an assumption.
-For the scope of this study, differences of activity along the month is assumed to be inexistent.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-day    & 1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days \\ \hline\hline
-1  & 3.19 & \multirow{7}{*}{23.05} & \multirow{14}{*}{45.63} & 3.01 & \multirow{7}{*}{25.16} & \multirow{14}{*}{48.08} & 3.34 & \multirow{7}{*}{23.06} &  \multirow{14}{*}{47.31} & 3.22 & \multirow{7}{*}{21.96} & \multirow{14}{*}{46.70} \\
-2  & 3.07 &                        &                         & 3.38  & & &  3.38 & & & 3.42 & & \\
-3  & 3.20 &                        &                         & 3.55  & & & 3.20 & & & 2.87 & & \\
-4  & 3.63 &                        &                         & 4.34  & & & 3.52 & & & 2.91 & & \\
-5  & 2.85 &                        &                         & 3.93  & & &  2.68 & & & 3.30 & & \\
-6  & 3.67 &                        &                         & 3.76  & & & 3.18 & & & 3.52 & & \\
-7  & 3.45 &                        &                         & 3.18  & & & 3.77 & & & 2.27 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-8  & 3.12 & \multirow{7}{*}{22.57} &                         & 3.36  & \multirow{7}{*}{22.92} & & 3.62 & \multirow{7}{*}{24.25} & & 3.72 & \multirow{7}{*}{24.73} &  \\
-9  & 2.57 &                        &                         & 3.44  & & & 3.82 & & & 3.97 & & \\
-10  & 2.92 &                        &                         & 3.17  & & & 3.06 & & & 3.77 & & \\
-11  & 3.54 &                        &                         & 3.88  & & & 3.11 & & & 3.27 & & \\
-12 & 3.23 &                        &                         & 2.94  & & & 3.40 & & & 2.75 & & \\
-13 & 3.39 &                        &                         & 3.29  & & & 3.55 & & & 3.34 & & \\
-14 & 3.81 &                        &                         & 2.83  & & & 3.69 & & & 3.93 & & \\\hline
-15 & 3.35 & \multirow{7}{*}{23.02} & \multirow{14}{*}{46.31} & 2.72  & \multirow{7}{*}{21.87} & \multirow{14}{*}{ 43.56} & 3.23 & \multirow{7}{*}{22.84} &  \multirow{14}{*}{ 44.01 } & 3.37 & \multirow{7}{*}{22.82} &  \multirow{14}{*}{46.00} \\
-16 & 3.77 &                        &                         & 2.96  & & & 2.94 & & & 3.37 & & \\
-17 & 3.45 &                        &                         & 3.01  & & & 3.02 & & & 2.95 & & \\
-18 & 3.47 &                        &                         & 3.39  & & & 3.63 & & & 3.22 & & \\
-19 & 2.90 &                        &                         & 3.42  & & & 3.16 & & & 3.59 & & \\
-20 & 2.80 &                        &                         & 3.09  & & & 3.25 & & & 3.21 & & \\
-21 & 3.29 &                        &                         & 3.27  & & & 3.61 & & & 3.13 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-22 & 2.88 & \multirow{7}{*}{23.29} &                         & 2.92  & \multirow{7}{*}{21.69} & & 3.80 & \multirow{7}{*}{21.17} & & 3.07 & \multirow{7}{*}{23.18} & \\
-23 & 4.01 &                        &                         & 3.27  & & & 3.03 & & & 3.06 & & \\
-24 & 3.13 &                        &                         & 2.92  & & & 2.31 & & & 2.72 & & \\
-25 & 3.57 &                        &                         & 2.83  & & & 2.38 & & & 3.16 & & \\
-26 & 3.27 &                        &                         & 2.97  & & & 3.49 & & & 3.57 & & \\
-27 & 3.27 &                        &                         & 3.41  & & & 2.92 & & & 3.92 & & \\
-28 & 3.17 &                        &                         & 3.36  & & & 3.26 & & & 3.69 & & \\\hline
-29 & 3.68 &  \multirow{3}{*}{8.06} &  \multirow{3}{*}{8.06}  & 2.93  & \multirow{3}{*}{8.36} & \multirow{3}{*}{8.36} & 3.34 & \multirow{3}{*}{8.68} & \multirow{3}{*}{8.68} & 3.15 & \multirow{3}{*}{7.30} & \multirow{3}{*}{7.30} \\
-30 & 2.76 &                        &                         & 3.14  & & & 3.75 & & & 2.71 & & \\
-31 & 1.63 &                        &                         & 2.29  & & & 1.60 & & & 1.45 & & \\\hline
-    \end{tabular}
-\end{center}
-\label{mes}
-\end{table*}
-
-\begin{table*}[t]
-\scriptsize
-    \caption{Activity along the year, in months, trimesters, quadrimesters and semesters.
-Engagement in list participation seem to concentrate in two periods: middle of the year (Jun-Aug, lists MET and LAD), and transition from years (Dec-Mar, lists CPP, LAU and LAD).
-Messages were considered as to complete 12 months slots, so every month has the same time of occurrences.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c|  c|c|c|c|c| c|c|c|c|c| c|c|c|c|c|}
-        \hline
-    & \multicolumn{5}{c|}{CPP} & \multicolumn{5}{c|}{MET} & \multicolumn{5}{c|}{LAU} & \multicolumn{5}{c|}{LAD}  \\ \hline
-    & m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. \\ \hline\hline
-Jan  & 8.70 & \multirow{2}{*}{17.00} & \multirow{3}{*}{\bf 27.23} &  \multirow{4}{*}{\bf 36.48} & \multirow{6}{*}{\bf 54.26} & 4.88 &  \multirow{2}{*}{11.01} & \multirow{3}{*}{16.90} & \multirow{4}{*}{23.32} & \multirow{6}{*}{47.74} & 10.22 & \multirow{2}{*}{\bf 19.56} & \multirow{3}{*}{\bf 28.23} & \multirow{4}{*}{\bf 35.09} & \multirow{6}{*}{49.17} & 11.23 & \multirow{2}{*}{18.49} & \multirow{3}{*}{26.43} & \multirow{4}{*}{36.04} & \multirow{6}{*}{\bf 57.95} \\
-Fev  & 8.29 &                        &                         & & & 6.13 & & & & & 9.34  & & & & & 7.26 & & & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Mar  & {\bf 10.23} & \multirow{2}{*}{\bf 19.49} &                         & & & 5.89 & \multirow{2}{*}{12.31} & & & & 8.67 & \multirow{2}{*}{15.52} & & & & 7.94 & \multirow{2}{*}{17.55} & & & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Apr  & 9.26 &                        &  \multirow{3}{*}{27.03} & & & 6.42 & & \multirow{3}{*}{30.84} & & & 6.85 & & \multirow{3}{*}{20.94} & & & 9.61 & & \multirow{3}{*}{\bf 31.51} & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Mai  & 9.41 & \multirow{2}{*}{17.78} &                         & \multirow{4}{*}{33.46} & & 10.46 & \multirow{2}{*}{\bf 24.42} & & \multirow{4}{*}{\bf 47.83} & & 7.27 & \multirow{2}{*}{14.09} & & \multirow{4}{*}{30.37} &  & 8.94 & \multirow{2}{*}{\bf 21.91} & & \multirow{4}{*}{\bf 37.56} & \\
-Jun & 8.37 &                        &                         & & & {\bf 13.96} & & & & & 6.81 & & & & & {\bf 12.97} & & & & \\\cline{3-3}\cline{4-4}\cline{6-6}\cline{8-9}\cline{11-11}\cline{13-14}\cline{16-16}\cline{18-19}\cline{21-21}
-Jul & 8.70 &  \multirow{2}{*}{15.68} & \multirow{3}{*}{22.94} & & \multirow{6}{*}{45.73} & 13.23 & \multirow{2}{*}{23.41} & \multirow{3}{*}{\bf 31.16} & & \multirow{6}{*}{\bf 52.26} & 8.96 & \multirow{2}{*}{16.28} & \multirow{3}{*}{24.47} & & \multirow{6}{*}{\bf 50.82} & 9.02 & \multirow{2}{*}{15.65} & \multirow{3}{*}{22.29} & & \multirow{6}{*}{42.05} \\
-Ago  & 6.98 &                       &                         & & & 10.28 & & & & & 7.31 & & & & & 6.63 & & & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Set  & 7.26 & \multirow{2}{*}{15.36} &                         & \multirow{4}{*}{30.06} & & 7.75 & \multirow{2}{*}{16.80} & & \multirow{4}{*}{28.86} & & 8.18 & \multirow{2}{*}{16.24} & & \multirow{4}{*}{34.54} & & 6.63 & \multirow{2}{*}{12.38} & & \multirow{4}{*}{26.40} & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Oct  & 8.10 &                        &  \multirow{3}{*}{22.80} & & & 9.05 & & \multirow{3}{*}{21.10} & & & 8.06 & & \multirow{3}{*}{26.36} & & & 5.74 & & \multirow{3}{*}{19.77} & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Nov  & 7.86 & \multirow{2}{*}{14.69} &                         & & & 7.46 & \multirow{2}{*}{12.06} & & & & 7.63 & \multirow{2}{*}{18.30} & & & & 7.63 & \multirow{2}{*}{14.02} & & & \\
-Dec & 6.81 &                        &                         &  & & 4.59 & & & & & {\bf 10.66} & & & & & 6.39 & & & & \\\hline
-    \end{tabular}
-\end{center}
-\label{ano}
-\end{table*}
-
-\clearpage
-\section{Figures of vertex classification fractions as the network evolves}\label{figures}
-
-Two lists are exhibited in this section, CPP and LAD. These structures are very similar in all
-four lists and laying extensively all figures is redundant. Window sizes of $ws =$ 10000, 5000,
-1000, 500, 250, 100 and 50 messages were used.
-
-\begin{figure*}[hb]
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification is plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp5000_}
-\end{figure*}
-
-%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp50_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad5000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad50_}
-\end{figure*}
-
-%
-%\begin{figure*}
-%    \centering
-%        \includegraphics[width=\textwidth]{pcm}
-%        \caption{Pulse Code Modulation (PCM) audio: an analogical signal is represented by 25 samples with 4 bits each.}
-%        \label{fig:PCM}
-%\end{figure*}
-%
+\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:pcat}
+\subsection{Betweenness, clustering and degree}
+\subsection{Betweenness, clustering, degrees and strengths}
+\subsection{Betweenness, clustering, degrees, strengths and symmetry measures}
 
 \nocite{*}
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
+\bibliography{supportingInformation}% Produces the bibliography via BibTeX.
 \end{document}

### 2015-04-27 supportingInformation.tex
added time tables to support information

@@
 }
+\usepackage{placeins}
 \usepackage{xr}
@@
 Theory presented in Section~\ref{sec:mtime} and results exposed in Section~\ref{constDisc} of the paper~\cite{tpaper}.
-\subsection{Circular measures}
+\subsection{Time circular measures}
 The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
@@
 Section~\ref{sec:mtime} describes the theoretical background of directional (or circular) statistics.
-\begin{table*}[t]
+\begin{table*}[!h]
+	\caption{LAU circular measures}
+\begin{center}
+    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
+        \hline
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeLAU}
+    \end{tabular}
+\end{center}
+\label{tab:circ}
+\end{table*}
+\begin{table*}[!h]
 	\caption{LAD circular measures}
@@
 \end{table*}
+\begin{table*}[!h]
+	\caption{MET circular measures}
+\begin{center}
+    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
+        \hline
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeMET}
+    \end{tabular}
+\end{center}
+\label{tab:circ}
+\end{table*}
+\begin{table*}[!h]
+	\caption{CPP circular measures}
+\begin{center}
+    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
+        \hline
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeCPP}
+    \end{tabular}
+\end{center}
+\label{tab:circ}
+\end{table*}
 
-\subsection{Histograms}
+\FloatBarrier
+\subsection{Time histograms}
+\subsection{Histograms of activity along the hours of the day}
+
+Activity percentages along the hours of the day.
+Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
+Around 2/3 of the whole activity takes place from noon to midnight.
+Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
+\begin{table}[!h]
+	\caption{LAU activity along the hours of the day}
+	\footnotesize
+	\input{tables/tabHoursLAU}
+\end{table}
+
+\begin{table}[!h]
+	\caption{LAD activity along the hours of the day}
+	\footnotesize
+	\input{tables/tabHoursLAD}
+\end{table}
+
+\begin{table}[!h]
+	\caption{MET activity along the hours of the day}
+	\footnotesize
+	\input{tables/tabHoursMET}
+\end{table}
+
+\begin{table}[!h]
+	\caption{CPP activity along the hours of the day}
+	\footnotesize
+	\input{tables/tabHoursCPP}
+\end{table}
+
+\FloatBarrier
+\subsection{Histograms of activity along the days of the week}
+
+Activity percentages along the days of the week. Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.
+
+\begin{table}[!h]
+\begin{center}
+    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
+        \hline
+        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
+	\input{tables/tabWeekdays}
+    \end{tabular}
+\end{center}
+\label{tab:win}
+\end{table}
+
+\FloatBarrier
+\subsection{Histograms of activity along the days of the month}
+	No significant variation of activity in the days along the month was observed.
+One cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
+The most important trait seems to be homogeneity.
+\begin{table}[!h]
+	\caption{LAU activity along the days of the month.}
+	\footnotesize
+	\input{tables/tabMonthdaysLAU}
+\label{tab:min}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD activity along the days of the month.}
+	\footnotesize
+	\input{tables/tabMonthdaysLAD}
+\label{tab:min}
+\end{table}
+\begin{table}[!h]
+	\caption{MET activity along the days of the month.}
+	\footnotesize
+	\input{tables/tabMonthdaysMET}
+\label{tab:min}
+\end{table}
+\begin{table}[!h]
+	\caption{CPP activity along the days of the month.}
+	\footnotesize
+	\input{tables/tabMonthdaysCPP}
+\label{tab:min}
+\end{table}
+
+\FloatBarrier
+\subsection{Histograms of activity along months of the year}
+	Activity percentages of the months along the year from LAD list messages.
+Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD.
+These observations fit academic calendars, vacations and end-of-year holidays.
+
+\begin{table}[!h]
+	\caption{LAU activity along the months of the year.}
+	\footnotesize
+	\input{tables/tabMonthsLAU}
+\label{tab:min2}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD activity along the months of the year.}
+	\footnotesize
+	\input{tables/tabMonthsLAD}
+\label{tab:min2}
+\end{table}
+\begin{table}[!h]
+	\caption{MET activity along the months of the year.}
+	\footnotesize
+	\input{tables/tabMonthsMET}
+\label{tab:min2}
+\end{table}
+
+\begin{table}[!h]
+	\caption{CPP activity along the months of the year.}
+	\footnotesize
+	\input{tables/tabMonthsCPP}
+\label{tab:min2}
+\end{table}
 
 \section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:frac}
-\subsection{LAD}
-\subsection{CPP}
 

### 2015-04-27 supportingInformation.tex
adicionadas figuras

@@
 \section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:frac}
+Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning}.
+Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown bellow, first for CPP, than for  LAD list.
 
-\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:pcat}
+Each step size takes two pages of plot.
+On the first page, the criteria are each centrality measure observed separately: in and out degrees and strengths.
+In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
+On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
+On the same graph, green counts the disconnected vertices.
+
+On the second page we show the fractions of participants with respect to each compound criteria for the Erd\"os sectioning.
+In the first plot, the fraction of vertices with unique classification is plotted in black: $\frac{\text{number of nodes uniquely classified}}{\text{number of nodes}}$.
+On the second plot, black represents the exceeding classifications for the given vertices : $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-50/CPP-W50-S200}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-50/CPP-W50-S200_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-100/CPP-W100-S200}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-100/CPP-W100-S200_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-250/CPP-W250-S250}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-250/CPP-W250-S250_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-500/CPP-W500-S500}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-500/CPP-W500-S500_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-1000/CPP-W1000-S1000}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-1000/CPP-W1000-S1000_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-5000/CPP-W5000-S5000}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-5000/CPP-W5000-S5000_}
+\end{figure*}
+
+\FloatBarrier
+\section{PCA of measures along the timeline}\label{sec:pcat}
 \subsection{Betweenness, clustering and degree}

### 2015-04-27 supportingInformation.tex
first version with all figures
first complete supporting information. Polishing up

@@
  reprint,%
+ floatfix,
 %author-year,%
@@
 }
-\usepackage{placeins}
+%\usepackage{placeins}
 \usepackage{xr}
 \externaldocument{paper}
-
-\maxdeadcycles=1000
+\usepackage[section] {placeins}
 
@@
 
+\subsection{CPP list}
 \begin{figure*}
@@
 \FloatBarrier
+\subsection{LAD list}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-50/LAD-W50-S200}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-50/LAD-W50-S200_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-100/LAD-W100-S200}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-100/LAD-W100-S200_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-250/LAD-W250-S250}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-250/LAD-W250-S250_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-500/LAD-W500-S500}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-500/LAD-W500-S500_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-1000/LAD-W1000-S1000}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-1000/LAD-W1000-S1000_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-5000/LAD-W5000-S5000}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-5000/LAD-W5000-S5000_}
+\end{figure*}
+
+\FloatBarrier
 \section{PCA of measures along the timeline}\label{sec:pcat}
+Loadings for the 14 metrics into the principal components for the MET list, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
+Note that the centrality measurements, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
+The clustering coefficient is only relevant for the third principal component.
+The three components have in average 80.36\% of the variance.
 \subsection{Betweenness, clustering and degree}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1LAU}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1LAD}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1MET}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1CPP}
+\label{tab:pcain}
+\end{table}
+
+\FloatBarrier
 \subsection{Betweenness, clustering, degrees and strengths}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2LAU}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2LAD}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2MET}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2CPP}
+\label{tab:pcain}
+\end{table}
+
+\FloatBarrier
 \subsection{Betweenness, clustering, degrees, strengths and symmetry measures}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3LAU}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3LAD}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3MET}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3CPP}
+\label{tab:pcain}
+\end{table}
 

### 2015-04-29 supportingInformation.tex
first version with corrected timelines
feita toda a revisao, menos os textos enviados por email e tipologia dos resultados

@@
  \altaffiliation{
-Centro de Ciências Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
+Centro de Ci\^encias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
 }%Lines break automatically or can be forced with \\
@@
         \hline
-scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
-	\input{tables/tab2TimeCPP}
+%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ \\ \hline\hline
+	\input{tables/tab2TimeCPP_}
     \end{tabular}
@@
    \centering
-        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-5000/CPP-W5000-S5000}
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-3300/CPP-W3300-S3300}
 \end{figure*}
@@
    \centering
-        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-5000/CPP-W5000-S5000_}
+   \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-3300/CPP-W3300-S3300_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-9900/CPP-W9900-S9900}
+\end{figure*}
+\begin{figure*}
+   \centering
+   \includegraphics[width=\textwidth]{evoTimelines/evoTimelineCPP-9900/CPP-W9900-S9900_}
 \end{figure*}
@@
    \centering
-        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-5000/LAD-W5000-S5000}
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-3300/LAD-W3300-S3300}
 \end{figure*}
@@
    \centering
-        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-5000/LAD-W5000-S5000_}
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-3300/LAD-W3300-S3300_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-9900/LAD-W9900-S9900}
+\end{figure*}
+\begin{figure*}
+   \centering
+        \includegraphics[width=\textwidth]{evoTimelines/evoTimelineLAD-9900/LAD-W9900-S9900_}
 \end{figure*}

### 2015-05-26 supportingInformation.tex
contemplando apontamentos do chu
melhorada escrita de algumas partes e as medidas ao longo do mes
adicionadas tabelas do particionamento e pca para as redes em geral

@@
         \hline
-%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
-scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ \\ \hline\hline
-	\input{tables/tab2TimeCPP_}
+scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+	\input{tables/tab2TimeCPP}
     \end{tabular}
@@
 
+\section{Stability in other networks: Twitter, Facebook, Participa.br}\label{sec:pcat}
+To further verify the hypothesis that such stability is a general property of human social networks,
+we analyzed networks driven from Twitter, Facebook and Participa.br. Selected networks are summarized in
+Table~\ref{tab:E}. Their Erd\"os sector relative sizes are given in Table~\ref{tab:secE}. PCA formations are given in
+Tables~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+
+%\begin{table}[!h]
+%	\caption{Description of selected networks analyzed to generalize results in email interaction networks.}
+%	\footnotesize
+%	\input{tables/tabExtra}
+%\label{tab:E}
+%\end{table}
+
+\begin{table}[!h]
+	\caption{Sec extra.}
+	\footnotesize
+\begin{center}
+\begin{tabular}{| l | c | c | c |}\hline
+	 & p. & i. & h. \\ \hline\hline
+	\input{tables/tabSectorsExtra}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:secE}
+\end{table}
+
+\begin{table*}[!h]
+	\caption{PCA1 of friendship networks}
+	\footnotesize
+\begin{center}
+%\begin{tabular}{| l | c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c | c | c | c |}\hline
+\begin{tabular}{| l ||  c |c |c |c |c || c | c | c | c | c || c |c |c |c |c |	}\cline{2-16}
+%	 & p. & i. & h. \\ \hline\hline
+\multicolumn{1}{c|}{} & \multicolumn{5}{c||}{PC1}          & \multicolumn{5}{c||}{PC2} & \multicolumn{5}{c|}{PC3}  \\\cline{2-16}
+	\multicolumn{1}{c|}{} & 
+	F1 & F2 & F3 & F4 & F5 &	
+	F1 & F2 & F3 & F4 & F5 &	
+	F1 & F2 & F3 & F4 & F5 	\\\hline
+	\input{tables/tabPCA1ExtraF}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:pcaE1F}
+\end{table*}
+
+\begin{table*}[!h]
+	\caption{PCA1 of interaction networks}
+	\footnotesize
+\begin{center}
+%\begin{tabular}{| l | c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c | c | c | c |}\hline
+\begin{tabular}{| l ||  c |c |c |c |c | c | c || c | c | c | c | c | c | c || c |c |c |c |c | c | c |	}\cline{2-22}
+%	 & p. & i. & h. \\ \hline\hline
+\multicolumn{1}{c|}{} & \multicolumn{7}{c||}{PC1}          & \multicolumn{7}{c||}{PC2} & \multicolumn{7}{c|}{PC3}  \\\cline{2-22}
+	\multicolumn{1}{c|}{} & 
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 \\\hline
+	\input{tables/tabPCA1ExtraI}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:pcaE1I}
+\end{table*}
+
+\begin{table*}[!h]
+	\caption{PCA2 of interaction networks}
+	\footnotesize
+\begin{center}
+%\begin{tabular}{| l | c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c | c | c | c |}\hline
+\begin{tabular}{| l ||  c |c |c |c |c | c | c || c | c | c | c | c | c | c || c |c |c |c |c | c | c |	}\cline{2-22}
+%	 & p. & i. & h. \\ \hline\hline
+\multicolumn{1}{c|}{} & \multicolumn{7}{c||}{PC1}          & \multicolumn{7}{c||}{PC2} & \multicolumn{7}{c|}{PC3}  \\\cline{2-22}
+	\multicolumn{1}{c|}{} & 
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 \\\hline
+	\input{tables/tabPCA2Extra}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:pcaE2}
+\end{table*}
+
+\begin{table*}[!h]
+	\caption{PCA3 of interaction networks}
+	\footnotesize
+\begin{center}
+%\begin{tabular}{| l | c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c |c | c | c | c |}\hline
+\begin{tabular}{| l ||  c |c |c |c |c | c | c || c | c | c | c | c | c | c || c |c |c |c |c | c | c |	}\cline{2-22}
+%	 & p. & i. & h. \\ \hline\hline
+\multicolumn{1}{c|}{} & \multicolumn{7}{c||}{PC1}          & \multicolumn{7}{c||}{PC2} & \multicolumn{7}{c|}{PC3}  \\\cline{2-22}
+	\multicolumn{1}{c|}{} & 
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 &
+	I1 & I2 & I3 & I4 & I5 & TT1 & TT2 \\\hline
+	\input{tables/tabPCA3Extra}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:pcaE3}
+\end{table*}
+
+%\begin{table}[!h]
+%	\caption{PCA2.}
+%	\footnotesize
+%	\input{tables/tabPCA2Extra}
+%\label{tab:pcaE2}
+%\end{table}
+%\begin{table}[!h]
+%	\caption{PCA3.}
+%	\footnotesize
+%	\input{tables/tabPCA3Extra}
+%\label{tab:pcaE3}
+%\end{table}
+
 \nocite{*}

### 2015-05-26 supportingInformation.tex
finished introduction of data from external human social networks

@@
 The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
-Note that the centrality measurements, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
-The clustering coefficient is only relevant for the third principal component.
+Note that the centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
+The clustering coefficient is only relevant for the third principal component, coupled with standard deviations of strengths and degrees.
 The three components have in average 80.36\% of the variance.
@@
 
-%\begin{table}[!h]
-%	\caption{Description of selected networks analyzed to generalize results in email interaction networks.}
-%	\footnotesize
-%	\input{tables/tabExtra}
-%\label{tab:E}
-%\end{table}
+\begin{table*}[!h]
+	\caption{Overview of selected networks analyzed in addition to email interaction networks.
+Three social platforms were the sources of network structures: Facebook, Twitter and Participa.br.
+Both friendship and interaction networks were observed, yielding undirected and directed networks, respectively.
+The number of agents $N$ and the number of edges $z$ are given on the last columns.
+The acronyms are used throughout Tables~\ref{tab:secE},~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+All these network data were collected in 2013 and 2014.}
+\begin{center}
+\begin{tabular}{| l | c | c | c | c | c | c | }\hline
+	acronym & provenance & edge & directed & description & $N$ & $z$ \\ \hline\hline
+	\input{tables/tabExtra}
+\hline
+\end{tabular}
+\end{center}
+\label{tab:E}
+\end{table*}
 
-\begin{table}[!h]
-	\caption{Sec extra.}
-	\footnotesize
+\begin{table*}[!h]
+	\caption{Percentage of agents in each Erd\"os sector in the friendship and interaction human networks of Table~\ref{tab:E}.
+The ratios found in email networks are preserved.
+Most notable exceptions are I1 and I4, which might be understood as outliers, A possible explanation for them is that they might be a system better characterized as a superposition of networks, rather than one coherent network.}
 \begin{center}
 \begin{tabular}{| l | c | c | c |}\hline
-	 & p. & i. & h. \\ \hline\hline
+	 & periphery & intermediary & hubs \\ \hline\hline
 	\input{tables/tabSectorsExtra}
@@
 \label{tab:secE}
-\end{table}
+\end{table*}
 
 \begin{table*}[!h]
-	\caption{PCA1 of friendship networks}
+	\caption{First three principal components formation and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Participa.br yields the networks that most resembles email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
+The friendship of Renato Fabbri (F1) is the only whose first component has more than 20\% of custering coefficient and second component has more than 40\% of degree centrality.}
 	\footnotesize
@@
 \begin{table*}[!h]
-	\caption{PCA1 of interaction networks}
+	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resembles email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.}
 	\footnotesize
@@
 \begin{table*}[!h]
-	\caption{PCA2 of interaction networks}
+	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resembles email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
+Important differences are: - the clustering coefficient was only important to the third component for two of the networks ($I2$, $I3$) and does not contribute significantly to any of the first three principal components in $I5$; - in the first component, I5 exhibited less contribution from in-strength, in-degree and betweenness, I4 exhibited less contribution from out-degree.}
 	\footnotesize
@@
 \begin{table*}[!h]
-	\caption{PCA3 of interaction networks}
+	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
+In comparison with the results found in email interaction networks, general characteristics are preserved.
+: first component is an average of degree and betweenness, second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
+Standard deviation of asymmetry and disequilibrium metrics are again coupled to clustering coefficient in the third component.
+Important differences are: - the first component is a less regular average of centrality measures and has a greater contribution of symmetry metrics; - The first component of I5 is formed mostly from symmetry, not centrality, metrics.}
 	\footnotesize

### 2015-05-27 supportingInformation.tex
pequenas melhoras de texto
SI partly revised

@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Stability of interaction networks, SUPPORTING INFORMATION]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics, SUPPORTING INFORMATION}% Force line breaks with \\
+\title[Interaction networks stability: SUPPORTING INFORMATION]{Time stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics (SUPPORTING INFORMATION)}% Force line breaks with \\
 
@@
 
-\author{Vilson V. da Silva Jr.}
-  \homepage{http://automata.cc/}
-  \email{vilson@void.cc}
-  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-
-\author{Ricardo Fabbri}
-  \homepage{http://www.lems.brown.edu/~rfabbri/}
-  \email{rfabbri@iprj.uerj.br}
- \altaffiliation{
-Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
-}%Lines break automatically or can be forced with \\
-
-\author{Deborah C. Antunes}
-  \homepage{http://lattes.cnpq.br/1065956470701739}
-  \email{deborahantunes@gmail.com}
-  \altaffiliation{
-Curso de Psicologia, Universidade Federal do Cer\'a (UFC)
-}%Lines break automatically or can be forced with \\
-
-\author{Marilia M. Pisani}
-  \homepage{http://lattes.cnpq.br/6738980149860322}
-  \email{marilia.m.pisani@gmail.com}
- \altaffiliation{
-Centro de Ci\^encias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
-}%Lines break automatically or can be forced with \\
+%\author{Vilson V. da Silva Jr.}
+%  \homepage{http://automata.cc/}
+%  \email{vilson@void.cc}
+%  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%
+%\author{Ricardo Fabbri}
+%  \homepage{http://www.lems.brown.edu/~rfabbri/}
+%  \email{rfabbri@iprj.uerj.br}
+% \altaffiliation{
+%Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
+%}%Lines break automatically or can be forced with \\
+%
+%\author{Deborah C. Antunes}
+%  \homepage{http://lattes.cnpq.br/1065956470701739}
+%  \email{deborahantunes@gmail.com}
+%  \altaffiliation{
+%Curso de Psicologia, Universidade Federal do Cer\'a (UFC)
+%}%Lines break automatically or can be forced with \\
+%
+%\author{Marilia M. Pisani}
+%  \homepage{http://lattes.cnpq.br/6738980149860322}
+%  \email{marilia.m.pisani@gmail.com}
+% \altaffiliation{
+%Centro de Ci\^encias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
+%}%Lines break automatically or can be forced with \\
 
@@
 \begin{abstract}
- This is the supporting information of the article that reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance.
+ This document presents supporting information for the article describing interaction networks stability.
+The sections exhibit activity distribution in time and among participants; the fraction of vertices in the peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance; the analysis of interaction networks from Facebook, Twitter and Participa.br.
+There is a focus on email list interaction networks for benchmarking.
 \end{abstract}
@@
 
-These results were produced with the Gmane public domain data and an open python package designed for attaining
-these, and related, results. The interested reader should follow Appendix~\ref{scripts} to access both data and rotines.
-Inline are results for 4 emails lists: LAD, LAU, MET and CPP, as described in Section~\ref{sec:data}.
+This documents exemplifies the stability of interaction networks by the observation of 4 emails lists (LAD, LAU, MET and CPP), described in Section~\ref{sec:data} of the main article.
+These results were produced with the Gmane public domain data and an open source python package designed for attaining
+these, and related, results. The interested reader should follow Appendix~\ref{scripts} of the main article to access both data and rotines.
 Similar results can be reproduced for any number of (Gmane) email lists.
-To avoid repeating text of each table for each list, the text is given inline.
+To avoid repeating text of each table for each list, the text is given inline, with the tables following the explanation.
 
@@
 Theory presented in Section~\ref{sec:mtime} and results exposed in Section~\ref{constDisc} of the paper~\cite{tpaper}.
-\subsection{Time circular measures}
+
+\subsection{Time circular measures}\label{si:circ}
 The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
 Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
-Section~\ref{sec:mtime} describes the theoretical background of directional (or circular) statistics.
+Section~\ref{sec:mtime} of the main document describes the theoretical background of directional (or circular) statistics.
 \begin{table*}[!h]
@@
 \subsection{Histograms of activity along the days of the month}
-	No significant variation of activity in the days along the month was observed.
-One cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait seems to be homogeneity.
+Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
 \begin{table}[!h]
@@
 \section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:frac}
-Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning}.
+Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning} of the mais document.
 Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown bellow, first for CPP, than for  LAD list.
@@
 Each step size takes two pages of plot.
-On the first page, the criteria are each centrality measure observed separately: in and out degrees and strengths.
+On the first page, the criteria are each centrality measure observed separately: in, out and total degrees and strengths.
 In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
+On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the greatest component.
 On the same graph, green counts the disconnected vertices.
@@
 Table~\ref{tab:E}. Their Erd\"os sector relative sizes are given in Table~\ref{tab:secE}. PCA formations are given in
-Tables~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}. Friendship networks considered are undirected and unweighted,
+therefore all measures of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
 
@@
 The ratios found in email networks are preserved.
-Most notable exceptions are I1 and I4, which might be understood as outliers, A possible explanation for them is that they might be a system better characterized as a superposition of networks, rather than one coherent network.}
+Most notable exceptions are I1 and I4, which might be understood as outliers, with a possible explanation that they be a system better characterized as a superposition of networks, rather than one coherent network.}
 \begin{center}
@@
 
+%\newpage
+%\pagebreak
+\clearpage
 \nocite{*}

### 2015-05-27 supportingInformation.tex
finished si review

@@
 \section{PCA of measures along the timeline}\label{sec:pcat}
-Loadings for the 14 metrics into the principal components for the MET list, $ws=1000$ messages in 20 disjoint positioning.
-The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
+Loadings for the 14 metrics into the principal components for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics, in accordance with the notation made of Section~\ref{measures} of the main document.
 Note that the centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
@@
 
-\section{Stability in other networks: Twitter, Facebook, Participa.br}\label{sec:pcat}
+\section{Stability in other networks: Twitter, Facebook, Participa.br}\label{si:ext}
 To further verify the hypothesis that such stability is a general property of human social networks,
@@
 therefore all measures of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
+Most important results from the consideration herein of this analysis are:
+
+\begin{itemize}
+	\item a further indicative that the stability reported with a focus on email interaction networks is valid for a broader class of phenomena.
+	\item The greater stability of email interaction networks with the same number of participants. This is specially important for benchmarking and probing general properties.
+\end{itemize}
 
@@
 The number of agents $N$ and the number of edges $z$ are given on the last columns.
-The acronyms are used throughout Tables~\ref{tab:secE},~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
-All these network data were collected in 2013 and 2014.}
+The acronyms, one for each network, are used throughout Tables~\ref{tab:secE},~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+All data collected in 2013 and 2014.}
 \begin{center}
@@
 The ratios found in email networks are preserved.
-Most notable exceptions are I1 and I4, which might be understood as outliers, with a possible explanation that they be a system better characterized as a superposition of networks, rather than one coherent network.}
+I1 and I4 are outliers, to which a possible explanation is that they are better characterized as a superposition of networks, rather than one coherent network.}
 \begin{center}
@@
 	\caption{First three principal components formation and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Participa.br yields the networks that most resembles email networks.
+Participa.br yields the networks that most resembles the email networks.
 Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
@@
 	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Twitter yields the networks that most resembles email networks.
+Twitter yields the networks that most resembles the email networks.
 Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.}
@@
 	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
-In comparison with the results found in email interaction networks, general characteristics are preserved.
-: first component is an average of degree and betweenness, second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
+In comparison with the results found in email interaction networks, general characteristics are preserved: first component is an average of degree and betweenness, second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
 Standard deviation of asymmetry and disequilibrium metrics are again coupled to clustering coefficient in the third component.

### 2015-05-29 supportingInformation.tex
atualizando para compartilhar

@@
 The ratios found in email networks are preserved.
-I1 and I4 are outliers, to which a possible explanation is that they are better characterized as a superposition of networks, rather than one coherent network.}
+I1 and I4 are outliers, to which a possible explanation is that they are better characterized as a superposition of networks, rather than one coherent network.
+Degree was used for sectorializing criteria.}
 \begin{center}

### 2015-06-02 tables/tabExtra.tex
added tables and figures used in latex files

F1 & Facebook  & friendship  & no  & the friendship network of Renato Fabbri (author)  & 1367  & 28606 \\\hline
F2 & Facebook  & friendship  & no  & the friendship network of Massimo Canevacci (senior anthropologist)  & 4764  & 59995 \\\hline
F3 & Facebook  & friendship  & no  & the friendship network of a brazilian direct democracy group  & 3599  & 59471 \\\hline
F4 & Facebook  & friendship  & no  & the friendship network of the Silicon Valley Global Network group  & 2026  & 15586 \\\hline
F5 & Participa.br  & friendship  & no  & the friendship network of a brazilian federal social participation portal  & 443  & 910 \\\hline
I1 & Facebook  & interaction  & yes  & the interaction network of the Silicon Valley Global Network group  & 104  & 154 \\\hline
I2 & Facebook  & interaction  & yes  & the interaction network of a Solidarity Economy group  & 64  & 120 \\\hline
I3 & Facebook  & interaction  & yes  & the interaction network of a brazilian direct democracy group  & 214  & 310 \\\hline
I4 & Facebook  & interaction  & yes  & the interaction network of the 'Cience with Frontiers' group  & 530  & 1658 \\\hline
I5 & Participa.br  & interaction  & yes  & the interaction network of a brazilian federal social participation portal  & 222  & 300 \\\hline
TT1 & Twitter  & retweet  & yes  & the retweet network of $\approx 22k$ tweets with the hashtag \#arenaNETmundial  & 2772  & 7222 \\\hline
TT2 & Twitter  & retweet  & yes  & same as TT1, but disconnected agents are not discarded  & 2975  & 7222 \\\hline

### 2015-06-06 supportingInformation.tex
terminada incorporacao das mudancas no artigo
finished integration of n x m x gamma
incorporadas todas as contribs da ultima revisao do chu
foo
adicionada localização do Supporting Information no artigo

@@
 \usepackage[section] {placeins}
+
+\newcommand{\beginsupplement}{%
+	\setcounter{table}{0}
+	\renewcommand{\thetable}{S\arabic{table}}%
+	\setcounter{figure}{0}
+	\renewcommand{\thefigure}{S\arabic{figure}}%
+}
+
+\beginsupplement
 
@@
 \end{center}
-\label{tab:circ}
+\label{tab:circLau}
 \end{table*}
@@
 \end{center}
-\label{tab:circ}
+\label{tab:circLad}
 \end{table*}
@@
 \end{center}
-\label{tab:circ}
+\label{tab:circMet}
 \end{table*}
@@
 \end{center}
-\label{tab:circ}
+\label{tab:circCPP}
 \end{table*}
@@
 \subsection{Time histograms}
-\subsection{Histograms of activity along the hours of the day}
+\subsection{Histograms of activity along the hours of the day}\label{si:hours}
 
@@
 \FloatBarrier
-\subsection{Histograms of activity along the days of the month}
+\subsection{Histograms of activity along the days of the month}\label{si:monthdays}
 Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
@@
 \FloatBarrier
-\subsection{Histograms of activity along months of the year}
+\subsection{Histograms of activity along months of the year}\label{si:months}
 	Activity percentages of the months along the year from LAD list messages.
@@
 
-\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{sec:frac}
+\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{si:frac}
 Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning} of the mais document.
@@
 \FloatBarrier
-\section{PCA of measures along the timeline}\label{sec:pcat}
+\section{PCA of measures along the timeline}\label{si:pcat}
 Loadings for the 14 metrics into the principal components for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.

### 2015-06-19 supportingInformationChu.tex
adicionado SI

--- 2015-06-06 supportingInformation.tex
@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Interaction networks stability: SUPPORTING INFORMATION]{Time stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics (SUPPORTING INFORMATION)}% Force line breaks with \\
+
+\title[Interaction networks stability: Supporting Information]{Time stability in human interaction networks, Supporting Information)}% Force line breaks with \\
 
@@
 
-\begin{abstract}
- This document presents supporting information for the article describing interaction networks stability.
-The sections exhibit activity distribution in time and among participants; the fraction of vertices in the peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance; the analysis of interaction networks from Facebook, Twitter and Participa.br.
-There is a focus on email list interaction networks for benchmarking.
-\end{abstract}
-
-\pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}
 \maketitle
 
-This documents exemplifies the stability of interaction networks by the observation of 4 emails lists (LAD, LAU, MET and CPP), described in Section~\ref{sec:data} of the main article.
-These results were produced with the Gmane public domain data and an open source python package designed for attaining
-these, and related, results. The interested reader should follow Appendix~\ref{scripts} of the main article to access both data and rotines.
-Similar results can be reproduced for any number of (Gmane) email lists.
-To avoid repeating text of each table for each list, the text is given inline, with the tables following the explanation.
+							\section{Activity along time}\label{sec:time}
 
-\section{Time tables in different scales}\label{sec:time}
-Theory presented in Section~\ref{sec:mtime} and results exposed in Section~\ref{constDisc} of the paper~\cite{tpaper}.
+							The activity along time obtained with circular statistics measurements is presented for the email lists.
 
-\subsection{Time circular measures}\label{si:circ}
-The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
-Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
-Section~\ref{sec:mtime} of the main document describes the theoretical background of directional (or circular) statistics.
 							\begin{table*}[!h]
-	\caption{LAU circular measures}
+								\caption{LAU circular measurements}
 								\begin{center}
@@
 							\begin{table*}[!h]
-	\caption{LAD circular measures}
+								\caption{LAD circular measurements}
 								\begin{center}
@@
 							\begin{table*}[!h]
-	\caption{MET circular measures}
+								\caption{MET circular measurements}
 								\begin{center}
@@
 							\begin{table*}[!h]
-	\caption{CPP circular measures}
+								\caption{CPP circular measurements}
 								\begin{center}
@@
 
-\FloatBarrier
-\subsection{Time histograms}
-\subsection{Histograms of activity along the hours of the day}\label{si:hours}
-
-Activity percentages along the hours of the day.
-Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
 							\begin{table}[!h]
@@
 
-\FloatBarrier
-\subsection{Histograms of activity along the days of the week}
-
-Activity percentages along the days of the week. Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.
-
 							\begin{table}[!h]
@@
 
-\FloatBarrier
-\subsection{Histograms of activity along the days of the month}\label{si:monthdays}
-Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
 							\begin{table}[!h]
@@
 							\end{table}
-
-\FloatBarrier
-\subsection{Histograms of activity along months of the year}\label{si:months}
-	Activity percentages of the months along the year from LAD list messages.
-Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD.
-These observations fit academic calendars, vacations and end-of-year holidays.
 
@@
 							\section{Fraction of participants in each Erd\"os Sector along the timeline}\label{si:frac}
-Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning} of the mais document.
-Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown bellow, first for CPP, than for  LAD list.
+							Here we present the fraction of participants in each Erd\"os sector with respect to each criterion defined in Section~\ref{sectioning}.
+Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown below, first for CPP, than for the LAD list.
 
 							Each step size takes two pages of plot.
-On the first page, the criteria are each centrality measure observed separately: in, out and total degrees and strengths.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
+On the first page, the criterion is based on each centrality metric observed separately: in, out and total degrees and strengths.
+In the first six plots, the code for the colors is as follows: red for hubs, green for the fraction of intermediary vertices and blue for the peripheral fraction.
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the greatest component.
-On the same graph, green counts the disconnected vertices.
+On the same graph, green represents the disconnected vertices.
 
-On the second page we show the fractions of participants with respect to each compound criteria for the Erd\"os sectioning.
+							On the second page we show the fractions of participants with respect to each compound criterion for the Erd\"os sectioning.
 In the first plot, the fraction of vertices with unique classification is plotted in black: $\frac{\text{number of nodes uniquely classified}}{\text{number of nodes}}$.
-On the second plot, black represents the exceeding classifications for the given vertices : $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+On the second plot, black represents the exceeding classifications for the given vertices: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 
@@
 							\FloatBarrier
-\section{PCA of measures along the timeline}\label{si:pcat}
-Loadings for the 14 metrics into the principal components for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
-The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics, in accordance with the notation made of Section~\ref{measures} of the main document.
+							\section{PCA for the metrics along the timeline}\label{si:pcat}
+							Loadings for the 14 metrics into the principal components are given for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
 Note that the centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
@@
 							To further verify the hypothesis that such stability is a general property of human social networks,
-we analyzed networks driven from Twitter, Facebook and Participa.br. Selected networks are summarized in
+							we analyzed networks from Twitter, Facebook and Participa.br. Selected networks are summarized in
 							Table~\ref{tab:E}. Their Erd\"os sector relative sizes are given in Table~\ref{tab:secE}. PCA formations are given in
-Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}. Friendship networks considered are undirected and unweighted,
-therefore all measures of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
-Most important results from the consideration herein of this analysis are:
+							Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+The friendship networks considered are undirected and unweighted, therefore all measurements of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
+							The most important results from this analysis are:
 
 							\begin{itemize}
-	\item a further indicative that the stability reported with a focus on email interaction networks is valid for a broader class of phenomena.
-	\item The greater stability of email interaction networks with the same number of participants. This is specially important for benchmarking and probing general properties.
+								\item the stability reported for email interaction networks is also valid for a broader class of phenomena.
+								\item the stability in email interaction networks is higher than for the other networks, considering the same number of participants.
+This is especially important for benchmarking and probing general properties.
 							\end{itemize}
@@
 The acronyms, one for each network, are used throughout Tables~\ref{tab:secE},~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
-All data collected in 2013 and 2014.}
+All the data were collected in 2013 and 2014.}
 								\begin{center}
@@
 The ratios found in email networks are preserved.
-I1 and I4 are outliers, to which a possible explanation is that they are better characterized as a superposition of networks, rather than one coherent network.
-Degree was used for sectorializing criteria.}
+I1 and I4 are outliers, probably because they should be better characterized as a superposition of networks, rather than one coherent network.
+The degree was used for establishing the sectors.}
 								\begin{center}
@@
 							\begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Participa.br yields the networks that most resembles the email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
-The friendship of Renato Fabbri (F1) is the only whose first component has more than 20\% of custering coefficient and second component has more than 40\% of degree centrality.}
+								\caption{First three principal components and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Participa.br yields the networks that most resemble the email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.
+The friendship network of Renato Fabbri (F1) is the only network whose first component has more than 20\% of clustering coefficient and second component has more than 40\% of degree centrality.}
 									\footnotesize
@@
 								\begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Twitter yields the networks that most resembles the email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.}
+									\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resemble the email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.}
 									\footnotesize
@@
 								\begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient and betweenness centrality.
-Twitter yields the networks that most resembles email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
+									\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resemble email networks.
+The general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.
 Important differences are: - the clustering coefficient was only important to the third component for two of the networks ($I2$, $I3$) and does not contribute significantly to any of the first three principal components in $I5$; - in the first component, I5 exhibited less contribution from in-strength, in-degree and betweenness, I4 exhibited less contribution from out-degree.}
@@
 								\begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
-In comparison with the results found in email interaction networks, general characteristics are preserved: first component is an average of degree and betweenness, second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
+									\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
+The characteristics found in email interaction networks are preserved: the first component is an average of degree and betweenness, the second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
 Standard deviation of asymmetry and disequilibrium metrics are again coupled to clustering coefficient in the third component.

### 2015-06-19 supportingInformation.tex
meia revisao do SI

@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Interaction networks stability: SUPPORTING INFORMATION]{Time stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics (SUPPORTING INFORMATION)}% Force line breaks with \\
+%\title[Interaction networks stability: SUPPORTING INFORMATION]{Time stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics (SUPPORTING INFORMATION)}% Force line breaks with \\
+
+\title[Interaction networks stability, Supporting Information]{Time stability in human interaction networks, Supporting Information}% Force line breaks with \\
 
@@
 
-\begin{abstract}
- This document presents supporting information for the article describing interaction networks stability.
-The sections exhibit activity distribution in time and among participants; the fraction of vertices in the peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance; the analysis of interaction networks from Facebook, Twitter and Participa.br.
-There is a focus on email list interaction networks for benchmarking.
-\end{abstract}
-
-\pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}
 \maketitle
 
-This documents exemplifies the stability of interaction networks by the observation of 4 emails lists (LAD, LAU, MET and CPP), described in Section~\ref{sec:data} of the main article.
-These results were produced with the Gmane public domain data and an open source python package designed for attaining
-these, and related, results. The interested reader should follow Appendix~\ref{scripts} of the main article to access both data and rotines.
-Similar results can be reproduced for any number of (Gmane) email lists.
-To avoid repeating text of each table for each list, the text is given inline, with the tables following the explanation.
+\tableofcontents
 
-\section{Time tables in different scales}\label{sec:time}
-Theory presented in Section~\ref{sec:mtime} and results exposed in Section~\ref{constDisc} of the paper~\cite{tpaper}.
+\vfill
+\newpage
+
+This document presents supporting information for the article describing interaction networks stability~\cite{tpaper}.
+Section~\ref{sec:time} exhibits activity statistics in time (circular statistics and histograms) and among participants;
+Section~\ref{si:frac} exposes the fraction of vertices in the peripheral, intermediary and hub sectors;
+Section~\ref{si:pcat} exemplifies the combination of basic topological measures into principal components with greater variance;
+There is a focus on email list interaction networks for benchmarking and
+Section~\ref{si:ext} reinforces the results with the analysis of networks from Facebook, Twitter and Participa.br.
+The interested reader should follow Section~\ref{scripts} of the main article~\cite{tpaper} to access both data and routines.
+
+\section{Time activity in different scales}\label{sec:time}
+Theory presented in Section~\ref{sec:mtime} and results in Section~\ref{constDisc} of the paper~\cite{tpaper}.
 
 \subsection{Time circular measures}\label{si:circ}
-The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
+The rescaled circular mean $\theta_\mu'$, standard deviation $S(z)$, variance $Var(z)$, circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence $\frac{max(incidence)}{min(incidence)}$ at each time unit.
 Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
-Section~\ref{sec:mtime} of the main document describes the theoretical background of directional (or circular) statistics.
+Greater dispersion is found on seconds and minutes, followed
+by days of the month. Greater localization if found in the hours of the day, followed by weekdays and months.
+
 \begin{table*}[!h]
@@
 \subsection{Time histograms}
-\subsection{Histograms of activity along the hours of the day}\label{si:hours}
+\subsubsection{Histograms of activity along the hours of the day}\label{si:hours}
 
@@
 \FloatBarrier
-\subsection{Histograms of activity along the days of the week}
+\subsubsection{Histograms of activity along the days of the week}
 
@@
 \FloatBarrier
-\subsection{Histograms of activity along the days of the month}\label{si:monthdays}
+\subsubsection{Histograms of activity along the days of the month}\label{si:monthdays}
 Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
@@
 \FloatBarrier
-\subsection{Histograms of activity along months of the year}\label{si:months}
-	Activity percentages of the months along the year from LAD list messages.
+\subsubsection{Histograms of activity along months of the year}\label{si:months}
+	Activity percentages in the months along the year.
 	Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD.
@@
 \section{Fraction of participants in each Erd\"os Sector along the timeline}\label{si:frac}
-Here we present the fraction of participants in each Erd\"os sector with respect to each criteria defined in Section~\ref{sectioning} of the mais document.
+	Here we present the fraction of participants in each Erd\"os sector with respect to each criterion defined in Section~\ref{sectioning} of the main document~\cite{tpaper}.
 Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown bellow, first for CPP, than for  LAD list.

### 2015-06-20 supportingInformation.tex
revisado SI

@@
 \begin{table*}[!h]
-	\caption{LAU circular measures}
+	\caption{LAU circular measurements.}
 \begin{center}
@@
 \begin{table*}[!h]
-	\caption{LAD circular measures}
+	\caption{LAD circular measurements.}
 \begin{center}
@@
 \begin{table*}[!h]
-	\caption{MET circular measures}
+	\caption{MET circular measurements.}
 \begin{center}
@@
 \begin{table*}[!h]
-	\caption{CPP circular measures}
+	\caption{CPP circular measurements.}
 \begin{center}
@@
 \begin{table}[!h]
-	\caption{LAU activity along the hours of the day}
+	\caption{LAU activity along the hours of the day.}
 	\footnotesize
@@
 \begin{table}[!h]
-	\caption{LAD activity along the hours of the day}
+	\caption{LAD activity along the hours of the day.}
 	\footnotesize
@@
 \begin{table}[!h]
-	\caption{MET activity along the hours of the day}
+	\caption{MET activity along the hours of the day.}
 	\footnotesize
@@
 \begin{table}[!h]
-	\caption{CPP activity along the hours of the day}
+	\caption{CPP activity along the hours of the day.}
 	\footnotesize
@@
 							Each step size takes two pages of plot.
-On the first page, the criteria are each centrality measure observed separately: in, out and total degrees and strengths.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
+On the first page, the criterion is based on each centrality metric observed separately: in, out and total degrees and strengths.
+In the first six plots, the code for the colors is as follows: red for hubs, green for the fraction of intermediary vertices and blue for the peripheral fraction.
 On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the greatest component.
-On the same graph, green counts the disconnected vertices.
+On the same graph, green represents the disconnected vertices.
 
-On the second page we show the fractions of participants with respect to each compound criteria for the Erd\"os sectioning.
+							On the second page we show the fractions of participants with respect to each compound criterion for the Erd\"os sectioning.
 In the first plot, the fraction of vertices with unique classification is plotted in black: $\frac{\text{number of nodes uniquely classified}}{\text{number of nodes}}$.
-On the second plot, black represents the exceeding classifications for the given vertices : $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
+On the second plot, black represents the exceeding classifications for the given vertices: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
 
@@
 \section{PCA of measures along the timeline}\label{si:pcat}
-Loadings for the 14 metrics into the principal components for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
-The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics, in accordance with the notation made of Section~\ref{measures} of the main document.
+Loadings for the 14 metrics into the principal components are given for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the tables, followed by 7 centrality metrics and 6 symmetry-related metrics.
 Note that the centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
@@
 The three components have in average 80.36\% of the variance.
+Further details are given in Sections~\ref{measures} and~\ref{prevalence} of the main document~\cite{tpaper}.
+
 \subsection{Betweenness, clustering and degree}
@@
 To further verify the hypothesis that such stability is a general property of human social networks,
-we analyzed networks driven from Twitter, Facebook and Participa.br. Selected networks are summarized in
+we analyzed networks from Twitter, Facebook and Participa.br. Selected networks are summarized in
 Table~\ref{tab:E}. Their Erd\"os sector relative sizes are given in Table~\ref{tab:secE}. PCA formations are given in
-Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}. Friendship networks considered are undirected and unweighted,
-therefore all measures of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
-Most important results from the consideration herein of this analysis are:
+Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
+The friendship networks considered are undirected and unweighted, therefore all measurements of strength, in- and out- centralities, asymmetry and disequilibrium have little or no meaning, which is why F1, F2, F3, F4 and F5 are only present in Table~\ref{tab:pcaE1F}.
+The most important results from this analysis are:
 
 \begin{itemize}
+%	\item the stability reported for email interaction networks is also valid for a broader class of phenomena.
 	\item a further indicative that the stability reported with a focus on email interaction networks is valid for a broader class of phenomena.
-	\item The greater stability of email interaction networks with the same number of participants. This is specially important for benchmarking and probing general properties.
+	\item The stability in email interaction networks is higher than for the other networks, considering the same number of participants.
+This is especially important for benchmarking and probing general properties.
 \end{itemize}
@@
 The acronyms, one for each network, are used throughout Tables~\ref{tab:secE},~\ref{tab:pcaE1I},~\ref{tab:pcaE1F},~\ref{tab:pcaE2} and~\ref{tab:pcaE3}.
-All data collected in 2013 and 2014.}
+All the data were collected in 2013 and 2014 within the anthropological physics framework~\cite{anPhy}.}
 \begin{center}
@@
 \begin{table*}[!h]
-	\caption{Percentage of agents in each Erd\"os sector in the friendship and interaction human networks of Table~\ref{tab:E}.
+	\caption{
+	Percentage of agents in each Erd\"os sector in the friendship and interaction human networks of Table~\ref{tab:E}.
 The ratios found in email networks are preserved.
-I1 and I4 are outliers, to which a possible explanation is that they are better characterized as a superposition of networks, rather than one coherent network.
-Degree was used for sectorializing criteria.}
+I1 and I4 are outliers, probably because they should be better characterized as a superposition of networks, rather than one coherent network.
+The degree was used for establishing the sectors.
+	}
 \begin{center}
@@
 \begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Participa.br yields the networks that most resembles the email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
-The friendship of Renato Fabbri (F1) is the only whose first component has more than 20\% of custering coefficient and second component has more than 40\% of degree centrality.}
+	\caption{First three principal components and variance concentration for each of the five friendship networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Participa.br yields the networks that most resemble the email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.
+The friendship network of Renato Fabbri (F1) is the only network whose first component has more than 20\% of clustering coefficient and second component has more than 40\% of degree centrality.}
 	\footnotesize
@@
 \begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
-Twitter yields the networks that most resembles the email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.}
+	\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} in the simplest case: dimensions correspond to degree, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resemble the email networks.
+Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.}
 	\footnotesize
@@
 \begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient and betweenness centrality.
-Twitter yields the networks that most resembles email networks.
-Overall, the general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for second component.
+	\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient and betweenness centrality.
+Twitter yields the networks that most resemble email networks.
+The general characteristic is preserved: first component is an average of degree and betweenness, while clustering is the most relevant for the second component.
 Important differences are: - the clustering coefficient was only important to the third component for two of the networks ($I2$, $I3$) and does not contribute significantly to any of the first three principal components in $I5$; - in the first component, I5 exhibited less contribution from in-strength, in-degree and betweenness, I4 exhibited less contribution from out-degree.}
@@
 \begin{table*}[!h]
-	\caption{First three principal components formation and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
-In comparison with the results found in email interaction networks, general characteristics are preserved: first component is an average of degree and betweenness, second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
+	\caption{First three principal components and variance concentration for each of the seven interaction networks of Table~\ref{tab:E} considering dimensions of in- and out- degrees and strengths, clustering coefficient, betweenness centrality and symmetry related metrics (see Section~\ref{measures}).
+The characteristics found in email interaction networks are preserved: the first component is an average of degree and betweenness, the second component is mostly governed by symmetry related metrics, and clustering coefficient is mostly relevant for the third component.
 Standard deviation of asymmetry and disequilibrium metrics are again coupled to clustering coefficient in the third component.
@@
 
-%\begin{table}[!h]
-%	\caption{PCA2.}
-%	\footnotesize
-%	\input{tables/tabPCA2Extra}
-%\label{tab:pcaE2}
-%\end{table}
-%\begin{table}[!h]
-%	\caption{PCA3.}
-%	\footnotesize
-%	\input{tables/tabPCA3Extra}
-%\label{tab:pcaE3}
-%\end{table}
-
 %\newpage
@@
 \clearpage
-\nocite{*}
+%\nocite{*}
 \bibliography{supportingInformation}% Produces the bibliography via BibTeX.

### 2015-06-21 supportingInformation.tex
incorporadas revisoes e sugestoes do chu

@@
 \section{Time activity in different scales}\label{sec:time}
-Theory presented in Section~\ref{sec:mtime} and results in Section~\ref{constDisc} of the paper~\cite{tpaper}.
+Here we complement the theory presented in Section~\ref{sec:mtime} and results in Section~\ref{constDisc} of the paper~\cite{tpaper}.
 
 \subsection{Time circular measures}\label{si:circ}
-The rescaled circular mean $\theta_\mu'$, standard deviation $S(z)$, variance $Var(z)$, circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence $\frac{max(incidence)}{min(incidence)}$ at each time unit.
+The measurements used were were the rescaled circular mean $\theta_\mu'$, standard deviation $S(z)$, variance $Var(z)$, circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence $\frac{max(incidence)}{min(incidence)}$ at each time unit.
 Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.
@@
 
-Activity percentages along the hours of the day.
 Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
 Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
+The activity peak occurs around midday, with a slight skew toward one hour before noon.
 \begin{table}[!h]
@@
 \FloatBarrier
+
 \subsubsection{Histograms of activity along the days of the week}
-
-Activity percentages along the days of the week. Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.
+Most notably, a decrease of activity on weekends of at least one third and at most two thirds.
 
@@
 \subsubsection{Histograms of activity along the days of the month}\label{si:monthdays}
-Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
+The most important feature seems to be the homogeneity made explicit by the high circular dispersion in the tables of Section~\ref{si:circ}.
+Slightly higher activity rates are found in the beginning of the month, although not statistically significant.
 \begin{table}[!h]
@@
 \subsubsection{Histograms of activity along months of the year}\label{si:months}
-	Activity percentages in the months along the year.
-	Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD.
-	These observations fit academic calendars, vacations and end-of-year holidays.
-
+	Activity is concentrated in Jun-Aug and/or in Dec-Mar.
+	These observations mostly fit academic calendars, vacations and end-of-year holidays.
 \begin{table}[!h]
@@
 The clustering coefficient (cc) appears as the first metric in the tables, followed by 7 centrality metrics and 6 symmetry-related metrics.
-Note that the centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
+The centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
 The clustering coefficient is only relevant for the third principal component, coupled with standard deviations of strengths and degrees.

### 2015-06-21 supportingInformation.tex
retoque de texto

@@
 \subsection{Time circular measures}\label{si:circ}
-The measurements used were were the rescaled circular mean $\theta_\mu'$, standard deviation $S(z)$, variance $Var(z)$, circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence $\frac{max(incidence)}{min(incidence)}$ at each time unit.
+The measurements used were the rescaled circular mean $\theta_\mu'$, standard deviation $S(z)$, variance $Var(z)$, circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence $\frac{max(incidence)}{min(incidence)}$ at each time unit.
 Also, $ \mu_{\frac{max(incidence')}{min(incidence')}} $ and $ \sigma_{\frac{max(incidence')}{min(incidence')} }$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples.

### 2015-06-22 supportingInformation.tex
detalhe de texto
pequenos retoques de escrita

@@
 
-\title[Interaction networks stability, Supporting Information]{Time stability in human interaction networks, Supporting Information}% Force line breaks with \\
+\title[Interaction networks stability (Supporting Information)]{Time stability in human interaction networks (Supporting Information)}% Force line breaks with \\
 
@@
 
-This document presents supporting information for the article describing interaction networks stability~\cite{tpaper}.
-Section~\ref{sec:time} exhibits activity statistics in time (circular statistics and histograms) and among participants;
-Section~\ref{si:frac} exposes the fraction of vertices in the peripheral, intermediary and hub sectors;
-Section~\ref{si:pcat} exemplifies the combination of basic topological measures into principal components with greater variance;
+The supporting information (this document) contains statistics for time activity (circular statistics and histograms) in Section~\ref{sec:time}, the fraction of vertices in the peripheral, intermediary and hub sectors in Section~\ref{si:frac} and the combination of basic topological measures into principal components with greater variance in Section~\ref{si:pcat}.
 There is a focus on email list interaction networks for benchmarking and
 Section~\ref{si:ext} reinforces the results with the analysis of networks from Facebook, Twitter and Participa.br.
-The interested reader should follow Section~\ref{scripts} of the main article~\cite{tpaper} to access both data and routines.
+More context (e.g. methods, discussion, data and scripts) is given in the main document~\cite{tpaper}.
 

### 2015-08-21 supportingInformation.tex
desinvertidas secoes de seccionamento e de pca

@@
 
-\title[Interaction networks stability (Supporting Information)]{Time stability in human interaction networks (Supporting Information)}% Force line breaks with \\
+\title[Interaction networks stability (Supporting Information)]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity (Supporting Information)}% Force line breaks with \\
 
@@
 
+\FloatBarrier
+\section{PCA of measures along the timeline}\label{si:pcat}
+Loadings for the 14 metrics into the principal components are given for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the tables, followed by 7 centrality metrics and 6 symmetry-related metrics.
+The centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
+The clustering coefficient is only relevant for the third principal component, coupled with standard deviations of strengths and degrees.
+The three components have in average 80.36\% of the variance.
+Further details are given in Sections~\ref{measures} and~\ref{prevalence} of the main document~\cite{tpaper}.
+
+\subsection{Betweenness, clustering and degree}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1LAU}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1LAD}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1MET}
+\label{tab:pcain}
+\end{table}
+
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA1CPP}
+\label{tab:pcain}
+\end{table}
+
+\FloatBarrier
+\subsection{Betweenness, clustering, degrees and strengths}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2LAU}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2LAD}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2MET}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA2CPP}
+\label{tab:pcain}
+\end{table}
+
+\FloatBarrier
+\subsection{Betweenness, clustering, degrees, strengths and symmetry measures}
+
+\begin{table}[!h]
+	\caption{LAU principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3LAU}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{LAD principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3LAD}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{MET principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3MET}
+\label{tab:pcain}
+\end{table}
+\begin{table}[!h]
+	\caption{CPP principal components formation and concentration of dispersion.}
+	\footnotesize
+	\input{tables/tabPCA3CPP}
+\label{tab:pcain}
+\end{table}
+
+\FloatBarrier
 \section{Fraction of participants in each Erd\"os Sector along the timeline}\label{si:frac}
 	Here we present the fraction of participants in each Erd\"os sector with respect to each criterion defined in Section~\ref{sectioning} of the main document~\cite{tpaper}.
-Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown bellow, first for CPP, than for  LAD list.
+Step sizes of 50, 100, 250, 500, 1000 and 5000 are shown below, first for CPP, then for  LAD list.
 
@@
 \FloatBarrier
-\section{PCA of measures along the timeline}\label{si:pcat}
-Loadings for the 14 metrics into the principal components are given for all LAD, LAU, MET, CPP, lists, $ws=1000$ messages in 20 disjoint positioning.
-The clustering coefficient (cc) appears as the first metric in the tables, followed by 7 centrality metrics and 6 symmetry-related metrics.
-The centrality metrics, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
-The clustering coefficient is only relevant for the third principal component, coupled with standard deviations of strengths and degrees.
-The three components have in average 80.36\% of the variance.
-Further details are given in Sections~\ref{measures} and~\ref{prevalence} of the main document~\cite{tpaper}.
-
-\subsection{Betweenness, clustering and degree}
-
-\begin{table}[!h]
-	\caption{LAU principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA1LAU}
-\label{tab:pcain}
-\end{table}
-
-\begin{table}[!h]
-	\caption{LAD principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA1LAD}
-\label{tab:pcain}
-\end{table}
-
-\begin{table}[!h]
-	\caption{MET principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA1MET}
-\label{tab:pcain}
-\end{table}
-
-\begin{table}[!h]
-	\caption{CPP principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA1CPP}
-\label{tab:pcain}
-\end{table}
-
-\FloatBarrier
-\subsection{Betweenness, clustering, degrees and strengths}
-
-\begin{table}[!h]
-	\caption{LAU principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA2LAU}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{LAD principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA2LAD}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{MET principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA2MET}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{CPP principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA2CPP}
-\label{tab:pcain}
-\end{table}
-
-\FloatBarrier
-\subsection{Betweenness, clustering, degrees, strengths and symmetry measures}
-
-\begin{table}[!h]
-	\caption{LAU principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA3LAU}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{LAD principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA3LAD}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{MET principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA3MET}
-\label{tab:pcain}
-\end{table}
-\begin{table}[!h]
-	\caption{CPP principal components formation and concentration of dispersion.}
-	\footnotesize
-	\input{tables/tabPCA3CPP}
-\label{tab:pcain}
-\end{table}
-
 \section{Stability in other networks: Twitter, Facebook, Participa.br}\label{si:ext}

### 2015-10-03 scripts/setoresTeoricoEmpirico2.py
feita figura com o seccionamento empirico para colocar no artigo

@@ -1,0 +1,94 @@
+#-*- coding: utf8 -*-
+from __future__ import division
+import numpy as n, pylab as p, networkx as x, random as r, collections as c, string
+from scipy import special # special.binom(x,y) ~ binom(x,y)
+
+__doc__="""Script to plot both scale-free network and random network distributions."""
+
+import gmane as g, os, pickle, time as T, numpy as n, pylab as p
+ENV=os.environ["PATH"]
+import  importlib
+from IPython.lib.deepreload import reload as dreload
+importlib.reload(g.pca)
+importlib.reload(g.loadMessages)
+importlib.reload(g.listDataStructures)
+importlib.reload(g.timeStatistics)
+importlib.reload(g.agentStatistics)
+importlib.reload(g.tableHelpers)
+importlib.reload(g.networkPartitioning)
+importlib.reload(g.networkEvolution)
+importlib.reload(g.evolutionTimelines)
+#importlib.reload(g.interactionNetwork)
+#importlib.reload(g.networkMeasures)
+dreload(g,exclude="pytz")
+os.environ["PATH"]=ENV
+
+labels={'gmane.comp.gcc.libstdc++.devel':"CPP", 'gmane.linux.audio.devel':"LAD", 'gmane.linux.audio.users':"LAU", 'gmane.politics.organizations.metareciclagem':"MET"}
+##print("initializing")
+dl=g.DownloadGmaneData('~/.gmane4/')
+TT=T.time()
+##print("{0:.2f} for initializing download dir initializing".format(T.time()-TT)); TT=T.time()
+def pDump(tobject,tfilename):
+    with open(tfilename,"wb") as f:
+        pickle.dump(tobject,f,-1)
+def pRead(tfilename):
+    with open(tfilename,"rb") as f:
+        tobject=pickle.load(f)
+    return tobject
+
+PDIR="/home/r/repos/gmane/tests/pickledir/"
+NEs=[] # for evolutions of the networks
+
+for lid in dl.downloaded_lists[:1]:
+    label=labels[lid]
+    NEs.append(pRead("{}neP{}.pickle".format(PDIR,label)))
+    print(label+"{0:.2f} for PICKLE loading evolved PCA structures".format(T.time()-TT)); TT=T.time()
[…]

### 2015-10-03 scripts/setoresTeoricoEmpirico2.py
incorporada figura melhorada do seccionamento

@@ -53,10 +53,10 @@ nm=ne.networks_measures[-1]
 np=g.NetworkPartitioning(nm)
 #p.subplot(212)
 p.figure(figsize=(10.,3.))
-p.subplots_adjust(left=0.08,bottom=0.28,right=0.99,top=0.87)
+p.subplots_adjust(left=0.05,bottom=0.28,right=0.99,top=0.87)
 
 
-p.bar([i[0]-0.5 for i in np.bins],
+p.bar([i[0]-0.4 for i in np.bins],
         np.empirical_distribution,
         alpha=.4, label="empirical distribution")
 #p.bar(n.log([i[0]-0.5 for i in np.bins]),
@@ -81,14 +81,28 @@ p.title(u"Three sections of a scale-free network",size=25)
 p.xlabel(r"k $\rightarrow$",size=20)
 #p.xticks(centers)
 p.yticks(())
-p.xticks((2,8),(r"$(k_L)$",r"$(k_R)$"),size=15)
+p.xticks((2,7),(r"$k_L$",r"$k_R$"),size=20)
 p.ylabel(r"P(k) $\rightarrow$",size=20)
 
 c_=n.array(centers)
 pk=c_[1:]**-1.5
 #p.plot(centers[1:],.4*pk,"ko")
 p.plot(centers[1:],.4*pk,"k-o",label="power-law distribution")
-p.ylim(0,0.4*pk[0]+0.1)
+p.ylim(0,0.4*pk[0]+0.03)
+
+# pontos de intersecção:
+x1=2.5; y=41.16-5
+p.plot((x1,x1),(-1000,y),"g--")
+x2=7.5; y=60.-5
+p.plot((x2,x2),(-1000,y),"g--")
+#p.xticks((x1,x2),(r"$(k_L)$",r"$(k_R)$"),size=25)
+
+p.text(-.28,0.12,"periphery"  ,size=20)
+p.text(4,.3,"intermediary",   size=20)
+p.text(10,.08,"hubs"      ,   size=20)
+
+
 
 p.legend()
-p.show()
+p.savefig("../figs/fser__.png")
+#p.show()

### 2015-10-05 paper.tex
rascunhada nova conclusao

--- 2015-04-17 paper.tex
@@
 
+\usepackage{xr}
+\externaldocument{supportingInformation}
 \maxdeadcycles=1000
 
+\usepackage{placeins}
 \begin{document}
@@
 %\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-\title[Stability of interaction networks]{Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics}% Force line breaks with \\
-
+%\title[Stability of interaction networks]{Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.}% Force line breaks with \\
+%\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
+\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
 \author{Renato Fabbri}%
@@
 \affiliation{ 
-S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP)%\\This line break forced with \textbackslash\textbackslash
+	S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP),
+	PO Box 369, 13560-970, S\~ao Carlos, SP, Brazil %\\This line break forced with \textbackslash\textbackslash
 }
-
-\author{Vilson V. da Silva Jr.}
-  \homepage{http://automata.cc/}
-  \email{vilson@void.cc}
-  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-
+%
+%\author{Vilson V. da Silva Jr.}
+%\homepage{http://automata.cc/}
+%\email{vilson@void.cc}
+%\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%
 \author{Ricardo Fabbri}
@@
 \altaffiliation{
-Centro de Ciências Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
+       %Centro de CiÃƒÂªncias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
 }%Lines break automatically or can be forced with \\
@@
 %  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-
-%\author{Osvaldo N. Oliveira Jr.}
-%  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
-%  \email{chu@ifsc.usp.br}
-% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+%
+\author{Leonardo Paulo Maia}
+  \homepage{http://www.ifsc.usp.br/~lpmaia/}
+  \email{lpmaia@ifsc.usp.br }
+ \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+
+\author{Osvaldo N. Oliveira Jr.}
+  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
+  \email{chu@ifsc.usp.br}
+ \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
@@
 \begin{abstract}
- This article reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance.
- We analyzed the temporal activity and topology evolution of networks in four email lists by considering window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline.
-Activity in terms of seconds and minutes exhibit an uniform pattern, while hours, days and months exhibit stable concentrations.
- Participant activity follows the expected distribution of scale-free networks. 
- We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
- Approximately 5\% of the vertices are hubs, 15-35\% are intermediary and the remainder belongs periphery. 
- The metrics that most contribute to data dispersion were found to be centrality-related (degree, strength and betweenness), followed by symmetry-related, and then clustering coefficient. 
- The results also embrace a sketch of a physics-based typology of agents, with proper social and psychological speculations.
- Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed are also present in general human interaction networks.
- Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments, audiovisual reconnaissance, and the report of quantitative differences of textual production as connectivity of agents changes.
+In this study we report on stable aspects of human interaction networks,
+with benchmarks derived from public email lists.
+Activity along time and topology evolution were observed by snapshots in a timeline 
+and in various scales.
+Notably, the activity in timescales, ranging from seconds to months, is practically the same for all networks.
+The most important metrics to the dispersion of participants in the topological measures space 
+are centrality measurements (degree, strength and betweenness),
+followed by symmetry-related metrics and then clustering coefficient.
+The activity of participants followed the expected scale-free trace,
+thus yielding three classes of vertices by comparison against the Erd\"os-R\'enyi model,
+namely hubs, intermediary and periphery.
+The relative size of these three sectors did not vary with time and was essentially the same for all email lists.
+Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices.
+Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
+Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on quantitative criteria.
 \end{abstract}
@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics}
+\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics, social psychology of big data}
 \maketitle
@@
 \begin{quotation}
-`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.
-The fact that human society has been up to now divided into classes affects more than the external relations of men.
-The marks of social repression are left within the individual soul.'
+	`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.' 
 	% `The conception of personality structure is the
@@
 \section{Introduction}\label{sec:into}
-Human interaction networks have been studied long before modern computers.
-Antecedents can be found in nineteenth century works, while the foundation of
-the social network analysis field is generally attributed to the psychiatrist Jacob Moreno~\cite{newmanBook}.
-More recently, the continuous increase of stored data related to human interaction has stimulated research,
-and overall characterization is scattered in literature.
-
-Reports on complex networks are usually focused on a particular measure or functionality, such
-as accessibility measure or detecting communities~\cite{access,newmanModularity}.
-Although there are research on networks arisen from specific phenomenon, such as social or protein networks, there are few attempts for a general characterization and even less endeavors that take in account the temporal dynamic.
-
-The present work is aimed at finding common and stable characteristics among (email) interaction networks.
-This is achieved through observations along time, which imply network evolution, a field that has received attention from the research community
-for more than a decade~\cite{barabasiEvo,newmanEvolving}.
-The evolution is represented herein by a timeline of activity snapshots with a constant number of contiguous messages.
-These snapshots yield a succession of networks.
-This approach, albeit coherent and intuitive, seems not to be explored to date~\cite{netWindowEvo}.
-
-We consider only directed, weighted and human interaction networks.
-Undirected and unweighted representations of such networks are also found in the literature and can be obtained by simplification~\cite{GMANE2}.
-Meaningful measures depend heavily on the model and system characteristics~\cite{newmanStru, newmanWeight}, therefore the most basic (and ubiquitous, and fundamental) metrics of centrality, clustering and symmetry are analyzed altogether in principal components formation.
-
-Typologies are the canon of scientific literature for classification of human agents~\cite{typCanon}.
-Consequently, it was considered central to benefit from human typologies.
-This is a complex matter carefully addressed in methodology, results and appendix.
-
-Although all networks considered originated from email lists,
- coherence with literature suggests that results hold for a more general class of interaction networks,
- such as observed in online platforms (e.g. LinkedIn, Facebook, Twitter).
-Indeed, text mining routines and typologies of online participants benefit from the results here presented~\cite{rcText,rcTipo}.
-
-Next subsection exposes related work.
-Section~\ref{sec:data} details data while Section~\ref{sec:carac} is dedicated to the characterization methods.
-Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work.
-Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:typ} further introduces the typological approach.
-Detailed tables and figures are bundled in Support Information, and are cited upon need.
+Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
+social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{moreno,newmanBook}.
+With the increasing availability of data related to human interactions, research about these networks has grown continuously.
+Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic.
+One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{barabasiHumanDyn,newmanFriendship}, with which
+several features of human interaction have been revealed.
+For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes.
+The dynamics of complex networks representing human interaction has also been addressed~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.
+
+In this paper we analyze the evolution of human interaction networks.
+Interaction in email lists was the most convenient for deriving results and for benchmarking,
+but networks from Facebook, Twitter and Participabr were also considered.
+Using a timeline of activity snapshots with a constant number of contiguous messages,
+we found remarkable stability for several of the network properties.
+Namely, the activity along different timescales exhibit pronounced patterns,
+the most basic topological measures always combine into very characteristic principal components,
+and the fractions of participants in each of the hubs, intermediady and periphery sectors
+are unshaken.
+This is not an intuitive result, given that participants transition in network structure constantly.
+Because these properties were shared by networks from various sources,
+and are consistent with the complex networks literature,
+we advocate that the conclusions might be valid for general classes of interaction networks.
+In particular, this allows us to 
+bridge the gap between data analysis 
+and social sciences in the discussion
+of types of networks and participants.
+Noteworthy, typologies are the canon of scientific literature for classification of human agents,
+with pragmatic standards~\cite{myers} and critical paradigms~\cite{adorno,typCanon}.
+
+Section~\ref{sec:related} describes related work,
+while details of the data and methods of analysis are given in Section~\ref{sec:data} and Section~\ref{sec:carac}.
+Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions.
+Subsidiary data analysis,
+including directions for video and sound mappings of network structures,
+and numeric results for networks from Twitter, Facebook and Participabr,
+are given in the Supporting Information.
 
 \subsection{Related work}\label{sec:related}
-Works on network evolution often consider solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}.
-Exceptions are reported in this section, with emphasis on those more closely related to the present article.
-
-The evolution of human interaction networks was addressed with a community focus, where the direction of edges was not taken into account~\cite{barabasiEvo}.
-In another article, two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email list networks, scale-free properties were reported with $\alpha=1$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
-
-Unreciprocated edges often exceed 50\% in the networks analyzed, which matches empirical evidence from literature~\cite{newmanEvolving}.
-No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were discarded in the present article.
-On the other hand, gender related behavior in mobile phone datasets has been reported~\cite{barabasiSex} and was not considered in this article as the email messages and addresses have no gender related metadata~\cite{GMANE}.
-
-There are research reports based on the results presented in this article, which are exposed in Section~\ref{sec:conc}.
-
-\section{Data description: email lists and messages}\label{sec:data}
+Research on network evolution is often restricted to network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}.
+Network types have been discussed with regard to the number of participants, intermittence of their activity and network longevity~\cite{barabasiEvo}.
+Two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email list networks, scale-free properties were reported with $\alpha \approx 1.8$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
+
+Unreciprocated edges often exceed 50\% in the analyzed networks, which matches empirical evidence from the literature~\cite{newmanEvolving} and motivated the inclusion of symmetry metrics in our analysis.
+No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were not considered in our study.
+Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}, but this was not considered in the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+
+\section{Data and scripts}\label{sec:data}\label{scripts}
 
 Email list messages were obtained from
-the GMANE email archive~\cite{GMANE}, which consists of more than 20,000 email lists and more than 130,000,000 messages~\cite{GMANEwikipedia}.
+the GMANE email archive~\cite{gmanePack}, which consists of more than $20\times 10^3$
+email lists and more than $130\times 10^6$ messages~\cite{GMANEwikipedia}.
 These lists cover a variety of topics, mostly technology-related.
@@
 
-Although many lists were analyzed, four email lists were selected for thorough observation, making it easier to infer general properties.
-These lists, picked as representing both a diverse set and usual lists, are:
+We analyzed many email lists (and data from Twitter, Facebook and Participabr)
+and selected four of them in order to make a thorough analysis,
+from which general properties can be inferred. These lists are:
 \begin{itemize}
-    \item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.}, with participants holding hybrid artistic and technological interests, from different countries.
+	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.},
+		with participants from different countries with artistic and technological interests.
+		English is the most used language. Abbreviated as LAU from now on.
+	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.},
+		with participants from different countries,
+		a more technical and less active version of LAU.
 		English is the language used the most. 
-Abbreviated as LAU from now on.
-    \item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.}, with participants are from different countries, and English is the language used the most.
-A more technical and less active version of LAU.
-Abbreviated LAD from now on.
-    \item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.}, with computer programmers from different countries.
-English is the language used the most.
+		Abbreviated as LAD from now on.
+	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.},
+		with computer programmers from different countries.
+		English is the most used language.
 		Abbreviated as CPP from now on.
-    \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.}, with Brazilian activists holding digital culture interests.
-Portuguese is the most used language, although Spanish and English are also incident.
-Abbreviated MET from now on.
+	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.},
+		a Brazilian digital culture interested email list.
+		Portuguese is the most used language,
+		although Spanish and English are also incident.
+		Abbreviated as MET from now on.
 \end{itemize}
@@
 The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
+We considered 140 additional email lists to
+report on the interdependence between the number of participants and 
+the number of discussion threads.
+Furthermore, 12 additional networks from Facebook (8), Twitter (2) and Participabr (2)
+are scrutinized in the Supporting Information document for better hypothesizing
+about the generality of the results.
 
@@
 	\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
-$N$ is the number of participants (number of different email addresses).
-$\Gamma$ is the number of threads (count of messages without antecedent).
-$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
-A relation holds for all lists carefully considered: as the number of participants increases, the number of threads decreases.
-This underpins a typology sketch of networks, as exposed in Section~\ref{sec:pty}.}
+		$N$ is the number of participants (number of different email addresses),
+		$\Gamma$ is the number of discussion threads (count of messages without antecedent),
+		$\overline{M}$ is the number of messages missing in the 20,000 collection
+		($100\frac{23}{20000}=0.115$ percent in the worst case).
+	}
 	\label{tab:genLists}
@@
 
-\section{Characterization methods}\label{sec:carac}
-The email lists and the networks generated from them were characterized by: 1) statistics of activity along time, from seconds to years; 2) sectioning of the networks in hubs, intermediary and periphery vertices; 3) topological measures prominence regarding dispersion; 4) iterative visualization and data mining; 5) typological speculation about networks and participants.
-Each of these procedures are described below.
-
-Distribution of activity among participants received little dedicated attention, but is considered through almost all results of this article, as degree and strength are highly correlated to activity.
+All data and scripts needed to derive results, figures, tables and this article itself
+are publicly available.
+Email messages are downloadable from the GMANE public database~\cite{GMANEwikipedia}.
+Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}.
+Data from Participabr was used from the linked data/semantic web RDF triples
+reported in~\cite{opa} and available in~\cite{datahub}.
+Script are delivered through a public domain Python PyPI package and
+an open Git repository~\cite{gmanePack}.
+This open approach to both data and scripts
+reinforces the scientific aspect of the contribution~\cite{openSci}
+and lightens ethical and moral issues of researching 
+systems constituted of
+human individuals~\cite{anPhy,ccs15}.
+
+\section{Methods}\label{sec:carac}
+%The networks were characterized with: 1) statistics of activity along time, in scales from seconds to years; 2) dispersion of basic topological metrics; 3) sectioning of the networks in hubs, intermediary and periphery; 4) iterative visualization, sonification and data inspection.
+%These procedures are described below.
 
 \subsection{Time activity statistics}\label{sec:mtime}
-  Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year.
-This resulted in histograms from which obvious patterns can be observed.
-Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution.
-
-  Measures of mean and dispersion were taken using circular statistics.
-For such, each $measure$ (data point) is represented as a complex number of unit magnitude $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measure\frac{2\pi}{period}$.
-Moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:
+Messages were counted over time as histograms in terms of seconds,
+minutes, hours, days of the week, days of the month, and months of the year.
+Most standard measures of location and dispersion, such as usual mean and
+standard deviation, hold little meaning in a compact Riemannian manifold.
+Equivalent measures were taken using circular statistics,
+in which each measurement $t$ (data point) is represented as a unity complex number,
+$z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=t\frac{2\pi}{T}$,
+and $T$ is the period in which the counting is repeated.
+For example, $\theta=12\frac{2\pi}{24}=\pi$ for a message sent at $t=12h$ and given $T=24h$ for days.
+The moments $m_n$, lengths of moments $R_n$, mean angles $\theta_\mu$, and rescaled mean angles $\theta_\mu'$ are defined as:
 
@@
 	\theta_\mu&=Arg(m_1) \nonumber \\
-	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
+	\theta_\mu'&=\frac{T}{2\pi} \theta_\mu \nonumber
 \end{align}
 
-The $\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:
+$\theta_\mu'$ is used as the measure of location.
+Dispersion is measured using the circular variance $Var(z)$,
+the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:
 
@@
 
-Although very simple, such measures enables both comparison of the email networks exposed in this article, and the deliverance of a benchmark to which other interaction networks can be compared.
-
-\subsection{Interaction network}\label{intNet}
-Interaction networks can be modeled both weighted or unweighted, both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
-Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e.
-we did not observe directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks.
-The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
-The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
-Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
-This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}.
+\noindent
+Also, the ratio $r=\frac{b_l}{b_h}$ between the lowest and the highest incidences on the histograms 
+served as a further clue of how close the distribution was to being uniform.
+As expected, a positive correlation was found in all $r, Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures,
+which can be noticed in the Section~\ref{si:circ} of the Supporting Information document.
+The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred in the discussion of results.
+
+\subsection{Interaction networks}\label{intNet}
+Interaction networks can be modeled both weighted or unweighted,
+both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
+Networks in this paper are directed and weighted, the most informative of trivial possibilities.
+We did not investigate directed unweighted, undirected weighted,
+and undirected unweighted representations of the interaction networks.
+
+The interaction networks were obtained as follows:
+a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
+The reasoning is: if B wrote a response to a message from A,
+he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
 Edges in both directions are allowed.
-Each time an interaction occurs, one is added to the edge weight.
-Self-loops were regarded as non-informative and discarded.
-These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}.
+Each time an interaction occurs, the value of one is added to the edge weight.
+Selfloops were regarded as non-informative and discarded.
+Inverting edge direction yields the status network:
+B read the message and considered what A wrote worth responding, giving status to A,
+thus $B\rightarrow A$.
+This paper considers by convention the information network as described above ($A\rightarrow B$) and depicted in Figure~\ref{formationNetwork}.
+These human social information interaction networks are reported in the literature
+as exhibiting scale-free and small-world properties,
+as expected for a number of social networks~\cite{bird,newmanBook}.
 
@@
 	\centering
-    \includegraphics[width=0.5\textwidth]{figs/criaRede_}
-    \caption{Formation of interaction network from email messages.
+	\includegraphics[width=0.44\textwidth]{figs/criaRede2_}
+	\caption{The formation of interaction networks from exchanged messages.
 	Each vertex represents a participant.
-A reply message from participant B to a message from participant A is regarded as evidence that B received information from A and yields a directed edge.
+	A reply message from author B to a message from author A is regarded as evidence
+	that B received information from A and yields a directed edge.
 	Multiple messages add ``weight'' to a directed edge.
@@
 
-Edges can be created from all antecedent message authors on the message-response thread to each message author.
-We only linked the immediate antecedent to the new message author, both for simplicity and for the valid objection that in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
+%Edges can be created from all antecedent message authors on the message-response thread to each message author.
+%We only linked the immediate antecedent to the new message author, both for simplicity and because in adding two edges, $x\rightarrow y$ and $y\rightarrow z$, there is also a weaker connection between $x$ and $z$.
 Potential interpretations for this weaker connection are: double length, half weight or with one more ``obstacles''.
-This suggests pertinence of centrality measures that account for the connectivity with all nodes, such as betweenness centrality and accessibility~\cite{luMeasures,access}.
-
-%\subsubsection{Sectioning of networks in periphery, intermediary and hubs sectors}\label{sectioning}
+This suggests the adequacy of centrality measurements to account for the connectivity of a node with all other nodes, such as betweenness centrality, eigenvector centrality and accessibility~\cite{luMeasures,access}.
+
+\subsubsection{Topological metrics}\label{measures}
+
+The topology of the networks was studied using a small selection of the most basic and fundamental measurements for each vertex, as follows:
+
+\begin{itemize}
+	\item Degree     $k_i$: number of edges linked to vertex $i$.
+	\item In-degree  $k_i^{in}$: number of edges ending at vertex $i$.
+	\item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
+	\item Strength $s$: sum of weights of all edges linked to vertex $i$.
+	\item In-strength $s_i^{in}$: sum of weights of all edges ending at vertex $i$.
+	\item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
+	\item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
+	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
+\end{itemize}
+
+In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$:
+
+\begin{itemize}
+	\item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
+	\item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
+$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
+	\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
+	\item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
+	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{dis}_i-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
+\end{itemize}
+
+These measures are used both for the Erd\"os sectioning (described in Secion~\ref{sectioning}) and for performing PCA (as described in Section~\ref{sec:pca}).
+
 \subsection{Erd\"os sectioning}\label{sectioning}
-In scale-free networks, the periphery, intermediary and hubs sectors can be derived from comparison with an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
-As a shorthand, this procedure is from now on called \emph{Erd\"os sectioning}, and the resulting sectors the \emph{Erd\"os sectors} or \emph{primitive sectors}.
+In a scale-free network, the peripheral, intermediary and hubs sectors can be derived from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
+We shall refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors being referred to as \emph{Erd\"os sectors} or \emph{primitive sectors}.
 
 The degree distribution $\widetilde{P}(k)$ of an ideal
-scale-free network $\mathcal{N}_f$ with $N$ vertices and $z$ edges has less
+scale-free network $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
 average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
@@
 
-If $\mathcal{N}_f$ is directed and has no self-loops, the probability
-of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$ (see Appendix~\ref{ap:ded}).
-A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertices and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability:
+If $\mathcal{N}_f(N,z)$ is directed and has no self-loops, the probability
+of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$.
+A vertex in the ideal Erd\"os-R\'enyi digraph with the same number of vertices and edges, and thus the same probability $p_e$ for the presence of an edge, will have degree $k$ with probability
 
@@
 
-The lower degree fat tail represents the border vertices, i.e.
+The lower degree fat tail corresponds to the border vertices, i.e.
 the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
@@
 $\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
-The reasoning for this classification is: 1) vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g.
-without preferential attachment) are correctly associated to the hubs sector.
+The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g.
+without preferential attachment and as in the Erd\"os R\'enyi model) are correctly associated to the hubs sector.
 Vertices with very few connections, which are way more abundant than expected by pure chance, are assigned to the periphery.
-Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in scale-free phenomena, are classified as intermediary.
+Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in the real network, are classified as intermediary.
 
@@
 	\centering
-    \includegraphics[width=0.5\textwidth]{figs/fser_}
-    \caption{Degree distribution of scale-free and Erd\"os-R\'enyi ideal networks. The latter has more
-        intermediary vertices, while the former has more peripheral and hub vertices. Sector borders are
-        given by the two intersections $k_\Leftarrow$ and $k_\Rightarrow$ of the connectivity distributions. Characteristic degrees
-    are in compact intervals of degree: $[0,k_\Leftarrow]$, $(k_\Leftarrow,k_\Rightarrow]$, $(k_\Rightarrow,k_{max}]$ for the Erd\"os sectors (periphery, intermediary and hubs).}
+	\includegraphics[width=0.5\textwidth]{figs/fser__}
+	\caption{Classification of vertices by comparing degree
+		distributions.
+		The binomial distribution of the Erd\"os R\'enyi network exhibit more intermediary vertices, while a scale-free network, associated with the power-law distribution,
+		has more peripheral and hub vertices.
+		The sector borders are defined with respect to the intersections of the distributions.
+Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the periphery,
+intermediary and hubs sectors, the ``Erd\"os sectors''.
+		The connectivity distribution of empirical interaction networks, e.g. derived from email lists,
+		can be sectioned by comparison against the associated binomial distribution with the same number of vertices and edges.
+		In this figure, a snapshot of 1000 messages from CPP list yield the degree distribution
+		of an interaction network of 98 nodes and 235 edges.
+		A throughout exposition of the method is exposed in Section~\ref{sectioning}.}
 		\label{fig:setores}
 \end{figure}
-
 To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network.
-Thus, each bin, starting at degree $k_i$, spans $\Delta_i=[k_{i},k_{j}]$ degree values, where $j$ is the smallest integer with which there are at least $\eta$ vertices with degree larger than or equal $k_i$, and less than or equal $k_{j}$.
-This changes equation~\ref{criterio} to:
+The range $\Delta$ of incident values should be partitioned in $m$ parts $\Delta=\cup_{i=1}^m \Delta_i$,
+with $\Delta_i\cap \Delta_j=\emptyset \; \forall\; i \neq j$:
+%\begin{equation}
+%\begin{split}
+\begin{align}
+\Delta_i =\Biggl\{ k\;\; & | & \overline{\Delta}_{i-1}< &\, k\leq j \text{ and }\;\;\;\;\;\;\;\;\;\;\;\;\nonumber\\
+		      & \Biggl[ & N - \sum_{k=0}^{\overline{\Delta}_{i-1}} & \eta_k< \eta \text{ or }\\
+&	\Bigl[ & \sum_{k=\overline{\Delta}_{i-1}+1}^j &  \eta_k \geq \eta \text{ and }\;\;\;\;\;\;\nonumber\\
+& \Bigl( &\sum_{k=\overline{\Delta}_{i-1}+1}^{j-1} &  \eta_k < \eta \text{ or }   j=\overline{\Delta}_{i-1}+1 \Bigr) \;\Bigr] \Biggr] \Biggr\}\nonumber
+\end{align}
+%\end{split}
+%\end{equation}
+
+\noindent where $\eta_k$ is the number of vertices with degree $k$, while $\overline{\Delta}_{i}=max(\Delta_{i})$, and $\overline{\Delta}_{0}=-1$.
+Equation~\ref{criterio} can now be written in the form:
 
 \begin{equation}\label{criterio2}
-    \sum_{x=k_i}^{k_j} \widetilde{P}(x) < \sum_{x=k_i}^{k_j} P(x) \Rightarrow \text{i is intermediary}
+	\begin{split}
+		\sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} P(x) \Rightarrow \\
+	\Rightarrow \Delta_i \text{ holds intermediary degree values.}
+	\end{split}
 \end{equation}
@@
 If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with:
+For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with
 \begin{equation}
-\hat{P}(k^{way})=\binom{N-1}{k^{way}}p_e^k(1-p_e)^{N-1-k^{way}}
+	\hat{P}(k^{way})=\binom{N-1}{k^{way}}p_e^k(1-p_e)^{N-1-k^{way}},
 \end{equation}
@@
 In and out strengths ($s^{in}$, $s^{out}$) are divided by $\overline{w}$ and compared also using $\hat{P}$.
-Notice that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with total degree.
-
-In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as:
+Note that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with the total degree.
+
+In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and each of the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as
 
@@
 	e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
-e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber
+	e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber,
 \end{alignat}
 
-\noindent where $\{\overline{k}_{\gamma,i}\}$ is:
+\noindent where $\{\overline{k}_{\gamma,i}\}$ is
 
@@
 		\overline{k}_{5,i}&=\frac{s_i^{in}}{\overline{w}} \\
-\overline{k}_{6,i}&=\frac{s_i^{out}}{\overline{w}} \\
+		\overline{k}_{6,i}&=\frac{s_i^{out}}{\overline{w}}
 	\end{split}
@@
 
-Since different metrics can be used to identify the three types of vertices, compound criteria can be defined.
+Since different metrics can be used to identify the three types of vertices, compound criteria were defined for conveniently analysing 
+networks with a low number of messages,
+such as in Section~\ref{si:frac} of the Supporting Information.
 For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
@@
 	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics.
-In this case, vertices classified (usually) does not reach 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
+In this case, vertices classified do not usually reach 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
 	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics.
-Therefore, a vertex may belong to more than one class, and total members may add more than 100\%, which is indicated by a black line in the figures of Appendix~\ref{figures}.
+Therefore, a vertex may belong to more than one class, and the total number of members may exceed 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
 	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics.
@@
 The remaining vertices are regarded as peripheral.
-    \item Inclusivist cascade $C_4$: vertices are hubs if they are classified as so according to any of the metrics.
+	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as such according to any of the metrics.
 The remaining vertices are classified as intermediary if they belong to this category for any of the metrics.
@@
 	\item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics.
-The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric.
+The remaining vertices are classified as peripheral if they fall into the periphery or hub classes for any metric.
 The rest of the nodes are classified as intermediary.
@@
 
-Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1<\delta<6$ can be described as:
+Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1\leq\delta\leq6$, can be described as:
 
@@
 
-The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests.
-Notice that the exclusivist cascade is the same sectioning that an inclusivist cascade from periphery to hubs (with inverted order of sectors precedence).
-Compound criteria can be used to examine network sections in the case of a low number of messages, such as in the last figures of Appendix~\ref{figures}.
-
-Results from applying the Erd\"os sectioning, i.e.
-the achievement of well defined sectors by comparison of the real network with the Erd\"os R\'enyi model, are reported in Section~\ref{subsec:pih} and Appendix~\ref{figures}.
-
-\subsection{Topology metrics for Principal Component Analysis}\label{measures}
-
-The topology of the networks was observed by Principal Component Analysis (PCA~\cite{pca}) outcomes
-with a small selection of the most basic and fundamental measurements for each vertex:
-
-\begin{itemize}
-    \item Degree     $k_i$: number of edges linked to vertex $i$.
-    \item In-degree  $k_i^{in}$: number of edges ending at vertex $i$.
-    \item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
-    \item Strength $s$: sum of weights of all edges linked to vertex $i$.
-    \item In-strength $s_i^{in}$: sum of weights of all edges ending at vertex $i$.
-    \item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
-    \item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
-    \item Betweenness centrality $bt_i$: fraction of geodesics that contain the vertex $i$. Betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
-\end{itemize}
-
-In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$ (see Section~\ref{prevalence}):
-
-\begin{itemize}
-    \item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-    \item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$.
-Where $e_{xy}$ is 1 if there is and edge from $x$ to $y$, $0$ otherwise.
-$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
-    \item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{asy} -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
-    \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-    \item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-    \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu_{dis}-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
-\end{itemize}
-
-\subsection{Evolution of the networks}
-   The evolution of the networks was observed within a fixed number of messages, the window size $ws$.
-This same number of contiguous messages is considered with different shifts in the message timeline to obtain snapshots.
-Each snapshot was used to both do the Erd\"os sectioning (see Section~\ref{sectioning}) and to perform PCA with topological measures (see Section~\ref{measures}).
-Furthermore, this approach to evolution was used for visualization (see Section~\ref{sec:viz}) and for typological speculations (see Section~\ref{subsec:typ}).
-The $ws$ used were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000. Within a same $ws$, the number of vertices and edges vary in time, as do other network characteristics.
-
-\subsection{Visualization of network evolution}\label{sec:viz}
-The evolution of the networks was visualized with animations, image galleries and online gadgets made for this research~\cite{animacoes,galGMANE,appGMANE}.
-Such visualizations were crucial to guide research into the most important features of network evolution, and prompted us to capture the prominence of topological metrics along time using mean and standard deviations (see Section~\ref{measures} and Appendix~\ref{sec:pcat}), in addition to the size of the three sectors in a timeline fashion (Appendix~\ref{figures}).
-Visualization of network structure was specially useful as part of the email lists data mining, from which parts of relevant structures and results were driven.
-See Appendix~\ref{sec:related} for further directions about visualization and text mining concerning the results herein presented, with dedicated articles.
-
-\subsection{Typology speculation}\label{subsec:typ}
-Qualitative, typological speculations were the result of all the methodology.
-More specifically, the Erd\"os sectors was regarded as yielding a \emph{primitive typology} of human agents in social contexts with at least dozens of participants.
-Visualization and data mining were efficient in fueling speculations about the qualities of each type. 
-The stability found in principal components of topological measures suggested that this primitive typology is general for (practically) all human contexts.
-Also, the first author is a developer with cultural interests and has been directly and indirectly part of these communities for more than a decade.
-This gave further safety about assumptions although care was taken not to regard this acquaintanceship as primary source or as definitive evidence.
-The interested reader should see Appendix~\ref{ap:typ} for further context, 
-as it might be considered audacious to bridge from a physics-based, 
-quantitative classification to a qualitative and speculative approach.
-
+Notice that the exclusivist cascade is the same sectioning of an inclusivist cascade from periphery to hubs, but with inverted order of sectors. 
+The simplification of all possible compound possibilities to the small set listed above might be formalized in strict mathematical terms, but this was considered out of the scope for current interests.
+
+%\subsubsection{Sectioning of networks in peripheral, intermediary and hubs sectors}\label{sectioning}
+\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
+Principal Component Analysis (PCA) is a well documented technique~\cite{pca} and was used to acquire knowledge about:
+1) which metrics contribute to each principal component and in which proportion; 2) how much of the dispersion is concentrated in each component; 3) expected values and dispersions for these quantities over various networks.
+
+Let $\mathbf{X}=\{X[i,j]\}$ be a matrix of all vertices $i$ and the respective
+values for each metric $j$, 
+$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ the mean of metric $j$ over all $I$ vertices, 
+$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
+and $\mathbf{X'}=\{X'[i,j]\}=\left\{\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}\right\}$ 
+the matrix with the \emph{z-score} of each metric. 
+Let $\mathbf{V}=\{V[j,k]\}$ be the matrix $J\times J$ of eigenvectors
+of the covariance matrix $\mathbf{C}$
+of $\mathbf{X'}$, one eigenvector per column.
+Each eigenvector combines the original measures into one principal component, therefore
+$V'[j,k]=100\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
+is the percentage of the principal component $k$
+that is directly proportional to the measure $j$.
+%With $k$ eigenvectors 
+%$D[k]$,
+%it is enough to 
+Be $\mathbf{D}=\{D[k]\}$ the eigenvalues associated to the eigenvectors $\mathbf{V}$,
+then $D'[k]=100*\frac{D[k]}{\sum_{k'}D[k']}$
+is the percentage of total dispersion of the system that the principal component $k$
+is responsible for.
+We consider, in general, the three greatest eigenvalues and
+the respective eigenvectors in percentages:
+$\{(D'[k],\;V'[j,k])\}$.
+These usually sum between 60 and 95\% of the dispersion
+and reveal patterns for a sound analysis.
+In particular, 
+given $L$ snapshots $l$ of the interaction network,
+we are interested in the mean
+$\mu_{V'}[j,k]$
+and the standard deviation $\sigma_{V'}[j,k]$ 
+of the contribution of metric $j$ to the principal component $k$,
+and the mean
+$\mu_{D'}[k]$
+and the standard deviation 
+$\sigma_{D'}[k]$
+of the contribution of the component $k$ to the dispersion
+of the system:
+
+\begin{align}\label{eq:pca}
+\mu_{V'}[j,k]   &=\frac{\sum_{l=1}^L V'[j,k,l]}{L}\nonumber\\
+\sigma_{V'}[j,k]&=\sqrt{\frac{\sum_{l=1}^L (\mu_{V'}-V'[j,k,l])^2}{L}}\\\nonumber
+\mu_{D'}[k]&=\frac{\sum_{l=1}^L D'[k,l]}{L}\\\nonumber
+\sigma_{D'}[k]&=\sqrt{\frac{\sum_{l=1}^L (\mu_{D'}-D'[k,l])^2}{L}}
+\end{align}
+
+The covariance matrix 
+$\mathbf{C}$ is the correlation matrix because $\mathbf{X'}$ is normalized.
+Therefore, $\mathbf{C}$ is also directly observed as a first clue for patterns
+by the most simple associations:
+low absolute values indicate low correlation (and a possible independence);
+high values indicate positive correlation;
+negative values with a high absolute value indicate negative correlation.
+\subsection{Evolution and visualization of the networks}\label{sec:viz}
+The evolution of the networks was observed within sequences of snapshots. with a fixed number of messages,
+Each of these sequences had a fixed number of messages, the window size $ws$,
+which was considered with different shifts in the message timeline to obtain snapshots.
+In each snapshot was performed both the PCA with topological metrics and the Erd\"os sectioning.  
+The values of $ws$ employed were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000.
+Variations in the number of vertices, edges
+and other network characteristics
+are exposed in Section~\ref{si:frac}
+of the Supporting Information document.
+
+Networks were visualized with animations, sonifications,
+image galleries and online gadgets developed specifically for this research~\cite{animacoes,galGMANE,appGMANE}.
+Such \emph{audiovisualizations} were crucial to guide the research into the most important features of network evolution.
+Furthermore, the size of the three Erd\"os sectors could be visualized in a timeline fashion.
+Visualization of network structure was especially useful in the initial inspection of data and of the structures derived from the email lists.
+
+%This is a way to enhance the reliability of the methods,
+%of algorithmic routines, of data consistency,
+%and of the results themselves.
+%Also, we believe that this practice raises the scientific contribution of the work,
+%handing not only the framework and the results, 
+%but the exact data and processes that render them~\cite{openSci}.
 \section{Results and discussion}\label{sec:results}
-
-Remarkable features from the analysis of the four email lists are:
-\begin{itemize}
-    \item The activity along time is practically the same for all lists (Section~\ref{constDisc}) and suggests stable patterns.
-    \item The fraction of participants in each Erd\"os sector is stable along time and can be observed even with very few messages (Section~\ref{subsec:pih}).
-    \item The topological measures (regarded as most basic and still representative for a first analysis) combine into principal components in the same way for all lists and all snapshots. 
-	    Furthermore, symmetry related measures, defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
-    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
-\end{itemize}
-
-The reminder of this section contains a detailed discussion of these results and Support Information holds further tables and figures from which such results can be better observed.
-The user activity exhibited in Table~\ref{autores} is
-
-\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
-One remarkable feature from the analysis is that the activity 
-along time is practically the same for all lists. Circular mean and standard deviation
-for each time scale are given in Table~\ref{tab:circ}.
-
-\begin{table*}[t]
-	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$.
-The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
+%Remarkable features from the analysis of the four email lists are:
+%\begin{itemize}
+%    \item The activity along time is practically the same for all lists, thus suggesting stable patterns.
+%    \item The fraction of participants in each Erd\"os sector is stable along time and can be determined even with very few messages 
+%    \item The topological metrics combine into principal components in PCA in the same way for all lists and all snapshots (). 
+%        \item Symmetry measures of the topology, as defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
+%    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
+%\end{itemize}
+\subsection{Activity along time}\label{constDisc}
+The observed activity along time, in terms of seconds, minutes, hours, days and months,  is practically the same for all lists.
+Histograms in each time scale were computed as were circular average and dispersion values.
+We provide detailed values in Table~\ref{tab:circ}-\ref{tab:min2} because they can
+be used for characterizing nodes (participants) in these and other networks,
+and networks themselves.
+For example, they may serve for identification of outliers in a community.
+
+\begin{table}
+	\caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$ described in Section~\ref{sec:mtime}.
+		This typical table was constructed using all LAD list messages, and the results are the same for other lists, as shown in Section~\ref{si:circ} of the Supporting Information document.
+	The most uniform distribution of activity was found in seconds and minutes, where the mean has little meaning.
+	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 14h and 15h ($\theta'=-9.61$).
+	Weekdays, month days and months have mean near zero (i.e. near the beginning of the week, month and year) and high dispersion.}
 	\begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c| }
+		\begin{tabular}{ |l|| c|c| }
 			\hline
-	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
-	\input{tables/tab2TimeCPP}
+			%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+			%& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
+			{\bf scale} & {\bf mean} \boldsymbol{$\theta_\mu'$} & {\bf dispersion} \boldsymbol{$\delta(z)$ }  \\ \hline\hline
+			\input{tables/tab2TimeLAD___}
 		\end{tabular}
@@
 	\label{tab:circ}
-\end{table*}
-
-\subsubsection{Seconds and minutes}
-Messages were slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}: for both seconds and minutes $\frac{max(incidence)}{min(incidence)} \in (1.26,1.275]$ in the lists, while Simulations reach these values but have in average more discrepant higher and lower peaks $\xi=\frac{max(incidence')}{min(incidence')} \Rightarrow \mu_\xi=1.2918 \text{ and } \sigma_\xi=0.04619$.
-Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform, i.e.
-no trend was detected and the pattern is the uniformity.
-Circular dispersion is maximized and mean has little meaning.
-
-\subsubsection{Hours of the day}
+\end{table}
+
+\begin{table}
+	\caption{Activity percentages along the hours of the day for the CPP list.
+Nearly identical distributions are found on other lists as shown in Section~\ref{si:hours} of the Supporting Information document.
 Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
 Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.
-See Table~\ref{dia} for the fraction of activity in 1h, 6h and 12h periods along the day.
-
-\begin{table}[!h]
-	\caption{.}
+Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.}
 	\footnotesize
@@
 
-\subsubsection{Days of the week}
-Higher activity was observed during weekdays, specially for the CPP and MET lists (see Table~\ref{semana}).
-The decrease of activity on weekends reaches at least one third, and two thirds in extreme cases.
-
-\begin{table}[h]
-	\caption{.}
+\begin{table}
+	\caption{Activity percentages along the days of the week for the four email lists.
+Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
 	\begin{center}
@@
 \end{table}
-
-\subsubsection{Days along the month}
-Variation of activity in the days along the month is less prominent, one cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active.
-The most important trait seems to be homogeneity.
-Last days of the month (29, 30 and 31) are not present in every month, and observed activity is proportional to incidence rates.
-See Table~\ref{mes} for further insights.
-
-\begin{table}[!h]
-	\caption{.}
+In the scale of seconds and minutes, activity obeys a homogeneous pattern, with the messages being slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
+In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks $\xi=\frac{min(incidence')}{max(incidence')} \Rightarrow \mu_\xi=0.7741 \text{ and } \sigma_\xi=0.02619$.
+Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform.
+In these cases, the circular dispersion is maximized and the mean has little meaning as indicated in Table~\ref{tab:circ}.
+As for the hours of the day, an abrupt peak appeared around 11am with the most active period being the afternoon.
+Days of the week revealed a decrease between one third and two thirds of activity on weekends.
+Days of the month were regarded as homogeneous with an inconclusive slight tendency of the first week being more active.
+Months of the year revealed patterns matching usual work and academic calendars.
+The time period examined here was not sufficient for the analysis of activity along the years.
+These patterns are exemplified in Tables~\ref{tab:hin}-\ref{tab:min2}.
+
+\FloatBarrier
+
+\begin{table}
+	\caption{Activity in the days along the month for the MET list.
+Nearly identical distributions are found on other lists as indicated in Section~\ref{si:monthdays} of the Supporting Information.
+Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.}
 	\footnotesize
-	\input{tables/tabMonthdaysCPP}
+	\input{tables/tabMonthdaysMET}
 	\label{tab:min}
@@
 
-\subsubsection{Months and larger divisions of the year}
- Activity is concentrated in Jun-Aug for MET and LAD, and in Dec-Mar for CPP, LAU and LAD (see Table~\ref{ano}). These observations fit academic calendars, vacations and end-of-year holidays.
-
-\begin{table}[!h]
-	\caption{.}
+\begin{table}
+	\caption{Activity percentages of the months along the year from LAU list.
+Activity is usually concentrated in Jun-Aug and/or in Dec-Mar (see Section~\ref{si:months} of the Supporting Information).
+These observations fit academic calendars, vacations and end-of-year holidays.}
 	\footnotesize
-	\input{tables/tabMonthsLAD}
-\label{tab:min}
+	\input{tables/tabMonthsLAU}
+	\label{tab:min2}
 \end{table}
 
-\subsubsection{Activity long years}
-Literature reports relation of number of participants and hubs stability to community longevity~\cite{barabasiEvo}.
-Also, activity should follow trends such as economic, climate and technological.
-Periodicity should be observed to some extent.
-Nevertheless, the time period examined was not sufficient for the analysis of activity along the years (see Section~\ref{sec:conc}).
-
-\subsection{Scalable fat-tail structure: constancy of membership fractions in the Erd\"os sectors}\label{subsec:pih}
-
-There is a concentration of hub activity and of vertex with few connections.
-Table~\ref{autores} exposes this expected distribution of activity among participants in a scale-free context.
-As we have seen, one way to observe such distribution of participants is through a comparison of the real network with an Erd\"os R\'enyi model.
-
-\begin{table}[!h!]
-    \caption{Distribution of activity among agents.
-First column presents the percentage of messages sent by the most active participant.
-Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $-10D$ column has maximum percentage of participants responsible for 10\% of messages.}
+\subsection{Stable fractions of participants in each Erd\"os sector}\label{subsec:pih}
+
+The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered. 
+Moreover, the same distribution applies to the networks of all email lists analyzed,
+as demonstrated in Figure~\ref{fig:sectIL} and in Section~\ref{si:frac} of the Supporting Information document. 
+Activity is highly concentrated on the hubs, while a very large number of peripheral vertices contribute to only a fraction of the activity.
+This is expected for a system with a scale-free profile, as confirmed by the data in Table~\ref{autores} for the distribution of activity among participants.
+
+Typically, $\approx [3-12]\%$ of the vertices are hubs, $\approx [15-45]\%$ are intermediary and $\approx [44-81]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
+These results hold for the total, in and out degrees and strengths.
+Stable distributions can also be obtained for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
+The networks hold their basic structure with as few as 10-50 messages,
+i.e. concentration of activity and the abundance of low-activity participants take place even with very few messages,
+which is highlighted in Section~\ref{si:frac} of the Supporting Information document.
+A minimum window size for the observation of more general properties might be inferred by monitoring 
+both the giant component and the degeneration of the Erd\"os sectors.
+
+In order to support hypotheses about the generality of these findings,
+we obtained the Erd\"os sectors of 12 networks from Facebook, Twitter and Participabr.
+The results are given in Table~\ref{tab:secE} of the Supporting Information,
+which indicate that the percentages of hubs, intermediary and periphery nodes are essentially the same as for the email lists.
+
+\begin{figure*} 
+	\centering
+	\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
+	\caption{Fractions of agents in each Erd\"os sector, where the fractions for hubs, intermediary and peripheral vertices are represented in red, green and blue, respectively.
+Fractions derived from degree and strength criteria are both on the left.
+	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right,
+with black lines representing the fraction of vertices without class and with more than one class, respectively. 
+See Section~\ref{si:frac} of Supporting Information document for a collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}.
+Table~\ref{tab:secE}, also from Supporting Information document, presents these fractions of agents in networks from Facebook, Twitter and Participabr.}
+	\label{fig:sectIL}
+\end{figure*}
+
+\begin{table}[h]
+	\caption{Distribution of activity among participants.
+The first column presents the percentage of messages sent by the most active participant.
+The column for the first quartile ($1Q$) shows the minimum percentage of participants responsible for at least 25\% of total messages.
+Similarly, the column for the first three quartiles $1-3Q$ gives the minimum percentage of participants responsible for 75\% of total messages.
+The last decile $-10D$ column exposes the maximum percentage of participants responsible for 10\% of messages.}
 	\begin{center}
@@
 
-The distribution of vertices in the  hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered.
-Moreover, the same distribution applies to the networks of all the four email lists, to which are dedicated the various figures in Appendix~\ref{figures}. 
-Typically, $\approx 5\%$ of the vertices are found to be hubs, $\approx [15-35]\%$ are intermediary and $\approx [60-80]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
-These results hold for total, in and out degrees and strengths.
-Stable distributions can also be obtained for 100 of less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
-The networks hold their basic structure with as few as 10-50 messages, concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of Appendix~\ref{figures}.
-A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the hub, intermediary and peripheral sections.
-
+\subsection{Stability of principal components and the prevalence of symmetry over clusterization for dispersion}\label{prevalence}
+%The topology was analyzed using standard, well-established metrics of centrality and clustering.
+%We also introduced symmetry metrics given the evidence of their importance in social contexts~\cite{newmanEvolving}.
+The contribution of each metric to the variance is very similar for all the networks and along time.
+In applying PCA to the snapshots, the contribution of each metric to the principal components presents very small standard deviation.
+Table~\ref{tab:pcain} exemplifies the formation of principal components for the MET email list.
+Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
+of the Supporting Information document for the other email lists (benchmarks) and other interaction networks,
+with the consideration of strategic combinations of metrics.
+
+\begin{table}[!h]
+	\caption{Loadings for the 14 metrics into the principal components for the MET list, $ws=1000$ messages in 20 disjoint positioning.
+The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
+Note that the centrality measurements, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
+The clustering coefficient is only relevant for the third principal component.
+The three components have in average 80\% of the variance.}
+		\footnotesize
+		\input{tables/tabPCA3CPP}
+		\label{tab:pcain}
+	\end{table}
+
+The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality.
+Therefore, all of these centrality measurements are equally important for characterizing the networks.
+On one hand, the relevance of all centrality metrics is not surprising since they are highly correlated, e.g.the degree and strength have Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
+	On the other hand, each of these metrics is related to a different participation characteristic, and their equal relevance is noticeable.
+The clustering coefficient is presented in almost perfect orthogonality to centrality metrics.
 \begin{figure*} 
 	\centering
-        \includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
-	\caption{Fraction of agents in each Erd\"os sector.
-Hubs, intermediary and periphery fractions are represented in red, green and blue.
-In this figure, simple criteria employed using degree and strength.
-Exclusivist and inclusivist compound criteria have black lines for fraction of vertices without class and with more than one class, respectively.
-See Appendix~\ref{figures} for a collection of such timeline figures with all simple and compound criteria and selected measures.}
-    \label{tab:sectIL}
-\end{figure*}
-
-For the histograms used in the classification process (see Section~\ref{sectioning}), the use of at least $\eta$ vertices for each bin did not yield significant differences.
-That was understood as a consequence of the observation scale (see Appendix~\ref{ap:ded}).
-
-\subsection{Variance prevalence of centrality over symmetry and symmetry over clusterization}\label{prevalence}
-The topology was analyzed using standard, well-established metrics of centrality and clustering.
-We also introduced symmetry measures because there is evidence that they are relevant in social contexts~\cite{newmanEvolving}.
-The contribution from the distinct metrics to network topology is very similar for all the networks, and did not vary with time.
-This stability in network behavior is remarkable, with small standard deviations of the contribution of each metric to principal components in all lists analyzed.
-Table~\ref{fig:pcain} exemplifies the principal components formation with all measures considered, while Tables~\ref{compPCA0}-\ref{compPCA2} of Support Information gives further assistance to the interested reader.
-
-\begin{table}[!h]
-	\caption{.}
-	\footnotesize
-	\input{tables/tabPCA3MET}
-\label{tab:min}
-\end{table}
-
-The principal component exhibited a pondered sum of centrality measures: degrees, strengths and betweenness centrality.
-This composition of the principal component suggests that all six degree and strength measures are equally important for system characterization, although it is known that they do not relate to the same participation characteristics.
-Clustering coefficient is presented in almost perfect orthogonality.
-Dispersion was more prevalent in symmetry related measures than clustering coefficient.
-
-Degree and strength are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-The typical PCA plot for the two first components exhibited in Table~\ref{compPCA} is shown in the second system of Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong.
-As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the second component (clustering related).
-The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects symmetry metrics relevance for the variance. 
-This can be observed in Table~\ref{compPCA2} where the clustering coefficient is only relevant for the third principal component (with contributions from out-degree and out-strength). 
-We concluded that the symmetry-related measurements can be more meaningful in characterizing interaction networks (and their participants) than the clustering coefficient, especially for hubs and intermediary vertices.
-
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr3PCA}
-%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of in- and out- degree and strength, betweenness centrality and clustering coefficient, as specified in Section~\ref{measures}.
-Principal component is a weighted average of centrality measures: degrees, strengths and betweenness centrality.
-Second component is mostly clustering coefficient.
-Table~\ref{compPCA} shows the composition of principal components.
-Similar plots were obtained for all window sizes $ws\;\in\;[500,10000]$, and for the networks of the other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.}
-%    \label{PCA}
-%\end{figure}
-
-\begin{figure*} 
-   \centering
-        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT}
-	\caption{Prevalence of symmetry measures over clustering for data dispersion.
-This was considered an important result of PCA analysis, together with stability on component formation.
-First plot holds the usual degree versus clustering coefficient layout.
-Second plot is similar, but the first component is an average of centrality measures (see Table~\ref{compPCA}).
-Third plot resulted from applying PCA with symmetry measures (see Table~\ref{compPCA2}), where greater dispersion can be observed in all sectors, although more pronounced in intermediary and hubs sectors.
-    For this figure, a window size of $ws = 1000$ messages was used from LAU list.
-The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
-Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.
-    In the third plot, clustering coefficient is omitted as it is only relevant for third component.
-In this case, second component is representative of symmetry measurements of vertex interactions.
-Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.
-    }
+	\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
+	\caption{The first plot shows degree versus clustering coefficient. This typical pattern is well known with high clustering more incident in vertices with lower degrees. 
+		The second plot is analogous but the abscissas is an average of centrality metrics.
+		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
+		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
+		for characterizing interaction networks than the clustering coefficient,
+		especially for hubs and intermediary vertices.
+		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
+		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
+		and in networks from Facebook, Twitter and Participabr,
+		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
+		the symmetry-related metrics and clustering coefficient.}
 	\label{fig:sym}
@@
 
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr11CC}
-%    \caption{Clustering coefficient versus degree of vertices with a window size of $ws = 1000$ email messages, LAU list.
-The general layout is consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.}
-%    \label{clust}
-%\end{figure}
-
-%\begin{figure} 
-%   \centering
-%        \includegraphics[width=\columnwidth]{figs/ev0pr1PCA}
-%    \caption{Scatter plot of vertices for the LAU list using two principal components from a PCA in the metrics space of (in-,  out- and total) degree, (in-,  out- and total) strength, betweenness centrality, clustering coefficient and symmetry-related measurements.
-The composition of the first three components are shown in Table~\ref{compPCA2} and metrics details are given in Section~\ref{measures}.
-Most importantly, clustering coefficient is only relevant for third component, being second component representative of symmetry measurements of vertex interactions.
-Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.}
-%    \label{PCA2}
-%\end{figure}
-
-    \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
-Such stability of activity along time, Erd\"os sector relative sizes, principal components formation and concentration of variance suggests that understanding types of networks and participants might benefit from a characterization within such framework.
-We considered the most trivial case, in which the sector to which a participant belongs is regarded as her/his type.
-This is useful in deriving more elaborate typologies, and yield a sketch of or a ``primitive'' typology.
-General considerations:
+Dispersion was more prevalent in symmetry-related metrics than for the clustering coefficient, as indicated in Table~\ref{tab:pcain}.
+This is also illustrated in Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong to.
+As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the third component (clustering related).
+The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects the relevance of the symmetry-related metrics for the variance.
+We conclude that the symmetry metrics is more useful, with respect to the dispersion in the topological space,
+in characterizing interaction networks (and their participants) than the clustering coefficient,
+especially for hubs and intermediary vertices.
+
+The relative importance of the topological metrics was also observed for the additional 12 networks from Facebook, Twitter and Participabr.
+With the exception of two of these networks, the overall behavior was maintained in that centrality measurements 
+were found prevalent in the first principal component,
+followed by symmetry-related metrics on the second principal component and then clustering coefficient on the third principal component.
+The results are given in Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2},~\ref{tab:pcaE3} of the Supporting Information document.
+Larger variability was found among such networks, which motivated the
+use of interaction networks derived from email lists for benchmarks.
+
+\subsection{Types from Erd\"os sectors}\label{sec:pty}
+The type of the Erd\"os sector to which a participant belongs 
+can be regarded as implying a type to the corresponding participant.
+Assigning a type to a participant raises important 
+issues about the potential for stigmatization and prejudice. 
+In the present case, the type of a participant
+changes both along time and as
+different networks are considered,
+and therefore the potential for prejudice of such
+participant typology is attenuated~\cite{adorno}. 
+In other words, an individual is a hub in a number of networks and peripheral in other networks, and even within the same network he/she probably changes type along time~\cite{4hubs}.
+Moreover, such proposed typology of participants bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.
+
+We analyzed the temporal evolution of the networks using visualization tools developed for this research~\cite{rcText,versinus} and inspected the raw data to infer the main characteristics of each type.
+Our main observations may be summarized as follows:
 \begin{itemize}
-	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
-	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
-	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for the compound inclusive and exclusive criterion given in Section~\ref{sectioning}).
-	\item The participant often transitions from one sector to another within a network.
-	\item Reported stability of network structure arises from activity with continuous change of the participant types.
-	\item Type often changes with the scale of the activity window. For example, two snapshots of 200 and 10000 messages will probably reveal participants with more than one type. 
-\end{itemize}
-
-This ephemerality of the human type is in accordance with canonical literature on psychological types.
-The observance of often transitions between types is even though of as an argument against prejudice~\cite{adorno}.
-Visualizations and raw data manipulations suggests further typological peculiarities. These are initial observations, which inspired this article and other ongoing research~\cite{rcText,versinus}:
-
-\begin{itemize}
-    \item Core hubs usually have intermittent activity.
-Very stable activity was found on MET hubs, which motivated its integration to this work.
-There are reports in the literature of greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
+	\item Core hubs usually have intermittent activity. Very stable activity was found only in smaller communities, which is consistent with the literature~\cite{barabasiEvo}.
 	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
@@
 Therefore, intermediary vertices seem responsible for the network structure.
-For example, intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
+Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
 	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
 These seem as authorities and contrast with participants that respond much more than receive responses.
-    \item The most obvious community structure, as observed by a high clustering coefficient, is found only in peripheral and intermediary sectors.
+	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
 \end{itemize}
 
-This ``primitive typology'', characterized by peripheral, intermediary and hub types, can be further scrutinized using concepts from other typologies, such as Meyer-Briggs, Pavlov or F-Scale.
-This is not a direct result of numeric analysis, it is a description refinement of the found structure, in typological terms, and coherent with complex networks literature.
-Although initial, this bridges human and exact sciences in the most pertinent way authors were able to, as is herein considered a result.
-
-Another typology suggested by results is about the networks themselves.
-This and paired articles~\cite{rcText,versinus} reports a dipole in human interaction network types: networks with few and stable agents, and (relative) many threads per amount of messages contrasts with networks with many agents of intermittent activity and (relative) few threads per amount of messages.
-This network typology is endorsed by the seminal Nature Letter by Palla, Barab{\'a}si and Vicsek~\cite{barabasiEvo}, which suggests that the smaller size of MET community is responsible for the stronger hubs observed.
-
-\section{Conclusions and further work}\label{sec:conc}
-The characterization of interaction networks resulted from stability observations.
-Along temporal activity statistics, this work reports the stability of the formation of principal components and of the relative sizes of periphery, intermediary and hubs Erd\"os sectors.
-These results suggested typologies for both agents and the networks.
-
-Observed systems were coherent with literature in different aspects, such as concentration of activity or such as the clusterization versus connectivity pattern.
-Even so, analysis of data from other virtual environments, such as Facebook, Twitter and LinkedIn, might help understanding how general are these structures and what are proper uses.
-
-Further work should inspect other topological measures in each sector.
-The subtraction $\widetilde{P}(k)-P(k)$ (see Section~\ref{subsec:typ}) results in two positive clusters for periphery and hubs, and a negative cluster for intermediary vertices.
-This might support classification of the three sectors by clustering, a more traditional approach to classification.
-Observance of attributes with greater contribution to principal components of LDA should reveal best chances to present these three sections as clusters in the network measurements space. 
-Another possibility, specially for a brute-force characterization of such sectors, is to remove vertices with degree close to $k_\Leftarrow$ or $k_\Rightarrow$ depicted in figure~\ref{fig:setores}.
-
-Stability here reported eases tipologization of both outliers and usual participation patterns (see Sections~\ref{sec:related} and~\ref{sec:pty}).
-A paired article with this one is reports significant differences in the textual production of each Ed\"os sector~\cite{rcText}. Resulting knowledge purposes networks and participant typologies.
-Another paired article describes a network time evolution visualization method~\cite{versinus}.
-The usage of such results are taking place in software packages, linked data, electronic government technologies, and anthropological physics experiments~\cite{gmanePack,ops,opa,ensaio,anPhy}.
+With focus on the networks as the whole objects of analysis,
+we were able to observe a peculiar correlation pattern 
+between the number of threads and the number of participants.
+When the number of participants exceeds a threshold,
+the number of threads displays a positive correlation with the
+number of participants.
+This finding is illustrated in Figure~\ref{fig:nmgamma3d} and can also be observed in Table~\ref{tab:genLists}.
+Obviously, network types can be derived from such results, which was not attempted here but 
+left for future work.
+
+\begin{figure}
+   \centering
+        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
+	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists.
+Highest numbers of threads are found in lists with few participants.
+The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise.
+This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}.
+For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants.
+All LAU, LAD, MET lists present smaller networks.}
+	\label{fig:nmgamma3d}
+\end{figure}
+
+\section{Conclusions}\label{sec:conc}
+The temporal stability reported in this article is very counterintuitive.
+These characteristics were expected from current scientific knowledge,
+such as the fat tail structure and the preponderance of centrality measures,
+but the very small standard deviations of principal components formation (Section~\ref{}),
+the presence of the Erd\"os sectors even in networks with
+few participants (Section~\ref{}),
+and the recurrent activity patterns along different timescales (Section~\ref{}),
+goes a step further in characterizing scale-free networks in the context
+of the interaction of human individuals.
+Maybe secondary, but nonetheless important, the importance of symmetry-related metrics,
+which surpassed that of clustering coefficient,
+with respect to dispersion of the system in the topological space,
+might add to current understanding of how Digraphs differ from
+undirected graphs.
+Noteworthy is the also very stable fraction participants in each Erd\"os sectors
+when the network reaches more than 200 participants (more that 1000 messages).
+Benchmarks were derived from email list networks
+and the analysis of
+networks from Facebook,
+Twitter and Participabr might ease hypothesizing
+about the generality of such results.
+Selected tables and figures sum dozens of pages
+and are gathered in
+the Supporting Information document.
+
+Further work should expand the analysis to include
+more types of networks and more metrics.
+The data and software needed to reach these results
+might also receive dedicated and in-depth
+documentation as they enable a greater level of transparency
+and work share,
+which is adequate to both benchmarking
+and the study of systems constituted
+by human individuals (see Section~\ref{sec:data}).
+The derived typology of hub, intermediary and peripheral participants
+was used for semantic web and participatory democracy efforts,
+and might be enhanced to yield scientific knowledge~\cite{pnud3,pnud4,pnud5}.
+Also, we should pursue the publication of the visualization and sonification
+processes used for this research~\cite{versinus,4hubs} and
+of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.
+
+% trabalhos de visualização (versinus), diferenciação de texto
+% aplicação da tipologia para democracia participativa e dados ligados
+% necessidade de publicar os dados em formatos ligados para comparacao da estrutura
+% Future work on including more measures, other networks, sharing
+% data in RDF for benchmarking, exploring larger timescales
+% and of the data and OWL ontologies 
+% 
+%
+
+%A systematic study of the activity of participants belonging to the three 
+%distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices,
+%while the network structure was governed by the intermediary vertices.
+%These properties were shared by all email lists and were time-independent,
+%which is consistent with the literature.
+%We may therefore consider the Erd\"os sectors as leading to a human typology which bridges exact sciences, with quantitative procedures for the classification, and human sciences, where there is a legacy in the observation of human types.
 
 \begin{acknowledgments}
-Renato Fabbri is grateful to CNPq (process: 140860/2013-4,
-project 870336/1997-5), United Nations Development Program (PNUD/ONU, contract: 2013/000566; project BRA/12/018)  and 
-the Postgraduate Committee of the IFSC/USP. This author is also grateful for
-the American Jewish Committee for maintaining an online copy of the Adorno book
-used on the epigraph~\cite{adorno}.
-Authors thank GMANE creators and maintainers, specifically: GMANE is run by Lars Magne Ingebrigtsen, and the administrators are Tom Koelman, Jason R.
-Mastaler, Steinar Bang, Jon Ericson, Wolfgang Schnerring, Sebastian D.B.
-Krause, Nicolas Bareil, Raymond Scholz, and Adam Sjøgren.
-Authors thank referred email lists communities and welcome feedback as core contribution to this, and similar, research.
+	Financial support was obtained from CNPq (140860/2013-4,
+	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
+	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers, and to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
+	We are also grateful to developers and users of Python scientific tools.
 \end{acknowledgments}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{More graph-related results}
-\subsection{Edge existence probability in a directed network without self-loops}\label{ap:ded}
-Be $\mathcal{N}$ a directed network without self-loops with $z$ edges and $N$ vertices.
-The probability that an edge exists between two arbitrary vertex is $p_e=\frac{z}{max( \text{number of edges} |\ \text{N vertices})}$, where $max( \text{number of edges} |\ \text{N vertices})=2[(N-1)+(N-2)+...+1]=2[\sum_1^{N-1}i]=2[\frac{N(N-1)}{2}]$ is the maximum number of edges for a network with $N$ vertices.
-Therefore:
-\begin{align}
-    p_e&=\frac{z}{max( \text{number of edges} |\ \text{N vertices})} \nonumber \\
-       &=\frac{z}{2[(N-1)+(N-2)+...+1]}=\frac{z}{2\frac{N(N-1)}{2}} \nonumber \\
-   p_e &=\frac{z}{N(N-1)}
-\end{align}
-
-\subsection{Effect of scale in histogram smoothing}\label{ap:ded}
-There are between 20 and 200 participants in the message window sizes used to derive most of the results ($ws \in [200,1500]$ messages).
-As peripheral vertices are abundant and span few degrees, there are more than $\eta$ vertices with each low degree value.
-For the case of higher degrees, one should consider that with the $ws$ used, each participant is $p \in [0.1\%,0.5\%]$ of all participants.
-Therefore, if incident connectivity is very improbable in an Erd\"os R\'enyi network (less than $p$, the probability that a single participant represents when the histogram is normalized to the density function), than it is not an intermediary connectivity, but a hub.
-Therefore, using at least $\eta$ vertices for each bin did not impact the results.
-
-\section{Typologies}\label{ap:typ}
-There are other ways to split and characterize networks.
-To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e.
-the radius is the minimum eccentricity).
-In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network).
-Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery.
-Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.
-
-Models of human dynamics can be used to predict and classify activity.
-In this case, human activity is commonly considered a Poisson process, as a consequence of the randomly distributed events in time.
-Even so, evidence-based models suggests that human activity patterns follow non-Poisson statistics, characterized by a long tail of inactivity with bursts of rapidly occurring events~\cite{barabasiHumanDyn,barabasiPhone}.
-
-Typologies can also be conveniently adapted from psychiatric, psychological and psychoanalytic theories.
-Concerning empirical research,
-Theodor Adorno is a core conceiver of an one-of-a-kind typology that resulted from observing authoritarian
-personality traces to detect Nazism adoption, antisemitism and potential fascists, depicted as an authoritarian syndrome~\cite{adorno}.
-
-Other classic typologies of interest include Jung's extroversion-introversion trait with four modes of orientation.
-This four modes are divided in two perceiving functions (sensation and intuition) and two judging functions (thinking and feeling), each individual manifesting one of these four modes as dominant, and each mode expressed primarily as introverted or extroverted~\cite{jung}.
-Myers-Briggs Type Indicator extrapolated Jungian theories into a questionnaire and added perceiving and judging as a fourth dipole~\cite{myers}.
-Even plain Freudian criteria, such as neurosis, psychosis, perversity and denegation, can be used directly for such categorization, as they have verbal and behavioral typical traces~\cite{freud,freud2}.
-
-It was considered central to benefit from key human typologies, both by describing types and by further characterizing classes in the terms encountered.
-A primitive physics-based typology is described in Section~\ref{sec:pty} as a consequence of the periphery, intermediary and hub sectors yielded by comparing the real networks with the Erd\"os R\'enyi model.
-Also, ethic and moral issues are developed by such legacy.
-For example, Adorno et al.
-conceptualized that personality is dynamic, not static or immutable, and that recognizing this was important for an ethic empirical study of human authoritarian traces~\cite{adorno}.
-Indeed, this dynamic typological approach is so vital to secure an ethic study of human systems that our epigraph is devoted to make this point explicit.
-
-\section{Data and scripts}\label{scripts}
-Messages are downloaded from the GMANE database by RSS in the mbox email text format. 
-They are requested one by one to avoid reaching maximum size of the requests accepted by
-GMANE API.
-
-Every message has about 30 fields, from which the following are crucial
-for the present work:
-\begin{itemize}
-    \item ``From'' field, as it specifies the sender of the message, in the usual format of ``First\_name Last\_Name $<email>$''.
-    \item ``Date'' field, which is given with the resolution of a second.
-    \item ``Message-ID'', important to state antecedent/consequent relation between messages and therefore from an author to a replier.
-    \item ``References'', has the ID of the message it is an answer to, if any, and earlier messages in the thread.
-\end{itemize}
-
-Field ``In-Reply-To'' has only the ID of the message it replies and can be sometimes
-a shortcut or an alternative to ``References''. Also, the textual content of the messages,
-accessed through ``payload'' method of the mbox message object, is of central interest and
-the authors dedicated an article to include the textual content of the messages to the analysis~\cite{rcText}.
-
-\subsection{Python scripts}\label{ap:os}
-Basic constructs for obtaining all results are the product of scripts written in the Python programming language.
-These are kept in a public git repository for backup and sharing with research community~\cite{scriptsFim}.
-Core scripts, for deriving structures and results exhibited in this article, are in the LEIAME file.
-
-\subsection{Third party libraries and software}
-The programming framework used 
-is mainly Python-based, with emphasis on usual
-scientific tools. More specifically,
-scripts where written for 2.7.3 version of Python,
-with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
-Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
-\clearpage
-\section{Tables}\label{sectables}
-\subsection{PCA tables}\label{sec:pcat}
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-The first component is a weighted average of degree and betweenness centrality.
-The second component is mostly clustering coefficient.
-The first and second components represent more than 95\% of total variance.
-The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
-$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
-$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
-$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
-  \end{tabular}
-  \label{compPCA0}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components composition in percentages.
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-First component is a weighted average of degree and strength and betweenness centrality.
-The second component is mostly related to the  clustering coefficient.
-The first and second components represent more than 90\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
-$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
-$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
-$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
-$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
-$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
-$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
-$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
-$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
-  \end{tabular}
-  \label{compPCA}
-\end{table}
-
-\begin{table}[!h]
-  \centering
-  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}).
-LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics.
-In this case, clusterization is pushed to the third principal component.
-The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation.
-Betweenness centrality again has a role similar to degree, but weaker.
-The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength.
-The three components have in average 80.36\% of the variance.}
-  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
- & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
-       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
-$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
-$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
-$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
-$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
-$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
-$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
-$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
-$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
-$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
-$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
-$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
-$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
-$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
-$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
-$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
-  \end{tabular}
-  \label{compPCA2}
-\end{table}
-
-\subsection{Tables for activity along time and among participants}\label{tabTime}
-
-\begin{table}[!h]
-    \caption{Distribution of activity among agents.
-First column is dedicated to percentage of messages sent by the most active participant.
-Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages.
-The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
-\begin{center}
-    \begin{tabular}{ | l ||  c | c | c | c | }
-        \hline
-        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
-    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
-    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
-    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
-    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
-    \end{tabular}
-\end{center}
-\label{autores}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Percentage of activity ($100\frac{\text{counted messages}}{\text{total messages}}$) in each hour, 6 hours and 12 hours.
-Maximum activity rates are in bold.
-In 1h columns, minimum activity is also bold.
-The less active period of the day is around 4-6h.
-Maximum activity is between 10-13h.
-Afternoon is most active in 6h division of the day.
-The noon has $\approx \frac{2}{3}$ of 24h activity.
-}\label{dia}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-    & 1h   & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h &   1h & 6h & 12h \\ \hline\hline
-0h  & 3.66 & \multirow{6}{*}{10.67} & \multirow{12}{*}{33.76} & 2.87  & \multirow{6}{*}{7.15} & \multirow{12}{*}{29.33} & 3.58 & \multirow{6}{*}{10.14} &  \multirow{12}{*}{36.88} & 4.00 & \multirow{6}{*}{10.77} & \multirow{12}{*}{33.13} \\
-1h  & 2.76 &                        &                         & 1.77  & & &  2.22 & & & 2.52 & & \\
-2h  & 1.79 &                        &                         & 1.04  & & & 1.63 & & & 1.79 & & \\
-3h  & 1.10 &                        &                         & 0.64  & & & 1.06 & & & 1.06 & & \\
-4h  & {\bf 0.68} &                        &                         & 0.47  & & &  0.84 & & & 0.75 & & \\
-5h  & 0.69 &                        &                         & {\bf 0.38}  & & & {\bf 0.82} & & & {\bf 0.66} & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-6h  & 0.83 & \multirow{6}{*}{23.09} &                         & 0.72  & \multirow{6}{*}{22.18} & & 1.17 & \multirow{6}{*}{26.74} & & 0.85 & \multirow{6}{*}{22.36} &  \\
-7h  & 1.24 &                        &                         & 1.33  & & & 2.37 & & & 1.56 & & \\
-8h  & 2.28 &                        &                         & 2.67  & & & 3.54 & & & 2.96 & & \\
-9h  & 4.52 &                        &                         & 4.40  & & & 6.04 & & & 4.68 & & \\
-10h & 6.62 &                        &                         & 6.29  & & & {\bf 6.83} & & & 5.93 & & \\
-11h & {\bf 7.61} &                        &                         & 6.78  & & & 6.79 & & & 6.40 & & \\\hline
-12h & 6.44 & \multirow{6}{*}{\bf 37.63} & \multirow{12}{*}{\bf 66.24} & {\bf 7.33}  & \multirow{6}{*}{\bf 42.22} & \multirow{12}{*}{ \bf 70.66} & 6.11 & \multirow{6}{*}{\bf 35.65} &  \multirow{12}{*}{ \bf 63.12} & {\bf 6.41} & \multirow{6}{*}{\bf 37.25} &  \multirow{12}{*}{\bf 66.87} \\
-13h & 6.04 &                        &                         & 7.08  & & & 6.26 & & & 6.12 & & \\
-14h & 6.47 &                        &                         & 7.09  & & & 6.38 & & & 6.33 & & \\
-15h & 6.10 &                        &                         & 7.14  & & & 5.93 & & & 5.98 & & \\
-16h & 6.22 &                        &                         & 6.68  & & & 5.52 & & & 6.40 & & \\
-17h & 6.36 &                        &                         & 6.89  & & & 5.46 & & & 6.02 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-18h & 6.01 & \multirow{6}{*}{28.61} &                         & 5.99  & \multirow{6}{*}{28.44} & & 5.24 & \multirow{6}{*}{27.46} & & 5.99 & \multirow{6}{*}{29.63} & \\
-19h & 5.02 &                        &                         & 5.23  & & & 4.52 & & & 5.03 & & \\
-20h & 4.85 &                        &                         & 4.98  & & & 4.55 & & & 4.63 & & \\
-21h & 4.38 &                        &                         & 4.37  & & & 4.42 & & & 4.59 & & \\
-22h & 4.06 &                        &                         & 4.24  & & & 4.51 & & & 4.88 & & \\
-23h & 4.30 &                        &                         & 3.64  & & & 4.23 & & & 4.53 & & \\\hline
-    \end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table}[h]
-    \caption{Percentage of activity on days along the week.
-Weekend days are at least $\frac{1}{3}$ less active and can reach $\frac{1}{3}$ of activity.
-MET concentrates activity in weekdays the most, leaving only 13.98\% of total activity to Saturday and Sunday.
-LAU is the one that less concentrates activity in weekdays, reaching 20.94\% of total activity in weekends.
-These might suggest professional relation of CPP and MET participants to the topics of interest, or a hobby relation of LAU and LAD participants.}
-\begin{center}
-    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
-        \hline
-        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
-        CPP & 17.06 & 17.43 & 17.61 & 17.13 & 16.30 & 6.81 & 7.67 \\ \hline
-        MET & 17.53 & 17.54 & 16.43 & 17.06 & 17.46 & 7.92 & 6.06 \\ \hline
-        LAU & 15.71 & 15.80 & 15.88 & 16.43 & 15.13 & 10.13 & 10.91 \\ \hline
-        LAD & 14.91 & 17.73 & 17.01 & 15.40 & 14.25 & 10.39 & 10.30 \\\hline
-    \end{tabular}
-\end{center}
-\label{semana}
-\end{table}
-
-\begin{table*}
-%\tiny
-    \caption{Activity along the days of the month.
-The pattern is to have no clear prevalent period.
-One might point a slight tendency for the first two weeks to be more active, although this table does not present statistical foundation for such an assumption.
-For the scope of this study, differences of activity along the month is assumed to be inexistent.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c| c|c|c| c|c|c| c|c|c|}
-        \hline
-    & \multicolumn{3}{c|}{CPP} & \multicolumn{3}{c|}{MET} & \multicolumn{3}{c|}{LAU} & \multicolumn{3}{c|}{LAD}  \\ \hline
-day    & 1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days &   1 day & 7 days & 14 days \\ \hline\hline
-1  & 3.19 & \multirow{7}{*}{23.05} & \multirow{14}{*}{45.63} & 3.01 & \multirow{7}{*}{25.16} & \multirow{14}{*}{48.08} & 3.34 & \multirow{7}{*}{23.06} &  \multirow{14}{*}{47.31} & 3.22 & \multirow{7}{*}{21.96} & \multirow{14}{*}{46.70} \\
-2  & 3.07 &                        &                         & 3.38  & & &  3.38 & & & 3.42 & & \\
-3  & 3.20 &                        &                         & 3.55  & & & 3.20 & & & 2.87 & & \\
-4  & 3.63 &                        &                         & 4.34  & & & 3.52 & & & 2.91 & & \\
-5  & 2.85 &                        &                         & 3.93  & & &  2.68 & & & 3.30 & & \\
-6  & 3.67 &                        &                         & 3.76  & & & 3.18 & & & 3.52 & & \\
-7  & 3.45 &                        &                         & 3.18  & & & 3.77 & & & 2.27 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-8  & 3.12 & \multirow{7}{*}{22.57} &                         & 3.36  & \multirow{7}{*}{22.92} & & 3.62 & \multirow{7}{*}{24.25} & & 3.72 & \multirow{7}{*}{24.73} &  \\
-9  & 2.57 &                        &                         & 3.44  & & & 3.82 & & & 3.97 & & \\
-10  & 2.92 &                        &                         & 3.17  & & & 3.06 & & & 3.77 & & \\
-11  & 3.54 &                        &                         & 3.88  & & & 3.11 & & & 3.27 & & \\
-12 & 3.23 &                        &                         & 2.94  & & & 3.40 & & & 2.75 & & \\
-13 & 3.39 &                        &                         & 3.29  & & & 3.55 & & & 3.34 & & \\
-14 & 3.81 &                        &                         & 2.83  & & & 3.69 & & & 3.93 & & \\\hline
-15 & 3.35 & \multirow{7}{*}{23.02} & \multirow{14}{*}{46.31} & 2.72  & \multirow{7}{*}{21.87} & \multirow{14}{*}{ 43.56} & 3.23 & \multirow{7}{*}{22.84} &  \multirow{14}{*}{ 44.01 } & 3.37 & \multirow{7}{*}{22.82} &  \multirow{14}{*}{46.00} \\
-16 & 3.77 &                        &                         & 2.96  & & & 2.94 & & & 3.37 & & \\
-17 & 3.45 &                        &                         & 3.01  & & & 3.02 & & & 2.95 & & \\
-18 & 3.47 &                        &                         & 3.39  & & & 3.63 & & & 3.22 & & \\
-19 & 2.90 &                        &                         & 3.42  & & & 3.16 & & & 3.59 & & \\
-20 & 2.80 &                        &                         & 3.09  & & & 3.25 & & & 3.21 & & \\
-21 & 3.29 &                        &                         & 3.27  & & & 3.61 & & & 3.13 & & \\\cline{3-3}\cline{6-6}\cline{9-9}\cline{12-12}
-
-22 & 2.88 & \multirow{7}{*}{23.29} &                         & 2.92  & \multirow{7}{*}{21.69} & & 3.80 & \multirow{7}{*}{21.17} & & 3.07 & \multirow{7}{*}{23.18} & \\
-23 & 4.01 &                        &                         & 3.27  & & & 3.03 & & & 3.06 & & \\
-24 & 3.13 &                        &                         & 2.92  & & & 2.31 & & & 2.72 & & \\
-25 & 3.57 &                        &                         & 2.83  & & & 2.38 & & & 3.16 & & \\
-26 & 3.27 &                        &                         & 2.97  & & & 3.49 & & & 3.57 & & \\
-27 & 3.27 &                        &                         & 3.41  & & & 2.92 & & & 3.92 & & \\
-28 & 3.17 &                        &                         & 3.36  & & & 3.26 & & & 3.69 & & \\\hline
-29 & 3.68 &  \multirow{3}{*}{8.06} &  \multirow{3}{*}{8.06}  & 2.93  & \multirow{3}{*}{8.36} & \multirow{3}{*}{8.36} & 3.34 & \multirow{3}{*}{8.68} & \multirow{3}{*}{8.68} & 3.15 & \multirow{3}{*}{7.30} & \multirow{3}{*}{7.30} \\
-30 & 2.76 &                        &                         & 3.14  & & & 3.75 & & & 2.71 & & \\
-31 & 1.63 &                        &                         & 2.29  & & & 1.60 & & & 1.45 & & \\\hline
-    \end{tabular}
-\end{center}
-\label{mes}
-\end{table*}
-
-\begin{table*}[t]
-\scriptsize
-    \caption{Activity along the year, in months, trimesters, quadrimesters and semesters.
-Engagement in list participation seem to concentrate in two periods: middle of the year (Jun-Aug, lists MET and LAD), and transition from years (Dec-Mar, lists CPP, LAU and LAD).
-Messages were considered as to complete 12 months slots, so every month has the same time of occurrences.}
-\begin{center}
-    \begin{tabular}{ |l|| c|c|c|c|c|  c|c|c|c|c| c|c|c|c|c| c|c|c|c|c|}
-        \hline
-    & \multicolumn{5}{c|}{CPP} & \multicolumn{5}{c|}{MET} & \multicolumn{5}{c|}{LAU} & \multicolumn{5}{c|}{LAD}  \\ \hline
-    & m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. &   m. & b. & t. & q. & s. \\ \hline\hline
-Jan  & 8.70 & \multirow{2}{*}{17.00} & \multirow{3}{*}{\bf 27.23} &  \multirow{4}{*}{\bf 36.48} & \multirow{6}{*}{\bf 54.26} & 4.88 &  \multirow{2}{*}{11.01} & \multirow{3}{*}{16.90} & \multirow{4}{*}{23.32} & \multirow{6}{*}{47.74} & 10.22 & \multirow{2}{*}{\bf 19.56} & \multirow{3}{*}{\bf 28.23} & \multirow{4}{*}{\bf 35.09} & \multirow{6}{*}{49.17} & 11.23 & \multirow{2}{*}{18.49} & \multirow{3}{*}{26.43} & \multirow{4}{*}{36.04} & \multirow{6}{*}{\bf 57.95} \\
-Fev  & 8.29 &                        &                         & & & 6.13 & & & & & 9.34  & & & & & 7.26 & & & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Mar  & {\bf 10.23} & \multirow{2}{*}{\bf 19.49} &                         & & & 5.89 & \multirow{2}{*}{12.31} & & & & 8.67 & \multirow{2}{*}{15.52} & & & & 7.94 & \multirow{2}{*}{17.55} & & & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Apr  & 9.26 &                        &  \multirow{3}{*}{27.03} & & & 6.42 & & \multirow{3}{*}{30.84} & & & 6.85 & & \multirow{3}{*}{20.94} & & & 9.61 & & \multirow{3}{*}{\bf 31.51} & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Mai  & 9.41 & \multirow{2}{*}{17.78} &                         & \multirow{4}{*}{33.46} & & 10.46 & \multirow{2}{*}{\bf 24.42} & & \multirow{4}{*}{\bf 47.83} & & 7.27 & \multirow{2}{*}{14.09} & & \multirow{4}{*}{30.37} &  & 8.94 & \multirow{2}{*}{\bf 21.91} & & \multirow{4}{*}{\bf 37.56} & \\
-Jun & 8.37 &                        &                         & & & {\bf 13.96} & & & & & 6.81 & & & & & {\bf 12.97} & & & & \\\cline{3-3}\cline{4-4}\cline{6-6}\cline{8-9}\cline{11-11}\cline{13-14}\cline{16-16}\cline{18-19}\cline{21-21}
-Jul & 8.70 &  \multirow{2}{*}{15.68} & \multirow{3}{*}{22.94} & & \multirow{6}{*}{45.73} & 13.23 & \multirow{2}{*}{23.41} & \multirow{3}{*}{\bf 31.16} & & \multirow{6}{*}{\bf 52.26} & 8.96 & \multirow{2}{*}{16.28} & \multirow{3}{*}{24.47} & & \multirow{6}{*}{\bf 50.82} & 9.02 & \multirow{2}{*}{15.65} & \multirow{3}{*}{22.29} & & \multirow{6}{*}{42.05} \\
-Ago  & 6.98 &                       &                         & & & 10.28 & & & & & 7.31 & & & & & 6.63 & & & & \\\cline{3-3}\cline{5-5}\cline{8-8}\cline{10-10}\cline{13-13}\cline{15-15}\cline{18-18}\cline{20-20}
-Set  & 7.26 & \multirow{2}{*}{15.36} &                         & \multirow{4}{*}{30.06} & & 7.75 & \multirow{2}{*}{16.80} & & \multirow{4}{*}{28.86} & & 8.18 & \multirow{2}{*}{16.24} & & \multirow{4}{*}{34.54} & & 6.63 & \multirow{2}{*}{12.38} & & \multirow{4}{*}{26.40} & \\\cline{4-4}\cline{9-9}\cline{14-14}\cline{19-19}
-Oct  & 8.10 &                        &  \multirow{3}{*}{22.80} & & & 9.05 & & \multirow{3}{*}{21.10} & & & 8.06 & & \multirow{3}{*}{26.36} & & & 5.74 & & \multirow{3}{*}{19.77} & & \\\cline{3-3}\cline{8-8}\cline{13-13}\cline{18-18}
-Nov  & 7.86 & \multirow{2}{*}{14.69} &                         & & & 7.46 & \multirow{2}{*}{12.06} & & & & 7.63 & \multirow{2}{*}{18.30} & & & & 7.63 & \multirow{2}{*}{14.02} & & & \\
-Dec & 6.81 &                        &                         &  & & 4.59 & & & & & {\bf 10.66} & & & & & 6.39 & & & & \\\hline
-    \end{tabular}
-\end{center}
-\label{ano}
-\end{table*}
-
-\clearpage
-\section{Figures of vertex classification fractions as the network evolves}\label{figures}
-
-Two lists are exhibited in this section, CPP and LAD. These structures are very similar in all
-four lists and laying extensively all figures is redundant. Window sizes of $ws =$ 10000, 5000,
-1000, 500, 250, 100 and 50 messages were used.
-
-\begin{figure*}[hb]
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification is plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp5000_}
-\end{figure*}
-
-%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-CPP Std library official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:cpp50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/CPP/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:cpp50_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad10000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/10000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad10000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad5000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/5000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad5000_}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad1000}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/1000_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad1000_}
-\end{figure*}
-
-%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad500}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/500_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad500_}
-\end{figure*}
-
-%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad250}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/250_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad250_}
-\end{figure*}
-
-%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad100}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/100_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad100_}
-\end{figure*}
-
-%%%%%%%%%%
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50}
-    \caption{Distribution of vertices with respect to each centrality measure: in and out degrees and strengths.
-Linux Audio Users (LAD) official mailing list.
-In the first six plots, red is fraction of hubs, green is the fraction of intermediary and blue is for peripheral fraction.
-On the last plot, red is the center (maximum distance to another vertex is equal to radius), blue is periphery (maximum distance equals to diameter) of the giant component.
-On the same graph, green counts the disconnected vertices.}
-    \label{fig:lad50}
-\end{figure*}
-
-\begin{figure*}[hbtp] 
-   \centering
-        \includegraphics[width=\textwidth]{figs/LAD/50_2}
-    \caption{Distribution of vertex with respect to compound criteria.
-Red, green and blue designate hubs, intermediary and border (peripheral) vertex fractions.
-The first two plots exhibit classifications that are not functions.
-Thus, in the first plot, the fraction of vertices with unique classification in plotted in black.
-On the second plot, black represents the fraction of vertices that has more than one class: $\frac{\text{number of classifications} - \text{number of nodes}}{\text{number of nodes}}$.
-Compound criteria is described in Section~\ref{sectioning}.}
-    \label{fig:lad50_}
-\end{figure*}
-
-%
-%\begin{figure*}
-%    \centering
-%        \includegraphics[width=\textwidth]{pcm}
-%        \caption{Pulse Code Modulation (PCM) audio: an analogical signal is represented by 25 samples with 4 bits each.}
-%        \label{fig:PCM}
-%\end{figure*}
-%
-
-\nocite{*}
+%\nocite{*}
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2015-10-05 paper.tex
polindo

@@
 such as the fat tail structure and the preponderance of centrality measures,
-but the very small standard deviations of principal components formation (Section~\ref{}),
+but the very small standard deviations of principal components formation
+(see Sections~\ref{sec:pca} and~\ref{prevalence}),
 the presence of the Erd\"os sectors even in networks with
-few participants (Section~\ref{}),
-and the recurrent activity patterns along different timescales (Section~\ref{}),
+few participants (see Sections~\ref{sectioning} and~\ref{subsec:pih}),
+and the recurrent activity patterns along different timescales (see Sections~\ref{sec:mtime} and~\ref{constDisc}),
 goes a step further in characterizing scale-free networks in the context
@@
 with respect to dispersion of the system in the topological space,
-might add to current understanding of how Digraphs differ from
-undirected graphs.
-Noteworthy is the also very stable fraction participants in each Erd\"os sectors
+might add to current understanding of key-differences between digraphs and
+undirected graphs in complex networks.
+Noteworthy is the also very stable fraction participants in each Erd\"os sector
 when the network reaches more than 200 participants (more that 1000 messages).
 Benchmarks were derived from email list networks
-and the analysis of
+and the supplied analysis of
 networks from Facebook,
 Twitter and Participabr might ease hypothesizing
-about the generality of such results.
+about the generality of these characteristics.
 Selected tables and figures sum dozens of pages

### 2015-10-05 paper.tex
arrumada secao dos tipos

@@
 \subsection{Types from Erd\"os sectors}\label{sec:pty}
+Assigning a type to a participant raises important 
+issues about the scientific cannon for human ypes
+and the potential for stigmatization and prejudice. 
 The type of the Erd\"os sector to which a participant belongs 
 can be regarded as implying a type to the corresponding participant.
-Assigning a type to a participant raises important 
-issues about the potential for stigmatization and prejudice. 
-In the present case, the type of a participant
+In this case, the type of a participant
 changes both along time and as
@@
 participant typology is attenuated~\cite{adorno}. 
-In other words, an individual is a hub in a number of networks and peripheral in other networks, and even within the same network he/she probably changes type along time~\cite{4hubs}.
-Moreover, such proposed typology of participants bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.
+In other words, an individual is a hub in a number of networks and peripheral in other networks,
+and even within the same network he/she most probably changes type along time~\cite{animacoes}.
 
-We analyzed the temporal evolution of the networks using visualization tools developed for this research~\cite{rcText,versinus} and inspected the raw data to infer the main characteristics of each type.
-Our main observations may be summarized as follows:
-\begin{itemize}
-	\item Core hubs usually have intermittent activity. Very stable activity was found only in smaller communities, which is consistent with the literature~\cite{barabasiEvo}.
-	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
-The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
-Therefore, intermediary vertices seem responsible for the network structure.
-Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
-	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
-These seem as authorities and contrast with participants that respond much more than receive responses.
-	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
-\end{itemize}
+The importance of this issue can be grasped by
+the consideration of static types derived from quantitative criteria.
+For example, in email lists with a small number of participants,
+the number of threads has a negative correlation
+with the number of participants.
+When the number of participants exceeds a threshold,
+the number of threads has a positive correlation
+with the number of participants.
+This finding is illustrated in Figure~\ref{fig:nmgamma3d}
+and can also be observed in Table~\ref{tab:genLists}.
+The attribution of types to individuals, derived from 
+this analysis of the systems, has more potential for prejudice because
+the derived participant type is static.
+The types, in this latter case, fails to acknowledge that
+human individuals are not immutable entities.
 
-With focus on the networks as the whole objects of analysis,
-we were able to observe a peculiar correlation pattern 
-between the number of threads and the number of participants.
-When the number of participants exceeds a threshold,
-the number of threads displays a positive correlation with the
-number of participants.
-This finding is illustrated in Figure~\ref{fig:nmgamma3d} and can also be observed in Table~\ref{tab:genLists}.
-Obviously, network types can be derived from such results, which was not attempted here but 
-left for future work.
+Beyond the results discussed above and in previous sections,
+our main observations regarding the Erd\"os sectors
+and the implicit participant types,
+consistent with the literature~\cite{barabasiEvo},
+are that
+1) hubs and intermediary participants usually have intermittent activity.
+Stable activity was found only in smaller communities.
+For example, MET list have stable hubs while all LAU, LAD and CPP
+present intermittent hubs.
+2) Network structure seems to be most influenced by the
+activity of intermediary participants as they have less extreme
+roles than hubs and peripheral participants and
+can therefore connect to the sectors and other participants 
+in a more selective and explicit manner.
 
+%Moreover, such typology of participants bridges exact and human sciences and may 
+%be enriched with concepts from other typologies,
+%such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.
+
+%We analyzed the temporal evolution of the networks
+%using visualization
+%tools developed for this research~\cite{rcText,versinus}
+%and inspected raw data.
+
+%dictated (or revealed) by the
+%(e.g. stable or intermittent patterns of activity and preferential communication
+%with hubs or periphery)
+
+%of both hubs and peripheral vertices
+%have the trivial facets of interacting 
+%
+%
+%
+%\begin{itemize}
+%	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
+%The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
+%Therefore, intermediary vertices seem responsible for the network structure.
+%Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
+%%	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
+%%These seem as authorities and contrast with participants that respond much more than receive responses.
+%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
+%\end{itemize}
+
+%Within networks as the whole objects of analysis,
+%we were able to observe a peculiar correlation pattern 
+%between the number of threads and the number of participants.
 \begin{figure}
@@
 was used for semantic web and participatory democracy efforts,
-and might be enhanced to yield scientific knowledge~\cite{pnud3,pnud4,pnud5}.
+and might be enhanced to yield scientific knowledge~\cite{opa}.
 Also, we should pursue the publication of the visualization and sonification
-processes used for this research~\cite{versinus,4hubs} and
+processes used for this research~\cite{versinus,animacoes} and
 of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.
@@
 	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
-	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers, and to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
+	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
 	We are also grateful to developers and users of Python scientific tools.
 \end{acknowledgments}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{document}
-%
-% ****** End of file aipsamp.tex ******

### 2015-10-05 paper.tex
arrumacoes em medidas e pca

@@
 	\item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-	\item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|=k_i}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
-$J_i$ is the set of neighbors of vertex $i$, and $|J_i|=k_i$ is the number of neighbors of vertex $i$.
-	\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{k_i}  }$.
+	\item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
+$J_i$ is the set of neighbors of vertex $i$, and $|J_i|$ is the number of neighbors of vertex $i$.
+	\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$.
 	\item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{k_i}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{dis}_i-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{k_i}}$.
+	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{dis}_i-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$.
 \end{itemize}
@@
 
-\subsection{Stability of principal components and the prevalence of symmetry over clusterization for dispersion}\label{prevalence}
+\subsection{Stability of principal components and the prevalence of symmetry over clusterization}\label{prevalence}
 %The topology was analyzed using standard, well-established metrics of centrality and clustering.
 %We also introduced symmetry metrics given the evidence of their importance in social contexts~\cite{newmanEvolving}.
-The contribution of each metric to the variance is very similar for all the networks and along time.
-In applying PCA to the snapshots, the contribution of each metric to the principal components presents very small standard deviation.
-Table~\ref{tab:pcain} exemplifies the formation of principal components for the MET email list.
-Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
-of the Supporting Information document for the other email lists (benchmarks) and other interaction networks,
-with the consideration of strategic combinations of metrics.
+%The contribution of each metric to the variance is very similar for all the networks and along time.
+
+The principal components of the participants are very stable in the topological space,
+i.e. in PCA of network measures.
+Table~\ref{tab:pcain} exemplifies the formation of principal components
+by exposing averages over non-overlapped activity snapshots of a network.
+The most important result of this application of PCA,
+the stability of principal components,
+is underpinned by the very small standard deviations
+of the contribution of
+each metric to each principal component.
+%The contribution of each metric to the
+%principal components presents
+%very small standard deviation.
 
@@
 
-The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality.
-Therefore, all of these centrality measurements are equally important for characterizing the networks.
-On one hand, the relevance of all centrality metrics is not surprising since they are highly correlated, e.g.the degree and strength have Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
-	On the other hand, each of these metrics is related to a different participation characteristic, and their equal relevance is noticeable.
-The clustering coefficient is presented in almost perfect orthogonality to centrality metrics.
-\begin{figure*} 
-	\centering
-	\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
-	\caption{The first plot shows degree versus clustering coefficient. This typical pattern is well known with high clustering more incident in vertices with lower degrees. 
-		The second plot is analogous but the abscissas is an average of centrality metrics.
-		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
-		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
-		for characterizing interaction networks than the clustering coefficient,
-		especially for hubs and intermediary vertices.
-		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
-		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
-		and in networks from Facebook, Twitter and Participabr,
-		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
-		the symmetry-related metrics and clustering coefficient.}
-	\label{fig:sym}
-\end{figure*}
+The first principal component is an average of centrality metrics:
+degrees, strengths and betweenness centrality.
+On one hand, the similar relevance of all centrality metrics is not surprising since they are highly correlated,
+e.g. degree and strength have Spearman correlation coefficient $\in [0.95,1]$ 
+and Pearson coefficient $\in [0.85,1)$ for window sizes greater than a thousand messages.
+On the other hand, each of these metrics is related to a different participation characteristic,
+and their equal relevance for variability,
+as measured by dispersion, is noticeable.
+Also, this suggests that these centrality metrics 
+are equally adequate for characterizing the networks
+and the participants.
 
-Dispersion was more prevalent in symmetry-related metrics than for the clustering coefficient, as indicated in Table~\ref{tab:pcain}.
-This is also illustrated in Figure~\ref{fig:sym}, where each vertex is colored according to the sector they belong to.
-As expected, peripheral vertices have very low values in the first component (centrality related) and greater dispersion in the third component (clustering related).
-The PCA plot in the third system of Figure~\ref{fig:sym}, where all metrics are considered, reflects the relevance of the symmetry-related metrics for the variance.
-We conclude that the symmetry metrics is more useful, with respect to the dispersion in the topological space,
-in characterizing interaction networks (and their participants) than the clustering coefficient,
+%\begin{figure*} 
+%	\centering
+%	\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
+%	\caption{The first plot shows degree versus clustering coefficient, a well known pattern characterized by the high clustering being more incident in vertices with lower degrees.
+%		In the second plot, the abscissas is an average of centrality metrics, an intermediary plot to ease understanding.
+%		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
+%		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
+%		for characterizing interaction networks than the clustering coefficient,
+%		especially for hubs and intermediary vertices.
+%		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
+%		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
+%		and in networks from Facebook, Twitter and Participabr,
+%		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
+%		the symmetry-related metrics and clustering coefficient.}
+%	\label{fig:sym}
+%\end{figure*}
+
+Evident in Table~\ref{tab:pcain},
+dispersion is greater in symmetry-related
+metrics than in clustering coefficient.
+As expected by basic complex network theory, 
+peripheral vertices have low values
+of centrality metrics
+and greater dispersion with respect to clustering coefficient.
+%The scatter plot in the third system of Figure~\ref{fig:sym},
+%where all metrics are considered and there is a greater dispersion
+%with respect to the ordinates,
+This reflects in the relevance of the symmetry-related metrics.
+We conclude that the symmetry metrics are more powerful,
+with respect to dispersion in the topological space,
+in characterizing interaction networks
+and their participants, than the clustering coefficient,
 especially for hubs and intermediary vertices.
 
-The relative importance of the topological metrics was also observed for the additional 12 networks from Facebook, Twitter and Participabr.
-With the exception of two of these networks, the overall behavior was maintained in that centrality measurements 
-were found prevalent in the first principal component,
-followed by symmetry-related metrics on the second principal component and then clustering coefficient on the third principal component.
-The results are given in Tables~\ref{tab:pcaE1F},~\ref{tab:pcaE1I},~\ref{tab:pcaE2},~\ref{tab:pcaE3} of the Supporting Information document.
-Larger variability was found among such networks, which motivated the
+%These results are also reported for 12 networks from Facebook, Twitter and Participabr
+%in Section~\ref{si:ext} of the Supporting Information document.
+Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
+of the Supporting Information document for other email lists and other interaction networks.
+Larger variability was found among the latter networks,
+which motivated the
 use of interaction networks derived from email lists for benchmarks.
+
+%the overall behavior was maintained in that centrality measurements 
+%were found prevalent in the first principal component,
+%followed by symmetry-related metrics on the second principal
+%component and then clustering coefficient on the third principal component.
+%Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
+%of the Supporting Information document for other email lists and other interaction networks,
+%with the consideration of strategic combinations of metrics.
 
@@
    \centering
-        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
-	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists.
-Highest numbers of threads are found in lists with few participants.
+   \includegraphics[trim={0 0 0 1cm},clip,width=\columnwidth]{figs/mpgamma____________}
+	\caption{A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.
+Highest $\Gamma$ are associated with low $N$.
 The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise.
 This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}.
-For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants.
-All LAU, LAD, MET lists present smaller networks.}
+Accordingly,
+for $M=20000$ messages, this inflection
+of correlation 
+was found to be around $N=1500$
+and LAU, LAD, MET lists 
+present smaller networks.}
 	\label{fig:nmgamma3d}

### 2015-10-05 paper.tex
arrumando medidas e resultados de pca

@@
 	\item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{s_i}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
-	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{dis}_i-\frac{(w_{ji}-w_{ij})}{s_i}]^2}{|J_i|}}$.
+	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{w_{ji}-w_{ij}}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}\left[\mu^{dis}_i-\frac{w_{ji}-w_{ij}}{w_{ji}+w_{ij}}\right]^2}{|J_i|}}$.
 \end{itemize}

### 2015-10-05 paper.tex
arrumadas metricas

@@
 
-The topology of the networks was studied using a small selection of the most basic and fundamental measurements for each vertex, as follows:
+The topology of the networks was characterized 
+using a small selection of the most basic and 
+fundamental measurements for each vertex, as follows:
 
@@
 	\item Out-strength $s_i^{out}$: sum of weights of all edges departing from vertex $i$.
-	\item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked.  The standard clustering coefficient for undirected graphs was used.
-	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index considered directions and weight, as specified in~\cite{faster}.
+	\item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked, i.e. the standard clustering coefficient metric for undirected graphs.
+	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index was calculated for weighted digraphs as specified in~\cite{faster}.
 \end{itemize}
 
-In order to capture symmetries in the activity of participants, the following metrics were introduced for a vertex $i$:
+The non-standard metrics bellow were elaborated
+in order to capture symmetries in the activity of participants:
 
 \begin{itemize}
-	\item Asymmetry: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-	\item Mean of asymmetry of edges: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
+	\item Asymmetry of vertex $i$: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
+	\item Mean of asymmetry of edges of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
 $J_i$ is the set of neighbors of vertex $i$, and $|J_i|$ is the number of neighbors of vertex $i$.
@@
 
-These measures are used both for the Erd\"os sectioning (described in Secion~\ref{sectioning}) and for performing PCA (as described in Section~\ref{sec:pca}).
+Both standard and non-standard metrics are used for the Erd\"os sectioning (described in Secion~\ref{sectioning}) and for performing PCA (as described in Section~\ref{sec:pca}).
 

### 2015-10-05 paper.tex
incorporada nova figura de symetria e clustering coef

@@
 
-%\begin{figure*} 
-%	\centering
-%	\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
-%	\caption{The first plot shows degree versus clustering coefficient, a well known pattern characterized by the high clustering being more incident in vertices with lower degrees.
-%		In the second plot, the abscissas is an average of centrality metrics, an intermediary plot to ease understanding.
-%		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
-%		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
-%		for characterizing interaction networks than the clustering coefficient,
-%		especially for hubs and intermediary vertices.
-%		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
-%		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
-%		and in networks from Facebook, Twitter and Participabr,
-%		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
-%		the symmetry-related metrics and clustering coefficient.}
-%	\label{fig:sym}
-%\end{figure*}
+\begin{figure} 
+	\centering
+	%\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT__}
+	\includegraphics[width=.45\textwidth]{figs/im13PCAPLOT__}
+	\caption{The first plot shows degree versus clustering coefficient, a well known pattern characterized by the high clustering being more incident in vertices with lower degrees.
+		In the second plot, the abscissas is an average of centrality metrics, an intermediary plot to ease understanding.
+		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
+		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
+		for characterizing interaction networks than the clustering coefficient,
+		especially for hubs and intermediary vertices.
+		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
+		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
+		and in networks from Facebook, Twitter and Participabr,
+		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
+		the symmetry-related metrics and clustering coefficient.}
+	\label{fig:sym}
+\end{figure}
 
@@
    \centering
-   \includegraphics[trim={0 0 0 1cm},clip,width=\columnwidth]{figs/mpgamma____________}
+   \includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma____________}
 	\caption{A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.

### 2015-10-06 scripts/setoresTeoricoEmpirico2.py
terminada revisao das figuras p pb

@@ -58,7 +58,7 @@ p.subplots_adjust(left=0.05,bottom=0.28,right=0.99,top=0.87)
 
 p.bar([i[0]-0.4 for i in np.bins],
         np.empirical_distribution,
-        alpha=.4, label="empirical distribution")
+        alpha=.4,color="k", label="empirical distribution")
 #p.bar(n.log([i[0]-0.5 for i in np.bins]),
 #      n.log(  np.empirical_distribution),
 #        alpha=.7, label="empirical distribution")
@@ -73,7 +73,7 @@ for i, bin_ in enumerate(np.bins):
     binomial_probs.append(binomial_prob)
 #p.plot(centers,np.binomial.cdf(centers))
 #p.plot(centers,binomial_probs,"ro")
-p.plot(centers,binomial_probs,"r-o",label="binomial distribution")
+p.plot(centers,binomial_probs,"k-.^",label="binomial distribution",linewidth=8,alpha=.8)
 #p.plot(n.log(centers),n.log(binomial_probs),"ro",label="binomial distribution")
 #p.title("Three sections of a scale-free network"+"\nCPP list, {} messages, {} participants, {} edges".format(
 #         ne.window_size, nm.N, nm.E),size=25)
@@ -87,8 +87,8 @@ p.ylabel(r"P(k) $\rightarrow$",size=20)
 c_=n.array(centers)
 pk=c_[1:]**-1.5
 #p.plot(centers[1:],.4*pk,"ko")
-p.plot(centers[1:],.4*pk,"k-o",label="power-law distribution")
-p.ylim(0,0.4*pk[0]+0.03)
+p.plot(centers[1:],.4*pk,"k-o",label="power-law distribution",linewidth=5,alpha=.8)
+p.ylim(0,0.4*pk[0]-0.03)
 
 # pontos de intersecção:
 x1=2.5; y=41.16-5
@@ -98,7 +98,7 @@ p.plot((x2,x2),(-1000,y),"g--")
 #p.xticks((x1,x2),(r"$(k_L)$",r"$(k_R)$"),size=25)
 
 p.text(-.28,0.12,"periphery"  ,size=20)
-p.text(4,.3,"intermediary",   size=20)
+p.text(4,.27,"intermediary",   size=20)
 p.text(10,.08,"hubs"      ,   size=20)
 
 

### 2015-10-06 paper.tex
revisada figura das tls
terminada revisao das figuras p pb
terminada revisao das figuras

@@
 	\centering
-	\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
-	\caption{Fractions of agents in each Erd\"os sector, where the fractions for hubs, intermediary and peripheral vertices are represented in red, green and blue, respectively.
-Fractions derived from degree and strength criteria are both on the left.
-	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right,
-with black lines representing the fraction of vertices without class and with more than one class, respectively. 
-See Section~\ref{si:frac} of Supporting Information document for a collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}.
-Table~\ref{tab:secE}, also from Supporting Information document, presents these fractions of agents in networks from Facebook, Twitter and Participabr.}
+	\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000_}
+	\caption{Stability of Erd\"os sector sizes.
+		Fractions of participants derived from degree and strength criteria, $E_1$ and $E_4$ in Section~\ref{sectioning}, are both on the left.
+	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right.
+	In the ordinates $\overline{e_{n,x}}=\frac{|e_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $E_n$
+	and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
+Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning} and present these fractions of participants in networks from Facebook, Twitter and Participabr.}
 	\label{fig:sectIL}
@@
 	\includegraphics[width=.45\textwidth]{figs/im13PCAPLOT__}
-	\caption{The first plot shows degree versus clustering coefficient, a well known pattern characterized by the high clustering being more incident in vertices with lower degrees.
-		In the second plot, the abscissas is an average of centrality metrics, an intermediary plot to ease understanding.
-		The third plot exhibits the greater dispersion of the symmetry-related ordinates.
+	\caption{The first plot exposes the wel known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
+		The second plot exhibits the greater dispersion of the symmetry-related ordinates.
 		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
@@
 		especially for hubs and intermediary vertices.
-		This figure was obtained with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
-		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
-		and in networks from Facebook, Twitter and Participabr,
-		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
-		the symmetry-related metrics and clustering coefficient.}
+	This figure was reflects a snapshot of the LAU list with $ws= 1000$.}
+
+%		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$, in networks derived from email lists,
+%		and in networks from Facebook, Twitter and Participabr,
+%		which suggests a common relationship between the metrics of degrees, strengths and betweenness centrality,
+%		the symmetry-related metrics and clustering coefficient.}
 	\label{fig:sym}
@@
    \centering
-   \includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma____________}
+   \includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma2}
 	\caption{A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.
@@
 of correlation 
-was found to be around $N=1500$
+was found around $N=1500$
 and LAU, LAD, MET lists 

### 2015-10-06 paper.tex
terminando de revisar resultados dos setores de erdos

@@
 
-\subsection{Stable fractions of participants in each Erd\"os sector}\label{subsec:pih}
+\subsection{Stable sizes of Erd\"os sectors}\label{subsec:pih}
 
@@
 These results hold for the total, in and out degrees and strengths.
-Stable distributions can also be obtained for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
-The networks hold their basic structure with as few as 10-50 messages,
+Stable sizes are also observed for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
+The networks often hold their basic structure with as few as 10-50 messages,
 i.e. concentration of activity and the abundance of low-activity participants take place even with very few messages,
@@
 In order to support hypotheses about the generality of these findings,
-we obtained the Erd\"os sectors of 12 networks from Facebook, Twitter and Participabr.
-The results are given in Table~\ref{tab:secE} of the Supporting Information,
-which indicate that the percentages of hubs, intermediary and periphery nodes are essentially the same as for the email lists.
+we expose the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr
+in Table~\ref{tab:secE} of the Supporting Information document.
+The fractions of hubs, intermediary and periphery nodes are essentially the same as for the email lists.
 
@@
 
-Evident in Table~\ref{tab:pcain},
+Evident in Table~\ref{tab:pcain} and Figure~\ref{fig:sym},
 dispersion is greater in symmetry-related
@@
 peripheral vertices have low values
-of centrality metrics
+of centrality measures
 and greater dispersion with respect to clustering coefficient.

### 2015-10-06 paper.tex
terminada revisao dos resultados nos setores de erdos

@@
 
-The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered. 
-Moreover, the same distribution applies to the networks of all email lists analyzed,
-as demonstrated in Figure~\ref{fig:sectIL} and in Section~\ref{si:frac} of the Supporting Information document. 
-Activity is highly concentrated on the hubs, while a very large number of peripheral vertices contribute to only a fraction of the activity.
-This is expected for a system with a scale-free profile, as confirmed by the data in Table~\ref{autores} for the distribution of activity among participants.
-
-Typically, $\approx [3-12]\%$ of the vertices are hubs, $\approx [15-45]\%$ are intermediary and $\approx [44-81]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
+The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors 
+is remarkably stable along time is samples of 200 or more messages,
+as evident in Figure~\ref{fig:sectIL} of current document
+and in Section~\ref{si:frac} of the Supporting Information document. 
+%Moreover, all email lists analyzed exhibit the same distribution profile.
+Activity is highly concentrated on the hubs,
+while a very large number of peripheral vertices contribute to only a fraction of the activity.
+This is expected for a system with a scale-free profile
+and confirmed by Table~\ref{autores} in the distribution of activity among participants.
+
+Typically, $\approx [3-12]\%$ of the vertices are hubs,
+$\approx [15-45]\%$ are intermediary and $\approx [44-81]\%$ are peripheral,
+which is consistent with the literature~\cite{secFree}.
 These results hold for the total, in and out degrees and strengths.
-Stable sizes are also observed for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
-The networks often hold their basic structure with as few as 10-50 messages,
+Stable sizes are also observed for 100 or less messages if classification 
+of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
+The networks often hold this basic structure with as few as 10-50 messages,
 i.e. concentration of activity and the abundance of low-activity participants take place even with very few messages,
@@
 in Table~\ref{tab:secE} of the Supporting Information document.
-The fractions of hubs, intermediary and periphery nodes are essentially the same as for the email lists.
+The fractions of hubs, intermediary and periphery nodes are
+essentially the same as for the email list networks but with exceptions and a greater variability.
 
@@
 The first column presents the percentage of messages sent by the most active participant.
-The column for the first quartile ($1Q$) shows the minimum percentage of participants responsible for at least 25\% of total messages.
-Similarly, the column for the first three quartiles $1-3Q$ gives the minimum percentage of participants responsible for 75\% of total messages.
+The column for the first quartile ($1Q$) exposes 
+the minimum percentage of participants
+responsible for at least 25\% of total messages with the actual number in parenthesis.
+Similarly, the column for the first three quartiles $3Q$
+gives the minimum percentage of participants responsible for 75\% of total messages.
 The last decile $-10D$ column exposes the maximum percentage of participants responsible for 10\% of messages.}
@@
 			\hline
-			list & hub & $ 1Q $ & $ 1-3Q $ & $-10D$ \\ \hline
+			list & hub & $ 1Q $ & $ 3Q $ & $-10D$ \\ \hline
 			\input{tables/userTab}

### 2015-10-06 paper.tex
arrumada tabela com os quantis

@@
 The first column presents the percentage of messages sent by the most active participant.
-The column for the first quartile ($1Q$) exposes 
+The column for the first quartile ($Q_1$) exposes 
 the minimum percentage of participants
 responsible for at least 25\% of total messages with the actual number in parenthesis.
-Similarly, the column for the first three quartiles $3Q$
+Similarly, the column for the first three quartiles $Q_3$
 gives the minimum percentage of participants responsible for 75\% of total messages.
-The last decile $-10D$ column exposes the maximum percentage of participants responsible for 10\% of messages.}
+The last decile $D_{-1}$ column exposes the maximum percentage of participants responsible for 10\% of messages.}
 	\begin{center}
@@
 			\hline
-			list & hub & $ 1Q $ & $ 3Q $ & $-10D$ \\ \hline
+			list & hub & $ Q_1 $ & $ Q_3 $ & $D_{-1}$ \\ \hline
 			\input{tables/userTab}

### 2015-10-06 paper.tex
terminada revisao dos resultados

@@
 \subsection{Time activity statistics}\label{sec:mtime}
-Messages were counted over time as histograms in terms of seconds,
+Messages were counted over time as histograms in the scales of seconds,
 minutes, hours, days of the week, days of the month, and months of the year.
@@
 \subsection{Activity along time}\label{constDisc}
-The observed activity along time, in terms of seconds, minutes, hours, days and months,  is practically the same for all lists.
-Histograms in each time scale were computed as were circular average and dispersion values.
-We provide detailed values in Table~\ref{tab:circ}-\ref{tab:min2} because they can
-be used for characterizing nodes (participants) in these and other networks,
-and networks themselves.
-For example, they may serve for identification of outliers in a community.
+Regular patterns of activity were observed along time
+in the scales of seconds, minutes, hours, days and months.
+%exhibited the same patterns in different networks.
+Histograms in each of the time scales were computed as were circular average and dispersion values.
+Values and dedicated expositions are provided in Tables~\ref{tab:circ}-\ref{tab:min2}.
+%they can be used for characterizing nodes (participants) in these and other networks,
+%and networks themselves.
+%For example, they may serve for identification of outliers in a community.
 
 \begin{table}
-	\caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$ described in Section~\ref{sec:mtime}.
-		This typical table was constructed using all LAD list messages, and the results are the same for other lists, as shown in Section~\ref{si:circ} of the Supporting Information document.
-	The most uniform distribution of activity was found in seconds and minutes, where the mean has little meaning.
+	\caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$, described in Section~\ref{sec:mtime}, for different timescales.
+		This example table was constructed using all LAD messages, and the results are the same for other lists, as shown in Section~\ref{si:circ} of the Supporting Information document.
+	The most uniform distribution of activity was found in seconds and minutes.
 	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 14h and 15h ($\theta'=-9.61$).
@@
 \begin{table}
-	\caption{Activity percentages along the hours of the day for the CPP list.
-Nearly identical distributions are found on other lists as shown in Section~\ref{si:hours} of the Supporting Information document.
-Higher activity was observed between noon and 6pm, followed by the time period between 6pm and midnight.
-Around 2/3 of the whole activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs around midday, with a slight skew toward one hour before noon.}
+	\caption{Activity percentages along the hours of the day. Nearly identical distributions are found on other systems as shown in Section~\ref{si:hours} of the Supporting Information document.
+	Higher activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight. Around 2/3 of the activity takes place from noon to midnight.
+Nevertheless, the activity peak occurs between 11 a.m. and 12 p.m. This example and empirical table corresponds to the activity in CPP.}
 	\footnotesize
@@
 \begin{table}
-	\caption{Activity percentages along the days of the week for the four email lists.
+	\caption{Activity percentages along the days of the week.
 	Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
@@
 \end{table}
-In the scale of seconds and minutes, activity obeys a homogeneous pattern, with the messages being slightly more evenly distributed in all lists than in simulations using uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
+For example, uniform activity is found with respect to seconds, minutes and days of the months,
+weekend days exhibit about half the activity of regular weekdays,
+and there is a peak of activity between 11am and 12pm.
+
+In the scales of seconds and minutes, activity in uniform,
+with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
 In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks $\xi=\frac{min(incidence')}{max(incidence')} \Rightarrow \mu_\xi=0.7741 \text{ and } \sigma_\xi=0.02619$.
@@
 In these cases, the circular dispersion is maximized and the mean has little meaning as indicated in Table~\ref{tab:circ}.
-As for the hours of the day, an abrupt peak appeared around 11am with the most active period being the afternoon.
+As for the hours of the day, an abrupt peak is found between 11am and 12pm with the most active period being the afternoon, with one third of total daily activity, and two thirds of activity is allocated in the second 12h of each day.
 Days of the week revealed a decrease between one third and two thirds of activity on weekends.
-Days of the month were regarded as homogeneous with an inconclusive slight tendency of the first week being more active.
+Days of the month were regarded as homogeneous with an inconclusive slight tendency of the first week to be more active.
 Months of the year revealed patterns matching usual work and academic calendars.
-The time period examined here was not sufficient for the analysis of activity along the years.
-These patterns are exemplified in Tables~\ref{tab:hin}-\ref{tab:min2}.
+The time period examined here was not sufficient for the analysis of activity along the years. These patterns are exemplified in Tables~\ref{tab:hin}-\ref{tab:min2}.
 
@@
 \begin{table}
-	\caption{Activity in the days along the month for the MET list.
-Nearly identical distributions are found on other lists as indicated in Section~\ref{si:monthdays} of the Supporting Information.
-Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.}
+	\caption{Activity along the days of the month cycle.
+	Nearly identical distributions are found in all systems
+as indicated in Section~\ref{si:monthdays} of the Supporting Information.
+Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.
+This example and empirical table corresponds to the activity of the MET email list.}
 	\footnotesize
@@
 \begin{table}
-	\caption{Activity percentages of the months along the year from LAU list.
-Activity is usually concentrated in Jun-Aug and/or in Dec-Mar (see Section~\ref{si:months} of the Supporting Information).
-These observations fit academic calendars, vacations and end-of-year holidays.}
+	\caption{Activity percentages on months along the year.
+		Activity is usually concentrated in Jun-Aug and/or in Dec-Mar,
+		potentially due to academic calendars, vacations and end-of-year holidays.
+	 This table corresponds to activity in LAU. 
+		Further evidence is in Section~\ref{si:months} of the Supporting Information document.
+	}
 	\footnotesize
@@
 especially for hubs and intermediary vertices.
+Interestingly, the clustering coefficient is always combined
+with the standard deviation of the asymmetry and disequilibrium
+of edges $\sigma^{asy}$ and $\sigma^{dis}$.
 

### 2015-10-07 paper.tex
revisada evolucao e audiovisualizacao dos dados

@@
 negative values with a high absolute value indicate negative correlation.
-\subsection{Evolution and visualization of the networks}\label{sec:viz}
-The evolution of the networks was observed within sequences of snapshots. with a fixed number of messages,
-Each of these sequences had a fixed number of messages, the window size $ws$,
-which was considered with different shifts in the message timeline to obtain snapshots.
-In each snapshot was performed both the PCA with topological metrics and the Erd\"os sectioning.  
-The values of $ws$ employed were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000.
+
+\subsection{Evolution and audiovisualization of the networks}\label{sec:viz}
+The evolution of the networks was observed within 
+sequences of snapshots.
+In each sequence,a fixed number of messages,
+the window size $ws$, was used for all snapshots.
+%was considered with different shifts in the message timeline to obtain snapshots.
+The snapshots are disjoint in the message timeline,
+and were used to perform both the PCA
+with topological metrics and the Erd\"os sectioning.  
+Figures and tables were usually inspected in
+$ws=\{50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000, 10000\}$ messages.
 Variations in the number of vertices, edges
-and other network characteristics
+and other network characteristics, within the same window size $ws$,
 are exposed in Section~\ref{si:frac}
@@
 
-Networks were visualized with animations, sonifications,
-image galleries and online gadgets developed specifically for this research~\cite{animacoes,galGMANE,appGMANE}.
-Such \emph{audiovisualizations} were crucial to guide the research into the most important features of network evolution.
-Furthermore, the size of the three Erd\"os sectors could be visualized in a timeline fashion.
-Visualization of network structure was especially useful in the initial inspection of data and of the structures derived from the email lists.
+Network structures were mapped to video animations, sound and musical structures,
+image galleries and online gadgets developed for this research~\cite{animacoes,galGMANE,appGMANE}.
+Such \emph{audiovisualizations} were crucial in the initial steps and
+to guide the research into the most important features of network evolution.
+%Furthermore, the size of the three Erd\"os sectors could be visualized in a timeline fashion.
+%Visualization of network structure was especially useful in the initial inspection of data and of the structures derived from the email lists.
 

### 2015-10-08 paper.tex
revisado corpo do artigo

@@
 The activity of participants followed the expected scale-free trace,
-thus yielding three classes of vertices by comparison against the Erd\"os-R\'enyi model,
-namely hubs, intermediary and periphery.
-The relative size of these three sectors did not vary with time and was essentially the same for all email lists.
-Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices.
-Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
-Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on quantitative criteria.
+thus yielding the hubs, intermediary and periphery classes of vertices by comparison against the Erd\"os-R\'enyi model.
+The relative sizes of these three sectors were essentially the same for all email lists and the same along time.
+Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and 44-81\% are peripheral vertices.
+Similar results for the distribution of participants in the three categories
+and for the relative importance of the topological metrics were obtained for 12 additional networks
+from Facebook, Twitter and Participabr.
+Consistent with expectations derived from the literature,
+these properties may be general for human interaction networks,
+which has important implications in establishing a typology of participants based on quantitative criteria.
 \end{abstract}
@@
 \section{Introduction}\label{sec:into}
-Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
+First studies explicitly about human interaction networks 
+date from the nineteenth century while the foundation of
 social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{moreno,newmanBook}.
@@
 the most basic topological measures always combine into very characteristic principal components,
-and the fractions of participants in each of the hubs, intermediady and periphery sectors
+and the fractions of participants in each of the hubs, intermediady and periphery Erd\"os sectors
 are unshaken.
@@
 of types of networks and participants.
-Noteworthy, typologies are the canon of scientific literature for classification of human agents,
+Noteworthy, typologies are the canon of scientific literature for the classification of human agents,
 with pragmatic standards~\cite{myers} and critical paradigms~\cite{adorno,typCanon}.
@@
 Section~\ref{sec:related} describes related work,
-while details of the data and methods of analysis are given in Section~\ref{sec:data} and Section~\ref{sec:carac}.
-Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions.
+while data, scripts and methods of analysis are given in Section~\ref{sec:data} and Section~\ref{sec:carac}.
+Section~\ref{sec:results} reports results and discussion, leading to Section~\ref{sec:conc} for conclusions.
 Subsidiary data analysis,
@@
 and numeric results for networks from Twitter, Facebook and Participabr,
-are given in the Supporting Information.
+are given in the Supporting Information document.
 
@@
 No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were not considered in our study.
-Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}, but this was not considered in the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex} but is not relevant for the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.
 
@@
 We analyzed many email lists (and data from Twitter, Facebook and Participabr)
-and selected four of them in order to make a thorough analysis,
+and selected four of them in order to expose a thorough analysis,
 from which general properties can be inferred. These lists are:
@@
 the number of discussion threads.
-Furthermore, 12 additional networks from Facebook (8), Twitter (2) and Participabr (2)
+Furthermore, 12 networks from Facebook (8), Twitter (2) and Participabr (2)
 are scrutinized in the Supporting Information document for better hypothesizing
@@
 reported in~\cite{opa} and available in~\cite{datahub}.
-Script are delivered through a public domain Python PyPI package and
+Computer scripts are delivered through a public domain Python PyPI package and
 an open Git repository~\cite{gmanePack}.
@@
 reinforces the scientific aspect of the contribution~\cite{openSci}
-and lightens ethical and moral issues of researching 
+and mitigates ethical and moral issues of researching 
 systems constituted of
@@
 Equivalent measures were taken using circular statistics,
-in which each measurement $t$ (data point) is represented as a unity complex number,
+in which each measurement $t$ is represented as a unity complex number,
 $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=t\frac{2\pi}{T}$,
@@
 which can be noticed in the Section~\ref{si:circ} of the Supporting Information document.
-The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred in the discussion of results.
+The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred 
+in the discussion of results.
 
@@
 both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
-Networks in this paper are directed and weighted, the most informative of trivial possibilities.
+Networks in this paper are directed and weighted, the most informative of the possibilities.
 We did not investigate directed unweighted, undirected weighted,
@@
 This paper considers by convention the information network as described above ($A\rightarrow B$) and depicted in Figure~\ref{formationNetwork}.
-These human social information interaction networks are reported in the literature
+These interaction networks are reported in the literature
 as exhibiting scale-free and small-world properties,
@@
 	\item Out-degree $k_i^{out}$: number of edges departing from vertex $i$.
-	\item Strength $s$: sum of weights of all edges linked to vertex $i$.
+	\item Strength $s_i$: sum of weights of all edges linked to vertex $i$.
 	\item In-strength $s_i^{in}$: sum of weights of all edges ending at vertex $i$.
@@
 In a scale-free network, the peripheral, intermediary and hubs sectors can be derived from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
-We shall refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors being referred to as \emph{Erd\"os sectors} or \emph{primitive sectors}.
-
-The degree distribution $\widetilde{P}(k)$ of an ideal
+We refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors regarded as \emph{Erd\"os sectors}.
+
+The degree distribution $\widetilde{P}(k)$ of a
 scale-free network $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
@@
 
-If $\mathcal{N}_f(N,z)$ is directed and has no self-loops, the probability
+If $\mathcal{N}_f(N,z)$ is directed and has no self-loops, the probability of the existence
 of an edge between two arbitrary vertices is $p_e=\frac{z}{N(N-1)}$.
@@
 
-The lower degree fat tail corresponds to the border vertices, i.e.
-the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
-The higher degree fat tail is the hub sector, i.e.
-$\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
-The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected at pure chance (e.g.
-without preferential attachment and as in the Erd\"os R\'enyi model) are correctly associated to the hubs sector.
-Vertices with very few connections, which are way more abundant than expected by pure chance, are assigned to the periphery.
-Vertices with degree values predicted as the most abundant if connections are created by pure chance, near the average, and less frequent in the real network, are classified as intermediary.
+The lower degree fat tail corresponds to the border vertices, i.e. the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
+The higher degree fat tail is the hub sector, i.e. $\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
+The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected without distinction of the vertices, i.e.
+without preferential attachment and as in the Erd\"os-R\'enyi model, are correctly associated to the hubs sector.
+Vertices with very few connections, which are way more abundant than expected in the Erd\"os-R\'enyi model, are assigned to the periphery.
+Vertices with degree values predicted as the most abundant in a Erd\"os-R\'enyi model,
+near the average, and less frequent in the real network, are classified as intermediary.
 
@@
 		distributions.
-		The binomial distribution of the Erd\"os R\'enyi network exhibit more intermediary vertices, while a scale-free network, associated with the power-law distribution,
+		The binomial distribution of the Erd\"os-R\'enyi network model exhibit more intermediary vertices,
+		while a scale-free network, associated with the power-law distribution,
 		has more peripheral and hub vertices.
@@
 The range $\Delta$ of incident values should be partitioned in $m$ parts $\Delta=\cup_{i=1}^m \Delta_i$,
-with $\Delta_i\cap \Delta_j=\emptyset \; \forall\; i \neq j$:
+with $\Delta_i\cap \Delta_j=\emptyset \; \forall\; i \neq j$ and:
 %\begin{equation}
@@
 
-If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used for comparison, with $\overline{w}=2\frac{z}{\sum_is_i}$ the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees ($k^{in}$, $k^{out}$) comparison of the real network should be made with
+If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ the strength of vertex $i$.
+For in and out degrees ($k^{in}$, $k^{out}$), comparison of the real network should be made against
 \begin{equation}
@@
 
-Since different metrics can be used to identify the three types of vertices, compound criteria were defined for conveniently analysing 
-networks with a low number of messages,
-such as in Section~\ref{si:frac} of the Supporting Information.
-For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
+Since different metrics can be used to identify the three types of vertices,
+more than one metric can be used simultaneously, which is convenient when analysing 
+small networks,
+such as the cases where $ws=50$ in Section~\ref{si:frac} of the Supporting Information.
+%For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
 After a careful consideration of possible combinations, these were reduced to six:
@@
 	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics.
-In this case, vertices classified do not usually reach 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
+		In this case, vertices classified do not usually reach $N$ (or 100\%),
+		which is indicated by a black line in Figure~\ref{fig:sectIL}.
 	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics.
-Therefore, a vertex may belong to more than one class, and the total number of members may exceed 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
+		Therefore, a vertex may belong to more than one class, and the total number of memberhips may exceed $N$ (or 100\%),
+		which is indicated by a black line in Figure~\ref{fig:sectIL}.
 	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics.
@@
 	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as such according to any of the metrics.
-The remaining vertices are classified as intermediary if they belong to this category for any of the metrics.
-Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics.
-	\item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics.
-The remaining vertices are classified as peripheral if they fall into the periphery or hub classes for any metric.
-The rest of the nodes are classified as intermediary.
+		The remaining vertices are intermediary if they belong to this category for any of the metrics.
+		Peripheral vertices are those which are classified as such with respect to all metrics.
+	\item Exclusivist externals $C_5$: vertices are hubs if they are classified as such according to all the metrics.
+		The remaining vertices are peripheral if they are peripheral or hubs for all metrics.
+		The rest of the nodes are intermediary.
 	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric.
-The remaining vertices will be peripheral if they are classified as such according to any metric.
-The rest of the vertices will be intermediary vertices.
+The remaining vertices are peripheral if they are classified as such according to any metric.
+The rest of the vertices are intermediary.
 \end{itemize}
 
-Using equations~\ref{eq:part}, these compound criteria $C_\delta$, with $\delta$ integer in the interval $1\leq\delta\leq6$, can be described as:
+Using equations~\ref{eq:part}, these \emph{compound criteria} $C_\delta$, with $\delta$ integer in the interval $1\leq\delta\leq6$, can be specified as:
 
@@
 Be $\mathbf{D}=\{D[k]\}$ the eigenvalues associated to the eigenvectors $\mathbf{V}$,
-then $D'[k]=100*\frac{D[k]}{\sum_{k'}D[k']}$
+then $D'[k]=100\frac{D[k]}{\sum_{k'}D[k']}$
 is the percentage of total dispersion of the system that the principal component $k$
@@
 These usually sum between 60 and 95\% of the dispersion
-and reveal patterns for a sound analysis.
+and reveal patterns for a first analysis.
 In particular, 

### 2015-10-08 paper.tex
adicionado boldface nas tabelas
retoque de escrita

@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, social network analysis, pattern recognition, statistics, anthropological physics, social psychology of big data}
+\keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data}
 \maketitle
@@
 was found around $N=1500$
-and LAU, LAD, MET lists 
+and CPP, LAU, LAD, MET lists 
 present smaller networks.}

### 2015-10-08 paper.tex
retoques de escrita

@@
 processes used for this research~\cite{versinus,animacoes} and
-of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.
+of the extreme linguistic differences found in each of the Erd\"os sectors~\cite{rcText}.
 

### 2015-10-08 paper.tex
retoques nas conclusoes

@@
 of the interaction of human individuals.
-Maybe secondary, but nonetheless important, the importance of symmetry-related metrics,
+Furthermore, the importance of symmetry-related metrics,
 which surpassed that of clustering coefficient,
@@
 Noteworthy is the also very stable fraction participants in each Erd\"os sector
-when the network reaches more than 200 participants (more that 1000 messages).
+when the network reaches more than 200 participants.
 Benchmarks were derived from email list networks
@@
 was used for semantic web and participatory democracy efforts,
-and might be enhanced to yield scientific knowledge~\cite{opa}.
+and these developments might be enhanced to yield scientific knowledge~\cite{opa}.
 Also, we should pursue the publication of the visualization and sonification

### 2015-10-15 paper.tex
arrumadas medidas circulares

@@
 In the scales of seconds and minutes, activity in uniform,
-with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.6.1, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
+with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.8.2, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
 In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks $\xi=\frac{min(incidence')}{max(incidence')} \Rightarrow \mu_\xi=0.7741 \text{ and } \sigma_\xi=0.02619$.

### 2015-10-15 paper.tex
pequena homogeneizacao das tabelas
arrumada notacao das medidas circulares

@@
 \noindent
-Also, the ratio $r=\frac{b_l}{b_h}$ between the lowest and the highest incidences on the histograms 
+Also, the ratio $r=\frac{b_l}{b_h}$ between the lowest $b_l$ and the highest $b_h$ incidences on the histograms 
 served as a further clue of how close the distribution was to being uniform.
@@
 			%& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
-			{\bf scale} & {\bf mean} \boldsymbol{$\theta_\mu'$} & {\bf dispersion} \boldsymbol{$\delta(z)$ }  \\ \hline\hline
+			scale & mean $\theta_\mu'$ & dispersion $\delta(z)$  \\ \hline
 			\input{tables/tab2TimeLAD___}
@@
 	\begin{center}
-		\begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
+		\begin{tabular}{ | l ||  c | c | c | c | c |   c | c |}
 			\hline

### 2015-10-15 paper.tex
feita revisão até a parte final das redes externas

@@
 \begin{table}
-	\caption{Activity percentages along the days of the week.
-	Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
+	\caption{Activity percentages along weekdays.
+	Higher activity was observed during workweek days, with a decrease of activity on weekend days of at least one third and at most two thirds.}
 	\begin{center}

### 2015-10-17 paper.tex
arrumado titulo, finalmente

@@
 %\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
-\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
+\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and activity along different timescales}% Force line breaks with \\
 \author{Renato Fabbri}%

### 2015-10-17 paper.tex
retoque no titulo

--- 2015-10-15 paper.tex
@@
 %\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
-\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
+\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and activity along diverse timescales}% Force line breaks with \\
 \author{Renato Fabbri}%

### 2015-11-02 paper.tex
incorporada revisao do ricardo ateh IIID

--- 2015-10-15 paper.tex
@@
 %\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
-\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
+\title[Stability of interaction networks]{Temporal stability in human interaction networks}
+% : sector sizes, topological prominence and activity along diverse timescales}% Force line breaks with \\
 \author{Renato Fabbri}%
@@
 \begin{abstract}
-In this study we report on stable aspects of human interaction networks,
+This paper reports on stable properties of human interaction networks,
 with benchmarks derived from public email lists.
 Activity along time and topology evolution were observed by snapshots in a timeline 
-and in various scales.
-Notably, the activity in timescales, ranging from seconds to months, is practically the same for all networks.
-The most important metrics to the dispersion of participants in the topological measures space 
+and at different scales.
+Our analysis show that the activity across timescales, ranging from seconds to months, is practically the same for all networks.
+The most important metrics to the dispersion of participants in the topological measures space are shown to be
 are centrality measurements (degree, strength and betweenness),
 followed by symmetry-related metrics and then clustering coefficient.
-The activity of participants followed the expected scale-free trace,
-thus yielding the hubs, intermediary and periphery classes of vertices by comparison against the Erd\"os-R\'enyi model.
-The relative sizes of these three sectors were essentially the same for all email lists and the same along time.
+The observed activity of participants follows the expected scale-free trace,
+thus yielding the hub, intermediary and periphery classes of vertices by comparison against the Erd\"os-R\'enyi model.
+The relative sizes of these three sectors are shown to be essentially the same for all email lists and the same along time.
 Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and 44-81\% are peripheral vertices.
-Similar results for the distribution of participants in the three categories
+Similar results for the distribution of participants in the three sectors
 and for the relative importance of the topological metrics were obtained for 12 additional networks
 from Facebook, Twitter and Participabr.
-Consistent with expectations derived from the literature,
-these properties may be general for human interaction networks,
-which has important implications in establishing a typology of participants based on quantitative criteria.
+These properties are consistent with expectations derived from the literature
+and may be general for human interaction networks,
+which has important implications for establishing a typology of participants based on quantitative criteria.
 \end{abstract}
@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data}
+\keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data, typology}
 \maketitle
@@
 In this paper we analyze the evolution of human interaction networks.
-Interaction in email lists was the most convenient for deriving results and for benchmarking,
-but networks from Facebook, Twitter and Participabr were also considered.
+Interaction networks from email lists were the most convenient for deriving results and for benchmarking
+while networks from Facebook, Twitter and Participabr enhanced the generalization.
 Using a timeline of activity snapshots with a constant number of contiguous messages,
-we found remarkable stability for several of the network properties.
-Namely, the activity along different timescales exhibit pronounced patterns,
-the most basic topological measures always combine into very characteristic principal components,
+we found remarkable stability for important network properties.
+For instance, the activity along different timescales exhibit pronounced patterns;
+the most basic topological measures always combine into very characteristic principal components;
 and the fractions of participants in each of the hubs, intermediady and periphery Erd\"os sectors
 are unshaken.
-This is not an intuitive result, given that participants transition in network structure constantly.
+This is not an intuitive result, given that participants constantly transition in network structure.
 Because these properties were shared by networks from various sources,
@@
 and social sciences in the discussion
-of types of networks and participants.
+of types of networks and of participants.
 Noteworthy, typologies are the canon of scientific literature for the classification of human agents,
@@
 Section~\ref{sec:results} reports results and discussion, leading to Section~\ref{sec:conc} for conclusions.
-Subsidiary data analysis,
+Supplementary data analysis,
 including directions for video and sound mappings of network structures,
-and numeric results for networks from Twitter, Facebook and Participabr,
-are given in the Supporting Information document.
+and numeric detailed results for networks from Twitter, Facebook and Participabr,
+are provided in the Supporting Information document.
 
@@
 Network types have been discussed with regard to the number of participants, intermittence of their activity and network longevity~\cite{barabasiEvo}.
-Two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email list networks, scale-free properties were reported with $\alpha \approx 1.8$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.
-
-Unreciprocated edges often exceed 50\% in the analyzed networks, which matches empirical evidence from the literature~\cite{newmanEvolving} and motivated the inclusion of symmetry metrics in our analysis.
-No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo}, therefore geographical positions were not considered in our study.
-Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex} but is not relevant for the present article because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+Two topologically different networks emerged from human interaction networks, depending on whether the frequency of interactions follows a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email list networks, scale-free properties were reported with $\alpha \approx 1.8$~\cite{bird} (as in web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{Gmane2}.
+
+The fact that unreciprocated edges often exceed 50\% in human interaction networks~\cite{newmanEvolving} motivated the inclusion of symmetry metrics in our analysis.
+No correlation of topological characteristics and geographical coordinates was found~\cite{barabasiGeo},
+therefore geographical positions were not considered in our study.
+Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}
+but is not relevant for the present manuscript because email messages and addresses have no gender related metadata~\cite{gmanePack}.
 
@@
 Email list messages were obtained from
-the GMANE email archive~\cite{gmanePack}, which consists of more than $20\times 10^3$
-email lists and more than $130\times 10^6$ messages~\cite{GMANEwikipedia}.
+the Gmane email archive~\cite{gmanePack}, which consists of more than $20,000$
+email lists (discussion groups) and more than $130\times 10^6$ messages~\cite{Gmanewikipedia}.
 These lists cover a variety of topics, mostly technology-related.
-The archive can be described as a corpus with metadata of its messages, including sent time, place, sender name, and sender email address.
-The GMANE usage in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{GMANE2,bird}.
-
-We analyzed many email lists (and data from Twitter, Facebook and Participabr)
-and selected four of them in order to expose a thorough analysis,
-from which general properties can be inferred. These lists are:
+The archive can be described as a corpus along with message metadata, including sent time, place, sender name, and sender email address.
+The usage of the Gmane database in scientific research is reported in studies of isolated lists and of lexical innovations~\cite{Gmane2,bird}.
+
+We analyzed many email lists together with data from Twitter, Facebook and Participabr
+and selected four email lists for a thorough analysis,
+from which general properties can be inferred:
 \begin{itemize}
-	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in GMANE.},
+	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in Gmane.},
 		with participants from different countries with artistic and technological interests.
-		English is the most used language. Abbreviated as LAU from now on.
-	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in GMANE.},
-		with participants from different countries,
+		English is the prevailing language. Abbreviated as LAU from now on.
+	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in Gmane.},
+		with participants from different countries;
 		a more technical and less active version of LAU.
-		English is the language used the most. 
+		English is the prevailing language. 
 		Abbreviated as LAD from now on.
-	\item Development list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in GMANE.},
+	\item Developer's list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in Gmane.},
 		with computer programmers from different countries.
-		English is the most used language.
+		English is the prevailing language.
 		Abbreviated as CPP from now on.
-	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in GMANE.},
+	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.},
 		a Brazilian digital culture interested email list.
-		Portuguese is the most used language,
+		Portuguese is the prevailing language,
 		although Spanish and English are also incident.
@@
 
-The first 20,000 messages of each list were considered, with total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
+The first 20,000 messages of each list were considered,
+with basic attributes of total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
 We considered 140 additional email lists to
@@
 \end{table}
-
+\subsection{Availability}
 All data and scripts needed to derive results, figures, tables and this article itself
 are publicly available.
-Email messages are downloadable from the GMANE public database~\cite{GMANEwikipedia}.
+Email messages are downloadable from the Gmane public database~\cite{Gmanewikipedia}.
 Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}.
 Data from Participabr was used from the linked data/semantic web RDF triples
-reported in~\cite{opa} and available in~\cite{datahub}.
+reported~\cite{opa} and available~\cite{datahub}.
 Computer scripts are delivered through a public domain Python PyPI package and
@@
 reinforces the scientific aspect of the contribution~\cite{openSci}
-and mitigates ethical and moral issues of researching 
+and mitigates ethical and moral issues involved in researching 
 systems constituted of
@@
 minutes, hours, days of the week, days of the month, and months of the year.
-Most standard measures of location and dispersion, such as usual mean and
-standard deviation, hold little meaning in a compact Riemannian manifold.
+Most standard measures of location and dispersion, e.g. the usual mean and
+standard deviation, hold little meaning in a compact Riemannian manifold,
+such as the recurrent time periods that we are interested in.
 Equivalent measures were taken using circular statistics,
-in which each measurement $t$ is represented as a unity complex number,
+in which each measurement $t$ is represented as a unit complex number,
 $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=t\frac{2\pi}{T}$,
@@
 \subsection{Interaction networks}\label{intNet}
-Interaction networks can be modeled both weighted or unweighted,
-both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
+Edges in interaction networks can be modeled both as weighted or unweighted,
+both as directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
 Networks in this paper are directed and weighted, the most informative of the possibilities.
@@
 Selfloops were regarded as non-informative and discarded.
-Inverting edge direction yields the status network:
+Inverting edge direction canonically yields the status network:
 B read the message and considered what A wrote worth responding, giving status to A,
@@
 The topology of the networks was characterized 
-using a small selection of the most basic and 
+from a small selection of the most basic and 
 fundamental measurements for each vertex, as follows:
@@
 	\item Clustering coefficient $cc_i$: fraction of pairs of neighbors of $i$ that are linked, i.e. the standard clustering coefficient metric for undirected graphs.
-	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index was calculated for weighted digraphs as specified in~\cite{faster}.
+	\item Betweenness centrality $bt_i$: fraction of geodesics that contain vertex $i$. The betweenness centrality index was computed for weighted digraphs as specified in~\cite{faster}.
 \end{itemize}
 
-The non-standard metrics bellow were elaborated
-in order to capture symmetries in the activity of participants:
+The non-standard metrics bellow were formulated
+to capture symmetries in the activity of participants:
 
@@
 	\item Asymmetry of vertex $i$: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-	\item Mean of asymmetry of edges of vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{xy}$ is 1 if there is an edge from $x$ to $y$, and $0$ otherwise.
-$J_i$ is the set of neighbors of vertex $i$, and $|J_i|$ is the number of neighbors of vertex $i$.
+	\item Average asymmetry of edges at vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{ij}$ is 1 if there is an edge from $i$ to $j$, and $0$ otherwise, and $J_i$ is the set of neighbors of vertex $i$.
 	\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$.
 	\item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-	\item Mean of disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{w_{ji}-w_{ij}}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+	\item Average disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{w_{ji}-w_{ij}}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
 	\item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}\left[\mu^{dis}_i-\frac{w_{ji}-w_{ij}}{w_{ji}+w_{ij}}\right]^2}{|J_i|}}$.
@@
 \subsection{Erd\"os sectioning}\label{sectioning}
-In a scale-free network, the peripheral, intermediary and hubs sectors can be derived from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices~\cite{3setores}, as depicted in Figure~\ref{fig:setores}.
-We refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors regarded as \emph{Erd\"os sectors}.
+It is often useful to think of
+vertices as
+hubs, peripheral and intermediary.
+Following this pattern,
+the peripheral, intermediary and hub sectors of the empirical networks were derived
+from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices,
+as depicted in Figure~\ref{fig:setores}.
+We refer to this procedure as \emph{Erd\"os sectioning},
+with the resulting sectors regarded as \emph{Erd\"os sectors}.
+The Erd\"os sectioning was theorized about by M. O. Jackson in his video lectures~\cite{3setores},
+but this might be the first time the method is reported to be applied.
 
@@
 The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected without distinction of the vertices, i.e.
-without preferential attachment and as in the Erd\"os-R\'enyi model, are correctly associated to the hubs sector.
+without preferential attachment and as in the Erd\"os-R\'enyi model, are correctly associated to the hub sector.
 Vertices with very few connections, which are way more abundant than expected in the Erd\"os-R\'enyi model, are assigned to the periphery.
@@
 	\caption{Classification of vertices by comparing degree
-		distributions.
-		The binomial distribution of the Erd\"os-R\'enyi network model exhibit more intermediary vertices,
+		distributions~\cite{3setores}.
+		The binomial distribution of the Erd\"os-R\'enyi network model exhibits more intermediary vertices,
 		while a scale-free network, associated with the power-law distribution,
@@
 Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the periphery,
-intermediary and hubs sectors, the ``Erd\"os sectors''.
+intermediary and hub sectors, the ``Erd\"os sectors''.
 		The connectivity distribution of empirical interaction networks, e.g. derived from email lists,
@@
 		of an interaction network of 98 nodes and 235 edges.
-		A throughout exposition of the method is exposed in Section~\ref{sectioning}.}
+		A through exposition of the method is exposed in Section~\ref{sectioning}.}
 		\label{fig:setores}
@@
 	\begin{split}
-		\sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} P(x) \Rightarrow \\
-	\Rightarrow \Delta_i \text{ holds intermediary degree values.}
+		\sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} P(x) \Leftrightarrow \\
+	\Leftrightarrow \Delta_i \text{ spans intermediary degree values.}
 	\end{split}
@@
 
-If strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ the strength of vertex $i$.
-For in and out degrees ($k^{in}$, $k^{out}$), comparison of the real network should be made against
+If the strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ the strength of vertex $i$.
+For in and out degrees ($k^{in}$, $k^{out}$), the real network should be compared against
 \begin{equation}
@@
 In and out strengths ($s^{in}$, $s^{out}$) are divided by $\overline{w}$ and compared also using $\hat{P}$.
-Note that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$ as with the total degree.
+Note that $p_e$ remains the same, as each edge yields an incoming (or outgoing) edge, and there are at most $N(N-1)$ incoming (or outgoing) edges, thus $p_e=\frac{z}{N(N-1)}$, as with the total degree.
 
@@
 		The remaining vertices are peripheral if they are peripheral or hubs for all metrics.
-		The rest of the nodes are intermediary.
+		The remaining nodes are intermediary.
 	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric.
@@
 
-Using equations~\ref{eq:part}, these \emph{compound criteria} $C_\delta$, with $\delta$ integer in the interval $1\leq\delta\leq6$, can be specified as:
+Using Equations~(\ref{eq:part}), these \emph{compound criteria} $C_\delta$, with $\delta$ integer in the interval $1\leq\delta\leq6$, can be specified as:
 
@@
 \subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
-Principal Component Analysis (PCA) is a well documented technique~\cite{pca} and was used to acquire knowledge about:
-1) which metrics contribute to each principal component and in which proportion; 2) how much of the dispersion is concentrated in each component; 3) expected values and dispersions for these quantities over various networks.
+Principal Component Analysis (PCA) is a well documented technique~\cite{pca}
+and was employed in the investigation to acquire knowledge about:
+1) which metrics contribute to each principal component and in what proportion;
+2) how much of the dispersion is concentrated in each component;
+3) expected values and dispersions for these quantities over various networks.
+This enables a characterization of the human interaction networks
+in terms of the relative importance of networks measures ad the way they combine.
 
@@
 Network structures were mapped to video animations, sound and musical structures,
-image galleries and online gadgets developed for this research~\cite{animacoes,galGMANE,appGMANE}.
+image galleries and online gadgets developed for this research~\cite{animacoes,galGmane,appGmane}.
 Such \emph{audiovisualizations} were crucial in the initial steps and
@@
 	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
-	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to GMANE creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
+	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
 	We are also grateful to developers and users of Python scientific tools.

### 2015-11-02 paper.tex
terminada incorporacao ateh o final dos metodos

@@
 
-Let $\mathbf{X}=\{X[i,j]\}$ be a matrix of all vertices $i$ and the respective
-values for each metric $j$, 
-$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ the mean of metric $j$ over all $I$ vertices, 
+Let $\mathbf{X}=\{X[i,j]\}$ be a matrix where each element is the value
+of of each metric $j$ at vertex $i$ .
+Let
+$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ be the mean of metric $j$ over all $I$ vertices, 
 $\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
@@
 %it is enough to 
-Be $\mathbf{D}=\{D[k]\}$ the eigenvalues associated to the eigenvectors $\mathbf{V}$,
+Let $\mathbf{D}=\{D[k]\}$ be the eigenvalues associated to the eigenvectors $\mathbf{V}$,
 then $D'[k]=100\frac{D[k]}{\sum_{k'}D[k']}$
@@
 $\{(D'[k],\;V'[j,k])\}$.
-These usually sum between 60 and 95\% of the dispersion
+These usually sum up between 60 and 95\% of the dispersion
 and reveal patterns for a first analysis.
@@
 sequences of snapshots.
-In each sequence,a fixed number of messages,
-the window size $ws$, was used for all snapshots.
+In each sequence, a fixed number of messages,
+i.e. the window size $ws$, was used for all snapshots.
 %was considered with different shifts in the message timeline to obtain snapshots.
 The snapshots are disjoint in the message timeline,
-and were used to perform both the PCA
-with topological metrics and the Erd\"os sectioning.  
+and were used to perform both PCA
+with topological metrics and Erd\"os sectioning.  
 Figures and tables were usually inspected in

### 2015-11-02 paper.tex
terminando incorporacao do ric nos resultados, commit intermediario para fazer figura

@@
 	The most uniform distribution of activity was found in seconds and minutes.
-	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 14h and 15h ($\theta'=-9.61$).
-	Weekdays, month days and months have mean near zero (i.e. near the beginning of the week, month and year) and high dispersion.}
+	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 2 p.m. and 3 p.m. ($\theta'=-9.61$).
+	Weekdays, month days and months have mean near zero (i.e. near the beginning of the week, month and year) and high dispersion.
+Notice that $\theta_u'$ has the dimensional unit of the corresponding time period while $\delta(z)$ is dimensionless.}
 	\begin{center}
@@
 \begin{table}
-	\caption{Activity percentages along the hours of the day. Nearly identical distributions are found on other systems as shown in Section~\ref{si:hours} of the Supporting Information document.
-	Higher activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight. Around 2/3 of the activity takes place from noon to midnight.
-Nevertheless, the activity peak occurs between 11 a.m. and 12 p.m. This example and empirical table corresponds to the activity in CPP.}
+	\caption{Activity percentages along the hours of the day. Nearly identical distributions
+		were observed on other social systems as shown in Section~\ref{si:hours} of the Supporting Information document.
+	Highest activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight.
+	Around 2/3 of the activity takes place from noon to midnight
+	but the activity peak occurs between 11 a.m. and 12 p.m.
+	This table shows results for the activity in CPP.}
 	\footnotesize
@@
 
-In the scales of seconds and minutes, activity in uniform,
+In the scales of seconds and minutes, activity is uniform,
 with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.8.2, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
@@
 Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.
-This example and empirical table corresponds to the activity of the MET email list.}
+This specific example and empirical table corresponds to the activity of the MET email list.}
 	\footnotesize
@@
 	 This table corresponds to activity in LAU. 
-		Further evidence is in Section~\ref{si:months} of the Supporting Information document.
-	}
+		Similar results are posted in Section~\ref{si:months} of the Supporting Information document.}
 	\footnotesize
@@
 
-The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors 
-is remarkably stable along time is samples of 200 or more messages,
+The distribution of vertices in the hub, intermediary, periphery Erd\"os sectors 
+is remarkably stable along time if the snapshots hold 200 or more messages,
 as evident in Figure~\ref{fig:sectIL} of current document
@@
 This is expected for a system with a scale-free profile
-and confirmed by Table~\ref{autores} in the distribution of activity among participants.
-
-Typically, $\approx [3-12]\%$ of the vertices are hubs,
-$\approx [15-45]\%$ are intermediary and $\approx [44-81]\%$ are peripheral,
-which is consistent with the literature~\cite{secFree}.
+and confirmed by Table~\ref{autores} of the distribution of activity among participants.
+
+Typically, $[3\%-12\%]$ of the vertices are hubs,
+$[15\%-45\%]$ are intermediary and $[44\%-81\%]$ are peripheral,
+which is consistent with other studies~\cite{secFree}.
 These results hold for the total, in and out degrees and strengths.
@@
 In order to support hypotheses about the generality of these findings,
-we expose the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr
+we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr
 in Table~\ref{tab:secE} of the Supporting Information document.
@@
 	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right.
-	In the ordinates $\overline{e_{n,x}}=\frac{|e_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $E_n$
+	The ordinates $\overline{e_{n,x}}=\frac{|e_{n,x}|}{N}$ denote the fraction of participants in sector $x$ through criterion $E_n$
 	and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
-Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning} and present these fractions of participants in networks from Facebook, Twitter and Participabr.}
+Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, and present these results in networks from Facebook, Twitter and Participabr.}
 	\label{fig:sectIL}
@@
 The principal components of the participants are very stable in the topological space,
-i.e. in PCA of network measures.
+i.e. in PCA space of network measures.
 Table~\ref{tab:pcain} exemplifies the formation of principal components
@@
 the stability of principal components,
-is underpinned by the very small standard deviations
+is underpinned by the very small dispersion 
 of the contribution of

### 2015-11-02 paper.tex
terminada incorporacao das contribs do ric

@@
 	and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
-Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, and present these results in networks from Facebook, Twitter and Participabr.}
+Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, and present these results for networks from Facebook, Twitter and Participabr.}
 	\label{fig:sectIL}
@@
 The clustering coefficient is only relevant for the third principal component.
-The three components have in average 80\% of the variance.}
+		The three components have in average more than 85\% of the variance.}
 		\footnotesize
@@
 	\includegraphics[width=.45\textwidth]{figs/im13PCAPLOT__}
-	\caption{The first plot exposes the wel known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
-		The second plot exhibits the greater dispersion of the symmetry-related ordinates.
-		This greater dispersion suggests that symmetry-related metrics are more powerful, with respect to dispersion,
+	\caption{The first plot exposes the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
+		The second plot shows the greater dispersion of the symmetry-related ordinates dominant in PC2.
+		This greater dispersion suggests that symmetry-related metrics are more powerful,
 		for characterizing interaction networks than the clustering coefficient,
@@
 We conclude that the symmetry metrics are more powerful,
-with respect to dispersion in the topological space,
+with respect to dispersion in the topological measures space,
 in characterizing interaction networks
@@
 Assigning a type to a participant raises important 
-issues about the scientific cannon for human ypes
+issues about the scientific cannon for human types
 and the potential for stigmatization and prejudice. 
 The type of the Erd\"os sector to which a participant belongs 
-can be regarded as implying a type to the corresponding participant.
+can be regarded as implying a social type to the corresponding participant.
 In this case, the type of a participant
@@
 different networks are considered,
+despite the stability of the network,
 and therefore the potential for prejudice of such
@@
 the derived participant type is static.
-The types, in this latter case, fails to acknowledge that
+The types, in this latter case, fail to acknowledge that
 human individuals are not immutable entities.
@@
 Stable activity was found only in smaller communities.
-For example, MET list have stable hubs while all LAU, LAD and CPP
+For instance, the MET list have stable hubs while all LAU, LAD and CPP
 present intermittent hubs.
@@
 The temporal stability reported in this article is very counterintuitive.
-These characteristics were expected from current scientific knowledge,
+Some of the observed characteristics were expected from current scientific knowledge,
 such as the fat tail structure and the preponderance of centrality measures,
@@
 which surpassed that of clustering coefficient,
-with respect to dispersion of the system in the topological space,
-might add to current understanding of key-differences between digraphs and
+with respect to dispersion of the system in the topological measures space,
+might add to the current understanding of key-differences between digraphs and
 undirected graphs in complex networks.
@@
 more types of networks and more metrics.
-The data and software needed to reach these results
-might also receive dedicated and in-depth
+The data and software needed to attain these results
+should also receive dedicated and in-depth
 documentation as they enable a greater level of transparency
 and work share,
-which is adequate to both benchmarking
-and the study of systems constituted
+which is adequate for both benchmarking
+and specifically for the study of systems constituted
 by human individuals (see Section~\ref{sec:data}).
 The derived typology of hub, intermediary and peripheral participants
-was used for semantic web and participatory democracy efforts,
+has been applied for semantic web and participatory democracy efforts,
 and these developments might be enhanced to yield scientific knowledge~\cite{opa}.
-Also, we should pursue the publication of the visualization and sonification
+Also, we plan to further apply and publish of the visualization and sonification
 processes used for this research~\cite{versinus,animacoes} and
-of the extreme linguistic differences found in each of the Erd\"os sectors~\cite{rcText}.
+the extreme linguistic differences found in each of the Erd\"os sectors~\cite{rcText}.
 

### 2015-11-02 paper.tex
pequeno retoque e anotacao

@@
 \subsection{Types from Erd\"os sectors}\label{sec:pty}
+
 Assigning a type to a participant raises important 
@@
 \end{figure}
+
+\section{Discussion}
+% given the results, and before reaching the conclusions
+% what to say?
+% --> what is the overall knowledge derived from the results
+% --> what are the limitations of this knowledge and of individual results
+% --> how should this results carry on is on the next sections.
+\subsection{Consecutive scientific research}
+% --> research
+% textual diferences
+% audiovisualization of data
+% typologies, sociological critical theory, social psychology
+
+\subsection{Technological applications}
+% --> technological 
+% resources categorization and recommendation
+% document creation
+% ontologies for the semantic web
+
+\subsection{Experimental and theoretical aspects of the research}
+% --> methods
+% Exploratory?
+% Hypothesis testing?
+% --> contributions
+% verifiable
+% knowledge
+% contextualization in the academic knowledge
+
+% applications 
+% experimental and theoretical aspects of the research
 
@@
 and these developments might be enhanced to yield scientific knowledge~\cite{opa}.
-Also, we plan to further apply and publish of the visualization and sonification
+Also, we plan to further explore and publish of the visualization and sonification
 processes used for this research~\cite{versinus,animacoes} and
@@
 	project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
-	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Brazilian Presidency of the Republic for keeping Participabr code and data open.
+	The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Presidency of the Brazilian Republic for keeping Participabr code and data open.
 	We are also grateful to developers and users of Python scientific tools.

### 2015-11-02 paper.tex
removidos titulos da secao extra de dicsussoes
arrumado nome, coisa pouca
arrumando resumo

@@
 %
-\author{Leonardo Paulo Maia}
+\author{Leonardo P. Maia}
   \homepage{http://www.ifsc.usp.br/~lpmaia/}
@@
 The most important metrics to the dispersion of participants in the topological measures space are shown to be
-are centrality measurements (degree, strength and betweenness),
+centrality metrics (degree, strength and betweenness),
 followed by symmetry-related metrics and then clustering coefficient.
@@
 
-\section{Discussion}
+%\section{Discussion}
 % given the results, and before reaching the conclusions
@@
 % --> how should this results carry on is on the next sections.
-\subsection{Consecutive scientific research}
+%\subsection{Consecutive scientific research}
 % --> research
@@
 
-\subsection{Technological applications}
+%\subsection{Technological applications}
 % --> technological 
@@
 
-\subsection{Experimental and theoretical aspects of the research}
+%\subsection{Experimental and theoretical aspects of the research}
 % --> methods
@@
 % contextualization in the academic knowledge
-
-% applications 
-% experimental and theoretical aspects of the research
 

### 2015-11-03 paper.tex
incorporadas contribuicoes do chu

@@
 
-%\title[Evolution of interaction networks]{On the evolution of interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
-%\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
-%\title[Stability of interaction networks]{Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.}% Force line breaks with \\
-%\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
 \title[Stability of interaction networks]{Temporal stability in human interaction networks}
@@
 \altaffiliation{
-	Curso de Psicologia, Universidade Federal do Cer\'a (UFC)
+	Curso de Psicologia, Universidade Federal do Cear\'a (UFC)
 }%Lines break automatically or can be forced with \\
@@
 %
-\author{Leonardo P. Maia}
+\author{Leonardo Paulo Maia}
 \homepage{http://www.ifsc.usp.br/~lpmaia/}
@@
 	This paper reports on stable properties of human interaction networks,
-with benchmarks derived from public email lists.
-Activity along time and topology evolution were observed by snapshots in a timeline 
+	with benchmarks derived from public email lists. Activity along time and topology evolution were observed in snapshots in a timeline, 
 	and at different scales.
-Our analysis show that the activity across timescales, ranging from seconds to months, is practically the same for all networks.
-The most important metrics to the dispersion of participants in the topological measures space are shown to be
-centrality metrics (degree, strength and betweenness),
-followed by symmetry-related metrics and then clustering coefficient.
-The observed activity of participants follows the expected scale-free trace,
-thus yielding the hub, intermediary and periphery classes of vertices by comparison against the Erd\"os-R\'enyi model.
-The relative sizes of these three sectors are shown to be essentially the same for all email lists and the same along time.
+Our analysis shows that activity is practically the same for all networks across timescales ranging from seconds to months.
+The most important metrics to the dispersion of participants in the topological measures space are related to centrality (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient.
+The activity of participants follows the expected scale-free trace, thus yielding the hub, intermediary and peripheral classes of vertices by comparison against the Erd\"os-R\'enyi model.
+The relative sizes of these three sectors are essentially the same for all email lists and the same along time.
 Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and 44-81\% are peripheral vertices.
-Similar results for the distribution of participants in the three sectors
-and for the relative importance of the topological metrics were obtained for 12 additional networks
-from Facebook, Twitter and Participabr.
-These properties are consistent with expectations derived from the literature
-and may be general for human interaction networks,
-which has important implications for establishing a typology of participants based on quantitative criteria.
+Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
+These properties are consistent with expectations from the literature and may be general for human interaction networks, which has important implications for establishing a typology of participants based on quantitative criteria.
 \end{abstract}
@@
 \section{Introduction}\label{sec:into}
-First studies explicitly about human interaction networks 
+The first studies dealing with human interaction networks explicitly
 date from the nineteenth century while the foundation of
@@
 In this paper we analyze the evolution of human interaction networks.
-Interaction networks from email lists were the most convenient for deriving results and for benchmarking
-while networks from Facebook, Twitter and Participabr enhanced the generalization.
-Using a timeline of activity snapshots with a constant number of contiguous messages,
-we found remarkable stability for important network properties.
-For instance, the activity along different timescales exhibit pronounced patterns;
-the most basic topological measures always combine into very characteristic principal components;
-and the fractions of participants in each of the hubs, intermediady and periphery Erd\"os sectors
-are unshaken.
+Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and Participabr were used for the sake of generalization.
+Using a timeline of activity snapshots with a constant number of contiguous messages, we found remarkable stability for important network properties.
+For instance, activity along different timescales follows specific patterns; the most basic topological metrics can always be combined into characteristic principal components; and the fractions of participants in different sectors do not vary with time.
 This is not an intuitive result, given that participants constantly transition in network structure.
-Because these properties were shared by networks from various sources,
-and are consistent with the complex networks literature,
-we advocate that the conclusions might be valid for general classes of interaction networks.
+Because these properties were shared by networks from various sources, and are consistent with the literature in complex networks, we advocate that the conclusions might be valid for general classes of interaction networks.
 In particular, this allows us to
-bridge the gap between data analysis 
-and social sciences in the discussion
-of types of networks and of participants.
-Noteworthy, typologies are the canon of scientific literature for the classification of human agents,
-with pragmatic standards~\cite{myers} and critical paradigms~\cite{adorno,typCanon}.
-
-Section~\ref{sec:related} describes related work,
-while data, scripts and methods of analysis are given in Section~\ref{sec:data} and Section~\ref{sec:carac}.
+bridge the gap between data analysis and social sciences in the discussion of types of networks and of participants.
+It is worth noting that typologies are the canon of scientific literature for the classification of human agents, with pragmatic standards~\cite{myers} and critical paradigms~\cite{adorno,typCanon}.
+
+This paper is organized as follows. Section~\ref{sec:related} describes related work, while data, scripts and methods of analysis are given in Section~\ref{sec:data} and Section~\ref{sec:carac}.
 Section~\ref{sec:results} reports results and discussion, leading to Section~\ref{sec:conc} for conclusions.
-Supplementary data analysis,
-including directions for video and sound mappings of network structures,
-and numeric detailed results for networks from Twitter, Facebook and Participabr,
-are provided in the Supporting Information document.
+Supplementary data analysis, including directions for video and sound mappings of network structures, and numeric detailed results for networks from Twitter, Facebook and Participabr, are provided in the Supporting Information document.
 
@@
 
-We analyzed many email lists together with data from Twitter, Facebook and Participabr
-and selected four email lists for a thorough analysis,
-from which general properties can be inferred:
+We analyzed many email lists together with data from Twitter, Facebook and Participabr and selected four email lists for a thorough analysis,
+from which general properties can be inferred. These lists are as follows:
+
 \begin{itemize}
-	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in Gmane.},
-		with participants from different countries with artistic and technological interests.
-		English is the prevailing language. Abbreviated as LAU from now on.
-	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in Gmane.},
-		with participants from different countries;
-		a more technical and less active version of LAU.
+	\item Linux Audio Users list\footnote{gmane.linux.audio.users is list ID in Gmane.},with participants from different countries with artistic and technological interests.
+English is the prevailing language.
+Abbreviated as LAU from now on.
+
+	\item Linux Audio Developers list\footnote{gmane.linux.audio.devel is list ID in Gmane.}, with participants from different countries; a more technical and less active version of LAU.
 English is the prevailing language.
 Abbreviated as LAD from now on.
-	\item Developer's list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in Gmane.},
-		with computer programmers from different countries.
+
+	\item Developer's list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in Gmane.}, with computer programmers from different countries.
 English is the prevailing language.
 Abbreviated as CPP from now on.
-	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.},
-		a Brazilian digital culture interested email list.
-		Portuguese is the prevailing language,
-		although Spanish and English are also incident.
+	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.}, a Brazilian email list for digital culture.
+	Portuguese is the prevailing language, although some messages are writte in Spanish and English.
 Abbreviated as MET from now on.
@@
 
-The first 20,000 messages of each list were considered,
-with basic attributes of total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
-We considered 140 additional email lists to
-report on the interdependence between the number of participants and 
-the number of discussion threads.
-Furthermore, 12 networks from Facebook (8), Twitter (2) and Participabr (2)
-are scrutinized in the Supporting Information document for better hypothesizing
-about the generality of the results.
+The first 20,000 messages of each list were considered, with basic attributes of total timespan, authors, threads and missing messages indicated in Table~\ref{tab:genLists}.
+We considered 140 additional email lists to report on the interdependence between the number of participants and the number of discussion threads.
+Furthermore, 12 networks from Facebook (8), Twitter (2) and Participabr (2) were scrutinized, and their analysis is given in the Supporting Information document for the purpose of testing the generality of the results.
 
@@
 \subsection{Availability}
-All data and scripts needed to derive results, figures, tables and this article itself
-are publicly available.
+The data and scripts used to derive the results, figures and tables, and this article itself are publicly available.
 Email messages are downloadable from the Gmane public database~\cite{Gmanewikipedia}.
 Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}.
-Data from Participabr was used from the linked data/semantic web RDF triples
-reported~\cite{opa} and available~\cite{datahub}.
-Computer scripts are delivered through a public domain Python PyPI package and
-an open Git repository~\cite{gmanePack}.
-This open approach to both data and scripts
-reinforces the scientific aspect of the contribution~\cite{openSci}
-and mitigates ethical and moral issues involved in researching 
-systems constituted of
-human individuals~\cite{anPhy,ccs15}.
+Data from Participabr were used from the linked data/semantic web RDF triples~\cite{opa}, available in~\cite{datahub}.
+Computer scripts are delivered through a public domain Python PyPI package and an open Git repository~\cite{gmanePack}.
+This open approach to both data and scripts reinforces the scientific aspect of the contribution~\cite{openSci} and mitigates ethical and moral issues involved in researching systems constituted of human individuals~\cite{anPhy,ccs15}.
 
@@
 Also, the ratio $r=\frac{b_l}{b_h}$ between the lowest $b_l$ and the highest $b_h$ incidences on the histograms 
-served as a further clue of how close the distribution was to being uniform.
-As expected, a positive correlation was found in all $r, Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures,
-which can be noticed in the Section~\ref{si:circ} of the Supporting Information document.
-The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred 
-in the discussion of results.
+served as a further clue of how close the distribution was to being uniform. As expected, a positive correlation was found in all $r, Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures,
+which can be noticed in Section~\ref{si:circ} of the Supporting Information. The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred in the discussion of results.
 
 \subsection{Interaction networks}\label{intNet}
-Edges in interaction networks can be modeled both as weighted or unweighted,
-both as directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
+Edges in interaction networks can be modeled both as weighted or unweighted, as directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
 Networks in this paper are directed and weighted, the most informative of the possibilities.
-We did not investigate directed unweighted, undirected weighted,
-and undirected unweighted representations of the interaction networks.
-
-The interaction networks were obtained as follows:
-a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
-The reasoning is: if B wrote a response to a message from A,
-he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
+We did not investigate directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks.
+
+The interaction networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B.
+The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$.
 Edges in both directions are allowed.
@@
 Selfloops were regarded as non-informative and discarded.
-Inverting edge direction canonically yields the status network:
-B read the message and considered what A wrote worth responding, giving status to A,
-thus $B\rightarrow A$.
+Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$.
 This paper considers by convention the information network as described above ($A\rightarrow B$) and depicted in Figure~\ref{formationNetwork}.
-These interaction networks are reported in the literature
-as exhibiting scale-free and small-world properties,
-as expected for a number of social networks~\cite{bird,newmanBook}.
+These interaction networks are reported in the literature as exhibiting scale-free and small-world properties, as expected for a number of social networks~\cite{bird,newmanBook}.
 
@@
 Each vertex represents a participant.
-	A reply message from author B to a message from author A is regarded as evidence
-	that B received information from A and yields a directed edge.
+A reply message from author B to a message from author A is regarded as evidence that B received information from A and yields a directed edge.
 	Multiple messages add ``weight'' to a directed edge.
@@
 
-The non-standard metrics bellow were formulated
-to capture symmetries in the activity of participants:
+The non-standard metrics below were formulated to capture symmetries in the activity of participants:
 
@@
 
-Both standard and non-standard metrics are used for the Erd\"os sectioning (described in Secion~\ref{sectioning}) and for performing PCA (as described in Section~\ref{sec:pca}).
+Both standard and non-standard metrics are used for the Erd\"os sectioning (described in Section~\ref{sectioning}) and for performing principal component analysis (PCA) (as described in Section~\ref{sec:pca}).
 
 \subsection{Erd\"os sectioning}\label{sectioning}
-It is often useful to think of
-vertices as
-hubs, peripheral and intermediary.
-Following this pattern,
-the peripheral, intermediary and hub sectors of the empirical networks were derived
-from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices,
+It is often useful to think of vertices as hubs, peripheral and intermediary.
+We have therefore derived the peripheral, intermediary and hub sectors of the empirical networks from a comparison against an Erd\"os-R\'enyi network with the same number of edges and vertices,
 as depicted in Figure~\ref{fig:setores}.
-We refer to this procedure as \emph{Erd\"os sectioning},
-with the resulting sectors regarded as \emph{Erd\"os sectors}.
-The Erd\"os sectioning was theorized about by M. O. Jackson in his video lectures~\cite{3setores},
-but this might be the first time the method is reported to be applied.
-
-The degree distribution $\widetilde{P}(k)$ of a
-scale-free network $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
+We refer to this procedure as \emph{Erd\"os sectioning}, with the resulting sectors being named as \emph{Erd\"os sectors}.
+The Erd\"os sectioning was recognized as a theoretical possibility by M.
+O.
+Jackson in his video lectures~\cite{3setores}, but to our knowledge it has not as yet been applied to empirical data.
+
+The degree distribution $\widetilde{P}(k)$ of a scale-free network $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
 average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
@@
 
-The lower degree fat tail corresponds to the border vertices, i.e. the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any intermediary sector value of $k$.
-The higher degree fat tail is the hub sector, i.e. $\widetilde{P}(k)>P(k)$ and $k$ is higher than any intermediary sector value of $k$.
+The lower degree fat tail corresponds to the border vertices, i.e. the peripheral sector or periphery where $\widetilde{P}(k)>P(k)$ and $k$ is lower than any value of $k$ in the intermediary sector.
+The higher degree fat tail is the hub sector, i.e.
+$\widetilde{P}(k)>P(k)$ and $k$ is higher than any value of $k$ in the intermediary sector.
 The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected without distinction of the vertices, i.e.
@@
 Vertices with very few connections, which are way more abundant than expected in the Erd\"os-R\'enyi model, are assigned to the periphery.
-Vertices with degree values predicted as the most abundant in a Erd\"os-R\'enyi model,
-near the average, and less frequent in the real network, are classified as intermediary.
+Vertices with degree values predicted as the most abundant in a Erd\"os-R\'enyi model, near the average, and less frequent in the real network, are classified as intermediary.
 
@@
 		distributions~\cite{3setores}.
-		The binomial distribution of the Erd\"os-R\'enyi network model exhibits more intermediary vertices,
-		while a scale-free network, associated with the power-law distribution,
-		has more peripheral and hub vertices.
+The binomial distribution of the Erd\"os-R\'enyi network model exhibits more intermediary vertices, while a scale-free network, associated with the power-law distribution, has more peripheral and hub vertices.
 The sector borders are defined with respect to the intersections of the distributions.
-Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the periphery,
-intermediary and hub sectors, the ``Erd\"os sectors''.
-		The connectivity distribution of empirical interaction networks, e.g. derived from email lists,
-		can be sectioned by comparison against the associated binomial distribution with the same number of vertices and edges.
-		In this figure, a snapshot of 1000 messages from CPP list yield the degree distribution
-		of an interaction network of 98 nodes and 235 edges.
-		A through exposition of the method is exposed in Section~\ref{sectioning}.}
+Characteristic degrees are in the compact intervals: $[0,k_L]$, $(k_L,k_R]$, $(k_R,k_{max}]$ for the periphery, intermediary and hub sectors, the ``Erd\"os sectors''.
+																																																																																	The connectivity distribution of empirical interaction networks, e.g.
+derived from email lists, can be sectioned by comparison against the associated binomial distribution with the same number of vertices and edges.
+In this figure, a snapshot of 1000 messages from CPP list yields the degree distribution of an interaction network of 98 nodes and 235 edges.
+A through explanation of the method is provided in Section~\ref{sectioning}.}
 																																																																																	\label{fig:setores}
 																																																																																\end{figure}
+
 																																																																																To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network.
@@
 
-If the strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ the strength of vertex $i$.
+																																																																																If the strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ is the strength of vertex $i$.
 For in and out degrees ($k^{in}$, $k^{out}$), the real network should be compared against
@@
 
-Since different metrics can be used to identify the three types of vertices,
-more than one metric can be used simultaneously, which is convenient when analysing 
-small networks,
+																																																																																Since different metrics can be used to identify the three types of vertices, more than one metric can be used simultaneously, which is convenient when analysing small networks,
 																																																																																such as the cases where $ws=50$ in Section~\ref{si:frac} of the Supporting Information.
@@
 																																																																																	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics.
-		In this case, vertices classified do not usually reach $N$ (or 100\%),
-		which is indicated by a black line in Figure~\ref{fig:sectIL}.
+In this case, vertices classified do not usually reach $N$ (or 100\%), which is indicated by a black line in Figure~\ref{fig:sectIL}.
+
 																																																																																	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics.
-		Therefore, a vertex may belong to more than one class, and the total number of memberhips may exceed $N$ (or 100\%),
-		which is indicated by a black line in Figure~\ref{fig:sectIL}.
+Therefore, a vertex may belong to more than one class, and the total number of memberships may exceed $N$ (or 100\%), which is indicated by a black line in Figure~\ref{fig:sectIL}.
+
 																																																																																	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics.
@@
 The remaining vertices are regarded as peripheral.
+
 																																																																																	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as such according to any of the metrics.
@@
 Peripheral vertices are those which are classified as such with respect to all metrics.
+
 																																																																																	\item Exclusivist externals $C_5$: vertices are hubs if they are classified as such according to all the metrics.
-		The remaining vertices are peripheral if they are peripheral or hubs for all metrics.
+Vertices are peripheral if they are peripheral or hubs for all metrics.
 The remaining nodes are intermediary.
+
 																																																																																	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric.
@@
 																																																																																This enables a characterization of the human interaction networks
-in terms of the relative importance of networks measures ad the way they combine.
+																																																																																in terms of the relative importance of networks measures and the way they combine.
 
 																																																																																Let $\mathbf{X}=\{X[i,j]\}$ be a matrix where each element is the value
-of of each metric $j$ at vertex $i$ .
+																																																																																of each metric $j$ at vertex $i$ .
 																																																																																Let
@@
 																																																																																is the percentage of the principal component $k$
-that is directly proportional to the measure $j$.
+																																																																																that is proportional to the metric $j$.
 																																																																																%With $k$ eigenvectors 
@@
 																																																																																%it is enough to 
-Let $\mathbf{D}=\{D[k]\}$ be the eigenvalues associated to the eigenvectors $\mathbf{V}$,
+																																																																																Let $\mathbf{D}=\{D[k]\}$ be the eigenvalues associated with the eigenvectors $\mathbf{V}$,
 																																																																																then $D'[k]=100\frac{D[k]}{\sum_{k'}D[k']}$
@@
 																																																																																The evolution of the networks was observed within 
-sequences of snapshots.
-In each sequence, a fixed number of messages,
+																																																																																sequences of snapshots. In each sequence, a fixed number of messages,
 																																																																																i.e. the window size $ws$, was used for all snapshots.
 																																																																																%was considered with different shifts in the message timeline to obtain snapshots.
-The snapshots are disjoint in the message timeline,
-and were used to perform both PCA
-with topological metrics and Erd\"os sectioning.  
-Figures and tables were usually inspected in
+																																																																																The snapshots are disjoint in the message timeline, and were used to perform both PCA with topological metrics and Erd\"os sectioning.
+																																																																																Figures and tables were usually inspected with 
 																																																																																$ws=\{50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000, 10000\}$ messages.
@@
 																																																																																and other network characteristics, within the same window size $ws$,
-are exposed in Section~\ref{si:frac}
-of the Supporting Information document.
-
-Network structures were mapped to video animations, sound and musical structures,
-image galleries and online gadgets developed for this research~\cite{animacoes,galGmane,appGmane}.
-Such \emph{audiovisualizations} were crucial in the initial steps and
+																																																																																are given in Section~\ref{si:frac} of the Supporting Information document.
+
+																																																																																Network structures were mapped to video animations, sound and musical structures, image galleries and online gadgets developed for this research~\cite{animacoes,galGmane,appGmane}.
+Such \emph{audiovisualizations} were crucial in the initial steps, and
 																																																																																to guide the research into the most important features of network evolution.
@@
 																																																																																in the scales of seconds, minutes, hours, days and months.
-%exhibited the same patterns in different networks.
-Histograms in each of the time scales were computed as were circular average and dispersion values.
-Values and dedicated expositions are provided in Tables~\ref{tab:circ}-\ref{tab:min2}.
-%they can be used for characterizing nodes (participants) in these and other networks,
-%and networks themselves.
-%For example, they may serve for identification of outliers in a community.
+																																																																																Histograms in each of the time scales were computed as were circular average and dispersion values, and The results are given in Tables~\ref{tab:circ}-\ref{tab:min2}.
+For example, uniform activity is found with respect to seconds, minutes and days of the months.
+Weekend days exhibit about half the activity of regular weekdays, and there is a peak of activity between 11am and 12pm.
 
@@
 The most uniform distribution of activity was found in seconds and minutes.
-	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 2 p.m. and 3 p.m. ($\theta'=-9.61$).
-	Weekdays, month days and months have mean near zero (i.e. near the beginning of the week, month and year) and high dispersion.
-Notice that $\theta_u'$ has the dimensional unit of the corresponding time period while $\delta(z)$ is dimensionless.}
+	Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 2 p.m.
+and 3 p.m.
+($\theta'=-9.61$).
+Weekdays, days of the month and months have mean near zero (i.e.
+near the beginning of the week, month and year) and high dispersion.
+Note that $\theta_u'$ has the dimensional unit of the corresponding time period while $\delta(z)$ is dimensionless.}
 																																																																																	\begin{center}
@@
 																																																																																\begin{table}
-	\caption{Activity percentages along the hours of the day. Nearly identical distributions
-		were observed on other social systems as shown in Section~\ref{si:hours} of the Supporting Information document.
+																																																																																	\caption{Activity percentages along the hours of the day.
+Nearly identical distributions were observed on other social systems as shown in Section~\ref{si:hours} of the Supporting Information document.
 																																																																																		Highest activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight.
@@
 																																																																																\end{table}
-For example, uniform activity is found with respect to seconds, minutes and days of the months,
-weekend days exhibit about half the activity of regular weekdays,
-and there is a peak of activity between 11am and 12pm.
 
@@
 																																																																																In these cases, the circular dispersion is maximized and the mean has little meaning as indicated in Table~\ref{tab:circ}.
-As for the hours of the day, an abrupt peak is found between 11am and 12pm with the most active period being the afternoon, with one third of total daily activity, and two thirds of activity is allocated in the second 12h of each day.
+																																																																																As for the hours of the day, an abrupt peak is found between 11am and 12pm with the most active period being the afternoon, with one third of total daily activity, and two thirds of activity are allocated in the second 12h of each day.
 Days of the week revealed a decrease between one third and two thirds of activity on weekends.
@@
 																																																																																Months of the year revealed patterns matching usual work and academic calendars.
-The time period examined here was not sufficient for the analysis of activity along the years. These patterns are exemplified in Tables~\ref{tab:hin}-\ref{tab:min2}.
+The time period examined here was not sufficient for the analysis of activity along the years.
+These patterns are exemplified in Tables~\ref{tab:hin}-\ref{tab:min2}.
 
@@
 Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.
-This specific example and empirical table corresponds to the activity of the MET email list.}
+																																																																																	This specific example and empirical table correspond to the activity of the MET email list.}
 																																																																																	\footnotesize
@@
 																																																																																	\caption{Activity percentages on months along the year.
-		Activity is usually concentrated in Jun-Aug and/or in Dec-Mar,
-		potentially due to academic calendars, vacations and end-of-year holidays.
+	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays.
 This table corresponds to activity in LAU.
-		Similar results are posted in Section~\ref{si:months} of the Supporting Information document.}
+Similar results are shown in Section~\ref{si:months} of the Supporting Information document.}
 																																																																																	\footnotesize
@@
 
-The distribution of vertices in the hub, intermediary, periphery Erd\"os sectors 
-is remarkably stable along time if the snapshots hold 200 or more messages,
-as evident in Figure~\ref{fig:sectIL} of current document
-and in Section~\ref{si:frac} of the Supporting Information document. 
+																																																																																The distribution of vertices in the hub, intermediary, periphery Erd\"os sectors is remarkably stable along time if the snapshots hold 200 or more messages, as it is clear in Figure~\ref{fig:sectIL} and in Section~\ref{si:frac} of the Supporting Information document.
 																																																																																%Moreover, all email lists analyzed exhibit the same distribution profile.
-Activity is highly concentrated on the hubs,
-while a very large number of peripheral vertices contribute to only a fraction of the activity.
-This is expected for a system with a scale-free profile
-and confirmed by Table~\ref{autores} of the distribution of activity among participants.
+																																																																																Activity is highly concentrated on the hubs, while a very large number of peripheral vertices contribute to only a fraction of the activity.
+																																																																																This is expected for a system with a scale-free profile, as confirmed with the distribution of activity among participants in Table~\ref{autores}.
 
@@
 																																																																																		of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}.
-The networks often hold this basic structure with as few as 10-50 messages,
-i.e. concentration of activity and the abundance of low-activity participants take place even with very few messages,
-which is highlighted in Section~\ref{si:frac} of the Supporting Information document.
+The networks often hold this basic structure with as few as 10-50 messages, i.e.
+concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in Section~\ref{si:frac} of the Supporting Information.
 A minimum window size for the observation of more general properties might be inferred by monitoring
@@
 
-In order to support hypotheses about the generality of these findings,
-we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr
-in Table~\ref{tab:secE} of the Supporting Information document.
+																																																																																		In order to support the generality of these findings,
+																																																																																		we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr in Table~\ref{tab:secE} of the Supporting Information document.
 The fractions of hubs, intermediary and periphery nodes are
@@
 																																																																																				Fractions of participants derived from degree and strength criteria, $E_1$ and $E_4$ in Section~\ref{sectioning}, are both on the left.
-	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right.
+																																																																																				Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are shown in the plots to the right.
 																																																																																				The ordinates $\overline{e_{n,x}}=\frac{|e_{n,x}|}{N}$ denote the fraction of participants in sector $x$ through criterion $E_n$
 																																																																																				and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
-Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information document hold a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, and present these results for networks from Facebook, Twitter and Participabr.}
+																																																																																			Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and Participabr.}
 																																																																																			\label{fig:sectIL}
@@
 																																																																																			\caption{Distribution of activity among participants.
-The first column presents the percentage of messages sent by the most active participant.
-The column for the first quartile ($Q_1$) exposes 
-the minimum percentage of participants
-responsible for at least 25\% of total messages with the actual number in parenthesis.
-Similarly, the column for the first three quartiles $Q_3$
-gives the minimum percentage of participants responsible for 75\% of total messages.
-The last decile $D_{-1}$ column exposes the maximum percentage of participants responsible for 10\% of messages.}
+																																																																																				The first column shows the percentage of messages sent by the most active participant.
+The column for the first quartile ($Q_1$) gives the minimum percentage of participants responsible for at least 25\% of total messages with the actual number in parentheses.
+Similarly, the column for the first three quartiles $Q_3$ gives the minimum percentage of participants responsible for 75\% of total messages.
+																																																																																				The last decile $D_{-1}$ column shows the maximum percentage of participants responsible for 10\% of messages.}
 																																																																																				\begin{center}
@@
 
-The principal components of the participants are very stable in the topological space,
-i.e. in PCA space of network measures.
-Table~\ref{tab:pcain} exemplifies the formation of principal components
-by exposing averages over non-overlapped activity snapshots of a network.
-The most important result of this application of PCA,
-the stability of principal components,
-is underpinned by the very small dispersion 
-of the contribution of
-each metric to each principal component.
+																																																																																			The principal components of the participants are very stable in the topological space, i.e.
+in PCA space of network measures.
+																																																																																			Table~\ref{tab:pcain} exemplifies the formation of principal components by providing the averages over non-overlapped activity snapshots of a network.
+The most important result of this application of PCA, the stability of principal components, is underpinned by the very small dispersion of the contribution of each metric to each principal component.
 																																																																																			%The contribution of each metric to the
@@
 
-Evident in Table~\ref{tab:pcain} and Figure~\ref{fig:sym},
-dispersion is greater in symmetry-related
-metrics than in clustering coefficient.
-As expected by basic complex network theory, 
-peripheral vertices have low values
-of centrality measures
-and greater dispersion with respect to clustering coefficient.
+																																																																																					According to Table~\ref{tab:pcain} and Figure~\ref{fig:sym},
+																																																																																					dispersion is larger in symmetry-related metrics than in clustering coefficient.
+As expected from basic complex network theory, peripheral vertices have low values of centrality metrics and larger dispersion with regard to the clustering coefficient.
 																																																																																					%The scatter plot in the third system of Figure~\ref{fig:sym},
@@
 																																																																																					This reflects in the relevance of the symmetry-related metrics.
-We conclude that the symmetry metrics are more powerful,
-with respect to dispersion in the topological measures space,
-in characterizing interaction networks
-and their participants, than the clustering coefficient,
-especially for hubs and intermediary vertices.
+																																																																																					We conclude that the symmetry metrics are more powerful, in terms of dispersion in the topological measures space, in characterizing interaction networks and their participants, than the clustering coefficient, especially for hubs and intermediary vertices.
 																																																																																					Interestingly, the clustering coefficient is always combined
@@
 																																																																																					Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
-of the Supporting Information document for other email lists and other interaction networks.
-Larger variability was found among the latter networks,
-which motivated the
-use of interaction networks derived from email lists for benchmarks.
+																																																																																					of the Supporting Information for other email lists and interaction networks.
+A larger variability was found for the latter networks,
+																																																																																					which motivated the use of interaction networks derived from email lists for benchmarking.
 
@@
 
-Assigning a type to a participant raises important 
-issues about the scientific cannon for human types
-and the potential for stigmatization and prejudice. 
-The type of the Erd\"os sector to which a participant belongs 
-can be regarded as implying a social type to the corresponding participant.
-In this case, the type of a participant
-changes both along time and as
-different networks are considered,
-despite the stability of the network,
-and therefore the potential for prejudice of such
-participant typology is attenuated~\cite{adorno}. 
-In other words, an individual is a hub in a number of networks and peripheral in other networks,
-and even within the same network he/she most probably changes type along time~\cite{animacoes}.
-
-The importance of this issue can be grasped by
-the consideration of static types derived from quantitative criteria.
-For example, in email lists with a small number of participants,
-the number of threads has a negative correlation
-with the number of participants.
-When the number of participants exceeds a threshold,
-the number of threads has a positive correlation
-with the number of participants.
+																																																																																					Assigning a type to a participant raises important issues about the scientific cannon for human types and the potential for stigmatization and prejudice.
+The Erd\"os sector to which a participant belongs can be regarded as implying a social type for this participant.
+																																																																																					In this case, the type of a participant changes both along time and as different networks are considered, despite the stability of the network.
+Therefore, the potential for prejudice of such participant typology is attenuated~\cite{adorno}.
+In other words, an individual is a hub in a number of networks and peripheral in other networks, and even within the same network he/she most probably changes type along time~\cite{animacoes}.
+
+																																																																																					The importance of this issue can be grasped by the consideration of static types derived from quantitative criteria.
+For example, in email lists with a small number of participants, the number of threads has a negative correlation with the number of participants.
+																																																																																					When the number of participants exceeds a threshold, the number of threads has a positive correlation with the number of participants.
 																																																																																					This finding is illustrated in Figure~\ref{fig:nmgamma3d}
 																																																																																					and can also be observed in Table~\ref{tab:genLists}.
-The attribution of types to individuals, derived from 
-this analysis of the systems, has more potential for prejudice because
-the derived participant type is static.
-The types, in this latter case, fail to acknowledge that
+																																																																																					The assignment of types to individuals, in this latter case,
+																																																																																					has more potential for prejudice because
+																																																																																					the derived participant type is static and
+																																																																																					one fails to acknowledge that
 																																																																																					human individuals are not immutable entities.
 
-Beyond the results discussed above and in previous sections,
-our main observations regarding the Erd\"os sectors
-and the implicit participant types,
-consistent with the literature~\cite{barabasiEvo},
-are that
-1) hubs and intermediary participants usually have intermittent activity.
-Stable activity was found only in smaller communities.
-For instance, the MET list have stable hubs while all LAU, LAD and CPP
-present intermittent hubs.
+																																																																																					Further observations regarding the Erd\"os sectors
+																																																																																					and the implicit participant types were made, which are consistent with the literature~\cite{barabasiEvo}: 1) hubs and intermediary participants usually have intermittent activity, and stable activity was found only in smaller communities.
+For instance, the MET list had stable hubs while LAU, LAD and CPP exhibited intermittent hubs.
 																																																																																					2) Network structure seems to be most influenced by the
@@
 																																																																																					%%These seem as authorities and contrast with participants that respond much more than receive responses.
-%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
+																																																																																					%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e.
+members know each other often, is found mostly in peripheral and intermediary sectors.
 																																																																																					%\end{itemize}
@@
 																																																																																							This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}.
-Accordingly,
-for $M=20000$ messages, this inflection
-of correlation 
-was found around $N=1500$
-and CPP, LAU, LAD, MET lists 
+																																																																																							Accordingly, for $M=20000$ messages, this inflection
+																																																																																							of correlation was found around $N=1500$, while CPP, LAU, LAD, MET lists 
 																																																																																						present smaller networks.}
@@
 																																																																																					\section{Conclusions}\label{sec:conc}
-The temporal stability reported in this article is very counterintuitive.
-Some of the observed characteristics were expected from current scientific knowledge,
-such as the fat tail structure and the preponderance of centrality measures,
-but the very small standard deviations of principal components formation
+																																																																																					The very small standard deviations of principal components formation
 																																																																																					(see Sections~\ref{sec:pca} and~\ref{prevalence}),
@@
 																																																																																					of the interaction of human individuals.
+
 																																																																																					Furthermore, the importance of symmetry-related metrics,
@@
 																																																																																					undirected graphs in complex networks.
-Noteworthy is the also very stable fraction participants in each Erd\"os sector
-when the network reaches more than 200 participants.
+																																																																																					Noteworthy is also the very stable fraction participants in each Erd\"os sector when the network reaches more than 200 participants.
 																																																																																					Benchmarks were derived from email list networks
@@
 																																																																																					about the generality of these characteristics.
-Selected tables and figures sum dozens of pages
-and are gathered in
-the Supporting Information document.
 
@@
 																																																																																					and these developments might be enhanced to yield scientific knowledge~\cite{opa}.
-Also, we plan to further explore and publish of the visualization and sonification
-processes used for this research~\cite{versinus,animacoes} and
-the extreme linguistic differences found in each of the Erd\"os sectors~\cite{rcText}.
+																																																																																					Also, we plan to further explore and publish the audiovisualizations
+																																																																																					used for this research~\cite{versinus,animacoes} and
+																																																																																					the linguistic differences found in each of the Erd\"os sectors~\cite{rcText}.
 

### 2015-11-03 paper.tex
incorporada revisao do chu

@@
 \item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.}, a Brazilian email list for digital culture.
-	Portuguese is the prevailing language, although some messages are writte in Spanish and English.
+	Portuguese is the prevailing language, although some messages are written in Spanish and English.
 Abbreviated as MET from now on.
@@
 \subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
-																																																																																Principal Component Analysis (PCA) is a well documented technique~\cite{pca}
-																																																																																and was employed in the investigation to acquire knowledge about:
-																																																																																1) which metrics contribute to each principal component and in what proportion;
-																																																																																2) how much of the dispersion is concentrated in each component;
-																																																																																3) expected values and dispersions for these quantities over various networks.
-																																																																																This enables a characterization of the human interaction networks
-																																																																																in terms of the relative importance of networks measures and the way they combine.
+Principal Component Analysis (PCA) is a well documented technique~\cite{pca}, used here to address the following questions:	1) which metrics contribute to each principal component and in what proportion;	2) how much of the dispersion is concentrated in each component;	3) which are the expected values and dispersions for these quantities over various networks.	This enables one to characterize human interaction networks in terms of the relative importance of network metrics and the way they combine.
 
@@
 Figures and tables were usually inspected with 
-																																																																																$ws=\{50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000, 10000\}$ messages.
-Variations in the number of vertices, edges
+$ws=\{50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000, 10000\}$ messages. Variations in the number of vertices, edges
 and other network characteristics, within the same window size $ws$,
@@
 in the scales of seconds, minutes, hours, days and months.
-																																																																																Histograms in each of the time scales were computed as were circular average and dispersion values, and The results are given in Tables~\ref{tab:circ}-\ref{tab:min2}.
+Histograms in each of the time scales were computed as were circular average and dispersion values, and the results are given in Tables~\ref{tab:circ}-\ref{tab:min2}.
 For example, uniform activity is found with respect to seconds, minutes and days of the months.
@@
 This table corresponds to activity in LAU.
-Similar results are shown in Section~\ref{si:months} of the Supporting Information document.}
+Similar results are shown for other lists in Section~\ref{si:months} of the Supporting Information document.}
 \footnotesize
@@
 
-																																																																																			The principal components of the participants are very stable in the topological space, i.e.
-in PCA space of network measures.
+The principal components of the participants are very stable in the topological space, i.e. in PCA space of network measures.
 Table~\ref{tab:pcain} exemplifies the formation of principal components by providing the averages over non-overlapped activity snapshots of a network.
@@
 Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
-																																																																																					of the Supporting Information for other email lists and interaction networks.
-A larger variability was found for the latter networks,
+of the Supporting Information for other email lists and interaction networks. A larger variability was found for the latter networks,
 which motivated the use of interaction networks derived from email lists for benchmarking.
@@
 %%These seem as authorities and contrast with participants that respond much more than receive responses.
-																																																																																					%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e.
-members know each other often, is found mostly in peripheral and intermediary sectors.
+%%	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
 %\end{itemize}

### 2015-11-11 paper.tex
incorporada seção de implicacoes

@@
 
+\subsection{Implications of the main findings}\label{sec:impl}
+The findings reported here arose from an exploratory procedure to inspect visually the networks and obtain considerable amounts of data from snapshots of interaction networks deriving mainly from email lists but also from other networks.
+While this procedure has certainly an ad hoc nature, the statistics in the data are sufficiently robust for important features from these interaction networks to be extracted.  Temporal stability, in the sense that interaction networks could be considered as stationary time series, is the most important feature.
+Also relevant is the significant stability found on the principal components, on the fraction of participants in each Erd\"os Sector and on the activity along different timescales.
+In fact, these findings confirm our initial hypothesis - based on the literature~\cite{newmanBook} - that interaction networks should exhibit some stability traces.
+The potential generality of these findings are suggested by the analysis of networks derived from diverse systems, with interaction networks from public email lists serving as proper benchmarks.
+Indeed, with such benchmarks one can compare any social network system.
+Furthermore, this analysis enables us to establish an outline of human interaction networks.
+It takes the hub, intermediary and periphery sectors out of the scientific folklore and into classes drawn from quantitative criteria.
+It enables the conception of non-static human types derived from natural properties.
+
+We envisage that the knowledge generated in the analysis may be exploited in applications where the type of each participant and the relative proportion of participants in each sector can be useful metadata.
+Just by way of illustration, this could be applied in semantic web initiatives, given that the Erd\"os sectorialization is static in a given snapshot.
+These results are also useful for classifying resources, e.g.
+in social media, and for resources recommendation to users~\cite{opa}.
+Finally, the knowledge acquired with a quantitative treatment of the whole data may help guide creation through collective processes of documents to assist in participatory democracy.
+
+Perhaps the most outreaching implications are related to sociological consequences.
+The results expose a classification of human individuals which is directly related to the concentration of wealth and based on natural laws.
+The derived human typology changes over different systems and over time in the same system, which implies a negation of the absolute concentration of wealth.
+Such concentration exists but changes across different wealth criteria and with time.
+Also, the hubs stand out as dedicated, sometimes enslaved,
+components of the social system.
+The peripheral participants have very limited interaction with the network.
+This suggests that intermediary participants tend to dictate structure, legitimate the hubs and stand out as authorities.
+
+With regard to the limitations of our study, one should emphasize that not all types of human interaction networks were analyzed.
+Therefore, the plausible generalization of properties has to be treated with caution, as a natural tendency of such systems and not as a rule.
+Also, the stable properties in the networks were not explored to the limit, which leaves many open questions.
+For example, what are the maximum and minimum sizes of the networks for which they hold?
+What is the outcome of PCA analysis when more metrics are considered?
+What is the granularity in which the activity along the timescales is preserved?
+Do the findings reported also apply to other systems, beyond human networks?
+
 \section{Conclusions}\label{sec:conc}

### 2015-11-11 paper.tex
terminados retoques sugeridos pelo chu

@@
 In this figure, a snapshot of 1000 messages from CPP list yields the degree distribution of an interaction network of 98 nodes and 235 edges.
-A through explanation of the method is provided in Section~\ref{sectioning}.}
+A thorough explanation of the method is provided in Section~\ref{sectioning}.}
 \label{fig:setores}
@@
 negative values with a high absolute value indicate negative correlation.
+Notice that in this case the variable $k$ is not the degree value
+but a principal component.
+In the results the principal components are numbered
+according to the magnitude of associated eigenvalue and $k$ is incorporated into
+the notation (e.g. PC2 for measures of $\mu_{V'}[j,2]$).
 
@@
 The findings reported here arose from an exploratory procedure to inspect visually the networks and obtain considerable amounts of data from snapshots of interaction networks deriving mainly from email lists but also from other networks.
-While this procedure has certainly an ad hoc nature, the statistics in the data are sufficiently robust for important features from these interaction networks to be extracted.  Temporal stability, in the sense that interaction networks could be considered as stationary time series, is the most important feature.
+While this procedure has certainly an ad hoc nature, the statistics in the data are sufficiently robust for important features from these interaction networks to be extracted.
+Temporal stability, in the sense that interaction networks could be considered as stationary time series, is the most important feature.
 Also relevant is the significant stability found on the principal components, on the fraction of participants in each Erd\"os Sector and on the activity along different timescales.

### 2016-08-01 paper.tex
adicionada tabela com as unidades dos dados triplificados
moved paulo mais to acknowledgements
minor revisions

@@
 %
-\author{Leonardo Paulo Maia}
-\homepage{http://www.ifsc.usp.br/~lpmaia/}
-\email{lpmaia@ifsc.usp.br }
-\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+% \author{Leonardo Paulo Maia}
+% \homepage{http://www.ifsc.usp.br/~lpmaia/}
+% \email{lpmaia@ifsc.usp.br }
+% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
 
@@
 \subsection{Implications of the main findings}\label{sec:impl}
-The findings reported here arose from an exploratory procedure to inspect visually the networks and obtain considerable amounts of data from snapshots of interaction networks deriving mainly from email lists but also from other networks.
+The findings reported here arose from an exploratory procedure to inspect visually the networks and obtain considerable amounts of interaction networks data deriving from email lists and also from other networks.
 While this procedure has certainly an ad hoc nature, the statistics in the data are sufficiently robust for important features from these interaction networks to be extracted.
@@
 The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Presidency of the Brazilian Republic for keeping Participabr code and data open.
-We are also grateful to developers and users of Python scientific tools.
+We are also grateful to developers and users of Python scientific tools
+and to Leonardo Paulo Maia (IFSC/USP) for valuable insights.
 \end{acknowledgments}

### 2016-08-02 paper.tex
finisehd a review of the article text

@@
 This paper reports on stable properties of human interaction networks,
-with benchmarks derived from public email lists. Activity along time and topology evolution were observed in snapshots in a timeline, 
+with benchmarks derived from public email lists. Activity along time and topology were observed in snapshots in a timeline, 
 and at different scales.
@@
 \pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data, typology}
+\keywords{complex networks, pattern recognition, statistics, social network analysis, typology}
+% \keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data, typology}
 \maketitle
@@
 
-\subsection{Time activity statistics}\label{sec:mtime}
+\subsection{Temporal activity statistics}\label{sec:mtime}
 Messages were counted over time as histograms in the scales of seconds,
@@
 such as the recurrent time periods that we are interested in.
-Equivalent measures were taken using circular statistics,
+Similar measures were taken using circular statistics,
 in which each measurement $t$ is represented as a unit complex number,
@@
 served as a further clue of how close the distribution was to being uniform. As expected, a positive correlation was found in all $r, Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures,
-which can be noticed in Section~\ref{si:circ} of the Supporting Information. The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred in the discussion of results.
+which can be noticed in Section~\ref*{si:circ} of the Supporting Information. The circular dispersion $\delta(z)$ was found more sensitive and therefore preferred in the discussion of results.
 
@@
 The topology of the networks was characterized 
-from a small selection of the most basic and 
-fundamental measurements for each vertex, as follows:
+from a small selection of the most basic 
+and fundamental measurements for each vertex, as follows:
 
@@
 \item Asymmetry of vertex $i$: $asy_i=\frac{k_i^{in}-k_i^{out}}{k_i}$.
-\item Average asymmetry of edges at vertex $i$: $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{ij}$ is 1 if there is an edge from $i$ to $j$, and $0$ otherwise, and $J_i$ is the set of neighbors of vertex $i$.
-\item Standard deviation of asymmetry of edges: $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$.
+\item Average asymmetry of edges at vertex $i$:\\ $\mu_i^{asy}=\frac{\sum_{j\in J_i} e_{ji}-e_{ij}}{|J_i|}$, where $e_{ij}$ is 1 if there is an edge from $i$ to $j$, and $0$ otherwise, and $J_i$ is the set of neighbors of vertex $i$.
+\item Standard deviation of asymmetry of edges:\\ $\sigma_i^{asy}=\sqrt{\frac{\sum_{j\in J_i}[\mu^{asy}_i -(e_{ji}-e_{ij}) ]^2  }{|J_i|}  }$.
 \item Disequilibrium: $dis_i=\frac{s_i^{in}-s_i^{out}}{s_i}$.
-\item Average disequilibrium of edges: $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{w_{ji}-w_{ij}}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
+\item Average disequilibrium of edges:\\ $\mu_i^{dis}=\frac{\sum_{j \in J_i}\frac{w_{ji}-w_{ij}}{w_{ji}+w_{ij}}}{|J_i|}$, where $w_{xy}$ is the weight of edge $x\rightarrow y$ and zero if there is no such edge.
 \item Standard deviation of disequilibrium of edges: $\sigma_i^{dis}=\sqrt{\frac{\sum_{j\in J_i}\left[\mu^{dis}_i-\frac{w_{ji}-w_{ij}}{w_{ji}+w_{ij}}\right]^2}{|J_i|}}$.
@@
 
-The degree distribution $\widetilde{P}(k)$ of a scale-free network $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
+The degree distribution $\widetilde{P}(k)$ of a real network with a scale-free profile $\mathcal{N}_f(N,z)$ with $N$ vertices and $z$ edges has less
 average degree nodes than the distribution $P(k)$ of an Erd\"os-R\'enyi
@@
 $\widetilde{P}(k)>P(k)$ and $k$ is higher than any value of $k$ in the intermediary sector.
-The reasoning for this classification is as follows: vertices so connected that they are virtually inexistent in networks connected without distinction of the vertices, i.e.
-without preferential attachment and as in the Erd\"os-R\'enyi model, are correctly associated to the hub sector.
-Vertices with very few connections, which are way more abundant than expected in the Erd\"os-R\'enyi model, are assigned to the periphery.
-Vertices with degree values predicted as the most abundant in a Erd\"os-R\'enyi model, near the average, and less frequent in the real network, are classified as intermediary.
-
+The reasoning for this classification is as follows: vertices so connected that they are virtually nonexistent in the Erd\"os-R\'enyi model, are coherently associated to the hub sector.
+Vertices with very few connections, which are way more abundant than expected in the Erd\"os-R\'enyi model,
+are assigned to the periphery.
+Vertices with degree values predicted as the most abundant in the Erd\"os-R\'enyi model,
+near the average, and less frequent in the real network, are classified as intermediary.
+
+\clearpage
 \begin{figure}[!h]
@@
 To ensure statistical validity of the histograms, bins can be chosen to contain at least $\eta$ vertices of the real network.
-The range $\Delta$ of incident values should be partitioned in $m$ parts $\Delta=\cup_{i=1}^m \Delta_i$,
+The range $\Delta$ of incident values of degree $k$ should be partitioned in $m$ parts $\Delta=\cup_{i=1}^m \Delta_i$,
 with $\Delta_i\cap \Delta_j=\emptyset \; \forall\; i \neq j$ and:
@@
 \begin{align}
-\Delta_i =\Biggl\{ k\;\; & | & \overline{\Delta}_{i-1}< &\, k\leq j \text{ and }\;\;\;\;\;\;\;\;\;\;\;\;\nonumber\\
-& \Biggl[ & N - \sum_{k=0}^{\overline{\Delta}_{i-1}} & \eta_k< \eta \text{ or }\\
-&	\Bigl[ & \sum_{k=\overline{\Delta}_{i-1}+1}^j &  \eta_k \geq \eta \text{ and }\;\;\;\;\;\;\nonumber\\
-& \Bigl( &\sum_{k=\overline{\Delta}_{i-1}+1}^{j-1} &  \eta_k < \eta \text{ or }   j=\overline{\Delta}_{i-1}+1 \Bigr) \;\Bigr] \Biggr] \Biggr\}\nonumber
+\Delta_i =\Biggl\{ k\;\; & | & \overline{\Delta}_{i-1}< &\, k\leq l \text{ and }\;\;\;\;\;\;\;\;\;\;\;\;\nonumber\\
+                         & \Biggl[ \Bigl[ & N - \sum_{k=0}^{\overline{\Delta}_{i-1}} & \eta_k< \eta \text{ and } l = \overline{\Delta} \Bigr] \text{ or }\\
+&	\Bigl[ & \sum_{k=\overline{\Delta}_{i-1}+1}^l &  \eta_k \geq \eta \text{ and }\;\;\;\;\;\;\nonumber\\
+& \Bigl( &\sum_{k=\overline{\Delta}_{i-1}+1}^{l-1} &  \eta_k < \eta \text{ or }   l=\overline{\Delta}_{i-1}+1 \Bigr) \;\Bigr] \Biggr] \Biggr\}\nonumber
 \end{align}
@@
 
-\noindent where $\eta_k$ is the number of vertices with degree $k$, while $\overline{\Delta}_{i}=max(\Delta_{i})$, and $\overline{\Delta}_{0}=-1$.
+\noindent where $\eta_k$ is the number of vertices with degree $k$,
+while $\overline{\Delta}_{(i)}=max(\Delta_{(i)})$, and $\overline{\Delta}_{0}=-1$.
+% and $\overline{\Delta}_{i}<l\leq max(\Delta)$.
 Equation~\ref{criterio} can now be written in the form:
@@
 
-If the strength $s$ is used for comparison, $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ is the strength of vertex $i$.
+If the strength $s$ is used for comparison of the real network against the Erdös-Rényi model,
+$P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ is the strength of vertex $i$.
 For in and out degrees ($k^{in}$, $k^{out}$), the real network should be compared against
@@
 e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
-e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber,
+e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}>\overline{k}_{\gamma,R}\} \nonumber,
 \end{alignat}
@@
 
-\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.
+\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above and exposed in Figure~\ref{fig:setores}.
 
 Since different metrics can be used to identify the three types of vertices, more than one metric can be used simultaneously, which is convenient when analysing small networks,
-such as the cases where $ws=50$ in Section~\ref{si:frac} of the Supporting Information.
+such as the cases where only 50 messages are considered in the Section~\ref*{si:frac} of the Supporting Information.
 %For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
@@
 Let $\mathbf{X}=\{X[i,j]\}$ be a matrix where each element is the value
-of each metric $j$ at vertex $i$ .
+of the metric $j$ at vertex $i$ .
 Let
@@
 %was considered with different shifts in the message timeline to obtain snapshots.
-The snapshots are disjoint in the message timeline, and were used to perform both PCA with topological metrics and Erd\"os sectioning.  
+The snapshots were made disjoint in the message timeline, and were used to perform both PCA with topological metrics and Erd\"os sectioning.  
 Figures and tables were usually inspected with 
@@
 and other network characteristics, within the same window size $ws$,
-are given in Section~\ref{si:frac} of the Supporting Information document.
+are given in Section~\ref*{si:frac} of the Supporting Information document.
 
 Network structures were mapped to video animations, sound and musical structures, image galleries and online gadgets developed for this research~\cite{animacoes,galGmane,appGmane}.
-Such \emph{audiovisualizations} were crucial in the initial steps, and
+Such \emph{audiovisualizations} were crucial in the initial steps and
 to guide the research into the most important features of network evolution.
@@
 \caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$, described in Section~\ref{sec:mtime}, for different timescales.
-This example table was constructed using all LAD messages, and the results are the same for other lists, as shown in Section~\ref{si:circ} of the Supporting Information document.
+This example table was constructed using all LAD messages, and the results are the same for other lists, as shown in Section~\ref*{si:circ} of the Supporting Information document.
 The most uniform distribution of activity was found in seconds and minutes.
@@
 \caption{Activity percentages along the hours of the day.
-Nearly identical distributions were observed on other social systems as shown in Section~\ref{si:hours} of the Supporting Information document.
+Nearly identical distributions were observed on other social systems as shown in Section~\ref*{si:hours} of the Supporting Information document.
 Highest activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight.
@@
 with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.8.2, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
-In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks $\xi=\frac{min(incidence')}{max(incidence')} \Rightarrow \mu_\xi=0.7741 \text{ and } \sigma_\xi=0.02619$.
+In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks, i.e.
+if $\xi=\frac{min(incidence')}{max(incidence')}$ than $\mu_\xi=0.7741 \text{ and } \sigma_\xi=0.02619$.
 Therefore, the incidence of messages at each second of a minute and at each minute of an hour was considered uniform.
@@
 Nearly identical distributions are found in all systems
-as indicated in Section~\ref{si:monthdays} of the Supporting Information.
+as indicated in Section~\ref*{si:monthdays} of the Supporting Information.
 Although slightly higher activity rates are found in the beginning of the month, the most important feature seems to be the homogeneity made explicit by the high circular dispersion in Table~\ref{tab:circ}.
@@
 This table corresponds to activity in LAU.
-Similar results are shown for other lists in Section~\ref{si:months} of the Supporting Information document.}
+Similar results are shown for other lists in Section~\ref*{si:months} of the Supporting Information document.}
 \footnotesize
@@
 
-The distribution of vertices in the hub, intermediary, periphery Erd\"os sectors is remarkably stable along time if the snapshots hold 200 or more messages, as it is clear in Figure~\ref{fig:sectIL} and in Section~\ref{si:frac} of the Supporting Information document.
+The distribution of vertices in the hub, intermediary, periphery Erd\"os sectors is remarkably stable along time if the snapshots hold 200 or more messages, as it is clear in Figure~\ref{fig:sectIL} and in Section~\ref*{si:frac} of the Supporting Information document.
 %Moreover, all email lists analyzed exhibit the same distribution profile.
@@
 
-Typically, $[3\%-12\%]$ of the vertices are hubs,
+Typically, $[3\%-12\%]$ of the vertices are hubs,\\
 $[15\%-45\%]$ are intermediary and $[44\%-81\%]$ are peripheral,
@@
 These results hold for the total, in and out degrees and strengths.
-Stable sizes are also observed for 100 or less messages if classification 
+Stable sizes are also observed for 100 or less messages if the classification 
 of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}.
 The networks often hold this basic structure with as few as 10-50 messages, i.e.
-concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in Section~\ref{si:frac} of the Supporting Information.
+concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in Section~\ref*{si:frac} of the Supporting Information.
 A minimum window size for the observation of more general properties might be inferred by monitoring
@@
 In order to support the generality of these findings,
-we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr in Table~\ref{tab:secE} of the Supporting Information document.
+we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr in Table~\ref*{tab:secE} of the Supporting Information document.
 The fractions of hubs, intermediary and periphery nodes are
@@
 and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
-Sections~\ref{si:frac} and~\ref{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and Participabr.}
+Sections~\ref*{si:frac} and~\ref*{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and Participabr.}
 \label{fig:sectIL}
@@
 
-The principal components of the participants are very stable in the topological space, i.e. in PCA space of network measures.
+The principal components of the participants are very stable in the topological space, i.e. in the space of principal components of network measures.
 Table~\ref{tab:pcain} exemplifies the formation of principal components by providing the averages over non-overlapped activity snapshots of a network.
@@
 and their equal relevance for variability,
-as measured by dispersion, is noticeable.
+as measured by the principal component, is noticeable.
 Also, this suggests that these centrality metrics 
@@
 dispersion is larger in symmetry-related metrics than in clustering coefficient.
-As expected from basic complex network theory, peripheral vertices have low values of centrality metrics and larger dispersion with regard to the clustering coefficient.
-%The scatter plot in the third system of Figure~\ref{fig:sym},
-%where all metrics are considered and there is a greater dispersion
-%with respect to the ordinates,
-This reflects in the relevance of the symmetry-related metrics.
-We conclude that the symmetry metrics are more powerful, in terms of dispersion in the topological measures space, in characterizing interaction networks and their participants, than the clustering coefficient, especially for hubs and intermediary vertices.
+% As expected from basic complex network theory, peripheral vertices have low values of centrality metrics and larger dispersion with regard to the clustering coefficient.
+% %The scatter plot in the third system of Figure~\ref{fig:sym},
+% %where all metrics are considered and there is a greater dispersion
+% %with respect to the ordinates,
+% This reflects in the relevance of the symmetry-related metrics.
+We conclude that the symmetry metrics are more powerful, in terms of dispersion in the topological measures space, in characterizing interaction networks and their participants, than the clustering coefficient, especially for hubs and intermediary vertices (peripheral vertices have larger dispersion with regard to the clustering coefficient).
 Interestingly, the clustering coefficient is always combined
@@
 %in Section~\ref{si:ext} of the Supporting Information document.
-Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
+Similar results are presented in Sections~\ref*{si:pcat} and~\ref*{si:ext}
 of the Supporting Information for other email lists and interaction networks. A larger variability was found for the latter networks,
@@
 \subsection{Implications of the main findings}\label{sec:impl}
-The findings reported here arose from an exploratory procedure to inspect visually the networks and obtain considerable amounts of interaction networks data deriving from email lists and also from other networks.
+The findings reported in this article arose from an exploratory procedure to visually inspect the networks and to analyze considerable amounts of interaction networks data.
+% deriving from email lists and also from other networks.
 While this procedure has certainly an ad hoc nature, the statistics in the data are sufficiently robust for important features from these interaction networks to be extracted.
@@
 in social media, and for resources recommendation to users~\cite{opa}.
-Finally, the knowledge acquired with a quantitative treatment of the whole data may help guide creation through collective processes of documents to assist in participatory democracy.
+Finally, the knowledge acquired with a quantitative treatment of the whole data may help guide the creation through collective processes of documents to assist in participatory democracy.
 
@@
 networks from Facebook,
-Twitter and Participabr might ease hypothesizing
+Twitter and Participabr in the Supporting Information might ease hypothesizing
 about the generality of these characteristics.

### 2016-08-02 paper.tex
finished a review of the figures and tables

@@
 \altaffiliation{
-%Centro de CiÃƒÂªncias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
+Centro de Ci\^encias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
 }%Lines break automatically or can be forced with \\
@@
 \centering
-\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000_}
+\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000__}
 \caption{Stability of Erd\"os sector sizes.
-Fractions of participants derived from degree and strength criteria, $E_1$ and $E_4$ in Section~\ref{sectioning}, are both on the left.
+Fractions of participants derived from degree and strength criteria, $E_1$ and $E_4$ described in Section~\ref{sectioning}, are both on the left.
 Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are shown in the plots to the right.
-The ordinates $\overline{e_{n,x}}=\frac{|e_{n,x}|}{N}$ denote the fraction of participants in sector $x$ through criterion $E_n$
-and, similarly, $\overline{c_{n,x}}=\frac{|c_{n,x}|}{N}$ denotes the fraction of participants in sector $x$ through criterion $C_n$.
+The ordinates $\overline{e_{\gamma,\phi}}=\frac{|e_{\gamma,\phi}|}{N}$ denote the fraction of participants in sector $\phi$ through criterion $E_\gamma$
+and, similarly, $\overline{c_{\delta,\phi}}=\frac{|c_{\delta,\phi}|}{N}$ denotes the fraction of participants in sector $\phi$ through criterion $C_\delta$.
 Sections~\ref*{si:frac} and~\ref*{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and Participabr.}
@@
 The first column shows the percentage of messages sent by the most active participant.
-The column for the first quartile ($Q_1$) gives the minimum percentage of participants responsible for at least 25\% of total messages with the actual number in parentheses.
+The column for the first quartile ($Q_1$) gives the minimum percentage of participants responsible for at least 25\% of total messages with the actual percentage in parentheses.
 Similarly, the column for the first three quartiles $Q_3$ gives the minimum percentage of participants responsible for 75\% of total messages.
@@
 \begin{table}[!h]
-\caption{Loadings for the 14 metrics into the principal components for the MET list, $ws=1000$ messages in 20 disjoint positioning.
-The clustering coefficient (cc) appears as the first metric in the Table, followed by 7 centrality metrics and 6 symmetry-related metrics.
+\caption{Loadings for the 14 metrics into the principal components for the MET list, $1000$ messages in 20 disjoint positions.
+The clustering coefficient (cc) appears as the first metric in the table, followed by 7 centrality metrics and 6 symmetry-related metrics.
 Note that the centrality measurements, including degrees, strength and betweenness centrality, are the most important contributors for the first principal component, while the second component is dominated by symmetry metrics.
 The clustering coefficient is only relevant for the third principal component.
-The three components have in average more than 85\% of the variance.}
+The three components have in average more than 85\% of the variance.
+The low standard deviation $\sigma$ implies that the principal components are considerably stable.}
 \footnotesize
@@
 \caption{The first plot exposes the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
-The second plot shows the greater dispersion of the symmetry-related ordinates dominant in PC2.
+    The second plot shows the greater dispersion of the symmetry-related ordinates dominant in the second principal component (PC2).
 This greater dispersion suggests that symmetry-related metrics are more powerful,
@@
 \centering
-\includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma2}
+\includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma2_}
 \caption{A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.
-Highest $\Gamma$ are associated with low $N$.
+Highest $\Gamma$ is associated with low $N$.
 The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise.

### 2016-08-12 paper.tex
minor correction
finished review of article with new tables

@@
 Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
-These properties are consistent with expectations from the literature and may be general for human interaction networks, which has important implications for establishing a typology of participants based on quantitative criteria.
+These properties are consistent with the literature and may be general for human interaction networks, which has important implications for establishing a typology of participants based on quantitative criteria.
 \end{abstract}
@@
 \section{Introduction}\label{sec:into}
-The first studies dealing with human interaction networks explicitly
+The first studies dealing explicitly with human interaction networks
 date from the nineteenth century while the foundation of
@@
 This is not an intuitive result, given that participants constantly transition in network structure.
-Because these properties were shared by networks from various sources, and are consistent with the literature in complex networks, we advocate that the conclusions might be valid for general classes of interaction networks.
+Because these properties were shared by networks from various sources, and are consistent with the literature in complex networks~\cite{newmanBook}, we advocate that the conclusions might be valid for general classes of interaction networks.
 In particular, this allows us to
@@
 Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}
-but is not relevant for the present manuscript because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+but is not relevant for the present work because email messages and addresses have no gender related metadata~\cite{gmanePack}.
 
@@
 
-We analyzed many email lists together with data from Twitter, Facebook and Participabr and selected four email lists for a thorough analysis,
+We observed various email lists and selected four of them together with data from Twitter, Facebook and Participabr for a thorough analysis,
 from which general properties can be inferred. These lists are as follows:
@@
 list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
-\input{tables/tab1Geral}
+\input{tables/tab1Overview}
 \end{tabular}
@@
 such as the recurrent time periods that we are interested in.
-Similar measures were taken using circular statistics,
+Similar measures were taken using circular statistics~\cite{directionalStats},
 in which each measurement $t$ is represented as a unit complex number,
@@
 \centering
-\includegraphics[width=0.44\textwidth]{figs/criaRede2_}
+\includegraphics[width=0.29\textwidth]{figs/criaRede3_}
 \caption{The formation of interaction networks from exchanged messages.
@@
 from a small selection of the most basic 
-and fundamental measurements for each vertex, as follows:
+and fundamental measurements for each vertex~\cite{newmanBook}, as follows:
 
@@
 
-\subsection{Stability of principal components and the prevalence of symmetry over clusterization}\label{prevalence}
+\subsection{Stability of principal components}\label{prevalence}
 %The topology was analyzed using standard, well-established metrics of centrality and clustering.
@@
 especially for hubs and intermediary vertices.
-This figure was reflects a snapshot of the LAU list with $ws= 1000$.}
+This figure reflects a snapshot of the LAU list with 1000 contiguous messages.}
 
@@
 In fact, these findings confirm our initial hypothesis - based on the literature~\cite{newmanBook} - that interaction networks should exhibit some stability traces.
-The potential generality of these findings are suggested by the analysis of networks derived from diverse systems, with interaction networks from public email lists serving as proper benchmarks.
+The potential generality of these findings is suggested by the analysis of networks derived from diverse systems, with interaction networks from public email lists serving as proper benchmarks.
 Indeed, with such benchmarks one can compare any social network system.
@@
 The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Presidency of the Brazilian Republic for keeping Participabr code and data open.
-We are also grateful to developers and users of Python scientific tools
-and to Leonardo Paulo Maia (IFSC/USP) for valuable insights.
+We are also grateful to developers and users of Python scientific tools, to Leonardo Paulo Maia (IFSC/USP) and to Francisco J. P. Lopes (UFRJ) for valuable insights.
 \end{acknowledgments}

### 2016-08-13 paper.tex
minor adds

@@
 Email list messages were obtained from
-the Gmane email archive~\cite{gmanePack}, which consists of more than $20,000$
+the Gmane email archive, which consists of more than $20,000$
 email lists (discussion groups) and more than $130\times 10^6$ messages~\cite{Gmanewikipedia}.
@@
 
-Network structures were mapped to video animations, sound and musical structures, image galleries and online gadgets developed for this research~\cite{animacoes,galGmane,appGmane}.
+Network structures were mapped to video animations, sound and musical structures developed for this research~\cite{animacoes}.% ,galGmane,appGmane}.
 Such \emph{audiovisualizations} were crucial in the initial steps and
@@
 In the scales of seconds and minutes, activity is uniform,
-with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.8.2, ``random.randint'' function, was used for simulations, algorithms in \url{https://pypi.python.org/pypi/gmane}.}.
+with the messages being slightly more evenly distributed in all lists than in simulations with the uniform distribution\footnote{Numpy version 1.8.2, ``random.randint'' function, was used for simulations, algorithms in \url{https://github.com/ttm/percolation}.}.
 In the networks, $\frac{min(incidence)}{max(incidence)} \in (0.784,.794)$ while simulations reach these values but have on average more discrepant higher and lower peaks, i.e.

### 2016-08-16 paper.tex
invariance and stability combined

@@
 \begin{abstract}
-This paper reports on stable properties of human interaction networks,
+    This paper reports on stable (or invariant) properties of human interaction networks,
 with benchmarks derived from public email lists. Activity along time and topology were observed in snapshots in a timeline, 
@@
 Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and Participabr were used for the sake of generalization.
-Using a timeline of activity snapshots with a constant number of contiguous messages, we found remarkable stability for important network properties.
+Using a timeline of activity snapshots with a constant number of contiguous messages, we found remarkable stability (or invariance) for important network properties.
 For instance, activity along different timescales follows specific patterns; the most basic topological metrics can always be combined into characteristic principal components; and the fractions of participants in different sectors do not vary with time.

### 2016-08-17 paper.tex
arrumado do clustering e simetria

@@
 with benchmarks derived from public email lists. Activity along time and topology were observed in snapshots in a timeline, 
-and at different scales.
-Our analysis shows that activity is practically the same for all networks across timescales ranging from seconds to months.
-The most important metrics to the dispersion of participants in the topological measures space are related to centrality (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient.
+and at different scales. Our analysis shows that activity is practically the same for all networks across timescales ranging from seconds to months.
+% The most important metrics to the dispersion of participants in the topological measures space are related to centrality (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient.
+The principal components of the participants in the topological measures space remain practically unchanged as different sets of messages are considered.
 The activity of participants follows the expected scale-free trace, thus yielding the hub, intermediary and peripheral classes of vertices by comparison against the Erd\"os-R\'enyi model.
@@
 with the standard deviation of the asymmetry and disequilibrium
-of edges $\sigma^{asy}$ and $\sigma^{dis}$.
+of edges $\sigma^{asy}$ and $\sigma^{dis}$ in the third principal component.
 

### 2016-08-17 paper.tex
incorporada revisao do Chu

@@
 % The most important metrics to the dispersion of participants in the topological measures space are related to centrality (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient.
-The principal components of the participants in the topological measures space remain practically unchanged as different sets of messages are considered.
+The principal components of the participants in the topological metrics space remain practically unchanged as different sets of messages are considered.
 The activity of participants follows the expected scale-free trace, thus yielding the hub, intermediary and peripheral classes of vertices by comparison against the Erd\"os-R\'enyi model.
@@
 Gender related behavior in mobile phone datasets was indeed reported~\cite{barabasiSex}
-but is not relevant for the present work because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+but it is not relevant for the present work because email messages and addresses have no gender related metadata~\cite{gmanePack}.
 
@@
 
-If the strength $s$ is used for comparison of the real network against the Erdös-Rényi model,
+If the strength $s$ is used for comparison of the real network against the Erd\"os-R\'enyi model,
 $P$ remains the same, but $P(\kappa_i)$ with $\kappa_i=\frac{s_i}{\overline{w}}$ should be used, where $\overline{w}=2\frac{z}{\sum_is_i}$ is the average weight of an edge and $s_i$ is the strength of vertex $i$.
@@
 
-\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above and exposed in Figure~\ref{fig:setores}.
+\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above and illustrated in Figure~\ref{fig:setores}.
 
 Since different metrics can be used to identify the three types of vertices, more than one metric can be used simultaneously, which is convenient when analysing small networks,
-such as the cases where only 50 messages are considered in the Section~\ref*{si:frac} of the Supporting Information.
+such as the cases where only 50 messages are considered in Section~\ref*{si:frac} of the Supporting Information.
 %For example, a very stringent criterion can be used, according to which a vertex is only regarded as pertaining to a sector if it is so for all the metrics.
@@
 of $\mathbf{X'}$, one eigenvector per column.
-Each eigenvector combines the original measures into one principal component, therefore
+Each eigenvector combines the original metrics into one principal component, therefore
 $V'[j,k]=100\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
@@
 is responsible for.
-We consider, in general, the three greatest eigenvalues and
+We consider, in general, the three largest eigenvalues and
 the respective eigenvectors in percentages:
@@
 according to the magnitude of associated eigenvalue and $k$ is incorporated into
-the notation (e.g. PC2 for measures of $\mu_{V'}[j,2]$).
+the notation (e.g. PC2 for metrics of $\mu_{V'}[j,2]$).
 
@@
 For example, uniform activity is found with respect to seconds, minutes and days of the months.
-Weekend days exhibit about half the activity of regular weekdays, and there is a peak of activity between 11am and 12pm.
+Weekend days exhibit about half the activity of regular weekdays, and there is a peak of activity between 11am and noon.
 
@@
 \includegraphics[width=.45\textwidth]{figs/im13PCAPLOT__}
-\caption{The first plot exposes the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
+\caption{The first plot highlights the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
     The second plot shows the greater dispersion of the symmetry-related ordinates dominant in the second principal component (PC2).
-This greater dispersion suggests that symmetry-related metrics are more powerful,
+This larger dispersion suggests that symmetry-related metrics are more powerful,
 for characterizing interaction networks than the clustering coefficient,
@@
 % This reflects in the relevance of the symmetry-related metrics.
-We conclude that the symmetry metrics are more powerful, in terms of dispersion in the topological measures space, in characterizing interaction networks and their participants, than the clustering coefficient, especially for hubs and intermediary vertices (peripheral vertices have larger dispersion with regard to the clustering coefficient).
+We conclude that the symmetry metrics are more powerful, in terms of dispersion in the topological metrics space, in characterizing interaction networks and their participants, than the clustering coefficient, especially for hubs and intermediary vertices (peripheral vertices have larger dispersion with regard to the clustering coefficient).
 Interestingly, the clustering coefficient is always combined
@@
 and the recurrent activity patterns along different timescales (see Sections~\ref{sec:mtime} and~\ref{constDisc}),
-goes a step further in characterizing scale-free networks in the context
+go a step further in characterizing scale-free networks in the context
 of the interaction of human individuals.
-
 Furthermore, the importance of symmetry-related metrics,

### 2017-03-15 paper.tex
added figure from backup
arrumado referencia da deborah
incorporada revisao do physica A

@@
 \altaffiliation{
-Curso de Psicologia, Universidade Federal do Cear\'a (UFC)
+Curso de Psicologia, Universidade Federal do Cear\'a (UFC), Campus Sobral
 }%Lines break automatically or can be forced with \\
@@
     This paper reports on stable (or invariant) properties of human interaction networks,
-with benchmarks derived from public email lists. Activity along time and topology were observed in snapshots in a timeline, 
+with benchmarks derived from public email lists. Activity, recognized through messages sent,
+	along time and topology were observed in snapshots in a timeline, 
 and at different scales. Our analysis shows that activity is practically the same for all networks across timescales ranging from seconds to months.
@@
 The relative sizes of these three sectors are essentially the same for all email lists and the same along time.
-Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and 44-81\% are peripheral vertices.
+Typically, <15\% of the vertices are hubs, 15-45\% are intermediary and >45\% are peripheral vertices.
 Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
@@
 In this paper we analyze the evolution of human interaction networks.
+Directed and weighted representations were built through the observation of replies as links.
 Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and Participabr were used for the sake of generalization.
@@
 \subsection{Related work}\label{sec:related}
-Research on network evolution is often restricted to network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}.
-Network types have been discussed with regard to the number of participants, intermittence of their activity and network longevity~\cite{barabasiEvo}.
-Two topologically different networks emerged from human interaction networks, depending on whether the frequency of interactions follows a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
-In email list networks, scale-free properties were reported with $\alpha \approx 1.8$~\cite{bird} (as in web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{Gmane2}.
-
 The fact that unreciprocated edges often exceed 50\% in human interaction networks~\cite{newmanEvolving} motivated the inclusion of symmetry metrics in our analysis.
@@
 but it is not relevant for the present work because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+
+Research on network evolution is often restricted to network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}.
+Network types have been discussed with regard to the number of participants, intermittence of their activity and network longevity~\cite{barabasiEvo}.
+Two topologically different networks emerged from human interaction networks, depending on whether the frequency of interactions follows a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}.
+In email list networks, scale-free properties were reported with $\alpha \approx 1.8$~\cite{bird} (as in web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{Gmane2}.
 
@@
 \end{table}
-\subsection{Availability}
 The data and scripts used to derive the results, figures and tables, and this article itself are publicly available.

### 2017-03-15 paper.tex
arrumado os simbolos < e > no abstract
feita versao do artigo com marcacao das alteracoes

@@
 \usepackage{amsmath}
+\usepackage{soul}
 \hypersetup{
@@
     This paper reports on stable (or invariant) properties of human interaction networks,
-with benchmarks derived from public email lists. Activity, recognized through messages sent,
+	with benchmarks derived from public email lists. Activity\hl{, recognized through messages sent,}
 	along time and topology were observed in snapshots in a timeline, 
@@
 The relative sizes of these three sectors are essentially the same for all email lists and the same along time.
-Typically, <15\% of the vertices are hubs, 15-45\% are intermediary and >45\% are peripheral vertices.
+	Typically, \hl{$<15\%$ of the vertices are hubs, 15-45\% are intermediary and $>45\%$ are peripheral vertices}.
 Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
@@
 In this paper we analyze the evolution of human interaction networks.
-Directed and weighted representations were built through the observation of replies as links.
+\hl{Directed and weighted representations were built through the observation of replies as links.}
 Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and Participabr were used for the sake of generalization.
@@
 but it is not relevant for the present work because email messages and addresses have no gender related metadata~\cite{gmanePack}.
+
+\hl{These two paragraphs were switched.}
 
@@
 \end{table}
+
+\hl{"Availability" subsection heading removed}
 The data and scripts used to derive the results, figures and tables, and this article itself are publicly available.

### 2017-03-17 physicaa/elsarticle-template.tex
adicionados arquivos para o physica a

--- 2017-03-15 paper.tex
@@
-% ****** Start of file aipsamp.tex ******
-%
-%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
-%   Version 4.1 of REVTeX, October 2009
-%
-%   Copyright (c) 2009 American Institute of Physics.
+\documentclass[review]{elsarticle}
 
-% Use this file as a source of example code for your aip document.
-% Use the file aiptemplate.tex as a template for your document.
-\documentclass[%
-aip,
-jmp,%
-amsmath,amssymb,
-%preprint,%
-reprint,%
-%author-year,%
-%author-numerical,%
-]{revtex4-1}
-\usepackage{graphicx}% Include figure files
-\usepackage{grffile}
-\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
-%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
-%\linenumbers\relax % Commence numbering lines
+\usepackage{lineno,hyperref}
+\usepackage{amsmath}
 \usepackage{multirow}
-\usepackage{color} % for the notes
-\usepackage{etex}
-\reserveinserts{58}
-%\usepackage{morefloats}
-\usepackage{hyperref}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\hypersetup{
-colorlinks,
-linkcolor={red!50!black},
-citecolor={blue!50!black},
-urlcolor={blue!80!black}
-}
+\usepackage{placeins}
+\modulolinenumbers[5]
+
+\journal{Physica A: Statistical Mechanics and its Applications}
 
 \usepackage{xr}
-\externaldocument{supportingInformation}
-\maxdeadcycles=1000
+\externaldocument{../supportingInformation}
 
-\usepackage{placeins}
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
 
-\preprint{XXXXX (preprint)}
+\begin{frontmatter}
 
-\title[Stability of interaction networks]{Temporal stability in human interaction networks}
-% : sector sizes, topological prominence and activity along diverse timescales}% Force line breaks with \\
-\author{Renato Fabbri}%
-\homepage{http://ifsc.usp.br/~fabbri/}
-\email{fabbri@usp.br}
-\affiliation{ 
-S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP),
-PO Box 369, 13560-970, S\~ao Carlos, SP, Brazil %\\This line break forced with \textbackslash\textbackslash
-}
-%
-%\author{Vilson V. da Silva Jr.}
-%\homepage{http://automata.cc/}
-%\email{vilson@void.cc}
-%\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-%
-\author{Ricardo Fabbri}
-\homepage{http://www.lems.brown.edu/~rfabbri/}
-\email{rfabbri@iprj.uerj.br}
-\altaffiliation{
-Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
-}%Lines break automatically or can be forced with \\
+\title{Temporal stability in human interaction networks\tnoteref{mytitlenote}}
+\tnotetext[mytitlenote]{The Supporting Information document supplies thorough tables and figures.}
 
-\author{Deborah C. Antunes}
-\homepage{http://lattes.cnpq.br/1065956470701739}
-\email{deborahantunes@gmail.com}
-\altaffiliation{
-Curso de Psicologia, Universidade Federal do Cear\'a (UFC), Campus Sobral
-}%Lines break automatically or can be forced with \\
+%% Group authors per affiliation:
+\author[add1]{Renato Fabbri\corref{mycorrespondingauthor}}
+\address[add1]{S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP)}
+\ead{fabbri@usp.br}
 
-\author{Marilia M. Pisani}
-\homepage{http://lattes.cnpq.br/6738980149860322}
-\email{marilia.m.pisani@gmail.com}
-\altaffiliation{
-Centro de Ci\^encias Naturais e Humanas, Universidade Federal do ABC (CCNH/UFABC)
-}%Lines break automatically or can be forced with \\
+\author[add2]{Ricardo Fabbri}%
+\address[add2]{Polytechnic Institute, Rio de Janeiro State University (IPRJ/UERJ)}
+\ead{rfabbri@iprj.uerj.br}
 
-%
-%%\author{Luciano da Fontoura Costa}
-%%  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
-%%  \email{ldfcosta@gmail.com}
-%  \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
-%
-% \author{Leonardo Paulo Maia}
-% \homepage{http://www.ifsc.usp.br/~lpmaia/}
-% \email{lpmaia@ifsc.usp.br }
-% \altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+\author[add3]{Deborah Christina Antunes}
+\ead{deborahantunes@gmail.com}
+\address[add3]{Psychology Course, Federal University of Ceará (UFC), Sobral Campus}
 
-\author{Osvaldo N. Oliveira Jr.}
-\homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
-\email{chu@ifsc.usp.br}
-\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
+\author[add4]{Marilia Mello Pisani}
+\ead{marilia.m.pisani@gmail.com}
+\address[add4]{Center for Natural and Human Sciences, Federal University of ABC (CCNH/UFABC)}
+%% or include affiliations in footnotes:
 
-\date{\today}% It is always \today, today,
-%  but any date may be explicitly specified
+\author[add1]{Osvaldo Novais de Oliveira Junior}
+\ead{chu@ifsc.usp.br}
+
+\cortext[mycorrespondingauthor]{Corresponding author}
 
@@
 and at different scales. Our analysis shows that activity is practically the same for all networks across timescales ranging from seconds to months.
-% The most important metrics to the dispersion of participants in the topological measures space are related to centrality (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient.
 The principal components of the participants in the topological metrics space remain practically unchanged as different sets of messages are considered.
@@
 The relative sizes of these three sectors are essentially the same for all email lists and the same along time.
-Typically, <15\% of the vertices are hubs, 15-45\% are intermediary and >45\% are peripheral vertices.
+	Typically, $<15\%$ of the vertices are hubs, 15-45\% are intermediary and $>45\%$ are peripheral vertices.
 Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
@@
 
-\pacs{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
-\keywords{complex networks, pattern recognition, statistics, social network analysis, typology}
-% \keywords{complex networks, pattern recognition, statistics, social network analysis, anthropological physics, social psychology of big data, typology}
-\maketitle
+\begin{keyword}
+complex networks \sep pattern recognition \sep statistics \sep social network analysis \sep human typology
+% \MSC[2010] 00-01\sep  99-00
+\PACS{89.75.Fb,05.65.+b,89.65.-s}% PACS, the Physics and Astronomy
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
 
@@
 `The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.' 
-% `The conception of personality structure is the
-%best safeguard against the
-% inclination to attribute persistent trends in the
-% individual to something
-% "innate" or "basic" or "racial" within him. The
-% Nazi allegation that natural, biological traits decide the total being of a % person
-% would not have been such
-% a successful political device
-% had it not been possible to point to numerous
-% instances of relative fixity in human behavior and to
-% challenge those who
-% thought to explain them on any basis other than a biological one.'
 \emph{- Adorno et al, 1969, p. 747}
@@
 \end{table}
+
 The data and scripts used to derive the results, figures and tables, and this article itself are publicly available.
@@
 \centering
-\includegraphics[width=0.29\textwidth]{figs/criaRede3_}
+\includegraphics[width=0.4\textwidth]{figs/criaRede3_}
 \caption{The formation of interaction networks from exchanged messages.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figs/fser__}
+% \includegraphics[trim={0 0 0 1cm},clip,width=\textwidth]{figs/fser__}
+\includegraphics[width=\textwidth]{figs/fser__}
 \caption{Classification of vertices by comparing degree
@@
 %\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT__}
-\includegraphics[width=.45\textwidth]{figs/im13PCAPLOT__}
+\includegraphics[width=.75\textwidth]{figs/im13PCAPLOT__}
 \caption{The first plot highlights the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
@@
 
-\begin{acknowledgments}
+\subsection{Acknowledgments}
 Financial support was obtained from CNPq (140860/2013-4,
@@
 We are also grateful to developers and users of Python scientific tools, to Leonardo Paulo Maia (IFSC/USP) and to Francisco J. P. Lopes (UFRJ) for valuable insights.
-\end{acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\nocite{*}
-\bibliography{paper}% Produces the bibliography via BibTeX.
+
+\section*{References}
+
+\bibliography{../paper}
 

### 2017-03-17 physicaa/elsarticle-template.tex
adicionados titulos das figuras e tabelas nos captions e removido das figuras

@@
 \centering
-\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
+	\caption{{\bf Overview of the email lists analyzed.} Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
 $N$ is the number of participants (number of different email addresses),
@@
 \label{tab:genLists}
-\begin{tabular}{|l|c|c|c|c|c|}\hline
+\begin{tabular}{l|c|c|c|c|c}\hline
 list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
@@
 \includegraphics[width=0.4\textwidth]{figs/criaRede3_}
-\caption{The formation of interaction networks from exchanged messages.
-Each vertex represents a participant.
+	\caption{{\bf The formation of interaction networks from exchanged messages.} Each vertex represents a participant.
 A reply message from author B to a message from author A is regarded as evidence that B received information from A and yields a directed edge.
@@
 % \includegraphics[trim={0 0 0 1cm},clip,width=\textwidth]{figs/fser__}
-\includegraphics[width=\textwidth]{figs/fser__}
-\caption{Classification of vertices by comparing degree
-distributions~\cite{3setores}.
+\includegraphics[width=\textwidth]{figs/fser___}
+	\caption{{\bf Three sectors of a scale-free networks.} This is a classification of vertices by comparing degree distributions~\cite{3setores}.
 The binomial distribution of the Erd\"os-R\'enyi network model exhibits more intermediary vertices, while a scale-free network, associated with the power-law distribution, has more peripheral and hub vertices.
@@
 Figures and tables were usually inspected with 
-$ws=\{50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000, 10000\}$ messages. Variations in the number of vertices, edges
+$ws=\{50, 100, 200, 400, 500, 800,$ $1000, 2000, 2500, 5000, 10000\}$ messages. Variations in the number of vertices, edges
 and other network characteristics, within the same window size $ws$,
@@
 \begin{table}
-\caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$, described in Section~\ref{sec:mtime}, for different timescales.
+	\caption{{\bf Time-related circular statistics.} The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$, described in Section~\ref{sec:mtime}, for different timescales.
 This example table was constructed using all LAD messages, and the results are the same for other lists, as shown in Section~\ref*{si:circ} of the Supporting Information document.
@@
 \begin{center}
-\begin{tabular}{ |l|| c|c| }
+\begin{tabular}{ l|| c|c }
 \hline
@@
 \begin{table}
-\caption{Activity percentages along the hours of the day.
-Nearly identical distributions were observed on other social systems as shown in Section~\ref*{si:hours} of the Supporting Information document.
+	\caption{{\bf Activity percentages along the hours of the day.} Nearly identical distributions were observed on other social systems as shown in Section~\ref*{si:hours} of the Supporting Information document.
 Highest activity was observed between noon and 6pm (with 1/3 of total day activity), followed by the time period between 6pm and midnight.
@@
 \begin{table}
-\caption{Activity percentages along weekdays.
+	\caption{{\bf Activity percentages along weekdays.}
 Higher activity was observed during workweek days, with a decrease of activity on weekend days of at least one third and at most two thirds.}
 \begin{center}
-\begin{tabular}{ | l ||  c | c | c | c | c |   c | c |}
+\begin{tabular}{ l ||  c | c | c | c | c |   c | c}
 \hline
@@
 \begin{table}
-\caption{Activity along the days of the month cycle.
+	\caption{{\bf Activity along the days of the month.}
 Nearly identical distributions are found in all systems
@@
 \begin{table}
-\caption{Activity percentages on months along the year.
-	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays.
+	\caption{{\bf Activity percentages on months along the year.} 	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays.
 This table corresponds to activity in LAU.
@@
 \includegraphics[width=\textwidth]{figs/InText-WLAU-S1000__}
-\caption{Stability of Erd\"os sector sizes.
+	\caption{{\bf Stability of Erd\"os sector sizes.}
 Fractions of participants derived from degree and strength criteria, $E_1$ and $E_4$ described in Section~\ref{sectioning}, are both on the left.
@@
 \begin{table}[h]
-\caption{Distribution of activity among participants.
+	\caption{{\bf Distribution of activity among participants.}
 The first column shows the percentage of messages sent by the most active participant.
@@
 \begin{center}
-\begin{tabular}{ | l ||  c | c | c | c | }
+\begin{tabular}{  l ||  c | c | c | c  }
 \hline
@@
 \begin{table}[!h]
-\caption{Loadings for the 14 metrics into the principal components for the MET list, $1000$ messages in 20 disjoint positions.
+	\caption{{\bf Invariance of principal components.} Loadings for the 14 metrics into the principal components for the MET list, $1000$ messages in 20 disjoint positions.
 The clustering coefficient (cc) appears as the first metric in the table, followed by 7 centrality metrics and 6 symmetry-related metrics.
@@
 %\includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT__}
-\includegraphics[width=.75\textwidth]{figs/im13PCAPLOT__}
-\caption{The first plot highlights the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
+\includegraphics[width=.75\textwidth]{figs/im13PCAPLOT___}
+	\caption{{\bf Symmetry-related and clustering coefficient components along connectivity.} The first plot highlights the well-known pattern of degree versus clustering coefficient, characterized by the higher clustering coefficient of lower degree vertices.
     The second plot shows the greater dispersion of the symmetry-related ordinates dominant in the second principal component (PC2).
@@
 \centering
-\includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma2_}
-\caption{A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.
+% \includegraphics[trim={0 0 0 1cm},clip,width=.7\columnwidth]{figs/mpgamma2_}
+\includegraphics[width=.7\columnwidth]{figs/mpgamma2__}
+	\caption{{\bf Threads against participants and messages.} A scatter plot of number of messages $M$ versus number of participants $N$ versus number of threads $\Gamma$ for 140 email lists.
 Highest $\Gamma$ is associated with low $N$.

### 2017-03-18 physicaa/elsarticle-template.tex
adicionado arquivo final gerado pelo Physica A

@@
 \usepackage{xr}
-\externaldocument{../supportingInformation}
+\externaldocument{supportingInformation}
 
@@
 \ead{deborahantunes@gmail.com}
-\address[add3]{Psychology Course, Federal University of Ceará (UFC), Sobral Campus}
+\address[add3]{Psychology Course, Federal University of Cear\'a (UFC), Sobral Campus}
 
@@
 Typically, $<15\%$ of the vertices are hubs, 15-45\% are intermediary and $>45\%$ are peripheral vertices.
-Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participabr.
+Similar results for the distribution of participants in the three sectors and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and ParticipaBR.
 These properties are consistent with the literature and may be general for human interaction networks, which has important implications for establishing a typology of participants based on quantitative criteria.
@@
 Directed and weighted representations were built through the observation of replies as links.
-Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and Participabr were used for the sake of generalization.
+Interaction networks from email lists were the most convenient for deriving results and for benchmarking while networks from Facebook, Twitter and ParticipaBR were used for the sake of generalization.
 Using a timeline of activity snapshots with a constant number of contiguous messages, we found remarkable stability (or invariance) for important network properties.
@@
 Section~\ref{sec:results} reports results and discussion, leading to Section~\ref{sec:conc} for conclusions.
-Supplementary data analysis, including directions for video and sound mappings of network structures, and numeric detailed results for networks from Twitter, Facebook and Participabr, are provided in the Supporting Information document.
+Supplementary data analysis, including directions for video and sound mappings of network structures, and numeric detailed results for networks from Twitter, Facebook and ParticipaBR, are provided in the Supporting Information document.
 
@@
 
-We observed various email lists and selected four of them together with data from Twitter, Facebook and Participabr for a thorough analysis,
+We observed various email lists and selected four of them together with data from Twitter, Facebook and ParticipaBR for a thorough analysis,
 from which general properties can be inferred. These lists are as follows:
@@
 We considered 140 additional email lists to report on the interdependence between the number of participants and the number of discussion threads.
-Furthermore, 12 networks from Facebook (8), Twitter (2) and Participabr (2) were scrutinized, and their analysis is given in the Supporting Information document for the purpose of testing the generality of the results.
+Furthermore, 12 networks from Facebook (8), Twitter (2) and ParticipaBR (2) were scrutinized, and their analysis is given in the Supporting Information document for the purpose of testing the generality of the results.
 
@@
 \centering
+\label{tab:genLists}
+\begin{tabular}{l|c|c|c|c|c}\hline
+list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
+\input{tables/tab1Overview}
+\end{tabular}
 	\caption{{\bf Overview of the email lists analyzed.} Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
@@
 }
-\label{tab:genLists}
-\begin{tabular}{l|c|c|c|c|c}\hline
-list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
-\input{tables/tab1Overview}
-\end{tabular}
 \end{table}
@@
 Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}.
-Data from Participabr were used from the linked data/semantic web RDF triples~\cite{opa}, available in~\cite{datahub}.
+Data from ParticipaBR were used from the linked data/semantic web RDF triples~\cite{opa}, available in~\cite{datahub}.
 Computer scripts are delivered through a public domain Python PyPI package and an open Git repository~\cite{gmanePack}.
@@
 \begin{table}
+\begin{center}
+\begin{tabular}{ l|| c|c }
+\hline
+%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
+%& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
+scale & mean $\theta_\mu'$ & dispersion $\delta(z)$  \\ \hline
+\input{tables/tab2TimeLAD___}
+\end{tabular}
+\end{center}
 	\caption{{\bf Time-related circular statistics.} The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$, described in Section~\ref{sec:mtime}, for different timescales.
@@
 Note that $\theta_u'$ has the dimensional unit of the corresponding time period while $\delta(z)$ is dimensionless.}
-\begin{center}
-\begin{tabular}{ l|| c|c }
-\hline
-%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
-%& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
-scale & mean $\theta_\mu'$ & dispersion $\delta(z)$  \\ \hline
-\input{tables/tab2TimeLAD___}
-\end{tabular}
-\end{center}
 \label{tab:circ}
@@
 \begin{table}
+\footnotesize
+\input{tables/tabHoursCPP_}
+\label{tab:hin}
 	\caption{{\bf Activity percentages along the hours of the day.} Nearly identical distributions were observed on other social systems as shown in Section~\ref*{si:hours} of the Supporting Information document.
@@
 This table shows results for the activity in CPP.}
-\footnotesize
-\input{tables/tabHoursCPP_}
-\label{tab:hin}
 \end{table}
@@
 \begin{table}
-	\caption{{\bf Activity percentages along weekdays.}
-Higher activity was observed during workweek days, with a decrease of activity on weekend days of at least one third and at most two thirds.}
 \begin{center}
@@
 \end{center}
+	\caption{{\bf Activity percentages along weekdays.}
+Higher activity was observed during workweek days, with a decrease of activity on weekend days of at least one third and at most two thirds.}
 \label{tab:win}
@@
 \begin{table}
+\footnotesize
+\input{tables/tabMonthdaysMET}
+\label{tab:min}
 	\caption{{\bf Activity along the days of the month.}
@@
 This specific example and empirical table correspond to the activity of the MET email list.}
-\footnotesize
-\input{tables/tabMonthdaysMET}
-\label{tab:min}
 \end{table}
@@
 \begin{table}
+\footnotesize
+\input{tables/tabMonthsLAU}
+\label{tab:min2}
 	\caption{{\bf Activity percentages on months along the year.} 	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays.
@@
 Similar results are shown for other lists in Section~\ref*{si:months} of the Supporting Information document.}
-\footnotesize
-\input{tables/tabMonthsLAU}
-\label{tab:min2}
 \end{table}
@@
 In order to support the generality of these findings,
-we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and Participabr in Table~\ref*{tab:secE} of the Supporting Information document.
+we list the Erd\"os sector sizes of 12 networks from Facebook, Twitter and ParticipaBR in Table~\ref*{tab:secE} of the Supporting Information document.
 The fractions of hubs, intermediary and periphery nodes are
@@
 and, similarly, $\overline{c_{\delta,\phi}}=\frac{|c_{\delta,\phi}|}{N}$ denotes the fraction of participants in sector $\phi$ through criterion $C_\delta$.
-Sections~\ref*{si:frac} and~\ref*{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and Participabr.}
+Sections~\ref*{si:frac} and~\ref*{si:ext} of the Supporting Information bring a systematic collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}, with results for networks from Facebook, Twitter and ParticipaBR.}
 \label{fig:sectIL}
@@
 \begin{table}[h]
-	\caption{{\bf Distribution of activity among participants.}
-The first column shows the percentage of messages sent by the most active participant.
-The column for the first quartile ($Q_1$) gives the minimum percentage of participants responsible for at least 25\% of total messages with the actual percentage in parentheses.
-Similarly, the column for the first three quartiles $Q_3$ gives the minimum percentage of participants responsible for 75\% of total messages.
-The last decile $D_{-1}$ column shows the maximum percentage of participants responsible for 10\% of messages.}
 \begin{center}
@@
 \end{center}
+	\caption{{\bf Distribution of activity among participants.}
+The first column shows the percentage of messages sent by the most active participant.
+The column for the first quartile ($Q_1$) gives the minimum percentage of participants responsible for at least 25\% of total messages with the actual percentage in parentheses.
+Similarly, the column for the first three quartiles $Q_3$ gives the minimum percentage of participants responsible for 75\% of total messages.
+The last decile $D_{-1}$ column shows the maximum percentage of participants responsible for 10\% of messages.}
 \label{autores}
@@
 \begin{table}[!h]
+\footnotesize
+\input{tables/tabPCA3CPP}
+\label{tab:pcain}
 	\caption{{\bf Invariance of principal components.} Loadings for the 14 metrics into the principal components for the MET list, $1000$ messages in 20 disjoint positions.
@@
 The low standard deviation $\sigma$ implies that the principal components are considerably stable.}
-\footnotesize
-\input{tables/tabPCA3CPP}
-\label{tab:pcain}
 \end{table}
@@
 networks from Facebook,
-Twitter and Participabr in the Supporting Information might ease hypothesizing
+Twitter and ParticipaBR in the Supporting Information might ease hypothesizing
 about the generality of these characteristics.
@@
 project 870336/1997-5), United Nations Development Program (contract: 2013/000566; project BRA/12/018) and FAPESP. 
-The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Presidency of the Brazilian Republic for keeping Participabr code and data open.
+The authors are grateful to the American Jewish Committee for maintaining an online copy of the Adorno book used on the epigraph~\cite{adorno}, to Gmane creators and maintainers for the public email list data, to the communities of the email lists and other groups used in the analysis, and to the Presidency of the Brazilian Republic for keeping ParticipaBR code and data open.
 We are also grateful to developers and users of Python scientific tools, to Leonardo Paulo Maia (IFSC/USP) and to Francisco J. P. Lopes (UFRJ) for valuable insights.
@@
 
-\bibliography{../paper}
+\bibliography{paper}
 
